---
title: "Cardiac and TOP2 Genes - Log2CPM Boxplots"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## **📌 Log2CPM Boxplots for Cardiac and TOP2 Genes**

This analysis generates **boxplots for cardiac genes and TOP2 genes** across different treatments and timepoints.

---

## **📌 Load Required Libraries**
```{r load_libraries, echo=TRUE, message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(org.Hs.eg.db)
library(clusterProfiler)
```

## **📌 Read Log2CPM Data**
```{r load_File Paths, echo=TRUE,results='hide',message=FALSE}
# Load feature count matrix
boxplot1 <- read.csv("data/Feature_count_Matrix_Log2CPM_filtered.csv") %>% as.data.frame()

# Ensure column names are cleaned
colnames(boxplot1) <- trimws(gsub("^X", "", colnames(boxplot1)))  
```


## **📌 Define Genes of Interest**
```{r Genes of Interest, echo=TRUE, message=FALSE}
# Define the genes of interest
top2_genes <- c("TOP2A", "TOP2B")
cardiac_genes <- c("ACTN2", "CALR", "MYBPC3", "MYH6", "MYH7", 
                   "MYL2", "RYR2", "SCN5A", "TNNI3", "TNNT2", "TTN")
dna_damage_genes <- c("TP53")  # Using correct gene symbol TP53
```


## **📌 Read and Process DEGs Data**
```{r load_Feature Count Matrix, echo=TRUE,results='hide',message=FALSE}
# Load Toptables
deg_files <- list.files("data/DEGs", pattern = "Toptable_.*\\.csv", full.names = TRUE)
deg_list <- lapply(deg_files, read.csv)
names(deg_list) <- gsub("data/DEGs/Toptable_|\\.csv", "", deg_files)  

# Function to check significance based on **Entrez_ID in the correct sample**
is_significant <- function(gene, drug, conc, timepoint) {
  condition <- paste(drug, conc, timepoint, sep = "_")
  if (!condition %in% names(deg_list)) return(FALSE)
  
  toptable <- deg_list[[condition]]
  gene_entrez <- boxplot1$ENTREZID[boxplot1$SYMBOL == gene]
  
  if (length(gene_entrez) == 0) return(FALSE)
  
  return(any(gene_entrez %in% toptable$Entrez_ID[toptable$adj.P.Val < 0.05]))
}
```


## **📌 Process Data for Plotting**
```{r Data for Plotting, echo=TRUE, message=FALSE}
process_gene_data <- function(gene) {
  # Filter log2CPM data for the gene
  gene_data <- boxplot1 %>% filter(SYMBOL == gene)
  
  # Reshape data
  long_data <- gene_data %>%
    pivot_longer(cols = -c(ENTREZID, SYMBOL, GENENAME), names_to = "Sample", values_to = "log2CPM") %>%
    mutate(
      Indv = case_when(
        grepl("75.1", Sample) ~ "1",
        grepl("78.1", Sample) ~ "2",
        grepl("87.1", Sample) ~ "3",
        grepl("17.3", Sample) ~ "4",
        grepl("84.1", Sample) ~ "5",
        grepl("90.1", Sample) ~ "6",
        TRUE ~ NA_character_
      ),
      Drug = case_when(
        grepl("CX.5461", Sample) ~ "CX",
        grepl("DOX", Sample) ~ "DOX",
        grepl("VEH", Sample) ~ "VEH",
        TRUE ~ NA_character_
      ),
      Conc. = case_when(
        grepl("_0.1_", Sample) ~ "0.1",
        grepl("_0.5_", Sample) ~ "0.5",
        TRUE ~ NA_character_
      ),
      Timepoint = case_when(
        grepl("_3$", Sample) ~ "3",
        grepl("_24$", Sample) ~ "24",
        grepl("_48$", Sample) ~ "48",
        TRUE ~ NA_character_
      ),
      Condition = paste(Drug, Conc., Timepoint, sep = "_")
    )

  # **Ensure Condition is Ordered Correctly**
  long_data$Condition <- factor(
    long_data$Condition, 
    levels = c(
      "CX_0.1_3", "CX_0.1_24", "CX_0.1_48", "CX_0.5_3", "CX_0.5_24", "CX_0.5_48",
      "DOX_0.1_3", "DOX_0.1_24", "DOX_0.1_48", "DOX_0.5_3", "DOX_0.5_24", "DOX_0.5_48",
      "VEH_0.1_3", "VEH_0.1_24", "VEH_0.1_48", "VEH_0.5_3", "VEH_0.5_24", "VEH_0.5_48"
    )
  )
  
  # Identify significant conditions **per Drug, Conc, and Timepoint**
  significance_labels <- long_data %>%
    distinct(Drug, Conc., Timepoint, Condition) %>%
    rowwise() %>%
    mutate(
      max_log2CPM = max(long_data$log2CPM[long_data$Condition == Condition], na.rm = TRUE),
      Significance = ifelse(is_significant(gene, Drug, Conc., Timepoint), "*", "")
    ) %>%
    filter(Significance != "") %>% ungroup()
  
  list(long_data = long_data, significance_labels = significance_labels)
}
```


## **📌Generate Boxplots for Cardiac Genes**
```{r Cardiac, echo=TRUE,message=FALSE}
for (gene in cardiac_genes) {
  data_info <- process_gene_data(gene)
  p <- ggplot(data_info$long_data, aes(x = Condition, y = log2CPM, fill = Drug)) +
    geom_boxplot(outlier.shape = NA) +
    scale_fill_manual(values = c("CX" = "#0000FF", "DOX" = "#e6d800", "VEH" = "#FF00FF")) +
    geom_point(aes(color = Indv), size = 2, alpha = 0.5, position = position_jitter(width = -1, height = 0)) +
    geom_text(data = data_info$significance_labels, aes(x = Condition, y = max_log2CPM + 0.5, label = Significance),
              inherit.aes = FALSE, size = 6, color = "black") +
    ggtitle(paste("Log2CPM Expression of", gene)) +
    labs(x = "Treatment", y = "log2CPM") +
    theme_bw() +
    theme(plot.title = element_text(size = rel(2), hjust = 0.5),
          axis.title = element_text(size = 15, color = "black"),
          axis.text.x = element_text(size = 10, color = "black", angle = 90, hjust = 1))
  
  print(p)
}
```


## **📌Generate Boxplots for TOP2 Genes**
```{r TOP2, echo=TRUE,message=FALSE}
for (gene in top2_genes) {
  data_info <- process_gene_data(gene)
  p <- ggplot(data_info$long_data, aes(x = Condition, y = log2CPM, fill = Drug)) +
    geom_boxplot(outlier.shape = NA) +
    scale_fill_manual(values = c("CX" = "#0000FF", "DOX" = "#e6d800", "VEH" = "#FF00FF")) +
    geom_point(aes(color = Indv), size = 2, alpha = 0.5, position = position_jitter(width = -1, height = 0)) +
    geom_text(data = data_info$significance_labels, aes(x = Condition, y = max_log2CPM + 0.5, label = Significance),
              inherit.aes = FALSE, size = 6, color = "black") +
    ggtitle(paste("Log2CPM Expression of", gene)) +
    labs(x = "Treatment", y = "log2CPM") +
    theme_bw() +
    theme(plot.title = element_text(size = rel(2), hjust = 0.5),
          axis.title = element_text(size = 15, color = "black"),
          axis.text.x = element_text(size = 10, color = "black", angle = 90, hjust = 1))
  
  print(p)
}
```

## **📌 Generate Boxplots for DNA Damage Genes**
```{r DNA Damage, echo=TRUE,message=FALSE}
for (gene in dna_damage_genes) {
  data_info <- process_gene_data(gene)
  p <- ggplot(data_info$long_data, aes(x = Condition, y = log2CPM, fill = Drug)) +
    geom_boxplot(outlier.shape = NA) +
    scale_fill_manual(values = c("CX" = "#0000FF", "DOX" = "#e6d800", "VEH" = "#FF00FF")) +
    geom_point(aes(color = Indv), size = 2, alpha = 0.5, position = position_jitter(width = -1, height = 0)) +
    geom_text(data = data_info$significance_labels, aes(x = Condition, y = max_log2CPM + 0.5, label = Significance),
              inherit.aes = FALSE, size = 6, color = "black") +
    ggtitle(paste("Log2CPM Expression of", gene)) +
    labs(x = "Treatment", y = "log2CPM") +
    theme_bw() +
    theme(plot.title = element_text(size = rel(2), hjust = 0.5),
          axis.title = element_text(size = 15, color = "black"),
          axis.text.x = element_text(size = 10, color = "black", angle = 90, hjust = 1))
  
  print(p)
}
```
##**📌 DNA Damage Proportion**

## **📌 Read and Process DEG Data**
```{r load_DEGs, echo=TRUE, results='hide', message=FALSE, warning=FALSE}
# Load DEGs Data
CX_0.1_3 <- read.csv("data/DEGs/Toptable_CX_0.1_3.csv")
CX_0.1_24 <- read.csv("data/DEGs/Toptable_CX_0.1_24.csv")
CX_0.1_48 <- read.csv("data/DEGs/Toptable_CX_0.1_48.csv")
CX_0.5_3 <- read.csv("data/DEGs/Toptable_CX_0.5_3.csv")
CX_0.5_24 <- read.csv("data/DEGs/Toptable_CX_0.5_24.csv")
CX_0.5_48 <- read.csv("data/DEGs/Toptable_CX_0.5_48.csv")

DOX_0.1_3 <- read.csv("data/DEGs/Toptable_DOX_0.1_3.csv")
DOX_0.1_24 <- read.csv("data/DEGs/Toptable_DOX_0.1_24.csv")
DOX_0.1_48 <- read.csv("data/DEGs/Toptable_DOX_0.1_48.csv")
DOX_0.5_3 <- read.csv("data/DEGs/Toptable_DOX_0.5_3.csv")
DOX_0.5_24 <- read.csv("data/DEGs/Toptable_DOX_0.5_24.csv")
DOX_0.5_48 <- read.csv("data/DEGs/Toptable_DOX_0.5_48.csv")

# Extract Significant DEGs
DEGs <- list(
  "CX_0.1_3" = CX_0.1_3$Entrez_ID[CX_0.1_3$adj.P.Val < 0.05],
  "CX_0.1_24" = CX_0.1_24$Entrez_ID[CX_0.1_24$adj.P.Val < 0.05],
  "CX_0.1_48" = CX_0.1_48$Entrez_ID[CX_0.1_48$adj.P.Val < 0.05],
  "CX_0.5_3" = CX_0.5_3$Entrez_ID[CX_0.5_3$adj.P.Val < 0.05],
  "CX_0.5_24" = CX_0.5_24$Entrez_ID[CX_0.5_24$adj.P.Val < 0.05],
  "CX_0.5_48" = CX_0.5_48$Entrez_ID[CX_0.5_48$adj.P.Val < 0.05],
  "DOX_0.1_3" = DOX_0.1_3$Entrez_ID[DOX_0.1_3$adj.P.Val < 0.05],
  "DOX_0.1_24" = DOX_0.1_24$Entrez_ID[DOX_0.1_24$adj.P.Val < 0.05],
  "DOX_0.1_48" = DOX_0.1_48$Entrez_ID[DOX_0.1_48$adj.P.Val < 0.05],
  "DOX_0.5_3" = DOX_0.5_3$Entrez_ID[DOX_0.5_3$adj.P.Val < 0.05],
  "DOX_0.5_24" = DOX_0.5_24$Entrez_ID[DOX_0.5_24$adj.P.Val < 0.05],
  "DOX_0.5_48" = DOX_0.5_48$Entrez_ID[DOX_0.5_48$adj.P.Val < 0.05]
)

# Extract Significant DEGs
DEG1 <- CX_0.1_3$Entrez_ID[CX_0.1_3$adj.P.Val < 0.05]
DEG2 <- CX_0.1_24$Entrez_ID[CX_0.1_24$adj.P.Val < 0.05]
DEG3 <- CX_0.1_48$Entrez_ID[CX_0.1_48$adj.P.Val < 0.05]
DEG4 <- CX_0.5_3$Entrez_ID[CX_0.5_3$adj.P.Val < 0.05]
DEG5 <- CX_0.5_24$Entrez_ID[CX_0.5_24$adj.P.Val < 0.05]
DEG6 <- CX_0.5_48$Entrez_ID[CX_0.5_48$adj.P.Val < 0.05]

DEG7 <- DOX_0.1_3$Entrez_ID[DOX_0.1_3$adj.P.Val < 0.05]
DEG8 <- DOX_0.1_24$Entrez_ID[DOX_0.1_24$adj.P.Val < 0.05]
DEG9 <- DOX_0.1_48$Entrez_ID[DOX_0.1_48$adj.P.Val < 0.05]
DEG10 <- DOX_0.5_3$Entrez_ID[DOX_0.5_3$adj.P.Val < 0.05]
DEG11 <- DOX_0.5_24$Entrez_ID[DOX_0.5_24$adj.P.Val < 0.05]
DEG12 <- DOX_0.5_48$Entrez_ID[DOX_0.5_48$adj.P.Val < 0.05]
```

## **📌 DNA Damage Repair Proportion with Chi-Square Test**
```{r DNA_Damage_Proportion, echo=TRUE, results='hide', message=FALSE, warning=FALSE}
# Load DNA Damage Genes List
DNA_damage <- read.csv("data/DNA_Damage.csv", stringsAsFactors = FALSE)
DNA_damage$Entrez_ID <- mapIds(org.Hs.eg.db,
                               keys = DNA_damage$Symbol,
                               column = "ENTREZID",
                               keytype = "SYMBOL",
                               multiVals = "first")

# Define CX-5461 DEG lists
CX_DEGs <- list(
  "CX_0.1_3" = DEG1, "CX_0.1_24" = DEG2, "CX_0.1_48" = DEG3,
  "CX_0.5_3" = DEG4, "CX_0.5_24" = DEG5, "CX_0.5_48" = DEG6
)

# Define DOX DEG lists
DOX_DEGs <- list(
  "DOX_0.1_3" = DEG7, "DOX_0.1_24" = DEG8, "DOX_0.1_48" = DEG9,
  "DOX_0.5_3" = DEG10, "DOX_0.5_24" = DEG11, "DOX_0.5_48" = DEG12
)

# Extract Entrez_IDs from DNA Damage gene dataset
DNA_damage_genes <- na.omit(DNA_damage$Entrez_ID)

# Combine CX-5461 DEGs into a dataframe with a "Drug" column
CX_DEGs_df <- bind_rows(
  lapply(CX_DEGs, function(ids) data.frame(Entrez_ID = ids, Drug = "CX-5461")),
  .id = "Sample"
)

# Combine DOX DEGs into a dataframe with a "Drug" column
DOX_DEGs_df <- bind_rows(
  lapply(DOX_DEGs, function(ids) data.frame(Entrez_ID = ids, Drug = "DOX")),
  .id = "Sample"
)

# Merge CX-5461 and DOX datasets
DEGs_df <- bind_rows(CX_DEGs_df, DOX_DEGs_df)

# Check if genes are in DNA Damage list
DEGs_df <- DEGs_df %>%
  mutate(Category = ifelse(Entrez_ID %in% DNA_damage_genes, "Yes", "No"))

# Count DNA damage genes in each sample
proportion_data <- DEGs_df %>%
  group_by(Sample, Drug, Category) %>%
  summarise(Count = n(), .groups = "drop") %>%
  group_by(Sample, Drug) %>%
  mutate(Percentage = (Count / sum(Count)) * 100)

# Normalize Percentages to Sum Exactly 100%
proportion_data <- proportion_data %>%
  group_by(Sample) %>%
  mutate(Percentage = round(Percentage, 2)) %>%
  mutate(Adjustment = 100 - sum(Percentage, na.rm = TRUE)) %>%
  mutate(Percentage = ifelse(Category == "No", Percentage + Adjustment, Percentage)) %>%
  mutate(Percentage = ifelse(Percentage < 0, 0, Percentage)) %>%
  mutate(Percentage = ifelse(Percentage > 100, 100, Percentage)) %>%
  ungroup() %>%
  replace_na(list(Percentage = 0))

# Ensure "Yes" is at the Bottom and "No" is at the Top
proportion_data$Category <- factor(proportion_data$Category, levels = c("Yes", "No"))

# **🔹 Maintain Correct X-Axis Order (3 → 24 → 48)**
sample_order <- c(
  "CX_0.1_3", "CX_0.1_24", "CX_0.1_48", 
  "CX_0.5_3", "CX_0.5_24", "CX_0.5_48",
  "DOX_0.1_3", "DOX_0.1_24", "DOX_0.1_48", 
  "DOX_0.5_3", "DOX_0.5_24", "DOX_0.5_48"
)
proportion_data$Sample <- factor(proportion_data$Sample, levels = sample_order, ordered = TRUE)

# **Perform Chi-Square Test for CX vs DOX Pairs**
chi_square_results <- data.frame(Sample = character(), P_Value = numeric())

for (i in seq(1, 6)) {  # Pairwise comparison (CX vs DOX)
  cx_sample <- sample_order[i]
  dox_sample <- sample_order[i + 6]  # Correctly pairs CX_0.1_3 with DOX_0.1_3, etc.
  
  cx_data <- filter(proportion_data, Sample == cx_sample)
  dox_data <- filter(proportion_data, Sample == dox_sample)
  
  # Construct contingency table for Chi-Square test
  contingency_table <- matrix(
    c(sum(cx_data$Count[cx_data$Category == "Yes"]), sum(cx_data$Count[cx_data$Category == "No"]),
      sum(dox_data$Count[dox_data$Category == "Yes"]), sum(dox_data$Count[dox_data$Category == "No"])),
    nrow = 2, byrow = TRUE
  )
  
  # Run Chi-Square Test
  test_result <- chisq.test(contingency_table)
  p_value <- test_result$p.value
  
  # Store results
  chi_square_results <- rbind(chi_square_results, data.frame(Sample = cx_sample, P_Value = p_value))
}

# Identify significant CX samples (p < 0.05)
chi_square_results$Significant <- ifelse(chi_square_results$P_Value < 0.05, "*", "")

# **🔹 Merge Chi-Square Results WITHOUT Modifying Order**
proportion_data <- left_join(proportion_data, chi_square_results, by = "Sample")

# **Reapply Factor Order to Prevent Changes**
proportion_data$Sample <- factor(proportion_data$Sample, levels = sample_order, ordered = TRUE)
```


## **📌 DNA Damage Repair Proportion Plot**
```{r ggplot, echo=TRUE, message=FALSE}

# **Generate Proportion Plot for CX-5461 and DOX**
ggplot(proportion_data, aes(x = Sample, y = Percentage, fill = Category)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(data = subset(proportion_data, Significant == "*"),
            aes(x = Sample, y = 102, label = "*"),
            size = 6, color = "black", fontface = "bold") +
  scale_y_continuous(labels = scales::percent_format(scale = 1), limits = c(0, 105)) +
  scale_fill_manual(values = c("Yes" = "#e41a1c", "No" = "#377eb8")) +
  labs(
    title = "Proportion of DNA Damage Repair Genes in\nCX-5461 and DOX DEGs with Significance",
    x = "Samples (CX-5461 and DOX)",
    y = "Percentage",
    fill = "Category"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = rel(1.5), hjust = 0.5),
    axis.title = element_text(size = 15, color = "black"),
    axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
    legend.title = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1.2),
    strip.background = element_blank(),
    strip.text = element_text(size = 12, face = "bold")
  )
```



