---
title: "Colorectal cancer dataset comparison"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## **📌 Proportion of Colorectal cancer DE genes in Corrmotif clusters**

## **📌 Load Required Libraries**
```{r load_libraries, echo=TRUE, message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(org.Hs.eg.db)
library(clusterProfiler)
library(biomaRt)
```

## **📌 Load Data**
```{r load_File Paths, echo=TRUE,results='hide',message=FALSE}
# Define correct timepoints
timepoints <- c("6hr", "24hr", "72hr")

# Initialize an empty dataframe to store chi-square results
all_chi_results <- data.frame()

# Loop through each timepoint and perform Chi-square test
for (time in timepoints) {
  
  # Load proportion data for the given timepoint
  file_path <- paste0("data/Colorectal/Proportion_data_", time, ".csv")
  
  # Check if the file exists before reading
  if (!file.exists(file_path)) {
    cat("\n🚨 Warning: File does not exist:", file_path, "\nSkipping this timepoint...\n")
    next  # Skip this iteration if the file is missing
  }
  
  proportion_data <- read.csv(file_path)
  
  # Extract counts for Non-Response group
  non_response_counts <- proportion_data %>%
    filter(Set == "Non Response") %>%
    dplyr::select(DEG, `Non.DEG`) %>%
    unlist(use.names = FALSE)  # Convert to numeric vector
  
  # Debugging: Print Non-Response counts
  cat("\nNon-Response Counts for", time, ":\n")
  print(non_response_counts)
  
  # Initialize a list to store chi-square test results for this timepoint
  chi_results <- list()
  
  # Perform chi-square test for each response group
  for (group in unique(proportion_data$Set)) {
    if (group == "Non Response") next  # Skip Non Response group
    
    # Extract counts for the current response group
    group_counts <- proportion_data %>%
      filter(Set == group) %>%
      dplyr::select(DEG, `Non.DEG`) %>%
      unlist(use.names = FALSE)  # Convert to numeric vector
    
    # Ensure valid counts for chi-square test
    if (length(group_counts) < 2) group_counts <- c(group_counts, 0)
    if (length(non_response_counts) < 2) non_response_counts <- c(non_response_counts, 0)
    
    # Create contingency table
    contingency_table <- matrix(c(
      group_counts[1], group_counts[2],  # Current response group counts
      non_response_counts[1], non_response_counts[2]  # Non-Response counts
    ), nrow = 2, byrow = TRUE)
    
    # Debugging: Print contingency table
    cat("\nProcessing Group:", group, "at", time, "\n")
    cat("Contingency Table:\n")
    print(contingency_table)
    
    # Perform chi-square test
    test_result <- chisq.test(contingency_table)
    p_value <- test_result$p.value
    significance <- ifelse(p_value < 0.05, "*", "")  # Mark * for p < 0.05
    
    # Store results
    chi_results[[group]] <- data.frame(
      Set = group,
      Timepoint = time,
      Chi2 = test_result$statistic,
      p_value = p_value,
      Significance = significance
    )
  }
  
  # Combine results for this timepoint into a single dataframe
  chi_results <- do.call(rbind, chi_results)
  
  # Append to the overall results dataframe
  all_chi_results <- rbind(all_chi_results, chi_results)
}

# Save final chi-square results
write.csv(all_chi_results, "data/Colorectal//Chi_Square_Results_All.csv", row.names = FALSE)

```

## **📌 Proportion of DE genes across response groups**
```{r load_data_tissue, echo=TRUE,results='hide', message=FALSE, fig.width=14,fig.height=8}

# Load the saved datasets
prob_all_1 <- read.csv("data/prob_all_1.csv")$Entrez_ID
prob_all_2 <- read.csv("data/prob_all_2.csv")$Entrez_ID
prob_all_3 <- read.csv("data/prob_all_3.csv")$Entrez_ID
prob_all_4 <- read.csv("data/prob_all_4.csv")$Entrez_ID

# Example Response Groups Data (Replace with actual data)
response_groups <- list(
  "Non Response" = prob_all_1, # Replace 'prob_all_1', 'prob_all_2', etc. with your actual response group dataframes
  "CX_DOX Shared Late Response" = prob_all_2,
  "DOX-Specific Response" = prob_all_3,
  "Late High Dose DOX-Specific Response" = prob_all_4
)

# Load the proportion data again for visualization
proportion_data <-read.csv("data/Colorectal/Proportion_data.csv")
# Merge chi-square results into proportion data for plotting
proportion_data <- proportion_data %>%
  left_join(all_chi_results %>% dplyr::select(Set, Timepoint, Significance), by = c("Set", "Timepoint"))

# Convert to factors for ordered display
proportion_data$Set <- factor(proportion_data$Set, levels = c(
  "Non Response",
  "CX_DOX Shared Late Response",
  "DOX-Specific Response",
  "Late High Dose DOX-Specific Response"
))
proportion_data$Timepoint <- factor(proportion_data$Timepoint, levels = timepoints)

# Plot proportions with significance stars
ggplot(proportion_data, aes(x = Set, y = Percentage, fill = Category)) +
  geom_bar(stat = "identity", position = "stack") +
  facet_wrap(~Timepoint, scales = "fixed") +
  scale_fill_manual(values = c("DEG" = "#e41a1c", "Non-DEG" = "#377eb8")) +
  geom_text(
    data = proportion_data %>% filter(Significance == "*") %>% distinct(Set, Timepoint, Significance),
    aes(x = Set, y = 105, label = Significance),
    inherit.aes = FALSE,
    size = 6,
    color = "black",
    hjust = 0.5
  ) +
  labs(
    title = "Proportion of Colorectal cancer DEGs and Non-DEGs\nAcross Response Groups (6hr, 24hr, 72hr)",
    x = "Response Groups",
    y = "Percentage",
    fill = "Category"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = rel(1.5), hjust = 0.5),
    axis.title = element_text(size = 15, color = "black"),
    axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
    legend.title = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, size = 1.2),
    strip.background = element_blank(),
    strip.text = element_text(size = 12, face = "bold")
  )
```

## **📌 Proportion of colorectal cancer genes in CX and DOX DEGs**

## **📌 Read and Process Data**
```{r load_DEGs, echo=TRUE, results='hide', message=FALSE, warning=FALSE}
# Load DEGs Data
CX_0.1_3 <- read.csv("data/DEGs/Toptable_CX_0.1_3.csv")
CX_0.1_24 <- read.csv("data/DEGs/Toptable_CX_0.1_24.csv")
CX_0.1_48 <- read.csv("data/DEGs/Toptable_CX_0.1_48.csv")
CX_0.5_3 <- read.csv("data/DEGs/Toptable_CX_0.5_3.csv")
CX_0.5_24 <- read.csv("data/DEGs/Toptable_CX_0.5_24.csv")
CX_0.5_48 <- read.csv("data/DEGs/Toptable_CX_0.5_48.csv")

DOX_0.1_3 <- read.csv("data/DEGs/Toptable_DOX_0.1_3.csv")
DOX_0.1_24 <- read.csv("data/DEGs/Toptable_DOX_0.1_24.csv")
DOX_0.1_48 <- read.csv("data/DEGs/Toptable_DOX_0.1_48.csv")
DOX_0.5_3 <- read.csv("data/DEGs/Toptable_DOX_0.5_3.csv")
DOX_0.5_24 <- read.csv("data/DEGs/Toptable_DOX_0.5_24.csv")
DOX_0.5_48 <- read.csv("data/DEGs/Toptable_DOX_0.5_48.csv")

# Extract Significant DEGs
DEG1 <- as.character(CX_0.1_3$Entrez_ID[CX_0.1_3$adj.P.Val < 0.05])
DEG2 <- as.character(CX_0.1_24$Entrez_ID[CX_0.1_24$adj.P.Val < 0.05])
DEG3 <- as.character(CX_0.1_48$Entrez_ID[CX_0.1_48$adj.P.Val < 0.05])
DEG4 <- as.character(CX_0.5_3$Entrez_ID[CX_0.5_3$adj.P.Val < 0.05])
DEG5 <- as.character(CX_0.5_24$Entrez_ID[CX_0.5_24$adj.P.Val < 0.05])
DEG6 <- as.character(CX_0.5_48$Entrez_ID[CX_0.5_48$adj.P.Val < 0.05])
DEG7 <- as.character(DOX_0.1_3$Entrez_ID[DOX_0.1_3$adj.P.Val < 0.05])
DEG8 <- as.character(DOX_0.1_24$Entrez_ID[DOX_0.1_24$adj.P.Val < 0.05])
DEG9 <- as.character(DOX_0.1_48$Entrez_ID[DOX_0.1_48$adj.P.Val < 0.05])
DEG10 <- as.character(DOX_0.5_3$Entrez_ID[DOX_0.5_3$adj.P.Val < 0.05])
DEG11 <- as.character(DOX_0.5_24$Entrez_ID[DOX_0.5_24$adj.P.Val < 0.05])
DEG12 <- as.character(DOX_0.5_48$Entrez_ID[DOX_0.5_48$adj.P.Val < 0.05])
```

## **📌 Proportion of colorectal cancer genes in CX and DOX DEGs datasets**
```{r proportion1, echo=TRUE,results='hide', message=FALSE, fig.width=14,fig.height=8}

data_6hr <- read.csv("data/Colorectal/6hr.csv", stringsAsFactors = FALSE)
data_24hr <- read.csv("data/Colorectal/24hr.csv", stringsAsFactors = FALSE)
data_72hr <- read.csv("data/Colorectal/72hr.csv", stringsAsFactors = FALSE)

# Map gene symbols to Entrez IDs using org.Hs.eg.db
data_6hr <- data_6hr %>%
  mutate(Entrez_ID = mapIds(org.Hs.eg.db,
                            keys = Symbol,
                            column = "ENTREZID",
                            keytype = "SYMBOL",
                            multiVals = "first"))

data_24hr <- data_24hr %>%
  mutate(Entrez_ID = mapIds(org.Hs.eg.db,
                            keys = Symbol,
                            column = "ENTREZID",
                            keytype = "SYMBOL",
                            multiVals = "first"))

data_72hr <- data_72hr %>%
  mutate(Entrez_ID = mapIds(org.Hs.eg.db,
                            keys = Symbol,
                            column = "ENTREZID",
                            keytype = "SYMBOL",
                            multiVals = "first"))

# Define CX-5461 DEG lists
CX_DEGs <- list(
  "CX_0.1_3" = DEG1, "CX_0.1_24" = DEG2, "CX_0.1_48" = DEG3,
  "CX_0.5_3" = DEG4, "CX_0.5_24" = DEG5, "CX_0.5_48" = DEG6
)

# Define DOX DEG lists
DOX_DEGs <- list(
  "DOX_0.1_3" = DEG7, "DOX_0.1_24" = DEG8, "DOX_0.1_48" = DEG9,
  "DOX_0.5_3" = DEG10, "DOX_0.5_24" = DEG11, "DOX_0.5_48" = DEG12
)

# Extract Entrez_IDs from cancer gene datasets for 6hr, 24hr, and 72hr
lit_genes <- list(
  "6hr" = na.omit(data_6hr$Entrez_ID),
  "24hr" = na.omit(data_24hr$Entrez_ID),
  "72hr" = na.omit(data_72hr$Entrez_ID)
)

# Combine CX-5461 DEGs into a dataframe with a "Drug" column
CX_DEGs_df <- bind_rows(
  lapply(CX_DEGs, function(ids) data.frame(Entrez_ID = ids, Drug = "CX-5461")),
  .id = "Sample"
)

# Combine DOX DEGs into a dataframe with a "Drug" column
DOX_DEGs_df <- bind_rows(
  lapply(DOX_DEGs, function(ids) data.frame(Entrez_ID = ids, Drug = "DOX")),
  .id = "Sample"
)

# Merge CX-5461 and DOX datasets
DEGs_df <- bind_rows(CX_DEGs_df, DOX_DEGs_df)

# Match Entrez_IDs with DEGs from 6hr, 24hr, and 72hr datasets
DEGs_df <- DEGs_df %>%
  mutate(
    Category_6hr = ifelse(Entrez_ID %in% lit_genes[["6hr"]], "Yes", "No"),
    Category_24hr = ifelse(Entrez_ID %in% lit_genes[["24hr"]], "Yes", "No"),
    Category_72hr = ifelse(Entrez_ID %in% lit_genes[["72hr"]], "Yes", "No")
  )

# Convert to long format for ggplot
proportion_data <- DEGs_df %>%
  pivot_longer(cols = starts_with("Category"), names_to = "Timepoint", values_to = "Category") %>%
  group_by(Sample, Drug, Timepoint, Category) %>%
  summarise(Count = n(), .groups = "drop") %>%
  group_by(Sample, Drug, Timepoint) %>%
  mutate(Percentage = (Count / sum(Count)) * 100)

# **Fix: Normalize Percentages to Sum Exactly 100%**
proportion_data <- proportion_data %>%
  group_by(Sample, Timepoint) %>%
  mutate(Percentage = round(Percentage, 2)) %>%  # Round each percentage
  mutate(Adjustment = 100 - sum(Percentage, na.rm = TRUE)) %>%  # Compute rounding difference
  mutate(Percentage = ifelse(Category == "No", Percentage + Adjustment, Percentage)) %>%  # Apply adjustment to "No"
  mutate(Percentage = ifelse(Percentage < 0, 0, Percentage)) %>%  # Prevent negatives
  mutate(Percentage = ifelse(Percentage > 100, 100, Percentage)) %>%  # Prevent values > 100
  ungroup() %>%  # Remove grouping to avoid select() errors
  replace_na(list(Percentage = 0))  # Ensure no NA values

# **Ensure "Yes" is at the Bottom and "No" is at the Top**
proportion_data$Category <- factor(proportion_data$Category, levels = c("Yes", "No"))

# Define correct factor orders for samples and timepoints
sample_order <- c(
  "CX_0.1_3", "CX_0.1_24", "CX_0.1_48", "CX_0.5_3", "CX_0.5_24", "CX_0.5_48",
  "DOX_0.1_3", "DOX_0.1_24", "DOX_0.1_48", "DOX_0.5_3", "DOX_0.5_24", "DOX_0.5_48"
)
proportion_data$Sample <- factor(proportion_data$Sample, levels = sample_order)

proportion_data$Timepoint <- factor(
  proportion_data$Timepoint,
  levels = c("Category_6hr", "Category_24hr", "Category_72hr"),
  labels = c("6hr", "24hr", "72hr")
)

# **Generate Combined Proportion Plot for CX-5461 and DOX**
ggplot(proportion_data, aes(x = Sample, y = Percentage, fill = Category)) +
  geom_bar(stat = "identity", position = "stack") +  # Stacked bars
  facet_wrap(~Timepoint, scales = "fixed") +  # Facet by Timepoint (6hr, 24hr, 72hr)
  scale_y_continuous(labels = scales::percent_format(scale = 1), limits = c(0, 100)) +  # Y-axis as percentage
  scale_fill_manual(values = c("Yes" = "#e41a1c", "No" = "#377eb8")) +  # Yes (Red) at Bottom, No (Blue) on Top
  labs(
    title = "Proportion of Human Colorectal Cancer Genes in CX-5461 and DOX DEGs",
    x = "Samples (CX-5461 and DOX)",
    y = "Percentage",
    fill = "Category"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = rel(1.5), hjust = 0.5),
    axis.title = element_text(size = 15, color = "black"),
    axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
    legend.title = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1.2),  # Updated for ggplot2 3.4.0+
    strip.background = element_blank(),
    strip.text = element_text(size = 12, face = "bold")
  )

```

## **📌 Correlation of colorectal cancer genes with CX and DOX expressed genes**
## **📌 6hr**
```{r Data_prep, echo=TRUE, message=FALSE, fig.height=8,fig.width=12}
# Import the CSV file
data_6hr_logFC <- read.csv("data/Colorectal/6hrlogFC.csv", stringsAsFactors = FALSE)

# Map gene symbols to Entrez IDs using org.Hs.eg.db
data_6hr_logFC <- data_6hr_logFC %>%
  mutate(Entrez_ID = mapIds(org.Hs.eg.db,
                            keys = Symbol,
                            column = "ENTREZID",
                            keytype = "SYMBOL",
                            multiVals = "first"))

# Merge datasets on Entrez_ID for CX and DOX at both concentrations (3 hours)
merged_CX_0.1 <- merge(data_6hr_logFC, CX_0.1_3, by = "Entrez_ID")
merged_CX_0.5 <- merge(data_6hr_logFC, CX_0.5_3, by = "Entrez_ID")
merged_DOX_0.1 <- merge(data_6hr_logFC, DOX_0.1_3, by = "Entrez_ID")
merged_DOX_0.5 <- merge(data_6hr_logFC, DOX_0.5_3, by = "Entrez_ID")

# Remove NA values
merged_CX_0.1 <- na.omit(merged_CX_0.1)
merged_CX_0.5 <- na.omit(merged_CX_0.5)
merged_DOX_0.1 <- na.omit(merged_DOX_0.1)
merged_DOX_0.5 <- na.omit(merged_DOX_0.5)

# Rename columns explicitly to avoid conflicts
colnames(merged_CX_0.1) <- colnames(merged_CX_0.5) <-
  colnames(merged_DOX_0.1) <- colnames(merged_DOX_0.5) <-
  c("Entrez_ID", "Symbol_6hr", "logFC_6hr", "logFC_Drug", "AveExpr_Drug", "t_Drug", "P.Value_Drug", "adj.P.Val_Drug", "B_Drug")

# Add drug and concentration labels for faceting
merged_CX_0.1$Drug <- "CX (3hr)"
merged_CX_0.5$Drug <- "CX (3hr)"
merged_DOX_0.1$Drug <- "DOX (3hr)"
merged_DOX_0.5$Drug <- "DOX (3hr)"

merged_CX_0.1$Concentration <- "0.1 µM"
merged_CX_0.5$Concentration <- "0.5 µM"
merged_DOX_0.1$Concentration <- "0.1 µM"
merged_DOX_0.5$Concentration <- "0.5 µM"

# Combine all datasets into a single data frame
merged_data <- rbind(
  merged_CX_0.1[, c("Entrez_ID", "logFC_Drug", "logFC_6hr", "Concentration", "Drug")],
  merged_CX_0.5[, c("Entrez_ID", "logFC_Drug", "logFC_6hr", "Concentration", "Drug")],
  merged_DOX_0.1[, c("Entrez_ID", "logFC_Drug", "logFC_6hr", "Concentration", "Drug")],
  merged_DOX_0.5[, c("Entrez_ID", "logFC_Drug", "logFC_6hr", "Concentration", "Drug")]
)

# Calculate correlations for each facet
cor_CX_0.1 <- cor.test(merged_CX_0.1$logFC_Drug, merged_CX_0.1$logFC_6hr, method = "pearson")
cor_CX_0.5 <- cor.test(merged_CX_0.5$logFC_Drug, merged_CX_0.5$logFC_6hr, method = "pearson")
cor_DOX_0.1 <- cor.test(merged_DOX_0.1$logFC_Drug, merged_DOX_0.1$logFC_6hr, method = "pearson")
cor_DOX_0.5 <- cor.test(merged_DOX_0.5$logFC_Drug, merged_DOX_0.5$logFC_6hr, method = "pearson")

# Data frame for r and p-values annotations
correlation_data <- data.frame(
  Drug = rep(c("CX (3hr)", "DOX (3hr)"), each = 2),
  Concentration = rep(c("0.1 µM", "0.5 µM"), times = 2),
  x = max(merged_data$logFC_Drug, na.rm = TRUE) * 0.75,  # Adjusted placement for better fit
  y = max(merged_data$logFC_6hr, na.rm = TRUE) * 0.85,
  label = c(
    paste0("r = ", round(cor_CX_0.1$estimate, 3), "\np = ", signif(cor_CX_0.1$p.value, 3)),
    paste0("r = ", round(cor_CX_0.5$estimate, 3), "\np = ", signif(cor_CX_0.5$p.value, 3)),
    paste0("r = ", round(cor_DOX_0.1$estimate, 3), "\np = ", signif(cor_DOX_0.1$p.value, 3)),
    paste0("r = ", round(cor_DOX_0.5$estimate, 3), "\np = ", signif(cor_DOX_0.5$p.value, 3))
  )
)

# Create scatter plot with facets for concentration (vertically) and drug (horizontally)
scatter_plot <- ggplot(merged_data, aes(x = logFC_Drug, y = logFC_6hr)) +
  geom_point(alpha = 0.6, color = "black") +  # Black scatter points
  geom_smooth(method = "lm", color = "black", se = FALSE) +  # Black regression line
  labs(
    title = "Correlation between Drug (3hr) and Colorectal 6hr logFC",
    x = "logFC (Drug_3hr)",
    y = "logFC (Colorectal 6hr)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),

    # Outer box around the entire facet plot
    panel.border = element_rect(color = "black", fill = NA, linewidth = 2),

    # Black box around each facet title
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1.5),
    strip.text = element_text(size = 12, face = "bold", color = "black")
  ) +

  # Facet by drug (horizontally) and concentration (vertically)
  facet_grid(Concentration ~ Drug) +

  # Add r and p-value in the upper-right corner of each facet with better fitting font size
  geom_text(data = correlation_data,
            aes(x = x, y = y, label = label),
            inherit.aes = FALSE, size = 3, fontface = "bold")  # Reduced font size for better fit

# Display plot
print(scatter_plot)
```

## **📌 24hr**
```{r Data_prep_24, echo=TRUE, message=FALSE, fig.height=8,fig.width=12}
# Import the CSV file
data_24hr_logFC <- read.csv("data/Colorectal/24hrlogFC.csv", stringsAsFactors = FALSE)

# Map gene symbols to Entrez IDs using org.Hs.eg.db
data_24hr_logFC <- data_24hr_logFC %>%
  mutate(Entrez_ID = mapIds(org.Hs.eg.db,
                            keys = Symbol,
                            column = "ENTREZID",
                            keytype = "SYMBOL",
                            multiVals = "first"))

# Merge datasets on Entrez_ID for CX and DOX at both concentrations (24 hours)
merged_CX_0.1_24 <- merge(data_24hr_logFC, CX_0.1_24, by = "Entrez_ID")
merged_CX_0.5_24 <- merge(data_24hr_logFC, CX_0.5_24, by = "Entrez_ID")
merged_DOX_0.1_24 <- merge(data_24hr_logFC, DOX_0.1_24, by = "Entrez_ID")
merged_DOX_0.5_24 <- merge(data_24hr_logFC, DOX_0.5_24, by = "Entrez_ID")

# Remove NA values
merged_CX_0.1_24 <- na.omit(merged_CX_0.1_24)
merged_CX_0.5_24 <- na.omit(merged_CX_0.5_24)
merged_DOX_0.1_24 <- na.omit(merged_DOX_0.1_24)
merged_DOX_0.5_24 <- na.omit(merged_DOX_0.5_24)

# Rename columns explicitly to avoid conflicts
colnames(merged_CX_0.1_24) <- colnames(merged_CX_0.5_24) <-
  colnames(merged_DOX_0.1_24) <- colnames(merged_DOX_0.5_24) <-
  c("Entrez_ID", "Symbol_24hr", "logFC_24hr", "logFC_Drug", "AveExpr_Drug", "t_Drug", "P.Value_Drug", "adj.P.Val_Drug", "B_Drug")

# Add drug and concentration labels for faceting
merged_CX_0.1_24$Drug <- "CX (24hr)"
merged_CX_0.5_24$Drug <- "CX (24hr)"
merged_DOX_0.1_24$Drug <- "DOX (24hr)"
merged_DOX_0.5_24$Drug <- "DOX (24hr)"

merged_CX_0.1_24$Concentration <- "0.1 µM"
merged_CX_0.5_24$Concentration <- "0.5 µM"
merged_DOX_0.1_24$Concentration <- "0.1 µM"
merged_DOX_0.5_24$Concentration <- "0.5 µM"

# Combine all datasets into a single data frame
merged_data_24hr <- rbind(
  merged_CX_0.1_24[, c("Entrez_ID", "logFC_Drug", "logFC_24hr", "Concentration", "Drug")],
  merged_CX_0.5_24[, c("Entrez_ID", "logFC_Drug", "logFC_24hr", "Concentration", "Drug")],
  merged_DOX_0.1_24[, c("Entrez_ID", "logFC_Drug", "logFC_24hr", "Concentration", "Drug")],
  merged_DOX_0.5_24[, c("Entrez_ID", "logFC_Drug", "logFC_24hr", "Concentration", "Drug")]
)

# Calculate correlations for each facet
cor_CX_0.1_24 <- cor.test(merged_CX_0.1_24$logFC_Drug, merged_CX_0.1_24$logFC_24hr, method = "pearson")
cor_CX_0.5_24 <- cor.test(merged_CX_0.5_24$logFC_Drug, merged_CX_0.5_24$logFC_24hr, method = "pearson")
cor_DOX_0.1_24 <- cor.test(merged_DOX_0.1_24$logFC_Drug, merged_DOX_0.1_24$logFC_24hr, method = "pearson")
cor_DOX_0.5_24 <- cor.test(merged_DOX_0.5_24$logFC_Drug, merged_DOX_0.5_24$logFC_24hr, method = "pearson")

# Data frame for r and p-values annotations
correlation_data_24hr <- data.frame(
  Drug = rep(c("CX (24hr)", "DOX (24hr)"), each = 2),
  Concentration = rep(c("0.1 µM", "0.5 µM"), times = 2),
  x = max(merged_data_24hr$logFC_Drug, na.rm = TRUE) * 0.75,  # Adjusted placement for better fit
  y = max(merged_data_24hr$logFC_24hr, na.rm = TRUE) * 0.85,
  label = c(
    paste0("r = ", round(cor_CX_0.1_24$estimate, 3), "\np = ", signif(cor_CX_0.1_24$p.value, 3)),
    paste0("r = ", round(cor_CX_0.5_24$estimate, 3), "\np = ", signif(cor_CX_0.5_24$p.value, 3)),
    paste0("r = ", round(cor_DOX_0.1_24$estimate, 3), "\np = ", signif(cor_DOX_0.1_24$p.value, 3)),
    paste0("r = ", round(cor_DOX_0.5_24$estimate, 3), "\np = ", signif(cor_DOX_0.5_24$p.value, 3))
  )
)

# Create scatter plot with facets for concentration (vertically) and drug (horizontally)
scatter_plot_24hr <- ggplot(merged_data_24hr, aes(x = logFC_Drug, y = logFC_24hr)) +
  geom_point(alpha = 0.6, color = "black") +  # Black scatter points
  geom_smooth(method = "lm", color = "black", se = FALSE) +  # Black regression line
  labs(
    title = "Correlation between Drug (24hr) and Colorectal 24hr logFC",
    x = "logFC (Drug_24hr)",
    y = "logFC (Colorectal 24hr)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),

    # Outer box around the entire facet plot
    panel.border = element_rect(color = "black", fill = NA, linewidth = 2),

    # Black box around each facet title
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1.5),
    strip.text = element_text(size = 12, face = "bold", color = "black")
  ) +

  # Facet by drug (horizontally) and concentration (vertically)
  facet_grid(Concentration ~ Drug) +

  # Add r and p-value in the upper-right corner of each facet with better fitting font size
  geom_text(data = correlation_data_24hr,
            aes(x = x, y = y, label = label),
            inherit.aes = FALSE, size = 3, fontface = "bold")  # Reduced font size for better fit

# Display plot
print(scatter_plot_24hr)
```

## **📌 72hr**
```{r Data_prep_72, echo=TRUE, message=FALSE, fig.height=8,fig.width=12}
# Import the CSV file
data_72hr_logFC <- read.csv("data/Colorectal/72hrlogFC.csv", stringsAsFactors = FALSE)

# Map gene symbols to Entrez IDs using org.Hs.eg.db
data_72hr_logFC <- data_72hr_logFC %>%
  mutate(Entrez_ID = mapIds(org.Hs.eg.db,
                            keys = Symbol,
                            column = "ENTREZID",
                            keytype = "SYMBOL",
                            multiVals = "first"))

# Merge datasets on Entrez_ID for CX and DOX at both concentrations (48 hours)
merged_CX_0.1_48 <- merge(data_72hr_logFC, CX_0.1_48, by = "Entrez_ID")
merged_CX_0.5_48 <- merge(data_72hr_logFC, CX_0.5_48, by = "Entrez_ID")
merged_DOX_0.1_48 <- merge(data_72hr_logFC, DOX_0.1_48, by = "Entrez_ID")
merged_DOX_0.5_48 <- merge(data_72hr_logFC, DOX_0.5_48, by = "Entrez_ID")

# Remove NA values
merged_CX_0.1_48 <- na.omit(merged_CX_0.1_48)
merged_CX_0.5_48 <- na.omit(merged_CX_0.5_48)
merged_DOX_0.1_48 <- na.omit(merged_DOX_0.1_48)
merged_DOX_0.5_48 <- na.omit(merged_DOX_0.5_48)

# Rename columns explicitly to avoid conflicts
colnames(merged_CX_0.1_48) <- colnames(merged_CX_0.5_48) <-
  colnames(merged_DOX_0.1_48) <- colnames(merged_DOX_0.5_48) <-
  c("Entrez_ID", "Symbol_72hr", "logFC_72hr", "logFC_Drug", "AveExpr_Drug", "t_Drug", "P.Value_Drug", "adj.P.Val_Drug", "B_Drug")

# Add drug and concentration labels for faceting
merged_CX_0.1_48$Drug <- "CX (48hr)"
merged_CX_0.5_48$Drug <- "CX (48hr)"
merged_DOX_0.1_48$Drug <- "DOX (48hr)"
merged_DOX_0.5_48$Drug <- "DOX (48hr)"

merged_CX_0.1_48$Concentration <- "0.1 µM"
merged_CX_0.5_48$Concentration <- "0.5 µM"
merged_DOX_0.1_48$Concentration <- "0.1 µM"
merged_DOX_0.5_48$Concentration <- "0.5 µM"

# Combine all datasets into a single data frame
merged_data_48hr <- rbind(
  merged_CX_0.1_48[, c("Entrez_ID", "logFC_Drug", "logFC_72hr", "Concentration", "Drug")],
  merged_CX_0.5_48[, c("Entrez_ID", "logFC_Drug", "logFC_72hr", "Concentration", "Drug")],
  merged_DOX_0.1_48[, c("Entrez_ID", "logFC_Drug", "logFC_72hr", "Concentration", "Drug")],
  merged_DOX_0.5_48[, c("Entrez_ID", "logFC_Drug", "logFC_72hr", "Concentration", "Drug")]
)

# Calculate correlations for each facet
cor_CX_0.1_48 <- cor.test(merged_CX_0.1_48$logFC_Drug, merged_CX_0.1_48$logFC_72hr, method = "pearson")
cor_CX_0.5_48 <- cor.test(merged_CX_0.5_48$logFC_Drug, merged_CX_0.5_48$logFC_72hr, method = "pearson")
cor_DOX_0.1_48 <- cor.test(merged_DOX_0.1_48$logFC_Drug, merged_DOX_0.1_48$logFC_72hr, method = "pearson")
cor_DOX_0.5_48 <- cor.test(merged_DOX_0.5_48$logFC_Drug, merged_DOX_0.5_48$logFC_72hr, method = "pearson")

# Data frame for r and p-values annotations
correlation_data_48hr <- data.frame(
  Drug = rep(c("CX (48hr)", "DOX (48hr)"), each = 2),
  Concentration = rep(c("0.1 µM", "0.5 µM"), times = 2),
  x = max(merged_data_48hr$logFC_Drug, na.rm = TRUE) * 0.75,  # Adjusted placement for better fit
  y = max(merged_data_48hr$logFC_72hr, na.rm = TRUE) * 0.85,
  label = c(
    paste0("r = ", round(cor_CX_0.1_48$estimate, 3), "\np = ", signif(cor_CX_0.1_48$p.value, 3)),
    paste0("r = ", round(cor_CX_0.5_48$estimate, 3), "\np = ", signif(cor_CX_0.5_48$p.value, 3)),
    paste0("r = ", round(cor_DOX_0.1_48$estimate, 3), "\np = ", signif(cor_DOX_0.1_48$p.value, 3)),
    paste0("r = ", round(cor_DOX_0.5_48$estimate, 3), "\np = ", signif(cor_DOX_0.5_48$p.value, 3))
  )
)

# Create scatter plot with facets for concentration (vertically) and drug (horizontally)
scatter_plot_48hr <- ggplot(merged_data_48hr, aes(x = logFC_Drug, y = logFC_72hr)) +
  geom_point(alpha = 0.6, color = "black") +  # Black scatter points
  geom_smooth(method = "lm", color = "black", se = FALSE) +  # Black regression line
  labs(
    title = "Correlation between Drug (48hr) and Colorectal 72hr logFC",
    x = "logFC (Drug_48hr)",
    y = "logFC (Colorectal 72hr)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),

    # Outer box around the entire facet plot
    panel.border = element_rect(color = "black", fill = NA, linewidth = 2),

    # Black box around each facet title
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1.5),
    strip.text = element_text(size = 12, face = "bold", color = "black")
  ) +

  # Facet by drug (horizontally) and concentration (vertically)
  facet_grid(Concentration ~ Drug) +

  # Add r and p-value in the upper-right corner of each facet with better fitting font size
  geom_text(data = correlation_data_48hr,
            aes(x = x, y = y, label = label),
            inherit.aes = FALSE, size = 3, fontface = "bold")  # Reduced font size for better fit

# Display plot
print(scatter_plot_48hr)
```

