---
title: "Function and Pathway Enrichment Analysis of Corrmotifs"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## **📌 Load Required Libraries**
```{r load_libraries, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse) 
library(ggfortify)
library(ggplot2)
library(cluster)
library(edgeR)
library(limma)
library(Homo.sapiens)
library(BiocParallel)
library(qvalue)
library(pheatmap)
library(clusterProfiler)
library(AnnotationDbi)
library(org.Hs.eg.db)
library(RColorBrewer)
library(readr)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(DOSE)

# Load UCSC transcript database
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
```

## **📌 Read and Process Data**
```{r load_DEGs, echo=TRUE, results='hide', message=FALSE, warning=FALSE}
# Load the saved datasets
prob_all_1 <- read.csv("data/prob_all_1.csv")$Entrez_ID
prob_all_2 <- read.csv("data/prob_all_2.csv")$Entrez_ID
prob_all_3 <- read.csv("data/prob_all_3.csv")$Entrez_ID
prob_all_4 <- read.csv("data/prob_all_4.csv")$Entrez_ID

CX_0.1_3 <- read.csv("data/DEGs/Toptable_CX_0.1_3.csv")
background<-as.character(CX_0.1_3$Entrez_ID)
```

## **📌 Non response**

## **📌 Non response GO Enrichment Clusterprofiler**
```{r GO1, echo=TRUE, message=FALSE, fig.width=12, fig.height=12}
# Perform GO enrichment analysis for BP, MF, and CC
go_enrichment_BP <- enrichGO(gene = prob_all_1,
                             OrgDb = org.Hs.eg.db,
                             keyType = "ENTREZID",
                             universe = background,
                             ont = "BP",
                             pvalueCutoff = 0.05)

go_enrichment_MF <- enrichGO(gene = prob_all_1,
                             OrgDb = org.Hs.eg.db,
                             keyType = "ENTREZID",
                             universe = background,
                             ont = "MF",
                             pvalueCutoff = 0.05)

go_enrichment_CC <- enrichGO(gene = prob_all_1,
                             OrgDb = org.Hs.eg.db,
                             keyType = "ENTREZID",
                             universe = background,
                             ont = "CC",
                             pvalueCutoff = 0.05)

# Convert each enrichment result to a tibble, add a category column, and select top 20 terms
process_enrichment_tibble <- function(enrichment, category) {
  if (is.null(enrichment) || nrow(as.data.frame(enrichment)) == 0) {
    return(tibble(Description = "No enriched terms", neglog = 0, Category = category))
  } else {
    enrichment %>%
      as_tibble() %>%
      mutate(Category = category,
             neglog = -log(p.adjust)) %>% # Add -log(p.adjust) column
      arrange(desc(neglog)) %>%          # Sort by -log(p.adjust)
      slice(1:20)                        # Select top 20 terms
  }
}

BP_Tibble <- process_enrichment_tibble(go_enrichment_BP, "Biological Process")
MF_Tibble <- process_enrichment_tibble(go_enrichment_MF, "Molecular Function")
CC_Tibble <- process_enrichment_tibble(go_enrichment_CC, "Cellular Component")

# Combine all tibbles
combined_GO_Tibble <- bind_rows(BP_Tibble, MF_Tibble, CC_Tibble)
# Function to generate enrichment plots
process_enrichment_plot <- function(tibble, title, color) {
  ggplot(data = tibble, aes(x = neglog, y = reorder(Description, neglog))) +
    geom_bar(stat = "identity", fill = color) +
    labs(x = "-log(p-adjust)",
         y = title,
         title = paste("Top 20", title, "GO Terms")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black"),
      panel.border = element_rect(colour = "black", fill = NA, size = 0.3)
    ) +
    xlim(c(0, max(tibble$neglog) + 1))
}

# Generate separate plots
plot_BP <- process_enrichment_plot(BP_Tibble, "Biological Process", "#2E86C1")
plot_MF <- process_enrichment_plot(MF_Tibble, "Molecular Function", "#28B463")
plot_CC <- process_enrichment_plot(CC_Tibble, "Cellular Component", "#D35400")

# Combine the plots using patchwork
combined_plot <- plot_BP / plot_MF / plot_CC

# Display the combined plot
combined_plot
```


## **📌 Non response GO Enrichment g:Profiler**
```{r gprofiler_GO, echo=TRUE, message=FALSE, fig.width=12, fig.height=12}

# Load the gprofiler2 package
library(gprofiler2)
library(ggplot2)
library(dplyr)
library(patchwork)

# Perform GO enrichment analysis with gprofiler2
gost_results <- gost(
  query = prob_all_1,           # List of input genes (prob_all_1)
  organism = "hsapiens",   # Human organism
  user_threshold = 0.05,   # Adjusted p-value cutoff
  correction_method = "fdr", # Multiple testing correction
  domain_scope = "custom", # Use custom background
  custom_bg = background,    # Background set of genes
  sources = c("GO:BP", "GO:MF", "GO:CC") # Analyze GO categories
)

# Check if enrichment results exist
if (is.null(gost_results$result) || nrow(gost_results$result) == 0) {
  # If no enriched terms, create a placeholder dataframe
  combined_results <- tibble(
    term_name = "No enriched terms",
    p.adjust = NA,
    source = "N/A",
    Category = "N/A"
  )
} else {
  # Convert results to a data frame
  gost_results_df <- gost_results$result

  # Add a column for adjusted p-value
  gost_results_df <- gost_results_df %>%
    rename(p.adjust = p_value)

  # Separate results for BP, MF, and CC
  BP_results <- gost_results_df %>%
    filter(source == "GO:BP") %>%
    mutate(Category = "Biological Process")

  MF_results <- gost_results_df %>%
    filter(source == "GO:MF") %>%
    mutate(Category = "Molecular Function")

  CC_results <- gost_results_df %>%
    filter(source == "GO:CC") %>%
    mutate(Category = "Cellular Component")

  # Select the top 20 terms by adjusted p-value for each category
  top_BP <- BP_results %>%
    arrange(p.adjust) %>%
    slice_head(n = 20)

  top_MF <- MF_results %>%
    arrange(p.adjust) %>%
    slice_head(n = 20)

  top_CC <- CC_results %>%
    arrange(p.adjust) %>%
    slice_head(n = 20)

  # Combine all categories
  combined_results <- bind_rows(top_BP, top_MF, top_CC)
}

# Ensure all columns are atomic types for CSV export
combined_results_clean <- combined_results %>%
  mutate(across(everything(), ~ if (is.list(.)) sapply(., toString) else .))

# Function for plotting top terms
plot_gprofiler_results <- function(data, title, color) {
  ggplot(data, aes(x = -log10(p.adjust), y = reorder(term_name, -log10(p.adjust)))) +
    geom_bar(stat = "identity", fill = color) +
    labs(
      x = "-log10(Adjusted p-value)",
      y = title,
      title = paste("Top 20", title, "GO Terms")
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black"),
      panel.border = element_rect(colour = "black", fill = NA, size = 0.3)
    )
}

# Check if there are enrichment terms to plot
if (nrow(combined_results) == 1 && combined_results$term_name == "No enriched terms") {
  message("No enriched GO terms found for the input gene set.")
} else {
  # Plot the top 20 terms for each category
  plot_BP <- plot_gprofiler_results(top_BP, "Biological Process", "#2E86C1")
  plot_MF <- plot_gprofiler_results(top_MF, "Molecular Function", "#28B463")
  plot_CC <- plot_gprofiler_results(top_CC, "Cellular Component", "#D35400")

  # Combine the plots using patchwork
  combined_plot <- plot_BP / plot_MF / plot_CC

  # Display the combined plot
  combined_plot
}
```

## **📌 Non response Pathway Enrichment**
```{r pathway_enrichment, echo=TRUE, message=FALSE, fig.width=12, fig.height=12}

# Load required libraries
library(clusterProfiler)
library(org.Hs.eg.db) # Required for enrichPathway
library(gprofiler2)
library(ggplot2)
library(dplyr)
library(patchwork)
library(ReactomePA)

# Function for ClusterProfiler Reactome & KEGG Analysis
process_clusterProfiler <- function(gene_set, background, category, color, y_title) {
  # Perform enrichment based on the selected category
  if (category == "Reactome") {
    enrichment <- enrichPathway(
      gene = gene_set,
      organism = "human",
      pvalueCutoff = 0.05,
      pAdjustMethod = "BH",
      universe = background
    )
  } else if (category == "KEGG") {
    enrichment <- enrichKEGG(
      gene = gene_set,
      organism = "hsa",
      pvalueCutoff = 0.05,
      pAdjustMethod = "BH",
      universe = background
    )
  }
  
  # Check if enrichment results exist
  if (is.null(enrichment) || nrow(as.data.frame(enrichment)) == 0) {
    message(paste("No significant enrichment found for", category, "in ClusterProfiler"))
    return(NULL)
  }
  
  # Convert results to tibble and process top 20 terms
  enrichment_tibble <- as_tibble(as.data.frame(enrichment)) %>%
    mutate(Category = category,
           neglog = -log10(p.adjust)) %>%  # Compute -log10(p.adjust)
    arrange(desc(neglog)) %>%
    slice_head(n = min(20, nrow(.)))  # Ensure safe slicing
  
  # Generate plot
  plot <- ggplot(enrichment_tibble, aes(x = neglog, y = reorder(Description, neglog))) +
    geom_bar(stat = "identity", fill = color) +
    labs(x = "-log10(Adjusted p-value)",
         y = y_title,
         title = paste("Enriched", category, "Pathways")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black")
    )
  
  return(plot)
}

# Function for gProfiler Reactome & KEGG Analysis
process_gprofiler <- function(gene_set, background, category, color, y_title) {
  # Perform enrichment using gprofiler2
  enrichment <- gost(
    query = gene_set,
    organism = "hsapiens",
    user_threshold = 0.05,
    correction_method = "fdr",
    domain_scope = "custom",
    custom_bg = background,
    sources = category # Either "REAC" or "KEGG"
  )
  
  # Check if enrichment results exist
  if (is.null(enrichment$result) || nrow(enrichment$result) == 0) {
    message(paste("No significant enrichment found for", category, "in gProfiler"))
    return(NULL)
  }
  
  # Convert results to tibble and process top 20 terms
  enrichment_tibble <- enrichment$result %>%
    as_tibble() %>%
    mutate(Category = category,
           neglog = -log10(p_value)) %>%  # Compute -log10(p-value)
    arrange(desc(neglog)) %>%
    slice_head(n = min(20, nrow(.)))  # Ensure safe slicing
  
  # Generate plot
  plot <- ggplot(enrichment_tibble, aes(x = neglog, y = reorder(term_name, neglog))) +
    geom_bar(stat = "identity", fill = color) +
    labs(x = "-log10(p-value)",
         y = y_title,
         title = paste("Enriched", category, "Pathways")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black")
    )
  
  return(plot)
}

# Perform analysis for Reactome and KEGG using ClusterProfiler
cluster_reactome <- process_clusterProfiler(
  gene_set = prob_all_1,
  background = background,
  category = "Reactome",
  color = "#2E86C1",
  y_title = "Reactome Pathways"
)

cluster_kegg <- process_clusterProfiler(
  gene_set = prob_all_1,
  background = background,
  category = "KEGG",
  color = "#28B463",
  y_title = "KEGG Pathways"
)

# Combine Reactome and KEGG for ClusterProfiler
if (!is.null(cluster_reactome) && !is.null(cluster_kegg)) {
  cluster_combined <- cluster_reactome / cluster_kegg
} else if (!is.null(cluster_reactome)) {
  cluster_combined <- cluster_reactome
} else if (!is.null(cluster_kegg)) {
  cluster_combined <- cluster_kegg
} else {
  cluster_combined <- NULL
}

# Perform analysis for Reactome and KEGG using GProfiler
gprofiler_reactome <- process_gprofiler(
  gene_set = prob_all_1,
  background = background,
  category = "REAC",  # Corrected category for Reactome in gProfiler
  color = "#D35400",
  y_title = "Reactome Pathways"
)

gprofiler_kegg <- process_gprofiler(
  gene_set = prob_all_1,
  background = background,
  category = "KEGG",
  color = "#F39C12",
  y_title = "KEGG Pathways"
)

# Combine Reactome and KEGG for GProfiler
if (!is.null(gprofiler_reactome) && !is.null(gprofiler_kegg)) {
  gprofiler_combined <- gprofiler_reactome / gprofiler_kegg
} else if (!is.null(gprofiler_reactome)) {
  gprofiler_combined <- gprofiler_reactome
} else if (!is.null(gprofiler_kegg)) {
  gprofiler_combined <- gprofiler_kegg
} else {
  gprofiler_combined <- NULL
}

# Display plots (if they are not NULL)
if (!is.null(cluster_combined)) print(cluster_combined)
if (!is.null(gprofiler_combined)) print(gprofiler_combined)
```


## **📌 CX_DOX shared late response**

## **📌 CX_DOX shared late response GO Enrichment Clusterprofiler**
```{r GO2, echo=TRUE, message=FALSE, fig.width=12, fig.height=12}
# Perform GO enrichment analysis for BP, MF, and CC
go_enrichment_BP <- enrichGO(gene = prob_all_2,
                             OrgDb = org.Hs.eg.db,
                             keyType = "ENTREZID",
                             universe = background,
                             ont = "BP",
                             pvalueCutoff = 0.05)

go_enrichment_MF <- enrichGO(gene = prob_all_2,
                             OrgDb = org.Hs.eg.db,
                             keyType = "ENTREZID",
                             universe = background,
                             ont = "MF",
                             pvalueCutoff = 0.05)

go_enrichment_CC <- enrichGO(gene = prob_all_2,
                             OrgDb = org.Hs.eg.db,
                             keyType = "ENTREZID",
                             universe = background,
                             ont = "CC",
                             pvalueCutoff = 0.05)

# Convert each enrichment result to a tibble, add a category column, and select top 20 terms
process_enrichment_tibble <- function(enrichment, category) {
  if (is.null(enrichment) || nrow(as.data.frame(enrichment)) == 0) {
    return(tibble(Description = "No enriched terms", neglog = 0, Category = category))
  } else {
    enrichment %>%
      as_tibble() %>%
      mutate(Category = category,
             neglog = -log(p.adjust)) %>% # Add -log(p.adjust) column
      arrange(desc(neglog)) %>%          # Sort by -log(p.adjust)
      slice(1:20)                        # Select top 20 terms
  }
}

BP_Tibble <- process_enrichment_tibble(go_enrichment_BP, "Biological Process")
MF_Tibble <- process_enrichment_tibble(go_enrichment_MF, "Molecular Function")
CC_Tibble <- process_enrichment_tibble(go_enrichment_CC, "Cellular Component")

# Combine all tibbles
combined_GO_Tibble <- bind_rows(BP_Tibble, MF_Tibble, CC_Tibble)
# Function to generate enrichment plots
process_enrichment_plot <- function(tibble, title, color) {
  ggplot(data = tibble, aes(x = neglog, y = reorder(Description, neglog))) +
    geom_bar(stat = "identity", fill = color) +
    labs(x = "-log(p-adjust)",
         y = title,
         title = paste("Top 20", title, "GO Terms")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black"),
      panel.border = element_rect(colour = "black", fill = NA, size = 0.3)
    ) +
    xlim(c(0, max(tibble$neglog) + 1))
}

# Generate separate plots
plot_BP <- process_enrichment_plot(BP_Tibble, "Biological Process", "#2E86C1")
plot_MF <- process_enrichment_plot(MF_Tibble, "Molecular Function", "#28B463")
plot_CC <- process_enrichment_plot(CC_Tibble, "Cellular Component", "#D35400")

# Combine the plots using patchwork
combined_plot <- plot_BP / plot_MF / plot_CC

# Display the combined plot
combined_plot
```


## **📌 CX_DOX shared late response GO Enrichment g:Profiler**
```{r gprofiler_GO2, echo=TRUE, message=FALSE, fig.width=12, fig.height=12}

# Load the gprofiler2 package
library(gprofiler2)
library(ggplot2)
library(dplyr)
library(patchwork)

# Perform GO enrichment analysis with gprofiler2
gost_results <- gost(
  query = prob_all_2,           # List of input genes (prob_all_2)
  organism = "hsapiens",   # Human organism
  user_threshold = 0.05,   # Adjusted p-value cutoff
  correction_method = "fdr", # Multiple testing correction
  domain_scope = "custom", # Use custom background
  custom_bg = background,    # Background set of genes
  sources = c("GO:BP", "GO:MF", "GO:CC") # Analyze GO categories
)

# Check if enrichment results exist
if (is.null(gost_results$result) || nrow(gost_results$result) == 0) {
  # If no enriched terms, create a placeholder dataframe
  combined_results <- tibble(
    term_name = "No enriched terms",
    p.adjust = NA,
    source = "N/A",
    Category = "N/A"
  )
} else {
  # Convert results to a data frame
  gost_results_df <- gost_results$result

  # Add a column for adjusted p-value
  gost_results_df <- gost_results_df %>%
    rename(p.adjust = p_value)

  # Separate results for BP, MF, and CC
  BP_results <- gost_results_df %>%
    filter(source == "GO:BP") %>%
    mutate(Category = "Biological Process")

  MF_results <- gost_results_df %>%
    filter(source == "GO:MF") %>%
    mutate(Category = "Molecular Function")

  CC_results <- gost_results_df %>%
    filter(source == "GO:CC") %>%
    mutate(Category = "Cellular Component")

  # Select the top 20 terms by adjusted p-value for each category
  top_BP <- BP_results %>%
    arrange(p.adjust) %>%
    slice_head(n = 20)

  top_MF <- MF_results %>%
    arrange(p.adjust) %>%
    slice_head(n = 20)

  top_CC <- CC_results %>%
    arrange(p.adjust) %>%
    slice_head(n = 20)

  # Combine all categories
  combined_results <- bind_rows(top_BP, top_MF, top_CC)
}

# Ensure all columns are atomic types for CSV export
combined_results_clean <- combined_results %>%
  mutate(across(everything(), ~ if (is.list(.)) sapply(., toString) else .))

# Function for plotting top terms
plot_gprofiler_results <- function(data, title, color) {
  ggplot(data, aes(x = -log10(p.adjust), y = reorder(term_name, -log10(p.adjust)))) +
    geom_bar(stat = "identity", fill = color) +
    labs(
      x = "-log10(Adjusted p-value)",
      y = title,
      title = paste("Top 20", title, "GO Terms")
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black"),
      panel.border = element_rect(colour = "black", fill = NA, size = 0.3)
    )
}

# Check if there are enrichment terms to plot
if (nrow(combined_results) == 1 && combined_results$term_name == "No enriched terms") {
  message("No enriched GO terms found for the input gene set.")
} else {
  # Plot the top 20 terms for each category
  plot_BP <- plot_gprofiler_results(top_BP, "Biological Process", "#2E86C1")
  plot_MF <- plot_gprofiler_results(top_MF, "Molecular Function", "#28B463")
  plot_CC <- plot_gprofiler_results(top_CC, "Cellular Component", "#D35400")

  # Combine the plots using patchwork
  combined_plot <- plot_BP / plot_MF / plot_CC

  # Display the combined plot
  combined_plot
}
```

## **📌 CX_DOX shared late response Pathway Enrichment**
```{r pathway_enrichment2, echo=TRUE, message=FALSE, fig.width=12, fig.height=12}

# Load required libraries
library(clusterProfiler)
library(org.Hs.eg.db) # Required for enrichPathway
library(gprofiler2)
library(ggplot2)
library(dplyr)
library(patchwork)
library(ReactomePA)

# Function for ClusterProfiler Reactome & KEGG Analysis
process_clusterProfiler <- function(gene_set, background, category, color, y_title) {
  # Perform enrichment based on the selected category
  if (category == "Reactome") {
    enrichment <- enrichPathway(
      gene = gene_set,
      organism = "human",
      pvalueCutoff = 0.05,
      pAdjustMethod = "BH",
      universe = background
    )
  } else if (category == "KEGG") {
    enrichment <- enrichKEGG(
      gene = gene_set,
      organism = "hsa",
      pvalueCutoff = 0.05,
      pAdjustMethod = "BH",
      universe = background
    )
  }
  
  # Check if enrichment results exist
  if (is.null(enrichment) || nrow(as.data.frame(enrichment)) == 0) {
    message(paste("No significant enrichment found for", category, "in ClusterProfiler"))
    return(NULL)
  }
  
  # Convert results to tibble and process top 20 terms
  enrichment_tibble <- as_tibble(as.data.frame(enrichment)) %>%
    mutate(Category = category,
           neglog = -log10(p.adjust)) %>%  # Compute -log10(p.adjust)
    arrange(desc(neglog)) %>%
    slice_head(n = min(20, nrow(.)))  # Ensure safe slicing
  
  # Generate plot
  plot <- ggplot(enrichment_tibble, aes(x = neglog, y = reorder(Description, neglog))) +
    geom_bar(stat = "identity", fill = color) +
    labs(x = "-log10(Adjusted p-value)",
         y = y_title,
         title = paste("Enriched", category, "Pathways")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black")
    )
  
  return(plot)
}

# Function for gProfiler Reactome & KEGG Analysis
process_gprofiler <- function(gene_set, background, category, color, y_title) {
  # Perform enrichment using gprofiler2
  enrichment <- gost(
    query = gene_set,
    organism = "hsapiens",
    user_threshold = 0.05,
    correction_method = "fdr",
    domain_scope = "custom",
    custom_bg = background,
    sources = category # Either "REAC" or "KEGG"
  )
  
  # Check if enrichment results exist
  if (is.null(enrichment$result) || nrow(enrichment$result) == 0) {
    message(paste("No significant enrichment found for", category, "in gProfiler"))
    return(NULL)
  }
  
  # Convert results to tibble and process top 20 terms
  enrichment_tibble <- enrichment$result %>%
    as_tibble() %>%
    mutate(Category = category,
           neglog = -log10(p_value)) %>%  # Compute -log10(p-value)
    arrange(desc(neglog)) %>%
    slice_head(n = min(20, nrow(.)))  # Ensure safe slicing
  
  # Generate plot
  plot <- ggplot(enrichment_tibble, aes(x = neglog, y = reorder(term_name, neglog))) +
    geom_bar(stat = "identity", fill = color) +
    labs(x = "-log10(p-value)",
         y = y_title,
         title = paste("Enriched", category, "Pathways")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black")
    )
  
  return(plot)
}

# Perform analysis for Reactome and KEGG using ClusterProfiler
cluster_reactome <- process_clusterProfiler(
  gene_set = prob_all_2,
  background = background,
  category = "Reactome",
  color = "#2E86C1",
  y_title = "Reactome Pathways"
)

cluster_kegg <- process_clusterProfiler(
  gene_set = prob_all_2,
  background = background,
  category = "KEGG",
  color = "#28B463",
  y_title = "KEGG Pathways"
)

# Combine Reactome and KEGG for ClusterProfiler
if (!is.null(cluster_reactome) && !is.null(cluster_kegg)) {
  cluster_combined <- cluster_reactome / cluster_kegg
} else if (!is.null(cluster_reactome)) {
  cluster_combined <- cluster_reactome
} else if (!is.null(cluster_kegg)) {
  cluster_combined <- cluster_kegg
} else {
  cluster_combined <- NULL
}

# Perform analysis for Reactome and KEGG using GProfiler
gprofiler_reactome <- process_gprofiler(
  gene_set = prob_all_2,
  background = background,
  category = "REAC",  # Corrected category for Reactome in gProfiler
  color = "#D35400",
  y_title = "Reactome Pathways"
)

gprofiler_kegg <- process_gprofiler(
  gene_set = prob_all_2,
  background = background,
  category = "KEGG",
  color = "#F39C12",
  y_title = "KEGG Pathways"
)

# Combine Reactome and KEGG for GProfiler
if (!is.null(gprofiler_reactome) && !is.null(gprofiler_kegg)) {
  gprofiler_combined <- gprofiler_reactome / gprofiler_kegg
} else if (!is.null(gprofiler_reactome)) {
  gprofiler_combined <- gprofiler_reactome
} else if (!is.null(gprofiler_kegg)) {
  gprofiler_combined <- gprofiler_kegg
} else {
  gprofiler_combined <- NULL
}

# Display plots (if they are not NULL)
if (!is.null(cluster_combined)) print(cluster_combined)
if (!is.null(gprofiler_combined)) print(gprofiler_combined)
```


## **📌 DOX-specific response**

## **📌 DOX-specific response GO Enrichment Clusterprofiler**
```{r GO3, echo=TRUE, message=FALSE, fig.width=12, fig.height=12}
# Perform GO enrichment analysis for BP, MF, and CC
go_enrichment_BP <- enrichGO(gene = prob_all_3,
                             OrgDb = org.Hs.eg.db,
                             keyType = "ENTREZID",
                             universe = background,
                             ont = "BP",
                             pvalueCutoff = 0.05)

go_enrichment_MF <- enrichGO(gene = prob_all_3,
                             OrgDb = org.Hs.eg.db,
                             keyType = "ENTREZID",
                             universe = background,
                             ont = "MF",
                             pvalueCutoff = 0.05)

go_enrichment_CC <- enrichGO(gene = prob_all_3,
                             OrgDb = org.Hs.eg.db,
                             keyType = "ENTREZID",
                             universe = background,
                             ont = "CC",
                             pvalueCutoff = 0.05)

# Convert each enrichment result to a tibble, add a category column, and select top 20 terms
process_enrichment_tibble <- function(enrichment, category) {
  if (is.null(enrichment) || nrow(as.data.frame(enrichment)) == 0) {
    return(tibble(Description = "No enriched terms", neglog = 0, Category = category))
  } else {
    enrichment %>%
      as_tibble() %>%
      mutate(Category = category,
             neglog = -log(p.adjust)) %>% # Add -log(p.adjust) column
      arrange(desc(neglog)) %>%          # Sort by -log(p.adjust)
      slice(1:20)                        # Select top 20 terms
  }
}

BP_Tibble <- process_enrichment_tibble(go_enrichment_BP, "Biological Process")
MF_Tibble <- process_enrichment_tibble(go_enrichment_MF, "Molecular Function")
CC_Tibble <- process_enrichment_tibble(go_enrichment_CC, "Cellular Component")

# Combine all tibbles
combined_GO_Tibble <- bind_rows(BP_Tibble, MF_Tibble, CC_Tibble)
# Function to generate enrichment plots
process_enrichment_plot <- function(tibble, title, color) {
  ggplot(data = tibble, aes(x = neglog, y = reorder(Description, neglog))) +
    geom_bar(stat = "identity", fill = color) +
    labs(x = "-log(p-adjust)",
         y = title,
         title = paste("Top 20", title, "GO Terms")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black"),
      panel.border = element_rect(colour = "black", fill = NA, size = 0.3)
    ) +
    xlim(c(0, max(tibble$neglog) + 1))
}

# Generate separate plots
plot_BP <- process_enrichment_plot(BP_Tibble, "Biological Process", "#2E86C1")
plot_MF <- process_enrichment_plot(MF_Tibble, "Molecular Function", "#28B463")
plot_CC <- process_enrichment_plot(CC_Tibble, "Cellular Component", "#D35400")

# Combine the plots using patchwork
combined_plot <- plot_BP / plot_MF / plot_CC

# Display the combined plot
combined_plot
```


## **📌 DOX-specific response GO Enrichment g:Profiler**
```{r gprofiler_GO3, echo=TRUE, message=FALSE, fig.width=12, fig.height=12}

# Load the gprofiler2 package
library(gprofiler2)
library(ggplot2)
library(dplyr)
library(patchwork)

# Perform GO enrichment analysis with gprofiler2
gost_results <- gost(
  query = prob_all_3,           # List of input genes (prob_all_3)
  organism = "hsapiens",   # Human organism
  user_threshold = 0.05,   # Adjusted p-value cutoff
  correction_method = "fdr", # Multiple testing correction
  domain_scope = "custom", # Use custom background
  custom_bg = background,    # Background set of genes
  sources = c("GO:BP", "GO:MF", "GO:CC") # Analyze GO categories
)

# Check if enrichment results exist
if (is.null(gost_results$result) || nrow(gost_results$result) == 0) {
  # If no enriched terms, create a placeholder dataframe
  combined_results <- tibble(
    term_name = "No enriched terms",
    p.adjust = NA,
    source = "N/A",
    Category = "N/A"
  )
} else {
  # Convert results to a data frame
  gost_results_df <- gost_results$result

  # Add a column for adjusted p-value
  gost_results_df <- gost_results_df %>%
    rename(p.adjust = p_value)

  # Separate results for BP, MF, and CC
  BP_results <- gost_results_df %>%
    filter(source == "GO:BP") %>%
    mutate(Category = "Biological Process")

  MF_results <- gost_results_df %>%
    filter(source == "GO:MF") %>%
    mutate(Category = "Molecular Function")

  CC_results <- gost_results_df %>%
    filter(source == "GO:CC") %>%
    mutate(Category = "Cellular Component")

  # Select the top 20 terms by adjusted p-value for each category
  top_BP <- BP_results %>%
    arrange(p.adjust) %>%
    slice_head(n = 20)

  top_MF <- MF_results %>%
    arrange(p.adjust) %>%
    slice_head(n = 20)

  top_CC <- CC_results %>%
    arrange(p.adjust) %>%
    slice_head(n = 20)

  # Combine all categories
  combined_results <- bind_rows(top_BP, top_MF, top_CC)
}

# Ensure all columns are atomic types for CSV export
combined_results_clean <- combined_results %>%
  mutate(across(everything(), ~ if (is.list(.)) sapply(., toString) else .))

# Function for plotting top terms
plot_gprofiler_results <- function(data, title, color) {
  ggplot(data, aes(x = -log10(p.adjust), y = reorder(term_name, -log10(p.adjust)))) +
    geom_bar(stat = "identity", fill = color) +
    labs(
      x = "-log10(Adjusted p-value)",
      y = title,
      title = paste("Top 20", title, "GO Terms")
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black"),
      panel.border = element_rect(colour = "black", fill = NA, size = 0.3)
    )
}

# Check if there are enrichment terms to plot
if (nrow(combined_results) == 1 && combined_results$term_name == "No enriched terms") {
  message("No enriched GO terms found for the input gene set.")
} else {
  # Plot the top 20 terms for each category
  plot_BP <- plot_gprofiler_results(top_BP, "Biological Process", "#2E86C1")
  plot_MF <- plot_gprofiler_results(top_MF, "Molecular Function", "#28B463")
  plot_CC <- plot_gprofiler_results(top_CC, "Cellular Component", "#D35400")

  # Combine the plots using patchwork
  combined_plot <- plot_BP / plot_MF / plot_CC

  # Display the combined plot
  combined_plot
}
```

## **📌 DOX-specific response Pathway Enrichment**
```{r pathway_enrichment3, echo=TRUE, message=FALSE, fig.width=12, fig.height=12}

# Load required libraries
library(clusterProfiler)
library(org.Hs.eg.db) # Required for enrichPathway
library(gprofiler2)
library(ggplot2)
library(dplyr)
library(patchwork)
library(ReactomePA)

# Function for ClusterProfiler Reactome & KEGG Analysis
process_clusterProfiler <- function(gene_set, background, category, color, y_title) {
  # Perform enrichment based on the selected category
  if (category == "Reactome") {
    enrichment <- enrichPathway(
      gene = gene_set,
      organism = "human",
      pvalueCutoff = 0.05,
      pAdjustMethod = "BH",
      universe = background
    )
  } else if (category == "KEGG") {
    enrichment <- enrichKEGG(
      gene = gene_set,
      organism = "hsa",
      pvalueCutoff = 0.05,
      pAdjustMethod = "BH",
      universe = background
    )
  }
  
  # Check if enrichment results exist
  if (is.null(enrichment) || nrow(as.data.frame(enrichment)) == 0) {
    message(paste("No significant enrichment found for", category, "in ClusterProfiler"))
    return(NULL)
  }
  
  # Convert results to tibble and process top 20 terms
  enrichment_tibble <- as_tibble(as.data.frame(enrichment)) %>%
    mutate(Category = category,
           neglog = -log10(p.adjust)) %>%  # Compute -log10(p.adjust)
    arrange(desc(neglog)) %>%
    slice_head(n = min(20, nrow(.)))  # Ensure safe slicing
  
  # Generate plot
  plot <- ggplot(enrichment_tibble, aes(x = neglog, y = reorder(Description, neglog))) +
    geom_bar(stat = "identity", fill = color) +
    labs(x = "-log10(Adjusted p-value)",
         y = y_title,
         title = paste("Enriched", category, "Pathways")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black")
    )
  
  return(plot)
}

# Function for gProfiler Reactome & KEGG Analysis
process_gprofiler <- function(gene_set, background, category, color, y_title) {
  # Perform enrichment using gprofiler2
  enrichment <- gost(
    query = gene_set,
    organism = "hsapiens",
    user_threshold = 0.05,
    correction_method = "fdr",
    domain_scope = "custom",
    custom_bg = background,
    sources = category # Either "REAC" or "KEGG"
  )
  
  # Check if enrichment results exist
  if (is.null(enrichment$result) || nrow(enrichment$result) == 0) {
    message(paste("No significant enrichment found for", category, "in gProfiler"))
    return(NULL)
  }
  
  # Convert results to tibble and process top 20 terms
  enrichment_tibble <- enrichment$result %>%
    as_tibble() %>%
    mutate(Category = category,
           neglog = -log10(p_value)) %>%  # Compute -log10(p-value)
    arrange(desc(neglog)) %>%
    slice_head(n = min(20, nrow(.)))  # Ensure safe slicing
  
  # Generate plot
  plot <- ggplot(enrichment_tibble, aes(x = neglog, y = reorder(term_name, neglog))) +
    geom_bar(stat = "identity", fill = color) +
    labs(x = "-log10(p-value)",
         y = y_title,
         title = paste("Enriched", category, "Pathways")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black")
    )
  
  return(plot)
}

# Perform analysis for Reactome and KEGG using ClusterProfiler
cluster_reactome <- process_clusterProfiler(
  gene_set = prob_all_3,
  background = background,
  category = "Reactome",
  color = "#2E86C1",
  y_title = "Reactome Pathways"
)

cluster_kegg <- process_clusterProfiler(
  gene_set = prob_all_3,
  background = background,
  category = "KEGG",
  color = "#28B463",
  y_title = "KEGG Pathways"
)

# Combine Reactome and KEGG for ClusterProfiler
if (!is.null(cluster_reactome) && !is.null(cluster_kegg)) {
  cluster_combined <- cluster_reactome / cluster_kegg
} else if (!is.null(cluster_reactome)) {
  cluster_combined <- cluster_reactome
} else if (!is.null(cluster_kegg)) {
  cluster_combined <- cluster_kegg
} else {
  cluster_combined <- NULL
}

# Perform analysis for Reactome and KEGG using GProfiler
gprofiler_reactome <- process_gprofiler(
  gene_set = prob_all_3,
  background = background,
  category = "REAC",  # Corrected category for Reactome in gProfiler
  color = "#D35400",
  y_title = "Reactome Pathways"
)

gprofiler_kegg <- process_gprofiler(
  gene_set = prob_all_3,
  background = background,
  category = "KEGG",
  color = "#F39C12",
  y_title = "KEGG Pathways"
)

# Combine Reactome and KEGG for GProfiler
if (!is.null(gprofiler_reactome) && !is.null(gprofiler_kegg)) {
  gprofiler_combined <- gprofiler_reactome / gprofiler_kegg
} else if (!is.null(gprofiler_reactome)) {
  gprofiler_combined <- gprofiler_reactome
} else if (!is.null(gprofiler_kegg)) {
  gprofiler_combined <- gprofiler_kegg
} else {
  gprofiler_combined <- NULL
}

# Display plots (if they are not NULL)
if (!is.null(cluster_combined)) print(cluster_combined)
if (!is.null(gprofiler_combined)) print(gprofiler_combined)
```

## **📌 Late high dose DOX-specific response**

## **📌 Late high dose DOX-specific response GO Enrichment Clusterprofiler**
```{r GO4, echo=TRUE, message=FALSE, fig.width=12, fig.height=12}
# Perform GO enrichment analysis for BP, MF, and CC
go_enrichment_BP <- enrichGO(gene = prob_all_4,
                             OrgDb = org.Hs.eg.db,
                             keyType = "ENTREZID",
                             universe = background,
                             ont = "BP",
                             pvalueCutoff = 0.05)

go_enrichment_MF <- enrichGO(gene = prob_all_4,
                             OrgDb = org.Hs.eg.db,
                             keyType = "ENTREZID",
                             universe = background,
                             ont = "MF",
                             pvalueCutoff = 0.05)

go_enrichment_CC <- enrichGO(gene = prob_all_4,
                             OrgDb = org.Hs.eg.db,
                             keyType = "ENTREZID",
                             universe = background,
                             ont = "CC",
                             pvalueCutoff = 0.05)

# Convert each enrichment result to a tibble, add a category column, and select top 20 terms
process_enrichment_tibble <- function(enrichment, category) {
  if (is.null(enrichment) || nrow(as.data.frame(enrichment)) == 0) {
    return(tibble(Description = "No enriched terms", neglog = 0, Category = category))
  } else {
    enrichment %>%
      as_tibble() %>%
      mutate(Category = category,
             neglog = -log(p.adjust)) %>% # Add -log(p.adjust) column
      arrange(desc(neglog)) %>%          # Sort by -log(p.adjust)
      slice(1:20)                        # Select top 20 terms
  }
}

BP_Tibble <- process_enrichment_tibble(go_enrichment_BP, "Biological Process")
MF_Tibble <- process_enrichment_tibble(go_enrichment_MF, "Molecular Function")
CC_Tibble <- process_enrichment_tibble(go_enrichment_CC, "Cellular Component")

# Combine all tibbles
combined_GO_Tibble <- bind_rows(BP_Tibble, MF_Tibble, CC_Tibble)
# Function to generate enrichment plots
process_enrichment_plot <- function(tibble, title, color) {
  ggplot(data = tibble, aes(x = neglog, y = reorder(Description, neglog))) +
    geom_bar(stat = "identity", fill = color) +
    labs(x = "-log(p-adjust)",
         y = title,
         title = paste("Top 20", title, "GO Terms")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black"),
      panel.border = element_rect(colour = "black", fill = NA, size = 0.3)
    ) +
    xlim(c(0, max(tibble$neglog) + 1))
}

# Generate separate plots
plot_BP <- process_enrichment_plot(BP_Tibble, "Biological Process", "#2E86C1")
plot_MF <- process_enrichment_plot(MF_Tibble, "Molecular Function", "#28B463")
plot_CC <- process_enrichment_plot(CC_Tibble, "Cellular Component", "#D35400")

# Combine the plots using patchwork
combined_plot <- plot_BP / plot_MF / plot_CC

# Display the combined plot
combined_plot
```


## **📌 Late high dose DOX-specific response GO Enrichment g:Profiler**
```{r gprofiler_GO4, echo=TRUE, message=FALSE, fig.width=12, fig.height=12}

# Load the gprofiler2 package
library(gprofiler2)
library(ggplot2)
library(dplyr)
library(patchwork)

# Perform GO enrichment analysis with gprofiler2
gost_results <- gost(
  query = prob_all_4,           # List of input genes (prob_all_4)
  organism = "hsapiens",   # Human organism
  user_threshold = 0.05,   # Adjusted p-value cutoff
  correction_method = "fdr", # Multiple testing correction
  domain_scope = "custom", # Use custom background
  custom_bg = background,    # Background set of genes
  sources = c("GO:BP", "GO:MF", "GO:CC") # Analyze GO categories
)

# Check if enrichment results exist
if (is.null(gost_results$result) || nrow(gost_results$result) == 0) {
  # If no enriched terms, create a placeholder dataframe
  combined_results <- tibble(
    term_name = "No enriched terms",
    p.adjust = NA,
    source = "N/A",
    Category = "N/A"
  )
} else {
  # Convert results to a data frame
  gost_results_df <- gost_results$result

  # Add a column for adjusted p-value
  gost_results_df <- gost_results_df %>%
    rename(p.adjust = p_value)

  # Separate results for BP, MF, and CC
  BP_results <- gost_results_df %>%
    filter(source == "GO:BP") %>%
    mutate(Category = "Biological Process")

  MF_results <- gost_results_df %>%
    filter(source == "GO:MF") %>%
    mutate(Category = "Molecular Function")

  CC_results <- gost_results_df %>%
    filter(source == "GO:CC") %>%
    mutate(Category = "Cellular Component")

  # Select the top 20 terms by adjusted p-value for each category
  top_BP <- BP_results %>%
    arrange(p.adjust) %>%
    slice_head(n = 20)

  top_MF <- MF_results %>%
    arrange(p.adjust) %>%
    slice_head(n = 20)

  top_CC <- CC_results %>%
    arrange(p.adjust) %>%
    slice_head(n = 20)

  # Combine all categories
  combined_results <- bind_rows(top_BP, top_MF, top_CC)
}

# Ensure all columns are atomic types for CSV export
combined_results_clean <- combined_results %>%
  mutate(across(everything(), ~ if (is.list(.)) sapply(., toString) else .))

# Function for plotting top terms
plot_gprofiler_results <- function(data, title, color) {
  ggplot(data, aes(x = -log10(p.adjust), y = reorder(term_name, -log10(p.adjust)))) +
    geom_bar(stat = "identity", fill = color) +
    labs(
      x = "-log10(Adjusted p-value)",
      y = title,
      title = paste("Top 20", title, "GO Terms")
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black"),
      panel.border = element_rect(colour = "black", fill = NA, size = 0.3)
    )
}

# Check if there are enrichment terms to plot
if (nrow(combined_results) == 1 && combined_results$term_name == "No enriched terms") {
  message("No enriched GO terms found for the input gene set.")
} else {
  # Plot the top 20 terms for each category
  plot_BP <- plot_gprofiler_results(top_BP, "Biological Process", "#2E86C1")
  plot_MF <- plot_gprofiler_results(top_MF, "Molecular Function", "#28B463")
  plot_CC <- plot_gprofiler_results(top_CC, "Cellular Component", "#D35400")

  # Combine the plots using patchwork
  combined_plot <- plot_BP / plot_MF / plot_CC

  # Display the combined plot
  combined_plot
}
```

## **📌 Late high dose DOX-specific response Pathway Enrichment**
```{r pathway_enrichment4, echo=TRUE, message=FALSE, fig.width=12, fig.height=12}

# Load required libraries
library(clusterProfiler)
library(org.Hs.eg.db) # Required for enrichPathway
library(gprofiler2)
library(ggplot2)
library(dplyr)
library(patchwork)
library(ReactomePA)

# Function for ClusterProfiler Reactome & KEGG Analysis
process_clusterProfiler <- function(gene_set, background, category, color, y_title) {
  # Perform enrichment based on the selected category
  if (category == "Reactome") {
    enrichment <- enrichPathway(
      gene = gene_set,
      organism = "human",
      pvalueCutoff = 0.05,
      pAdjustMethod = "BH",
      universe = background
    )
  } else if (category == "KEGG") {
    enrichment <- enrichKEGG(
      gene = gene_set,
      organism = "hsa",
      pvalueCutoff = 0.05,
      pAdjustMethod = "BH",
      universe = background
    )
  }
  
  # Check if enrichment results exist
  if (is.null(enrichment) || nrow(as.data.frame(enrichment)) == 0) {
    message(paste("No significant enrichment found for", category, "in ClusterProfiler"))
    return(NULL)
  }
  
  # Convert results to tibble and process top 20 terms
  enrichment_tibble <- as_tibble(as.data.frame(enrichment)) %>%
    mutate(Category = category,
           neglog = -log10(p.adjust)) %>%  # Compute -log10(p.adjust)
    arrange(desc(neglog)) %>%
    slice_head(n = min(20, nrow(.)))  # Ensure safe slicing
  
  # Generate plot
  plot <- ggplot(enrichment_tibble, aes(x = neglog, y = reorder(Description, neglog))) +
    geom_bar(stat = "identity", fill = color) +
    labs(x = "-log10(Adjusted p-value)",
         y = y_title,
         title = paste("Enriched", category, "Pathways")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black")
    )
  
  return(plot)
}

# Function for gProfiler Reactome & KEGG Analysis
process_gprofiler <- function(gene_set, background, category, color, y_title) {
  # Perform enrichment using gprofiler2
  enrichment <- gost(
    query = gene_set,
    organism = "hsapiens",
    user_threshold = 0.05,
    correction_method = "fdr",
    domain_scope = "custom",
    custom_bg = background,
    sources = category # Either "REAC" or "KEGG"
  )
  
  # Check if enrichment results exist
  if (is.null(enrichment$result) || nrow(enrichment$result) == 0) {
    message(paste("No significant enrichment found for", category, "in gProfiler"))
    return(NULL)
  }
  
  # Convert results to tibble and process top 20 terms
  enrichment_tibble <- enrichment$result %>%
    as_tibble() %>%
    mutate(Category = category,
           neglog = -log10(p_value)) %>%  # Compute -log10(p-value)
    arrange(desc(neglog)) %>%
    slice_head(n = min(20, nrow(.)))  # Ensure safe slicing
  
  # Generate plot
  plot <- ggplot(enrichment_tibble, aes(x = neglog, y = reorder(term_name, neglog))) +
    geom_bar(stat = "identity", fill = color) +
    labs(x = "-log10(p-value)",
         y = y_title,
         title = paste("Enriched", category, "Pathways")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black")
    )
  
  return(plot)
}

# Perform analysis for Reactome and KEGG using ClusterProfiler
cluster_reactome <- process_clusterProfiler(
  gene_set = prob_all_4,
  background = background,
  category = "Reactome",
  color = "#2E86C1",
  y_title = "Reactome Pathways"
)

cluster_kegg <- process_clusterProfiler(
  gene_set = prob_all_4,
  background = background,
  category = "KEGG",
  color = "#28B463",
  y_title = "KEGG Pathways"
)

# Combine Reactome and KEGG for ClusterProfiler
if (!is.null(cluster_reactome) && !is.null(cluster_kegg)) {
  cluster_combined <- cluster_reactome / cluster_kegg
} else if (!is.null(cluster_reactome)) {
  cluster_combined <- cluster_reactome
} else if (!is.null(cluster_kegg)) {
  cluster_combined <- cluster_kegg
} else {
  cluster_combined <- NULL
}

# Perform analysis for Reactome and KEGG using GProfiler
gprofiler_reactome <- process_gprofiler(
  gene_set = prob_all_4,
  background = background,
  category = "REAC",  # Corrected category for Reactome in gProfiler
  color = "#D35400",
  y_title = "Reactome Pathways"
)

gprofiler_kegg <- process_gprofiler(
  gene_set = prob_all_4,
  background = background,
  category = "KEGG",
  color = "#F39C12",
  y_title = "KEGG Pathways"
)

# Combine Reactome and KEGG for GProfiler
if (!is.null(gprofiler_reactome) && !is.null(gprofiler_kegg)) {
  gprofiler_combined <- gprofiler_reactome / gprofiler_kegg
} else if (!is.null(gprofiler_reactome)) {
  gprofiler_combined <- gprofiler_reactome
} else if (!is.null(gprofiler_kegg)) {
  gprofiler_combined <- gprofiler_kegg
} else {
  gprofiler_combined <- NULL
}

# Display plots (if they are not NULL)
if (!is.null(cluster_combined)) print(cluster_combined)
if (!is.null(gprofiler_combined)) print(gprofiler_combined)
```


