---
title: "Corrmotif Analysis"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## **📌 Fit Limma Model Functions**
```{r fit_limma_model_functions, echo=TRUE, results='hide',message=FALSE, warning=FALSE}
## Fit limma model using code as it is found in the original cormotif code. It has
## only been modified to add names to the matrix of t values, as well as the
## limma fits

limmafit.default <- function(exprs,groupid,compid) {
  limmafits  <- list()
  compnum    <- nrow(compid)
  genenum    <- nrow(exprs)
  limmat     <- matrix(0,genenum,compnum)
  limmas2    <- rep(0,compnum)
  limmadf    <- rep(0,compnum)
  limmav0    <- rep(0,compnum)
  limmag1num <- rep(0,compnum)
  limmag2num <- rep(0,compnum)

  rownames(limmat)  <- rownames(exprs)
  colnames(limmat)  <- rownames(compid)
  names(limmas2)    <- rownames(compid)
  names(limmadf)    <- rownames(compid)
  names(limmav0)    <- rownames(compid)
  names(limmag1num) <- rownames(compid)
  names(limmag2num) <- rownames(compid)

  for(i in 1:compnum) {
    selid1 <- which(groupid == compid[i,1])
    selid2 <- which(groupid == compid[i,2])
    eset   <- new("ExpressionSet", exprs=cbind(exprs[,selid1],exprs[,selid2]))
    g1num  <- length(selid1)
    g2num  <- length(selid2)
    designmat <- cbind(base=rep(1,(g1num+g2num)), delta=c(rep(0,g1num),rep(1,g2num)))
    fit <- lmFit(eset,designmat)
    fit <- eBayes(fit)
    limmat[,i] <- fit$t[,2]
    limmas2[i] <- fit$s2.prior
    limmadf[i] <- fit$df.prior
    limmav0[i] <- fit$var.prior[2]
    limmag1num[i] <- g1num
    limmag2num[i] <- g2num
    limmafits[[i]] <- fit

    # log odds
    # w<-sqrt(1+fit$var.prior[2]/(1/g1num+1/g2num))
    # log(0.99)+dt(fit$t[1,2],g1num+g2num-2+fit$df.prior,log=TRUE)-log(0.01)-dt(fit$t[1,2]/w, g1num+g2num-2+fit$df.prior, log=TRUE)+log(w)
  }
  names(limmafits) <- rownames(compid)
  limmacompnum<-nrow(compid)
  result<-list(t       = limmat,
               v0      = limmav0,
               df0     = limmadf,
               s20     = limmas2,
               g1num   = limmag1num,
               g2num   = limmag2num,
               compnum = limmacompnum,
               fits    = limmafits)
}

limmafit.counts <-
  function (exprs, groupid, compid, norm.factor.method = "TMM", voom.normalize.method = "none")
  {
    limmafits  <- list()
    compnum    <- nrow(compid)
    genenum    <- nrow(exprs)
    limmat     <- matrix(NA,genenum,compnum)
    limmas2    <- rep(0,compnum)
    limmadf    <- rep(0,compnum)
    limmav0    <- rep(0,compnum)
    limmag1num <- rep(0,compnum)
    limmag2num <- rep(0,compnum)

    rownames(limmat)  <- rownames(exprs)
    colnames(limmat)  <- rownames(compid)
    names(limmas2)    <- rownames(compid)
    names(limmadf)    <- rownames(compid)
    names(limmav0)    <- rownames(compid)
    names(limmag1num) <- rownames(compid)
    names(limmag2num) <- rownames(compid)

    for (i in 1:compnum) {
      message(paste("Running limma for comparision",i,"/",compnum))
      selid1 <- which(groupid == compid[i, 1])
      selid2 <- which(groupid == compid[i, 2])
      # make a new count data frame
      counts <- cbind(exprs[, selid1], exprs[, selid2])

      # remove NAs
      not.nas <- which(apply(counts, 1, function(x) !any(is.na(x))) == TRUE)

      # runn voom/limma
      d <- DGEList(counts[not.nas,])
      d <- calcNormFactors(d, method = norm.factor.method)
      g1num <- length(selid1)
      g2num <- length(selid2)
      designmat <- cbind(base = rep(1, (g1num + g2num)), delta = c(rep(0,
                                                                       g1num), rep(1, g2num)))

      y <- voom(d, designmat, normalize.method = voom.normalize.method)
      fit <- lmFit(y, designmat)
      fit <- eBayes(fit)

      limmafits[[i]] <- fit
      limmat[not.nas, i] <- fit$t[, 2]
      limmas2[i] <- fit$s2.prior
      limmadf[i] <- fit$df.prior
      limmav0[i] <- fit$var.prior[2]
      limmag1num[i] <- g1num
      limmag2num[i] <- g2num
    }
    limmacompnum <- nrow(compid)
    names(limmafits) <- rownames(compid)
    result <- list(t       = limmat,
                   v0      = limmav0,
                   df0     = limmadf,
                   s20     = limmas2,
                   g1num   = limmag1num,
                   g2num   = limmag2num,
                   compnum = limmacompnum,
                   fits    = limmafits)
  }

limmafit.list <-
  function (fitlist, cmp.idx=2)
  {
    compnum    <- length(fitlist)

    genes <- c()
    for (i in 1:compnum) genes <- unique(c(genes, rownames(fitlist[[i]])))

    genenum    <- length(genes)
    limmat     <- matrix(NA,genenum,compnum)
    limmas2    <- rep(0,compnum)
    limmadf    <- rep(0,compnum)
    limmav0    <- rep(0,compnum)
    limmag1num <- rep(0,compnum)
    limmag2num <- rep(0,compnum)

    rownames(limmat)  <- genes
    colnames(limmat)  <- names(fitlist)
    names(limmas2)    <- names(fitlist)
    names(limmadf)    <- names(fitlist)
    names(limmav0)    <- names(fitlist)
    names(limmag1num) <- names(fitlist)
    names(limmag2num) <- names(fitlist)

    for (i in 1:compnum) {
      this.t <- fitlist[[i]]$t[,cmp.idx]
      limmat[names(this.t),i] <- this.t

      limmas2[i]    <- fitlist[[i]]$s2.prior
      limmadf[i]    <- fitlist[[i]]$df.prior
      limmav0[i]    <- fitlist[[i]]$var.prior[cmp.idx]
      limmag1num[i] <- sum(fitlist[[i]]$design[,cmp.idx]==0)
      limmag2num[i] <- sum(fitlist[[i]]$design[,cmp.idx]==1)
    }

    limmacompnum <- compnum
    result <- list(t       = limmat,
                   v0      = limmav0,
                   df0     = limmadf,
                   s20     = limmas2,
                   g1num   = limmag1num,
                   g2num   = limmag2num,
                   compnum = limmacompnum,
                   fits    = limmafits)

  }

## Rank genes based on statistics
generank<-function(x) {
  xcol<-ncol(x)
  xrow<-nrow(x)
  result<-matrix(0,xrow,xcol)
  z<-(1:1:xrow)
  for(i in 1:xcol) {
    y<-sort(x[,i],decreasing=TRUE,na.last=TRUE)
    result[,i]<-match(x[,i],y)
    result[,i]<-order(result[,i])
  }
  result
}

## Log-likelihood for moderated t under H0
modt.f0.loglike<-function(x,df) {
  a<-dt(x, df, log=TRUE)
  result<-as.vector(a)
  flag<-which(is.na(result)==TRUE)
  result[flag]<-0
  result
}

## Log-likelihood for moderated t under H1
## param=c(df,g1num,g2num,v0)
modt.f1.loglike<-function(x,param) {
  df<-param[1]
  g1num<-param[2]
  g2num<-param[3]
  v0<-param[4]
  w<-sqrt(1+v0/(1/g1num+1/g2num))
  dt(x/w, df, log=TRUE)-log(w)
  a<-dt(x/w, df, log=TRUE)-log(w)
  result<-as.vector(a)
  flag<-which(is.na(result)==TRUE)
  result[flag]<-0
  result
}

## Correlation Motif Fit
cmfit.X<-function(x, type, K=1, tol=1e-3, max.iter=100) {
  ## initialize
  xrow <- nrow(x)
  xcol <- ncol(x)
  loglike0 <- list()
  loglike1 <- list()
  p <- rep(1, K)/K
  q <- matrix(runif(K * xcol), K, xcol)
  q[1, ] <- rep(0.01, xcol)
  for (i in 1:xcol) {
    f0 <- type[[i]][[1]]
    f0param <- type[[i]][[2]]
    f1 <- type[[i]][[3]]
    f1param <- type[[i]][[4]]
    loglike0[[i]] <- f0(x[, i], f0param)
    loglike1[[i]] <- f1(x[, i], f1param)
  }
  condlike <- list()
  for (i in 1:xcol) {
    condlike[[i]] <- matrix(0, xrow, K)
  }
  loglike.old <- -1e+10
  for (i.iter in 1:max.iter) {
    if ((i.iter%%50) == 0) {
      print(paste("We have run the first ", i.iter, " iterations for K=",
                  K, sep = ""))
    }
    err <- tol + 1
    clustlike <- matrix(0, xrow, K)
    #templike <- matrix(0, xrow, 2)
    templike1 <- rep(0, xrow)
    templike2 <- rep(0, xrow)
    for (j in 1:K) {
      for (i in 1:xcol) {
        templike1 <- log(q[j, i]) + loglike1[[i]]
        templike2 <- log(1 - q[j, i]) + loglike0[[i]]
        tempmax <- Rfast::Pmax(templike1, templike2)

        templike1 <- exp(templike1 - tempmax)
        templike2 <- exp(templike2 - tempmax)

        tempsum <- templike1 + templike2
        clustlike[, j] <- clustlike[, j] + tempmax +
          log(tempsum)
        condlike[[i]][, j] <- templike1/tempsum
      }
      clustlike[, j] <- clustlike[, j] + log(p[j])
    }
    #tempmax <- apply(clustlike, 1, max)
    tempmax <- Rfast::rowMaxs(clustlike, value=TRUE)
    for (j in 1:K) {
      clustlike[, j] <- exp(clustlike[, j] - tempmax)
    }
    #tempsum <- apply(clustlike, 1, sum)
    tempsum <- Rfast::rowsums(clustlike)
    for (j in 1:K) {
      clustlike[, j] <- clustlike[, j]/tempsum
    }
    #p.new <- (apply(clustlike, 2, sum) + 1)/(xrow + K)
    p.new <- (Rfast::colsums(clustlike) + 1)/(xrow + K)
    q.new <- matrix(0, K, xcol)
    for (j in 1:K) {
      clustpsum <- sum(clustlike[, j])
      for (i in 1:xcol) {
        q.new[j, i] <- (sum(clustlike[, j] * condlike[[i]][,
                                                           j]) + 1)/(clustpsum + 2)
      }
    }
    err.p <- max(abs(p.new - p)/p)
    err.q <- max(abs(q.new - q)/q)
    err <- max(err.p, err.q)
    loglike.new <- (sum(tempmax + log(tempsum)) + sum(log(p.new)) +
                      sum(log(q.new) + log(1 - q.new)))/xrow
    p <- p.new
    q <- q.new
    loglike.old <- loglike.new
    if (err < tol) {
      break
    }
  }
  clustlike <- matrix(0, xrow, K)
  for (j in 1:K) {
    for (i in 1:xcol) {
      templike1 <- log(q[j, i]) + loglike1[[i]]
      templike2 <- log(1 - q[j, i]) + loglike0[[i]]
      tempmax <- Rfast::Pmax(templike1, templike2)

      templike1 <- exp(templike1 - tempmax)
      templike2 <- exp(templike2 - tempmax)

      tempsum <- templike1 + templike2
      clustlike[, j] <- clustlike[, j] + tempmax + log(tempsum)
      condlike[[i]][, j] <- templike1/tempsum
    }
    clustlike[, j] <- clustlike[, j] + log(p[j])
  }
  #tempmax <- apply(clustlike, 1, max)
  tempmax <- Rfast::rowMaxs(clustlike, value=TRUE)
  for (j in 1:K) {
    clustlike[, j] <- exp(clustlike[, j] - tempmax)
  }
  #tempsum <- apply(clustlike, 1, sum)
  tempsum <- Rfast::rowsums(clustlike)
  for (j in 1:K) {
    clustlike[, j] <- clustlike[, j]/tempsum
  }
  p.post <- matrix(0, xrow, xcol)
  for (j in 1:K) {
    for (i in 1:xcol) {
      p.post[, i] <- p.post[, i] + clustlike[, j] * condlike[[i]][,
                                                                  j]
    }
  }
  loglike.old <- loglike.old - (sum(log(p)) + sum(log(q) +
                                                    log(1 - q)))/xrow
  loglike.old <- loglike.old * xrow
  result <- list(p.post = p.post, motif.prior = p, motif.q = q,
                 loglike = loglike.old, clustlike=clustlike, condlike=condlike)
}

## Fit using (0,0,...,0) and (1,1,...,1)
cmfitall<-function(x, type, tol=1e-3, max.iter=100) {
  ## initialize
  xrow<-nrow(x)
  xcol<-ncol(x)
  loglike0<-list()
  loglike1<-list()
  p<-0.01

  ## compute loglikelihood
  L0<-matrix(0,xrow,1)
  L1<-matrix(0,xrow,1)
  for(i in 1:xcol) {
    f0<-type[[i]][[1]]
    f0param<-type[[i]][[2]]
    f1<-type[[i]][[3]]
    f1param<-type[[i]][[4]]
    loglike0[[i]]<-f0(x[,i],f0param)
    loglike1[[i]]<-f1(x[,i],f1param)
    L0<-L0+loglike0[[i]]
    L1<-L1+loglike1[[i]]
  }


  ## EM algorithm to get MLE of p and q
  loglike.old <- -1e10
  for(i.iter in 1:max.iter) {
    if((i.iter%%50) == 0) {
      print(paste("We have run the first ", i.iter, " iterations",sep=""))
    }
    err<-tol+1

    ## compute posterior cluster membership
    clustlike<-matrix(0,xrow,2)
    clustlike[,1]<-log(1-p)+L0
    clustlike[,2]<-log(p)+L1

    tempmax<-apply(clustlike,1,max)
    for(j in 1:2) {
      clustlike[,j]<-exp(clustlike[,j]-tempmax)
    }
    tempsum<-apply(clustlike,1,sum)


    ## update motif occurrence rate
    for(j in 1:2) {
      clustlike[,j]<-clustlike[,j]/tempsum
    }

    p.new<-(sum(clustlike[,2])+1)/(xrow+2)

    ## evaluate convergence
    err<-abs(p.new-p)/p


    ## evaluate whether the log.likelihood increases
    loglike.new<-(sum(tempmax+log(tempsum))+log(p.new)+log(1-p.new))/xrow

    loglike.old<-loglike.new
    p<-p.new

    if(err<tol) {
      break;
    }
  }

  ## compute posterior p
  clustlike<-matrix(0,xrow,2)
  clustlike[,1]<-log(1-p)+L0
  clustlike[,2]<-log(p)+L1

  tempmax<-apply(clustlike,1,max)
  for(j in 1:2) {
    clustlike[,j]<-exp(clustlike[,j]-tempmax)
  }
  tempsum<-apply(clustlike,1,sum)

  for(j in 1:2) {
    clustlike[,j]<-clustlike[,j]/tempsum
  }

  p.post<-matrix(0,xrow,xcol)
  for(i in 1:xcol) {
    p.post[,i]<-clustlike[,2]
  }

  ## return

  #calculate back loglikelihood
  loglike.old<-loglike.old-(log(p)+log(1-p))/xrow
  loglike.old<-loglike.old*xrow
  result<-list(p.post=p.post, motif.prior=p, loglike=loglike.old)
}

## Fit each dataset separately
cmfitsep<-function(x, type, tol=1e-3, max.iter=100) {
  ## initialize
  xrow<-nrow(x)
  xcol<-ncol(x)
  loglike0<-list()
  loglike1<-list()
  p<-0.01*rep(1,xcol)
  loglike.final<-rep(0,xcol)

  ## compute loglikelihood
  for(i in 1:xcol) {
    f0<-type[[i]][[1]]
    f0param<-type[[i]][[2]]
    f1<-type[[i]][[3]]
    f1param<-type[[i]][[4]]
    loglike0[[i]]<-f0(x[,i],f0param)
    loglike1[[i]]<-f1(x[,i],f1param)
  }

  p.post<-matrix(0,xrow,xcol)

  ## EM algorithm to get MLE of p
  for(coli in 1:xcol) {
    loglike.old <- -1e10
    for(i.iter in 1:max.iter) {
      if((i.iter%%50) == 0) {
        print(paste("We have run the first ", i.iter, " iterations",sep=""))
      }
      err<-tol+1

      ## compute posterior cluster membership
      clustlike<-matrix(0,xrow,2)
      clustlike[,1]<-log(1-p[coli])+loglike0[[coli]]
      clustlike[,2]<-log(p[coli])+loglike1[[coli]]

      tempmax<-apply(clustlike,1,max)
      for(j in 1:2) {
        clustlike[,j]<-exp(clustlike[,j]-tempmax)
      }
      tempsum<-apply(clustlike,1,sum)

      ## evaluate whether the log.likelihood increases
      loglike.new<-sum(tempmax+log(tempsum))/xrow

      ## update motif occurrence rate
      for(j in 1:2) {
        clustlike[,j]<-clustlike[,j]/tempsum
      }

      p.new<-(sum(clustlike[,2]))/(xrow)

      ## evaluate convergence
      err<-abs(p.new-p[coli])/p[coli]
      loglike.old<-loglike.new
      p[coli]<-p.new

      if(err<tol) {
        break;
      }
    }

    ## compute posterior p
    clustlike<-matrix(0,xrow,2)
    clustlike[,1]<-log(1-p[coli])+loglike0[[coli]]
    clustlike[,2]<-log(p[coli])+loglike1[[coli]]

    tempmax<-apply(clustlike,1,max)
    for(j in 1:2) {
      clustlike[,j]<-exp(clustlike[,j]-tempmax)
    }
    tempsum<-apply(clustlike,1,sum)

    for(j in 1:2) {
      clustlike[,j]<-clustlike[,j]/tempsum
    }

    p.post[,coli]<-clustlike[,2]
    loglike.final[coli]<-loglike.old
  }


  ## return
  loglike.final<-loglike.final*xrow
  result<-list(p.post=p.post, motif.prior=p, loglike=loglike.final)
}

## Fit the full model
cmfitfull<-function(x, type, tol=1e-3, max.iter=100) {
  ## initialize
  xrow<-nrow(x)
  xcol<-ncol(x)
  loglike0<-list()
  loglike1<-list()
  K<-2^xcol
  p<-rep(1,K)/K
  pattern<-rep(0,xcol)
  patid<-matrix(0,K,xcol)

  ## compute loglikelihood
  for(i in 1:xcol) {
    f0<-type[[i]][[1]]
    f0param<-type[[i]][[2]]
    f1<-type[[i]][[3]]
    f1param<-type[[i]][[4]]
    loglike0[[i]]<-f0(x[,i],f0param)
    loglike1[[i]]<-f1(x[,i],f1param)
  }
  L<-matrix(0,xrow,K)
  for(i in 1:K)
  {
    patid[i,]<-pattern
    for(j in 1:xcol) {
      if(pattern[j] < 0.5) {
        L[,i]<-L[,i]+loglike0[[j]]
      } else {
        L[,i]<-L[,i]+loglike1[[j]]
      }
    }

    if(i < K) {
      pattern[xcol]<-pattern[xcol]+1
      j<-xcol
      while(pattern[j] > 1) {
        pattern[j]<-0
        j<-j-1
        pattern[j]<-pattern[j]+1
      }
    }
  }

  ## EM algorithm to get MLE of p and q
  loglike.old <- -1e10
  for(i.iter in 1:max.iter) {
    if((i.iter%%50) == 0) {
      print(paste("We have run the first ", i.iter, " iterations",sep=""))
    }
    err<-tol+1

    ## compute posterior cluster membership
    clustlike<-matrix(0,xrow,K)
    for(j in 1:K) {
      clustlike[,j]<-log(p[j])+L[,j]
    }

    tempmax<-apply(clustlike,1,max)
    for(j in 1:K) {
      clustlike[,j]<-exp(clustlike[,j]-tempmax)
    }
    tempsum<-apply(clustlike,1,sum)


    ## update motif occurrence rate
    for(j in 1:K) {
      clustlike[,j]<-clustlike[,j]/tempsum
    }

    p.new<-(apply(clustlike,2,sum)+1)/(xrow+K)

    ## evaluate convergence
    err<-max(abs(p.new-p)/p)

    ## evaluate whether the log.likelihood increases
    loglike.new<-(sum(tempmax+log(tempsum))+sum(log(p.new)))/xrow

    loglike.old<-loglike.new
    p<-p.new

    if(err<tol) {
      break;
    }
  }

  ## compute posterior p
  clustlike<-matrix(0,xrow,K)
  for(j in 1:K) {
    clustlike[,j]<-log(p[j])+L[,j]
  }

  tempmax<-apply(clustlike,1,max)
  for(j in 1:K) {
    clustlike[,j]<-exp(clustlike[,j]-tempmax)
  }
  tempsum<-apply(clustlike,1,sum)

  for(j in 1:K) {
    clustlike[,j]<-clustlike[,j]/tempsum
  }

  p.post<-matrix(0,xrow,xcol)
  for(j in 1:K) {
    for(i in 1:xcol) {
      if(patid[j,i] > 0.5) {
        p.post[,i]<-p.post[,i]+clustlike[,j]
      }
    }
  }

  ## return
  #calculate back loglikelihood
  loglike.old<-loglike.old-sum(log(p))/xrow
  loglike.old<-loglike.old*xrow
  result<-list(p.post=p.post, motif.prior=p, loglike=loglike.old)
}

generatetype<-function(limfitted)
{
  jtype<-list()
  df<-limfitted$g1num+limfitted$g2num-2+limfitted$df0
  for(j in 1:limfitted$compnum)
  {
    jtype[[j]]<-list(f0=modt.f0.loglike, f0.param=df[j], f1=modt.f1.loglike, f1.param=c(df[j],limfitted$g1num[j],limfitted$g2num[j],limfitted$v0[j]))
  }
  jtype
}

cormotiffit <- function(exprs, groupid=NULL, compid=NULL, K=1, tol=1e-3,
                        max.iter=100, BIC=TRUE, norm.factor.method="TMM",
                        voom.normalize.method = "none", runtype=c("logCPM","counts","limmafits"), each=3)
{
  # first I want to do some typechecking. Input can be either a normalized
  # matrix, a count matrix, or a list of limma fits. Dispatch the correct
  # limmafit accordingly.
  # todo: add some typechecking here
  limfitted <- list()
  if (runtype=="counts") {
    limfitted <- limmafit.counts(exprs,groupid,compid, norm.factor.method, voom.normalize.method)
  } else if (runtype=="logCPM") {
    limfitted <- limmafit.default(exprs,groupid,compid)
  } else if (runtype=="limmafits") {
    limfitted <- limmafit.list(exprs)
  } else {
    stop("runtype must be one of 'logCPM', 'counts', or 'limmafits'")
  }


  jtype<-generatetype(limfitted)
  fitresult<-list()
  ks <- rep(K, each = each)
  fitresult <- bplapply(1:length(ks), function(i, x, type, ks, tol, max.iter) {
    cmfit.X(x, type, K = ks[i], tol = tol, max.iter = max.iter)
  }, x=limfitted$t, type=jtype, ks=ks, tol=tol, max.iter=max.iter)

  best.fitresults <- list()
  for (i in 1:length(K)) {
    w.k <- which(ks==K[i])
    this.bic <- c()
    for (j in w.k) this.bic[j] <- -2 * fitresult[[j]]$loglike + (K[i] - 1 + K[i] * limfitted$compnum) * log(dim(limfitted$t)[1])
    w.min <- which(this.bic == min(this.bic, na.rm = TRUE))[1]
    best.fitresults[[i]] <- fitresult[[w.min]]
  }
  fitresult <- best.fitresults

  bic <- rep(0, length(K))
  aic <- rep(0, length(K))
  loglike <- rep(0, length(K))
  for (i in 1:length(K)) loglike[i] <- fitresult[[i]]$loglike
  for (i in 1:length(K)) bic[i] <- -2 * fitresult[[i]]$loglike + (K[i] - 1 + K[i] * limfitted$compnum) * log(dim(limfitted$t)[1])
  for (i in 1:length(K)) aic[i] <- -2 * fitresult[[i]]$loglike + 2 * (K[i] - 1 + K[i] * limfitted$compnum)
  if(BIC==TRUE) {
    bestflag=which(bic==min(bic))
  }
  else {
    bestflag=which(aic==min(aic))
  }
  result<-list(bestmotif=fitresult[[bestflag]],bic=cbind(K,bic),
               aic=cbind(K,aic),loglike=cbind(K,loglike), allmotifs=fitresult)

}

cormotiffitall<-function(exprs,groupid,compid, tol=1e-3, max.iter=100)
{
  limfitted<-limmafit(exprs,groupid,compid)
  jtype<-generatetype(limfitted)
  fitresult<-cmfitall(limfitted$t,type=jtype,tol=1e-3,max.iter=max.iter)
}

cormotiffitsep<-function(exprs,groupid,compid, tol=1e-3, max.iter=100)
{
  limfitted<-limmafit(exprs,groupid,compid)
  jtype<-generatetype(limfitted)
  fitresult<-cmfitsep(limfitted$t,type=jtype,tol=1e-3,max.iter=max.iter)
}

cormotiffitfull<-function(exprs,groupid,compid, tol=1e-3, max.iter=100)
{
  limfitted<-limmafit(exprs,groupid,compid)
  jtype<-generatetype(limfitted)
  fitresult<-cmfitfull(limfitted$t,type=jtype,tol=1e-3,max.iter=max.iter)
}

plotIC<-function(fitted_cormotif)
{
  oldpar<-par(mfrow=c(1,2))
  plot(fitted_cormotif$bic[,1], fitted_cormotif$bic[,2], type="b",xlab="Motif Number", ylab="BIC", main="BIC")
  plot(fitted_cormotif$aic[,1], fitted_cormotif$aic[,2], type="b",xlab="Motif Number", ylab="AIC", main="AIC")
}

plotMotif<-function(fitted_cormotif,title="")
{
  layout(matrix(1:2,ncol=2))
  u<-1:dim(fitted_cormotif$bestmotif$motif.q)[2]
  v<-1:dim(fitted_cormotif$bestmotif$motif.q)[1]
  image(u,v,t(fitted_cormotif$bestmotif$motif.q),
        col=gray(seq(from=1,to=0,by=-0.1)),xlab="Study",yaxt = "n",
        ylab="Corr. Motifs",main=paste(title,"pattern",sep=" "))
  axis(2,at=1:length(v))
  for(i in 1:(length(u)+1))
  {
    abline(v=(i-0.5))
  }
  for(i in 1:(length(v)+1))
  {
    abline(h=(i-0.5))
  }
  Ng=10000
  if(is.null(fitted_cormotif$bestmotif$p.post)!=TRUE)
    Ng=nrow(fitted_cormotif$bestmotif$p.post)
  genecount=floor(fitted_cormotif$bestmotif$motif.p*Ng)
  NK=nrow(fitted_cormotif$bestmotif$motif.q)
  plot(0,0.7,pch=".",xlim=c(0,1.2),ylim=c(0.75,NK+0.25),
       frame.plot=FALSE,axes=FALSE,xlab="No. of genes",ylab="", main=paste(title,"frequency",sep=" "))
  segments(0,0.7,fitted_cormotif$bestmotif$motif.p[1],0.7)
  rect(0,1:NK-0.3,fitted_cormotif$bestmotif$motif.p,1:NK+0.3,
       col="dark grey")
  mtext(1:NK,at=1:NK,side=2,cex=0.8)
  text(fitted_cormotif$bestmotif$motif.p+0.15,1:NK,
       labels=floor(fitted_cormotif$bestmotif$motif.p*Ng))
}
```

## **📌 Load Required Libraries**
```{r load_libraries, echo=TRUE, results='hide',message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(tidyr)
library(dplyr)
library(reshape2)
library(clusterProfiler)
library(AnnotationDbi)
library(org.Hs.eg.db)
library(RColorBrewer)
library(gprofiler2)
library(pheatmap)
library(ggpubr)
library(corrplot)
library(Cormotif)
library(Rfast)
library(BiocParallel)
```

## **📌 Load Corrmotif Data**
```{r load_data, echo=TRUE, message=FALSE}
# Read the Corrmotif Results
Corrmotif <- read.csv("data/Corrmotif/CX5461.csv")
Corrmotif_df <- data.frame(Corrmotif)

rownames(Corrmotif_df) <- Corrmotif_df$Gene
exprs.corrmotif <- as.matrix(Corrmotif_df[,2:109])

# Read group and comparison IDs
groupid_df <- read.csv("data/Corrmotif/groupid.csv")
compid_df <- read.csv("data/Corrmotif/Compid.csv")
```

## **📌 Fit Corrmotif Model (K=1:8)**
```{r Model, eval=FALSE,echo=TRUE, message=FALSE}
set.seed(11111)
# Fit Corrmotif Model (K = 1 to 8)
set.seed(11111)
motif.fitted <- cormotiffit(
  exprs = exprs.corrmotif,
  groupid = groupid_df,
  compid = compid_df,
  K = 1:8,
  max.iter = 1000,
  BIC = TRUE,
  runtype = "logCPM"
)

gene_prob_all <- motif.fitted$bestmotif$p.post
rownames(gene_prob_all) <- rownames(Corrmotif_df)
motif_prob <- motif.fitted$bestmotif$clustlike
rownames(motif_prob) <- rownames(gene_prob_all)
write.csv(motif_prob,"data/cormotif_probability_genelist_all.csv")
```

## **📌 Plot motif**
```{r Motif, echo=TRUE, message=FALSE}
cormotif_all <- readRDS("data/Corrmotif/cormotif_all.RDS")
cormotif_all$bic

plotIC(cormotif_all)

plotMotif(cormotif_all)
```

## **📌 Extract Gene Probabilities**
```{r Prob, echo=TRUE,results='hide', message=FALSE}
# Extract posterior probabilities for genes
gene_prob_all <- cormotif_all$bestmotif$p.post
rownames(gene_prob_all) <- rownames(Corrmotif_df)

# Define gene probability groups
prob_all_1  <- rownames(gene_prob_all[(gene_prob_all[,1] <0.5 & gene_prob_all[,2] <0.5 & gene_prob_all[,3] <0.5 & gene_prob_all[,4] <0.5 & gene_prob_all[,5] < 0.5 & gene_prob_all[,6]<0.5 & gene_prob_all[,7]<0.5 & gene_prob_all[,8]<0.5 & gene_prob_all[,9]<0.5 & gene_prob_all[,10]<0.5 & gene_prob_all[,11]<0.5 & gene_prob_all[,12]<0.5),])

length(prob_all_1)

prob_all_2  <- rownames(gene_prob_all[(gene_prob_all[,1] <0.5 & gene_prob_all[,2] <0.5 & gene_prob_all[,3] >0.5 & gene_prob_all[,4] >0.5 & gene_prob_all[,5] > 0.5 & gene_prob_all[,6]>0.5 & gene_prob_all[,7]<0.5 & gene_prob_all[,8]<0.5 & gene_prob_all[,9]>0.5 & gene_prob_all[,10]>0.1 & gene_prob_all[,11]>0.5 & gene_prob_all[,12]>0.1),])

length(prob_all_2)

prob_all_3  <- rownames(gene_prob_all[(gene_prob_all[,1] <0.5 & gene_prob_all[,2] <0.5 & gene_prob_all[,3] <0.5 & gene_prob_all[,4] <0.5 & gene_prob_all[,5] < 0.5 & gene_prob_all[,6]<0.5 & gene_prob_all[,7]<0.5 & gene_prob_all[,8]>=0.02 & gene_prob_all[,9]>0.5 & gene_prob_all[,10]>0.5 & gene_prob_all[,11]>0.5 & gene_prob_all[,12]>0.5),])

length(prob_all_3)

prob_all_4  <- rownames(gene_prob_all[(gene_prob_all[,1] <0.5 & gene_prob_all[,2] <0.5 & gene_prob_all[,3] <0.5 & gene_prob_all[,4] <0.5 & gene_prob_all[,5] < 0.5 & gene_prob_all[,6]<0.5 & gene_prob_all[,7]<0.5 & gene_prob_all[,8]<0.5 & gene_prob_all[,9]<0.5 & gene_prob_all[,10]>0.5 & gene_prob_all[,11]<0.5 & gene_prob_all[,12]>0.5),])

length(prob_all_4)
```


## **📌 Distribution of Gene Clusters Identified by Corrmotif**
```{r pie-chart, echo=TRUE, message=FALSE, warning=FALSE}

# Load necessary library
library(ggplot2)

# Data
data <- data.frame(
  Category = c("Non response", "CX_DOX shared late response",
               "Dox specific response", "Late high dose DOX specific response"),
  Value = c(7031, 414, 1531, 4874)
)

# Define custom colors
custom_colors <- c("Non response" = "#FF9999",
                   "CX_DOX shared late response" = "#66B2FF",
                   "Dox specific response" = "#99FF99",
                   "Late high dose DOX specific response" = "#FFD700")

# Create pie chart
ggplot(data, aes(x = "", y = Value, fill = Category)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start = 0) +
  geom_text(aes(label = Value),
            position = position_stack(vjust = 0.5),
            size = 4, color = "black") +
  labs(title = "Distribution of Gene Clusters Identified by Corrmotif", x = NULL, y = NULL) +
  theme_void() +
  scale_fill_manual(values = custom_colors)
```

## **📌 Corr-motif Boxplots**
## **📌 Non response (MRPL20)**
```{r box1, echo=TRUE, message=FALSE, warning=FALSE}
# Load the dataset from the data folder
# Load the dataset from the data folder
boxplot1 <- read.csv("data/boxplot1.csv", check.names = FALSE)

boxplot1 <- as.data.frame(boxplot1)

# Print first few column names to check structure
print(colnames(boxplot1))

# Print first few rows to check format
print(head(boxplot1[, 1:10]))

# Filter data for the gene with Entrez ID: 55052
gene_data <- boxplot1[boxplot1$ENTREZID == 55052, ]

# Check if gene_data is empty
if(nrow(gene_data) == 0) {
  stop("No data found for the selected gene ENTIREZID 55052.")
}

# Reshape data from wide to long format
gene_data_long <- melt(gene_data,
                       id.vars = c("ENTREZID", "SYMBOL", "GENENAME"),
                       variable.name = "Sample",
                       value.name = "log2CPM")

# Print some sample names for debugging
print(unique(gene_data_long$Sample))

# Extract details from sample names
gene_data_long <- gene_data_long %>%
  mutate(
    Time = sub(".*_(\\d+)$", "\\1", Sample),
    Concentration = sub(".*_(0\\.\\d)_\\d+$", "\\1", Sample),
    Drug = sub(".*_(CX\\.5461|DOX|VEH)_.*", "\\1", Sample),
    Indv = sub("^([0-9]+\\.[0-9]+)_.*", "\\1", Sample)  # Ensure correct pattern extraction
  )

# Print extracted `Indv` column for debugging
print(unique(gene_data_long$Indv))

# Convert extracted columns to factors
gene_data_long$Time <- factor(gene_data_long$Time, levels = c("3", "24", "48"))
gene_data_long$Concentration <- factor(gene_data_long$Concentration, levels = c("0.1", "0.5"))

# Ensure each concentration facet only shows relevant data
gene_data_long <- gene_data_long %>%
  group_by(Concentration) %>%
  filter(all(Concentration == "0.1") | all(Concentration == "0.5")) %>%
  ungroup()

# Map individual IDs
indv_mapping <- c("75.1" = "1", "78.1" = "2", "87.1" = "3", "17.3" = "4", "84.1" = "5", "90.1" = "6")

# Ensure `Indv` values match the keys in `indv_mapping`
gene_data_long <- gene_data_long %>%
  mutate(
    Indv = ifelse(Indv %in% names(indv_mapping), indv_mapping[Indv], NA)
  )

# Print `Indv` column again to confirm mapping worked
print(unique(gene_data_long$Indv))

# Check for missing values and handle them
gene_data_long <- gene_data_long %>%
  mutate(Indv = ifelse(is.na(Indv), "Unknown", Indv))  # Assign "Unknown" to unidentified individuals

# Define drug colors
drug_palette <- c("CX.5461" = "blue", "DOX" = "red", "VEH" = "green")

# Extract the gene symbol for labeling
gene_symbol <- unique(gene_data_long$SYMBOL)[1]

# Create the boxplot
ggplot(gene_data_long, aes(x = Drug, y = log2CPM, fill = Drug)) +
  geom_boxplot(outlier.shape = NA) +
  scale_fill_manual(values = drug_palette) +
  facet_grid(Concentration ~ Time, labeller = label_both) +
  geom_point(aes(color = Indv), size = 2, alpha = 0.5,
             position = position_jitter(width = -0.3, height = 0)) +
  ggtitle("Non response") +
  labs(
    title = "Non response",
    x = "Drugs",
    y = paste(gene_symbol, " log2CPM")
  ) +
  ylim(0, NA) +
  theme_bw() +
  theme(
    plot.title = element_text(size = rel(1.5), hjust = 0.5),
    axis.title = element_text(size = 15, color = "black"),
    axis.ticks = element_line(linewidth = 1.5),
    axis.line = element_line(linewidth = 1.5),
    axis.text.y = element_text(size = 10, color = "black"),
    axis.text.x = element_text(size = 10, color = "black", angle = 45, hjust = 1),
    strip.text = element_text(size = 12, face = "bold")
  )
```


## **📌 Non response (FAAP20)**
```{r box2, echo=TRUE, message=FALSE, warning=FALSE}
# Load the dataset from the data folder
# Load the dataset from the data folder

# Filter data for the gene with Entrez ID: 199990
gene_data <- boxplot1[boxplot1$ENTREZID == 199990, ]

# Check if gene_data is empty
if(nrow(gene_data) == 0) {
  stop("No data found for the selected gene ENTIREZID 199990.")
}

# Reshape data from wide to long format
gene_data_long <- melt(gene_data,
                       id.vars = c("ENTREZID", "SYMBOL", "GENENAME"),
                       variable.name = "Sample",
                       value.name = "log2CPM")

# Print some sample names for debugging
print(unique(gene_data_long$Sample))

# Extract details from sample names
gene_data_long <- gene_data_long %>%
  mutate(
    Time = sub(".*_(\\d+)$", "\\1", Sample),
    Concentration = sub(".*_(0\\.\\d)_\\d+$", "\\1", Sample),
    Drug = sub(".*_(CX\\.5461|DOX|VEH)_.*", "\\1", Sample),
    Indv = sub("^([0-9]+\\.[0-9]+)_.*", "\\1", Sample)  # Ensure correct pattern extraction
  )

# Print extracted `Indv` column for debugging
print(unique(gene_data_long$Indv))

# Convert extracted columns to factors
gene_data_long$Time <- factor(gene_data_long$Time, levels = c("3", "24", "48"))
gene_data_long$Concentration <- factor(gene_data_long$Concentration, levels = c("0.1", "0.5"))

# Ensure each concentration facet only shows relevant data
gene_data_long <- gene_data_long %>%
  group_by(Concentration) %>%
  filter(all(Concentration == "0.1") | all(Concentration == "0.5")) %>%
  ungroup()

# Map individual IDs
indv_mapping <- c("75.1" = "1", "78.1" = "2", "87.1" = "3", "17.3" = "4", "84.1" = "5", "90.1" = "6")

# Ensure `Indv` values match the keys in `indv_mapping`
gene_data_long <- gene_data_long %>%
  mutate(
    Indv = ifelse(Indv %in% names(indv_mapping), indv_mapping[Indv], NA)
  )

# Print `Indv` column again to confirm mapping worked
print(unique(gene_data_long$Indv))

# Check for missing values and handle them
gene_data_long <- gene_data_long %>%
  mutate(Indv = ifelse(is.na(Indv), "Unknown", Indv))  # Assign "Unknown" to unidentified individuals

# Define drug colors
drug_palette <- c("CX.5461" = "blue", "DOX" = "red", "VEH" = "green")

# Extract the gene symbol for labeling
gene_symbol <- unique(gene_data_long$SYMBOL)[1]

# Create the boxplot
ggplot(gene_data_long, aes(x = Drug, y = log2CPM, fill = Drug)) +
  geom_boxplot(outlier.shape = NA) +
  scale_fill_manual(values = drug_palette) +
  facet_grid(Concentration ~ Time, labeller = label_both) +
  geom_point(aes(color = Indv), size = 2, alpha = 0.5,
             position = position_jitter(width = -0.3, height = 0)) +
  ggtitle("Non response") +
  labs(
    title = "Non response",
    x = "Drugs",
    y = paste(gene_symbol, " log2CPM")
  ) +
  ylim(0, NA) +
  theme_bw() +
  theme(
    plot.title = element_text(size = rel(1.5), hjust = 0.5),
    axis.title = element_text(size = 15, color = "black"),
    axis.ticks = element_line(linewidth = 1.5),
    axis.line = element_line(linewidth = 1.5),
    axis.text.y = element_text(size = 10, color = "black"),
    axis.text.x = element_text(size = 10, color = "black", angle = 45, hjust = 1),
    strip.text = element_text(size = 12, face = "bold")
  )
```


## **📌 CX_DOX shared late response (MPHOSPH9)**
```{r box3, echo=TRUE, message=FALSE, warning=FALSE}
# Load the dataset from the data folder
# Load the dataset from the data folder

# Filter data for the gene with Entrez ID: 10198
gene_data <- boxplot1[boxplot1$ENTREZID == 10198, ]

# Check if gene_data is empty
if(nrow(gene_data) == 0) {
  stop("No data found for the selected gene ENTIREZID 10198.")
}

# Reshape data from wide to long format
gene_data_long <- melt(gene_data,
                       id.vars = c("ENTREZID", "SYMBOL", "GENENAME"),
                       variable.name = "Sample",
                       value.name = "log2CPM")

# Print some sample names for debugging
print(unique(gene_data_long$Sample))

# Extract details from sample names
gene_data_long <- gene_data_long %>%
  mutate(
    Time = sub(".*_(\\d+)$", "\\1", Sample),
    Concentration = sub(".*_(0\\.\\d)_\\d+$", "\\1", Sample),
    Drug = sub(".*_(CX\\.5461|DOX|VEH)_.*", "\\1", Sample),
    Indv = sub("^([0-9]+\\.[0-9]+)_.*", "\\1", Sample)  # Ensure correct pattern extraction
  )

# Print extracted `Indv` column for debugging
print(unique(gene_data_long$Indv))

# Convert extracted columns to factors
gene_data_long$Time <- factor(gene_data_long$Time, levels = c("3", "24", "48"))
gene_data_long$Concentration <- factor(gene_data_long$Concentration, levels = c("0.1", "0.5"))

# Ensure each concentration facet only shows relevant data
gene_data_long <- gene_data_long %>%
  group_by(Concentration) %>%
  filter(all(Concentration == "0.1") | all(Concentration == "0.5")) %>%
  ungroup()

# Map individual IDs
indv_mapping <- c("75.1" = "1", "78.1" = "2", "87.1" = "3", "17.3" = "4", "84.1" = "5", "90.1" = "6")

# Ensure `Indv` values match the keys in `indv_mapping`
gene_data_long <- gene_data_long %>%
  mutate(
    Indv = ifelse(Indv %in% names(indv_mapping), indv_mapping[Indv], NA)
  )

# Print `Indv` column again to confirm mapping worked
print(unique(gene_data_long$Indv))

# Check for missing values and handle them
gene_data_long <- gene_data_long %>%
  mutate(Indv = ifelse(is.na(Indv), "Unknown", Indv))  # Assign "Unknown" to unidentified individuals

# Define drug colors
drug_palette <- c("CX.5461" = "blue", "DOX" = "red", "VEH" = "green")

# Extract the gene symbol for labeling
gene_symbol <- unique(gene_data_long$SYMBOL)[1]

# Create the boxplot
ggplot(gene_data_long, aes(x = Drug, y = log2CPM, fill = Drug)) +
  geom_boxplot(outlier.shape = NA) +
  scale_fill_manual(values = drug_palette) +
  facet_grid(Concentration ~ Time, labeller = label_both) +
  geom_point(aes(color = Indv), size = 2, alpha = 0.5,
             position = position_jitter(width = -0.3, height = 0)) +
  ggtitle("CX_DOX shared late response") +
  labs(
    title = "CX_DOX shared late response",
    x = "Drugs",
    y = paste(gene_symbol, " log2CPM")
  ) +
  ylim(0, NA) +
  theme_bw() +
  theme(
    plot.title = element_text(size = rel(1.5), hjust = 0.5),
    axis.title = element_text(size = 15, color = "black"),
    axis.ticks = element_line(linewidth = 1.5),
    axis.line = element_line(linewidth = 1.5),
    axis.text.y = element_text(size = 10, color = "black"),
    axis.text.x = element_text(size = 10, color = "black", angle = 45, hjust = 1),
    strip.text = element_text(size = 12, face = "bold")
  )
```


## **📌 CX_DOX shared late response (SVIP)**
```{r box5, echo=TRUE, message=FALSE, warning=FALSE}
# Load the dataset from the data folder
# Load the dataset from the data folder

# Filter data for the gene with Entrez ID: 258010
gene_data <- boxplot1[boxplot1$ENTREZID == 258010, ]

# Check if gene_data is empty
if(nrow(gene_data) == 0) {
  stop("No data found for the selected gene ENTIREZID 258010.")
}

# Reshape data from wide to long format
gene_data_long <- melt(gene_data,
                       id.vars = c("ENTREZID", "SYMBOL", "GENENAME"),
                       variable.name = "Sample",
                       value.name = "log2CPM")

# Print some sample names for debugging
print(unique(gene_data_long$Sample))

# Extract details from sample names
gene_data_long <- gene_data_long %>%
  mutate(
    Time = sub(".*_(\\d+)$", "\\1", Sample),
    Concentration = sub(".*_(0\\.\\d)_\\d+$", "\\1", Sample),
    Drug = sub(".*_(CX\\.5461|DOX|VEH)_.*", "\\1", Sample),
    Indv = sub("^([0-9]+\\.[0-9]+)_.*", "\\1", Sample)  # Ensure correct pattern extraction
  )

# Print extracted `Indv` column for debugging
print(unique(gene_data_long$Indv))

# Convert extracted columns to factors
gene_data_long$Time <- factor(gene_data_long$Time, levels = c("3", "24", "48"))
gene_data_long$Concentration <- factor(gene_data_long$Concentration, levels = c("0.1", "0.5"))

# Ensure each concentration facet only shows relevant data
gene_data_long <- gene_data_long %>%
  group_by(Concentration) %>%
  filter(all(Concentration == "0.1") | all(Concentration == "0.5")) %>%
  ungroup()

# Map individual IDs
indv_mapping <- c("75.1" = "1", "78.1" = "2", "87.1" = "3", "17.3" = "4", "84.1" = "5", "90.1" = "6")

# Ensure `Indv` values match the keys in `indv_mapping`
gene_data_long <- gene_data_long %>%
  mutate(
    Indv = ifelse(Indv %in% names(indv_mapping), indv_mapping[Indv], NA)
  )

# Print `Indv` column again to confirm mapping worked
print(unique(gene_data_long$Indv))

# Check for missing values and handle them
gene_data_long <- gene_data_long %>%
  mutate(Indv = ifelse(is.na(Indv), "Unknown", Indv))  # Assign "Unknown" to unidentified individuals

# Define drug colors
drug_palette <- c("CX.5461" = "blue", "DOX" = "red", "VEH" = "green")

# Extract the gene symbol for labeling
gene_symbol <- unique(gene_data_long$SYMBOL)[1]

# Create the boxplot
ggplot(gene_data_long, aes(x = Drug, y = log2CPM, fill = Drug)) +
  geom_boxplot(outlier.shape = NA) +
  scale_fill_manual(values = drug_palette) +
  facet_grid(Concentration ~ Time, labeller = label_both) +
  geom_point(aes(color = Indv), size = 2, alpha = 0.5,
             position = position_jitter(width = -0.3, height = 0)) +
  ggtitle("CX_DOX shared late response") +
  labs(
    title = "CX_DOX shared late response",
    x = "Drugs",
    y = paste(gene_symbol, " log2CPM")
  ) +
  ylim(0, NA) +
  theme_bw() +
  theme(
    plot.title = element_text(size = rel(1.5), hjust = 0.5),
    axis.title = element_text(size = 15, color = "black"),
    axis.ticks = element_line(linewidth = 1.5),
    axis.line = element_line(linewidth = 1.5),
    axis.text.y = element_text(size = 10, color = "black"),
    axis.text.x = element_text(size = 10, color = "black", angle = 45, hjust = 1),
    strip.text = element_text(size = 12, face = "bold")
  )
```


## **📌 DOX-Specific Response (ESPN)**
```{r box6, echo=TRUE, message=FALSE, warning=FALSE}
# Load the dataset from the data folder
# Load the dataset from the data folder

# Filter data for the gene with Entrez ID: 	83715
gene_data <- boxplot1[boxplot1$ENTREZID == 	83715, ]

# Check if gene_data is empty
if(nrow(gene_data) == 0) {
  stop("No data found for the selected gene ENTIREZID 	
83715.")
}

# Reshape data from wide to long format
gene_data_long <- melt(gene_data,
                       id.vars = c("ENTREZID", "SYMBOL", "GENENAME"),
                       variable.name = "Sample",
                       value.name = "log2CPM")

# Print some sample names for debugging
print(unique(gene_data_long$Sample))

# Extract details from sample names
gene_data_long <- gene_data_long %>%
  mutate(
    Time = sub(".*_(\\d+)$", "\\1", Sample),
    Concentration = sub(".*_(0\\.\\d)_\\d+$", "\\1", Sample),
    Drug = sub(".*_(CX\\.5461|DOX|VEH)_.*", "\\1", Sample),
    Indv = sub("^([0-9]+\\.[0-9]+)_.*", "\\1", Sample)  # Ensure correct pattern extraction
  )

# Print extracted `Indv` column for debugging
print(unique(gene_data_long$Indv))

# Convert extracted columns to factors
gene_data_long$Time <- factor(gene_data_long$Time, levels = c("3", "24", "48"))
gene_data_long$Concentration <- factor(gene_data_long$Concentration, levels = c("0.1", "0.5"))

# Ensure each concentration facet only shows relevant data
gene_data_long <- gene_data_long %>%
  group_by(Concentration) %>%
  filter(all(Concentration == "0.1") | all(Concentration == "0.5")) %>%
  ungroup()

# Map individual IDs
indv_mapping <- c("75.1" = "1", "78.1" = "2", "87.1" = "3", "17.3" = "4", "84.1" = "5", "90.1" = "6")

# Ensure `Indv` values match the keys in `indv_mapping`
gene_data_long <- gene_data_long %>%
  mutate(
    Indv = ifelse(Indv %in% names(indv_mapping), indv_mapping[Indv], NA)
  )

# Print `Indv` column again to confirm mapping worked
print(unique(gene_data_long$Indv))

# Check for missing values and handle them
gene_data_long <- gene_data_long %>%
  mutate(Indv = ifelse(is.na(Indv), "Unknown", Indv))  # Assign "Unknown" to unidentified individuals

# Define drug colors
drug_palette <- c("CX.5461" = "blue", "DOX" = "red", "VEH" = "green")

# Extract the gene symbol for labeling
gene_symbol <- unique(gene_data_long$SYMBOL)[1]

# Create the boxplot
ggplot(gene_data_long, aes(x = Drug, y = log2CPM, fill = Drug)) +
  geom_boxplot(outlier.shape = NA) +
  scale_fill_manual(values = drug_palette) +
  facet_grid(Concentration ~ Time, labeller = label_both) +
  geom_point(aes(color = Indv), size = 2, alpha = 0.5,
             position = position_jitter(width = -0.3, height = 0)) +
  ggtitle("DOX-Specific Response") +
  labs(
    title = "DOX-Specific Response",
    x = "Drugs",
    y = paste(gene_symbol, " log2CPM")
  ) +
  ylim(0, NA) +
  theme_bw() +
  theme(
    plot.title = element_text(size = rel(1.5), hjust = 0.5),
    axis.title = element_text(size = 15, color = "black"),
    axis.ticks = element_line(linewidth = 1.5),
    axis.line = element_line(linewidth = 1.5),
    axis.text.y = element_text(size = 10, color = "black"),
    axis.text.x = element_text(size = 10, color = "black", angle = 45, hjust = 1),
    strip.text = element_text(size = 12, face = "bold")
  )
```

## **📌 DOX-Specific Response (FBXO2)**
```{r box7, echo=TRUE, message=FALSE, warning=FALSE}
# Load the dataset from the data folder
# Load the dataset from the data folder

# Filter data for the gene with Entrez ID: 	26232
gene_data <- boxplot1[boxplot1$ENTREZID == 	26232, ]

# Check if gene_data is empty
if(nrow(gene_data) == 0) {
  stop("No data found for the selected gene ENTIREZID 	
26232.")
}

# Reshape data from wide to long format
gene_data_long <- melt(gene_data,
                       id.vars = c("ENTREZID", "SYMBOL", "GENENAME"),
                       variable.name = "Sample",
                       value.name = "log2CPM")

# Print some sample names for debugging
print(unique(gene_data_long$Sample))

# Extract details from sample names
gene_data_long <- gene_data_long %>%
  mutate(
    Time = sub(".*_(\\d+)$", "\\1", Sample),
    Concentration = sub(".*_(0\\.\\d)_\\d+$", "\\1", Sample),
    Drug = sub(".*_(CX\\.5461|DOX|VEH)_.*", "\\1", Sample),
    Indv = sub("^([0-9]+\\.[0-9]+)_.*", "\\1", Sample)  # Ensure correct pattern extraction
  )

# Print extracted `Indv` column for debugging
print(unique(gene_data_long$Indv))

# Convert extracted columns to factors
gene_data_long$Time <- factor(gene_data_long$Time, levels = c("3", "24", "48"))
gene_data_long$Concentration <- factor(gene_data_long$Concentration, levels = c("0.1", "0.5"))

# Ensure each concentration facet only shows relevant data
gene_data_long <- gene_data_long %>%
  group_by(Concentration) %>%
  filter(all(Concentration == "0.1") | all(Concentration == "0.5")) %>%
  ungroup()

# Map individual IDs
indv_mapping <- c("75.1" = "1", "78.1" = "2", "87.1" = "3", "17.3" = "4", "84.1" = "5", "90.1" = "6")

# Ensure `Indv` values match the keys in `indv_mapping`
gene_data_long <- gene_data_long %>%
  mutate(
    Indv = ifelse(Indv %in% names(indv_mapping), indv_mapping[Indv], NA)
  )

# Print `Indv` column again to confirm mapping worked
print(unique(gene_data_long$Indv))

# Check for missing values and handle them
gene_data_long <- gene_data_long %>%
  mutate(Indv = ifelse(is.na(Indv), "Unknown", Indv))  # Assign "Unknown" to unidentified individuals

# Define drug colors
drug_palette <- c("CX.5461" = "blue", "DOX" = "red", "VEH" = "green")

# Extract the gene symbol for labeling
gene_symbol <- unique(gene_data_long$SYMBOL)[1]

# Create the boxplot
ggplot(gene_data_long, aes(x = Drug, y = log2CPM, fill = Drug)) +
  geom_boxplot(outlier.shape = NA) +
  scale_fill_manual(values = drug_palette) +
  facet_grid(Concentration ~ Time, labeller = label_both) +
  geom_point(aes(color = Indv), size = 2, alpha = 0.5,
             position = position_jitter(width = -0.3, height = 0)) +
  ggtitle("DOX-Specific Response") +
  labs(
    title = "DOX-Specific Response",
    x = "Drugs",
    y = paste(gene_symbol, " log2CPM")
  ) +
  ylim(0, NA) +
  theme_bw() +
  theme(
    plot.title = element_text(size = rel(1.5), hjust = 0.5),
    axis.title = element_text(size = 15, color = "black"),
    axis.ticks = element_line(linewidth = 1.5),
    axis.line = element_line(linewidth = 1.5),
    axis.text.y = element_text(size = 10, color = "black"),
    axis.text.x = element_text(size = 10, color = "black", angle = 45, hjust = 1),
    strip.text = element_text(size = 12, face = "bold")
  )
```


## **📌 Late high dose DOX-specific response (EPHX1)**
```{r box8, echo=TRUE, message=FALSE, warning=FALSE}
# Load the dataset from the data folder
# Load the dataset from the data folder

# Filter data for the gene with Entrez ID: 	2052
gene_data <- boxplot1[boxplot1$ENTREZID == 	2052, ]

# Check if gene_data is empty
if(nrow(gene_data) == 0) {
  stop("No data found for the selected gene ENTIREZID 	
2052.")
}

# Reshape data from wide to long format
gene_data_long <- melt(gene_data,
                       id.vars = c("ENTREZID", "SYMBOL", "GENENAME"),
                       variable.name = "Sample",
                       value.name = "log2CPM")

# Print some sample names for debugging
print(unique(gene_data_long$Sample))

# Extract details from sample names
gene_data_long <- gene_data_long %>%
  mutate(
    Time = sub(".*_(\\d+)$", "\\1", Sample),
    Concentration = sub(".*_(0\\.\\d)_\\d+$", "\\1", Sample),
    Drug = sub(".*_(CX\\.5461|DOX|VEH)_.*", "\\1", Sample),
    Indv = sub("^([0-9]+\\.[0-9]+)_.*", "\\1", Sample)  # Ensure correct pattern extraction
  )

# Print extracted `Indv` column for debugging
print(unique(gene_data_long$Indv))

# Convert extracted columns to factors
gene_data_long$Time <- factor(gene_data_long$Time, levels = c("3", "24", "48"))
gene_data_long$Concentration <- factor(gene_data_long$Concentration, levels = c("0.1", "0.5"))

# Ensure each concentration facet only shows relevant data
gene_data_long <- gene_data_long %>%
  group_by(Concentration) %>%
  filter(all(Concentration == "0.1") | all(Concentration == "0.5")) %>%
  ungroup()

# Map individual IDs
indv_mapping <- c("75.1" = "1", "78.1" = "2", "87.1" = "3", "17.3" = "4", "84.1" = "5", "90.1" = "6")

# Ensure `Indv` values match the keys in `indv_mapping`
gene_data_long <- gene_data_long %>%
  mutate(
    Indv = ifelse(Indv %in% names(indv_mapping), indv_mapping[Indv], NA)
  )

# Print `Indv` column again to confirm mapping worked
print(unique(gene_data_long$Indv))

# Check for missing values and handle them
gene_data_long <- gene_data_long %>%
  mutate(Indv = ifelse(is.na(Indv), "Unknown", Indv))  # Assign "Unknown" to unidentified individuals

# Define drug colors
drug_palette <- c("CX.5461" = "blue", "DOX" = "red", "VEH" = "green")

# Extract the gene symbol for labeling
gene_symbol <- unique(gene_data_long$SYMBOL)[1]

# Create the boxplot
ggplot(gene_data_long, aes(x = Drug, y = log2CPM, fill = Drug)) +
  geom_boxplot(outlier.shape = NA) +
  scale_fill_manual(values = drug_palette) +
  facet_grid(Concentration ~ Time, labeller = label_both) +
  geom_point(aes(color = Indv), size = 2, alpha = 0.5,
             position = position_jitter(width = -0.3, height = 0)) +
  ggtitle("Late high dose DOX-specific response") +
  labs(
    title = "Late high dose DOX-specific response",
    x = "Drugs",
    y = paste(gene_symbol, " log2CPM")
  ) +
  ylim(0, NA) +
  theme_bw() +
  theme(
    plot.title = element_text(size = rel(1.5), hjust = 0.5),
    axis.title = element_text(size = 15, color = "black"),
    axis.ticks = element_line(linewidth = 1.5),
    axis.line = element_line(linewidth = 1.5),
    axis.text.y = element_text(size = 10, color = "black"),
    axis.text.x = element_text(size = 10, color = "black", angle = 45, hjust = 1),
    strip.text = element_text(size = 12, face = "bold")
  )
```

## **📌 Late high dose DOX-specific response (OBSCN)**
```{r box9, echo=TRUE, message=FALSE, warning=FALSE}
# Load the dataset from the data folder
# Load the dataset from the data folder

# Filter data for the gene with Entrez ID: 	84033
gene_data <- boxplot1[boxplot1$ENTREZID == 	84033, ]

# Check if gene_data is empty
if(nrow(gene_data) == 0) {
  stop("No data found for the selected gene ENTIREZID 	
84033.")
}

# Reshape data from wide to long format
gene_data_long <- melt(gene_data,
                       id.vars = c("ENTREZID", "SYMBOL", "GENENAME"),
                       variable.name = "Sample",
                       value.name = "log2CPM")

# Print some sample names for debugging
print(unique(gene_data_long$Sample))

# Extract details from sample names
gene_data_long <- gene_data_long %>%
  mutate(
    Time = sub(".*_(\\d+)$", "\\1", Sample),
    Concentration = sub(".*_(0\\.\\d)_\\d+$", "\\1", Sample),
    Drug = sub(".*_(CX\\.5461|DOX|VEH)_.*", "\\1", Sample),
    Indv = sub("^([0-9]+\\.[0-9]+)_.*", "\\1", Sample)  # Ensure correct pattern extraction
  )

# Print extracted `Indv` column for debugging
print(unique(gene_data_long$Indv))

# Convert extracted columns to factors
gene_data_long$Time <- factor(gene_data_long$Time, levels = c("3", "24", "48"))
gene_data_long$Concentration <- factor(gene_data_long$Concentration, levels = c("0.1", "0.5"))

# Ensure each concentration facet only shows relevant data
gene_data_long <- gene_data_long %>%
  group_by(Concentration) %>%
  filter(all(Concentration == "0.1") | all(Concentration == "0.5")) %>%
  ungroup()

# Map individual IDs
indv_mapping <- c("75.1" = "1", "78.1" = "2", "87.1" = "3", "17.3" = "4", "84.1" = "5", "90.1" = "6")

# Ensure `Indv` values match the keys in `indv_mapping`
gene_data_long <- gene_data_long %>%
  mutate(
    Indv = ifelse(Indv %in% names(indv_mapping), indv_mapping[Indv], NA)
  )

# Print `Indv` column again to confirm mapping worked
print(unique(gene_data_long$Indv))

# Check for missing values and handle them
gene_data_long <- gene_data_long %>%
  mutate(Indv = ifelse(is.na(Indv), "Unknown", Indv))  # Assign "Unknown" to unidentified individuals

# Define drug colors
drug_palette <- c("CX.5461" = "blue", "DOX" = "red", "VEH" = "green")

# Extract the gene symbol for labeling
gene_symbol <- unique(gene_data_long$SYMBOL)[1]

# Create the boxplot
ggplot(gene_data_long, aes(x = Drug, y = log2CPM, fill = Drug)) +
  geom_boxplot(outlier.shape = NA) +
  scale_fill_manual(values = drug_palette) +
  facet_grid(Concentration ~ Time, labeller = label_both) +
  geom_point(aes(color = Indv), size = 2, alpha = 0.5,
             position = position_jitter(width = -0.3, height = 0)) +
  ggtitle("Late high dose DOX-specific response") +
  labs(
    title = "Late high dose DOX-specific response",
    x = "Drugs",
    y = paste(gene_symbol, " log2CPM")
  ) +
  ylim(0, NA) +
  theme_bw() +
  theme(
    plot.title = element_text(size = rel(1.5), hjust = 0.5),
    axis.title = element_text(size = 15, color = "black"),
    axis.ticks = element_line(linewidth = 1.5),
    axis.line = element_line(linewidth = 1.5),
    axis.text.y = element_text(size = 10, color = "black"),
    axis.text.x = element_text(size = 10, color = "black", angle = 45, hjust = 1),
    strip.text = element_text(size = 12, face = "bold")
  )
```

## **📌 Corr-motif set Abs logFC boxplots**
## **📌 Read and Process DEG Data**
```{r load_DEGs, echo=TRUE, results='hide', message=FALSE, warning=FALSE}
# Load DEGs Data
CX_0.1_3 <- read.csv("data/DEGs/Toptable_CX_0.1_3.csv")
CX_0.1_24 <- read.csv("data/DEGs/Toptable_CX_0.1_24.csv")
CX_0.1_48 <- read.csv("data/DEGs/Toptable_CX_0.1_48.csv")
CX_0.5_3 <- read.csv("data/DEGs/Toptable_CX_0.5_3.csv")
CX_0.5_24 <- read.csv("data/DEGs/Toptable_CX_0.5_24.csv")
CX_0.5_48 <- read.csv("data/DEGs/Toptable_CX_0.5_48.csv")

DOX_0.1_3 <- read.csv("data/DEGs/Toptable_DOX_0.1_3.csv")
DOX_0.1_24 <- read.csv("data/DEGs/Toptable_DOX_0.1_24.csv")
DOX_0.1_48 <- read.csv("data/DEGs/Toptable_DOX_0.1_48.csv")
DOX_0.5_3 <- read.csv("data/DEGs/Toptable_DOX_0.5_3.csv")
DOX_0.5_24 <- read.csv("data/DEGs/Toptable_DOX_0.5_24.csv")
DOX_0.5_48 <- read.csv("data/DEGs/Toptable_DOX_0.5_48.csv")

Toptable_CX_0.1_3_df<- data.frame(CX_0.1_3)
Toptable_CX_0.1_24_df<- data.frame(CX_0.1_24)
Toptable_CX_0.1_48_df<- data.frame(CX_0.1_48)
Toptable_CX_0.5_3_df<- data.frame(CX_0.5_3)
Toptable_CX_0.5_24_df<- data.frame(CX_0.5_24)
Toptable_CX_0.5_48_df<- data.frame(CX_0.5_48)
Toptable_DOX_0.1_3_df<- data.frame(DOX_0.1_3)
Toptable_DOX_0.1_24_df<- data.frame(DOX_0.1_24)
Toptable_DOX_0.1_48_df<- data.frame(DOX_0.1_48)
Toptable_DOX_0.5_3_df<- data.frame(DOX_0.5_3)
Toptable_DOX_0.5_24_df<- data.frame(DOX_0.5_24)
Toptable_DOX_0.5_48_df<- data.frame(DOX_0.5_48)
```

## **📌 Non response abs logFC boxplots**
```{r box10, echo=TRUE, message=FALSE, warning=FALSE}
# Extract Entrez_IDs from set1
non_response_entrez_ids <- prob_all_1

all_toptables <- bind_rows(
  Toptable_CX_0.1_3_df %>% mutate(Drug = "CX.5461", Concentration = 0.1, Timepoint = "3"),
  Toptable_CX_0.1_24_df %>% mutate(Drug = "CX.5461", Concentration = 0.1, Timepoint = "24"),
  Toptable_CX_0.1_48_df %>% mutate(Drug = "CX.5461", Concentration = 0.1, Timepoint = "48"),
  Toptable_CX_0.5_3_df %>% mutate(Drug = "CX.5461", Concentration = 0.5, Timepoint = "3"),
  Toptable_CX_0.5_24_df %>% mutate(Drug = "CX.5461", Concentration = 0.5, Timepoint = "24"),
  Toptable_CX_0.5_48_df %>% mutate(Drug = "CX.5461", Concentration = 0.5, Timepoint = "48"),
  Toptable_DOX_0.1_3_df %>% mutate(Drug = "DOX", Concentration = 0.1, Timepoint = "3"),
  Toptable_DOX_0.1_24_df %>% mutate(Drug = "DOX", Concentration = 0.1, Timepoint = "24"),
  Toptable_DOX_0.1_48_df %>% mutate(Drug = "DOX", Concentration = 0.1, Timepoint = "48"),
  Toptable_DOX_0.5_3_df %>% mutate(Drug = "DOX", Concentration = 0.5, Timepoint = "3"),
  Toptable_DOX_0.5_24_df %>% mutate(Drug = "DOX", Concentration = 0.5, Timepoint = "24"),
  Toptable_DOX_0.5_48_df %>% mutate(Drug = "DOX", Concentration = 0.5, Timepoint = "48")
)

# Filter and visualize data for shared late response genes
filtered_toptables <- all_toptables %>%
  filter(Entrez_ID %in% non_response_entrez_ids)

# Filter and visualize data with exact same font size and bold styling for concentration and timepoints
filtered_toptables %>%
  mutate(absFC = abs(logFC)) %>%  # Compute absolute log fold change
  mutate(
    Drug = factor(Drug, levels = c("CX.5461", "DOX")),  # Factorize Drug with CX.5461 first
    Timepoint = factor(Timepoint, levels = c("3", "24", "48"), labels = c("Timepoint: 3 hours", "Timepoint: 24 hours", "Timepoint: 48 hours")),
    Concentration = factor(Concentration, levels = c(0.1, 0.5), labels = c("Concentration: 0.1", "Concentration: 0.5"))
  ) %>% 
  ggplot(aes(x = Drug, y = absFC)) +
  geom_boxplot(aes(fill = Drug)) +
  scale_fill_manual(values = c("CX.5461" = "blue", "DOX" = "red")) +  # Updated color palette
  guides(fill = guide_legend(title = "Treatment")) +
  facet_grid(Concentration ~ Timepoint) +  # Facet by Concentration and Timepoint
  theme_bw() +
  xlab("Drugs") +
  ylab("|Log Fold Change|") +
  ggtitle("|Log Fold| for Non-Response Genes") +
  theme(
    plot.title = element_text(size = rel(1.5), hjust = 0.5),
    axis.title = element_text(size = 15, color = "black"),
    axis.line = element_line(linewidth = 1.5),
    strip.background = element_rect(fill = "gray"),  # Set strip background to gray
    strip.text = element_text(size = 12, color = "black", face = "bold"),  # Unified styling for all facet labels
    axis.text.x = element_text(size = 8, color = "black", angle = 15)
  )
```

## **📌 CX_DOX shared late response abs logFC boxplots**
```{r box11, echo=TRUE, message=FALSE, warning=FALSE}
# Extract Entrez_IDs from set2
shared_late_response_entrez_ids <- prob_all_2

# Combine all subset_toptables into a single data frame with additional metadata columns
all_toptables <- bind_rows(
  Toptable_CX_0.1_3_df %>% mutate(Drug = "CX.5461", Concentration = 0.1, Timepoint = "3"),
  Toptable_CX_0.1_24_df %>% mutate(Drug = "CX.5461", Concentration = 0.1, Timepoint = "24"),
  Toptable_CX_0.1_48_df %>% mutate(Drug = "CX.5461", Concentration = 0.1, Timepoint = "48"),
  Toptable_CX_0.5_3_df %>% mutate(Drug = "CX.5461", Concentration = 0.5, Timepoint = "3"),
  Toptable_CX_0.5_24_df %>% mutate(Drug = "CX.5461", Concentration = 0.5, Timepoint = "24"),
  Toptable_CX_0.5_48_df %>% mutate(Drug = "CX.5461", Concentration = 0.5, Timepoint = "48"),
  Toptable_DOX_0.1_3_df %>% mutate(Drug = "DOX", Concentration = 0.1, Timepoint = "3"),
  Toptable_DOX_0.1_24_df %>% mutate(Drug = "DOX", Concentration = 0.1, Timepoint = "24"),
  Toptable_DOX_0.1_48_df %>% mutate(Drug = "DOX", Concentration = 0.1, Timepoint = "48"),
  Toptable_DOX_0.5_3_df %>% mutate(Drug = "DOX", Concentration = 0.5, Timepoint = "3"),
  Toptable_DOX_0.5_24_df %>% mutate(Drug = "DOX", Concentration = 0.5, Timepoint = "24"),
  Toptable_DOX_0.5_48_df %>% mutate(Drug = "DOX", Concentration = 0.5, Timepoint = "48")
)

# Filter and visualize data for shared late response genes
filtered_toptables <- all_toptables %>%
  filter(Entrez_ID %in% shared_late_response_entrez_ids)

filtered_toptables %>%
  mutate(absFC = abs(logFC)) %>%  # Compute absolute log fold change
  mutate(
    Drug = factor(Drug, levels = c("CX.5461", "DOX")),  # Factorize Drug with CX.5461 first
    Timepoint = factor(Timepoint, levels = c("3", "24", "48"), labels = c("Timepoint: 3 hours", "Timepoint: 24 hours", "Timepoint: 48 hours")),
    Concentration = factor(Concentration, levels = c(0.1, 0.5), labels = c("Concentration: 0.1", "Concentration: 0.5"))
  ) %>%
  ggplot(aes(x = Drug, y = absFC)) +
  geom_boxplot(aes(fill = Drug)) +
  scale_fill_manual(values = c("CX.5461" = "blue", "DOX" = "red")) +  # Updated color palette
  guides(fill = guide_legend(title = "Treatment")) +
  facet_grid(Concentration ~ Timepoint) +  # Facet by Concentration and Timepoint
  theme_bw() +
  xlab("Drugs") +
  ylab("|Log Fold Change|") +
  ggtitle("|Log Fold| for CX_DOX shared late response Genes") +
  theme(
    plot.title = element_text(size = rel(1.5), hjust = 0.5),
    axis.title = element_text(size = 15, color = "black"),
    axis.line = element_line(linewidth = 1.5),
    strip.background = element_rect(fill = "gray"),  # Set strip background to gray
    strip.text = element_text(size = 12, color = "black", face = "bold"),  # Unified styling for all facet labels
    axis.text.x = element_text(size = 8, color = "black", angle = 15)
  )
```

## **📌 DOX Specific response abs logFC boxplots**
```{r box12, echo=TRUE, message=FALSE, warning=FALSE}
# Extract Entrez_IDs from set3
dox_specific_response_entrez_ids <- prob_all_3

# Combine all subset_toptables into a single data frame with additional metadata columns
all_toptables <- bind_rows(
  Toptable_CX_0.1_3_df %>% mutate(Drug = "CX.5461", Concentration = 0.1, Timepoint = "3"),
  Toptable_CX_0.1_24_df %>% mutate(Drug = "CX.5461", Concentration = 0.1, Timepoint = "24"),
  Toptable_CX_0.1_48_df %>% mutate(Drug = "CX.5461", Concentration = 0.1, Timepoint = "48"),
  Toptable_CX_0.5_3_df %>% mutate(Drug = "CX.5461", Concentration = 0.5, Timepoint = "3"),
  Toptable_CX_0.5_24_df %>% mutate(Drug = "CX.5461", Concentration = 0.5, Timepoint = "24"),
  Toptable_CX_0.5_48_df %>% mutate(Drug = "CX.5461", Concentration = 0.5, Timepoint = "48"),
  Toptable_DOX_0.1_3_df %>% mutate(Drug = "DOX", Concentration = 0.1, Timepoint = "3"),
  Toptable_DOX_0.1_24_df %>% mutate(Drug = "DOX", Concentration = 0.1, Timepoint = "24"),
  Toptable_DOX_0.1_48_df %>% mutate(Drug = "DOX", Concentration = 0.1, Timepoint = "48"),
  Toptable_DOX_0.5_3_df %>% mutate(Drug = "DOX", Concentration = 0.5, Timepoint = "3"),
  Toptable_DOX_0.5_24_df %>% mutate(Drug = "DOX", Concentration = 0.5, Timepoint = "24"),
  Toptable_DOX_0.5_48_df %>% mutate(Drug = "DOX", Concentration = 0.5, Timepoint = "48")
)

# Filter and visualize data for DOX-specific response genes
filtered_toptables <- all_toptables %>%
  filter(Entrez_ID %in% dox_specific_response_entrez_ids)

filtered_toptables %>%
  mutate(absFC = abs(logFC)) %>%  # Compute absolute log fold change
  mutate(
    Drug = factor(Drug, levels = c("CX.5461", "DOX")),  # Factorize Drug with CX.5461 first
    Timepoint = factor(Timepoint, levels = c("3", "24", "48"), labels = c("Timepoint: 3 hours", "Timepoint: 24 hours", "Timepoint: 48 hours")),
    Concentration = factor(Concentration, levels = c(0.1, 0.5), labels = c("Concentration: 0.1", "Concentration: 0.5"))
  ) %>%
  ggplot(aes(x = Drug, y = absFC)) +
  geom_boxplot(aes(fill = Drug)) +
  scale_fill_manual(values = c("CX.5461" = "blue", "DOX" = "red")) +  # Updated color palette
  guides(fill = guide_legend(title = "Treatment")) +
  facet_grid(Concentration ~ Timepoint) +  # Facet by Concentration and Timepoint
  theme_bw() +
  xlab("Drugs") +
  ylab("|Log Fold Change|") +
  ggtitle("|Log Fold| for DOX-Specific Response Genes") +
  theme(
    plot.title = element_text(size = rel(1.5), hjust = 0.5),
    axis.title = element_text(size = 15, color = "black"),
    axis.line = element_line(linewidth = 1.5),
    strip.background = element_rect(fill = "gray"),  # Set strip background to gray
    strip.text = element_text(size = 12, color = "black", face = "bold"),  # Unified styling for all facet labels
    axis.text.x = element_text(size = 8, color = "black", angle = 15)
  )
```

## **📌 Late high dose DOX-specific response abs logFC boxplots**
```{r box13, echo=TRUE, message=FALSE, warning=FALSE}
# Extract Entrez_IDs from set4
late_high_dose_dox_entrez_ids <- prob_all_4

# Combine all subset_toptables into a single data frame with additional metadata columns
all_toptables <- bind_rows(
  Toptable_CX_0.1_3_df %>% mutate(Drug = "CX.5461", Concentration = 0.1, Timepoint = "3"),
  Toptable_CX_0.1_24_df %>% mutate(Drug = "CX.5461", Concentration = 0.1, Timepoint = "24"),
  Toptable_CX_0.1_48_df %>% mutate(Drug = "CX.5461", Concentration = 0.1, Timepoint = "48"),
  Toptable_CX_0.5_3_df %>% mutate(Drug = "CX.5461", Concentration = 0.5, Timepoint = "3"),
  Toptable_CX_0.5_24_df %>% mutate(Drug = "CX.5461", Concentration = 0.5, Timepoint = "24"),
  Toptable_CX_0.5_48_df %>% mutate(Drug = "CX.5461", Concentration = 0.5, Timepoint = "48"),
  Toptable_DOX_0.1_3_df %>% mutate(Drug = "DOX", Concentration = 0.1, Timepoint = "3"),
  Toptable_DOX_0.1_24_df %>% mutate(Drug = "DOX", Concentration = 0.1, Timepoint = "24"),
  Toptable_DOX_0.1_48_df %>% mutate(Drug = "DOX", Concentration = 0.1, Timepoint = "48"),
  Toptable_DOX_0.5_3_df %>% mutate(Drug = "DOX", Concentration = 0.5, Timepoint = "3"),
  Toptable_DOX_0.5_24_df %>% mutate(Drug = "DOX", Concentration = 0.5, Timepoint = "24"),
  Toptable_DOX_0.5_48_df %>% mutate(Drug = "DOX", Concentration = 0.5, Timepoint = "48")
)

# Filter and visualize data for late high dose DOX-specific response genes
filtered_toptables <- all_toptables %>%
  filter(Entrez_ID %in% late_high_dose_dox_entrez_ids)

filtered_toptables %>%
  mutate(absFC = abs(logFC)) %>%  # Compute absolute log fold change
  mutate(
    Drug = factor(Drug, levels = c("CX.5461", "DOX")),  # Factorize Drug with CX.5461 first
    Timepoint = factor(Timepoint, levels = c("3", "24", "48"), labels = c("Timepoint: 3 hours", "Timepoint: 24 hours", "Timepoint: 48 hours")),
    Concentration = factor(Concentration, levels = c(0.1, 0.5), labels = c("Concentration: 0.1", "Concentration: 0.5"))
  ) %>%
  ggplot(aes(x = Drug, y = absFC)) +
  geom_boxplot(aes(fill = Drug)) +
  scale_fill_manual(values = c("CX.5461" = "blue", "DOX" = "red")) +  # Updated color palette
  guides(fill = guide_legend(title = "Treatment")) +
  facet_grid(Concentration ~ Timepoint) +  # Facet by Concentration and Timepoint
  theme_bw() +
  xlab("Drugs") +
  ylab("|Log Fold Change|") +
  ggtitle("|Log Fold| for Late High Dose DOX-Specific Response Genes") +
  theme(
    plot.title = element_text(size = rel(1.5), hjust = 0.5),
    axis.title = element_text(size = 15, color = "black"),
    axis.line = element_line(linewidth = 1.5),
    strip.background = element_rect(fill = "gray"),  # Set strip background to gray
    strip.text = element_text(size = 12, color = "black", face = "bold"),  # Unified styling for all facet labels
    axis.text.x = element_text(size = 8, color = "black", angle = 15)
  )
```

## **📌 Corr-motif set mean (Abs logFC) across timepoints**
```{r box14, echo=TRUE, message=FALSE, warning=FALSE, fig.height=12, fig.width=10}
all_toptables <- bind_rows(
  Toptable_CX_0.1_3_df %>% mutate(Drug = "CX.5461", Concentration = 0.1, Timepoint = "3"),
  Toptable_CX_0.1_24_df %>% mutate(Drug = "CX.5461", Concentration = 0.1, Timepoint = "24"),
  Toptable_CX_0.1_48_df %>% mutate(Drug = "CX.5461", Concentration = 0.1, Timepoint = "48"),
  Toptable_CX_0.5_3_df %>% mutate(Drug = "CX.5461", Concentration = 0.5, Timepoint = "3"),
  Toptable_CX_0.5_24_df %>% mutate(Drug = "CX.5461", Concentration = 0.5, Timepoint = "24"),
  Toptable_CX_0.5_48_df %>% mutate(Drug = "CX.5461", Concentration = 0.5, Timepoint = "48"),
  Toptable_DOX_0.1_3_df %>% mutate(Drug = "DOX", Concentration = 0.1, Timepoint = "3"),
  Toptable_DOX_0.1_24_df %>% mutate(Drug = "DOX", Concentration = 0.1, Timepoint = "24"),
  Toptable_DOX_0.1_48_df %>% mutate(Drug = "DOX", Concentration = 0.1, Timepoint = "48"),
  Toptable_DOX_0.5_3_df %>% mutate(Drug = "DOX", Concentration = 0.5, Timepoint = "3"),
  Toptable_DOX_0.5_24_df %>% mutate(Drug = "DOX", Concentration = 0.5, Timepoint = "24"),
  Toptable_DOX_0.5_48_df %>% mutate(Drug = "DOX", Concentration = 0.5, Timepoint = "48")
)

# Convert Entrez_ID in `all_toptables` to character for proper matching
all_toptables <- all_toptables %>% mutate(Entrez_ID = as.character(Entrez_ID))

# Define datasets for filtering (lists of Entrez IDs)
data_sets <- list(
  "Non Response" = prob_all_1,
  "CX_DOX Shared\nLate Response" = prob_all_2,
  "DOX-Specific Response" = prob_all_3,
  "Late High Dose\nDOX-Specific Response" = prob_all_4
)

# Filter datasets and add labels
data_combined <- bind_rows(lapply(names(data_sets), function(label) {
  all_toptables %>%
    filter(Entrez_ID %in% data_sets[[label]]) %>%  # Use list directly for filtering
    mutate(Dataset = label)
}))

# Ensure `logFC` column exists before computing mean absolute logFC
if (!"logFC" %in% colnames(data_combined)) {
  stop("Column `logFC` is missing in `data_combined`. Ensure that `all_toptables` has `logFC`.")
}

# Compute mean absolute logFC grouped by Dataset, Drug, Concentration, and Timepoint
data_summary <- data_combined %>%
  mutate(abs_logFC = abs(logFC)) %>%  # Take absolute log fold change
  group_by(Dataset, Drug, Concentration, Timepoint) %>%
  dplyr::summarize(mean_abs_logFC = mean(abs_logFC, na.rm = TRUE), .groups = "drop") %>%
  as.data.frame()

# Convert factors for plotting
data_summary <- data_summary %>%
  mutate(
    Timepoint = factor(Timepoint, levels = c("3", "24", "48"), labels = c("3 hours", "24 hours", "48 hours")),
    Concentration = factor(Concentration, levels = c(0.1, 0.5)),
    Dataset = factor(Dataset, levels = c("Non Response", "CX_DOX Shared\nLate Response", "DOX-Specific Response", "Late High Dose\nDOX-Specific Response"))
  )

# Define custom drug palette
drug_palette <- c("CX.5461" = "blue", "DOX" = "red")

# Plot the data for absolute logFC
ggplot(data_summary, aes(x = Timepoint, y = mean_abs_logFC, group = Drug, color = Drug)) +
  geom_point(size = 3) +
  geom_line(size = 1.2) +
  scale_color_manual(values = drug_palette) +
  ylim(0, 2) +  # Adjust the Y-axis for better visualization
  facet_grid(Dataset ~ Concentration) +
  theme_bw() +
  labs(
    x = "Timepoints",
    y = "Mean |Log Fold Change|",
    title = "Average |Log Fold Change| Across Motifs and Concentrations",
    color = "Drug"
  ) +
  theme(
    plot.title = element_text(size = rel(1.5), hjust = 0.5),
    axis.title = element_text(size = 15, color = "black"),
    axis.text = element_text(size = 12, color = "black"),
    strip.text = element_text(size = 12, color = "black", face = "bold"),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12)
  )
```

## **📌 Corr-motif set logFC boxplots**

## **📌 Non response logFC boxplots**
```{r box15, echo=TRUE, message=FALSE, warning=FALSE}
# Extract Entrez_IDs from set1
non_response_entrez_ids <- prob_all_1

# Combine all subset_toptables into a single data frame with additional metadata columns
all_toptables <- bind_rows(
  Toptable_CX_0.1_3_df %>% mutate(Drug = "CX.5461", Concentration = 0.1, Timepoint = "3"),
  Toptable_CX_0.1_24_df %>% mutate(Drug = "CX.5461", Concentration = 0.1, Timepoint = "24"),
  Toptable_CX_0.1_48_df %>% mutate(Drug = "CX.5461", Concentration = 0.1, Timepoint = "48"),
  Toptable_CX_0.5_3_df %>% mutate(Drug = "CX.5461", Concentration = 0.5, Timepoint = "3"),
  Toptable_CX_0.5_24_df %>% mutate(Drug = "CX.5461", Concentration = 0.5, Timepoint = "24"),
  Toptable_CX_0.5_48_df %>% mutate(Drug = "CX.5461", Concentration = 0.5, Timepoint = "48"),
  Toptable_DOX_0.1_3_df %>% mutate(Drug = "DOX", Concentration = 0.1, Timepoint = "3"),
  Toptable_DOX_0.1_24_df %>% mutate(Drug = "DOX", Concentration = 0.1, Timepoint = "24"),
  Toptable_DOX_0.1_48_df %>% mutate(Drug = "DOX", Concentration = 0.1, Timepoint = "48"),
  Toptable_DOX_0.5_3_df %>% mutate(Drug = "DOX", Concentration = 0.5, Timepoint = "3"),
  Toptable_DOX_0.5_24_df %>% mutate(Drug = "DOX", Concentration = 0.5, Timepoint = "24"),
  Toptable_DOX_0.5_48_df %>% mutate(Drug = "DOX", Concentration = 0.5, Timepoint = "48")
)


# Filter and visualize data for shared late response genes
filtered_toptables <- all_toptables %>%
  filter(Entrez_ID %in% non_response_entrez_ids)

# Visualization of logFC for non-response genes
filtered_toptables %>%
  mutate(
    Drug = factor(Drug, levels = c("CX.5461", "DOX")),  # Factorize Drug with CX.5461 first
    Timepoint = factor(Timepoint, levels = c("3", "24", "48"), labels = c("Timepoint: 3 hours", "Timepoint: 24 hours", "Timepoint: 48 hours")),
    Concentration = factor(Concentration, levels = c(0.1, 0.5), labels = c("Concentration: 0.1", "Concentration: 0.5"))
  ) %>% 
  ggplot(aes(x = Drug, y = logFC)) +  # Plot logFC directly
  geom_boxplot(aes(fill = Drug)) +
  scale_fill_manual(values = c("CX.5461" = "blue", "DOX" = "red")) +  # Updated color palette
  guides(fill = guide_legend(title = "Treatment")) +
  facet_grid(Concentration ~ Timepoint) +  # Facet by Concentration and Timepoint
  theme_bw() +
  xlab("Drugs") +
  ylab("Log Fold Change") +  # Updated y-axis label
  ggtitle("Log Fold Change for Non-Response Genes") +
  theme(
    plot.title = element_text(size = rel(1.5), hjust = 0.5),
    axis.title = element_text(size = 15, color = "black"),
    axis.line = element_line(linewidth = 1.5),
    strip.background = element_rect(fill = "gray"),  # Set strip background to gray
    strip.text = element_text(size = 12, color = "black", face = "bold"),  # Unified styling for all facet labels
    axis.text.x = element_text(size = 8, color = "black", angle = 15)
  )
```

## **📌 CX_DOX Shared Late Response logFC boxplots**
```{r box16, echo=TRUE, message=FALSE, warning=FALSE}
# Extract Entrez_IDs from set2
cx_dox_shared_entrez_ids <- prob_all_2

# Combine all subset_toptables into a single data frame with additional metadata columns
all_toptables <- bind_rows(
  Toptable_CX_0.1_3_df %>% mutate(Drug = "CX.5461", Concentration = 0.1, Timepoint = "3"),
  Toptable_CX_0.1_24_df %>% mutate(Drug = "CX.5461", Concentration = 0.1, Timepoint = "24"),
  Toptable_CX_0.1_48_df %>% mutate(Drug = "CX.5461", Concentration = 0.1, Timepoint = "48"),
  Toptable_CX_0.5_3_df %>% mutate(Drug = "CX.5461", Concentration = 0.5, Timepoint = "3"),
  Toptable_CX_0.5_24_df %>% mutate(Drug = "CX.5461", Concentration = 0.5, Timepoint = "24"),
  Toptable_CX_0.5_48_df %>% mutate(Drug = "CX.5461", Concentration = 0.5, Timepoint = "48"),
  Toptable_DOX_0.1_3_df %>% mutate(Drug = "DOX", Concentration = 0.1, Timepoint = "3"),
  Toptable_DOX_0.1_24_df %>% mutate(Drug = "DOX", Concentration = 0.1, Timepoint = "24"),
  Toptable_DOX_0.1_48_df %>% mutate(Drug = "DOX", Concentration = 0.1, Timepoint = "48"),
  Toptable_DOX_0.5_3_df %>% mutate(Drug = "DOX", Concentration = 0.5, Timepoint = "3"),
  Toptable_DOX_0.5_24_df %>% mutate(Drug = "DOX", Concentration = 0.5, Timepoint = "24"),
  Toptable_DOX_0.5_48_df %>% mutate(Drug = "DOX", Concentration = 0.5, Timepoint = "48")
)

# Filter and visualize data for CX_DOX shared late response genes
filtered_toptables <- all_toptables %>%
  filter(Entrez_ID %in% cx_dox_shared_entrez_ids)

# Visualization of logFC for CX_DOX shared late response genes
filtered_toptables %>%
  mutate(
    Drug = factor(Drug, levels = c("CX.5461", "DOX")),  # Factorize Drug with CX.5461 first
    Timepoint = factor(Timepoint, levels = c("3", "24", "48"), labels = c("Timepoint: 3 hours", "Timepoint: 24 hours", "Timepoint: 48 hours")),
    Concentration = factor(Concentration, levels = c(0.1, 0.5), labels = c("Concentration: 0.1", "Concentration: 0.5"))
  ) %>% 
  ggplot(aes(x = Drug, y = logFC)) +  # Plot logFC directly
  geom_boxplot(aes(fill = Drug)) +
  scale_fill_manual(values = c("CX.5461" = "blue", "DOX" = "red")) +  # Updated color palette
  guides(fill = guide_legend(title = "Treatment")) +
  facet_grid(Concentration ~ Timepoint) +  # Facet by Concentration and Timepoint
  theme_bw() +
  theme_bw() +
  xlab("Drugs") +
  ylab("Log Fold Change") +  # Updated y-axis label
  ggtitle("Log Fold Change for CX_DOX Shared Late Response Genes") +
  theme(
    plot.title = element_text(size = rel(1.5), hjust = 0.5),
    axis.title = element_text(size = 15, color = "black"),
    axis.line = element_line(linewidth = 1.5),
    strip.background = element_rect(fill = "gray"),  # Set strip background to gray
    strip.text = element_text(size = 12, color = "black", face = "bold"),  # Unified styling for all facet labels
    axis.text.x = element_text(size = 8, color = "black", angle = 15)
  )
```

## **📌 DOX-Specific Response logFC boxplots**
```{r box17, echo=TRUE, message=FALSE, warning=FALSE}
# Extract Entrez_IDs from set3
dox_specific_entrez_ids <- prob_all_3

# Combine all subset_toptables into a single data frame with additional metadata columns
all_toptables <- bind_rows(
  Toptable_CX_0.1_3_df %>% mutate(Drug = "CX.5461", Concentration = 0.1, Timepoint = "3"),
  Toptable_CX_0.1_24_df %>% mutate(Drug = "CX.5461", Concentration = 0.1, Timepoint = "24"),
  Toptable_CX_0.1_48_df %>% mutate(Drug = "CX.5461", Concentration = 0.1, Timepoint = "48"),
  Toptable_CX_0.5_3_df %>% mutate(Drug = "CX.5461", Concentration = 0.5, Timepoint = "3"),
  Toptable_CX_0.5_24_df %>% mutate(Drug = "CX.5461", Concentration = 0.5, Timepoint = "24"),
  Toptable_CX_0.5_48_df %>% mutate(Drug = "CX.5461", Concentration = 0.5, Timepoint = "48"),
  Toptable_DOX_0.1_3_df %>% mutate(Drug = "DOX", Concentration = 0.1, Timepoint = "3"),
  Toptable_DOX_0.1_24_df %>% mutate(Drug = "DOX", Concentration = 0.1, Timepoint = "24"),
  Toptable_DOX_0.1_48_df %>% mutate(Drug = "DOX", Concentration = 0.1, Timepoint = "48"),
  Toptable_DOX_0.5_3_df %>% mutate(Drug = "DOX", Concentration = 0.5, Timepoint = "3"),
  Toptable_DOX_0.5_24_df %>% mutate(Drug = "DOX", Concentration = 0.5, Timepoint = "24"),
  Toptable_DOX_0.5_48_df %>% mutate(Drug = "DOX", Concentration = 0.5, Timepoint = "48")
)

# Filter and visualize data for DOX-specific response genes
filtered_toptables <- all_toptables %>%
  filter(Entrez_ID %in% dox_specific_entrez_ids)

# Visualization of logFC for DOX-specific response genes
filtered_toptables %>%
  mutate(
    Drug = factor(Drug, levels = c("CX.5461", "DOX")),  # Factorize Drug with CX.5461 first
    Timepoint = factor(Timepoint, levels = c("3", "24", "48"), labels = c("Timepoint: 3 hours", "Timepoint: 24 hours", "Timepoint: 48 hours")),
    Concentration = factor(Concentration, levels = c(0.1, 0.5), labels = c("Concentration: 0.1", "Concentration: 0.5"))
  ) %>% 
  ggplot(aes(x = Drug, y = logFC)) +  # Plot logFC directly
  geom_boxplot(aes(fill = Drug)) +
  scale_fill_manual(values = c("CX.5461" = "blue", "DOX" = "red")) +  # Updated color palette
  guides(fill = guide_legend(title = "Treatment")) +
  facet_grid(Concentration ~ Timepoint) +  # Facet by Concentration and Timepoint
  theme_bw() +
  xlab("Drugs") +
  ylab("Log Fold Change") +  # Updated y-axis label
  ggtitle("Log Fold Change for DOX-Specific Response Genes") +
  theme(
    plot.title = element_text(size = rel(1.5), hjust = 0.5),
    axis.title = element_text(size = 15, color = "black"),
    axis.line = element_line(linewidth = 1.5),
    strip.background = element_rect(fill = "gray"),  # Set strip background to gray
    strip.text = element_text(size = 12, color = "black", face = "bold"),  # Unified styling for all facet labels
    axis.text.x = element_text(size = 8, color = "black", angle = 15)
  )
```

## **📌 Late High Dose DOX-Specific Response logFC boxplots**
```{r box18, echo=TRUE, message=FALSE, warning=FALSE}
# Extract Entrez_IDs from set4
late_high_dose_dox_specific_entrez_ids <- prob_all_4

# Combine all subset_toptables into a single data frame with additional metadata columns
all_toptables <- bind_rows(
  Toptable_CX_0.1_3_df %>% mutate(Drug = "CX.5461", Concentration = 0.1, Timepoint = "3"),
  Toptable_CX_0.1_24_df %>% mutate(Drug = "CX.5461", Concentration = 0.1, Timepoint = "24"),
  Toptable_CX_0.1_48_df %>% mutate(Drug = "CX.5461", Concentration = 0.1, Timepoint = "48"),
  Toptable_CX_0.5_3_df %>% mutate(Drug = "CX.5461", Concentration = 0.5, Timepoint = "3"),
  Toptable_CX_0.5_24_df %>% mutate(Drug = "CX.5461", Concentration = 0.5, Timepoint = "24"),
  Toptable_CX_0.5_48_df %>% mutate(Drug = "CX.5461", Concentration = 0.5, Timepoint = "48"),
  Toptable_DOX_0.1_3_df %>% mutate(Drug = "DOX", Concentration = 0.1, Timepoint = "3"),
  Toptable_DOX_0.1_24_df %>% mutate(Drug = "DOX", Concentration = 0.1, Timepoint = "24"),
  Toptable_DOX_0.1_48_df %>% mutate(Drug = "DOX", Concentration = 0.1, Timepoint = "48"),
  Toptable_DOX_0.5_3_df %>% mutate(Drug = "DOX", Concentration = 0.5, Timepoint = "3"),
  Toptable_DOX_0.5_24_df %>% mutate(Drug = "DOX", Concentration = 0.5, Timepoint = "24"),
  Toptable_DOX_0.5_48_df %>% mutate(Drug = "DOX", Concentration = 0.5, Timepoint = "48")
)

# Filter and visualize data for Late High Dose DOX-Specific Response Genes
filtered_toptables <- all_toptables %>%
  filter(Entrez_ID %in% late_high_dose_dox_specific_entrez_ids)

# Visualization of logFC for Late High Dose DOX-Specific Response Genes
filtered_toptables %>%
  mutate(
    Drug = factor(Drug, levels = c("CX.5461", "DOX")),  # Factorize Drug with CX.5461 first
    Timepoint = factor(Timepoint, levels = c("3", "24", "48"), labels = c("Timepoint: 3 hours", "Timepoint: 24 hours", "Timepoint: 48 hours")),
    Concentration = factor(Concentration, levels = c(0.1, 0.5), labels = c("Concentration: 0.1", "Concentration: 0.5"))
  ) %>% 
  ggplot(aes(x = Drug, y = logFC)) +  # Plot logFC directly
  geom_boxplot(aes(fill = Drug)) +
  scale_fill_manual(values = c("CX.5461" = "blue", "DOX" = "red")) +  # Updated color palette
  guides(fill = guide_legend(title = "Treatment")) +
  facet_grid(Concentration ~ Timepoint) +  # Facet by Concentration and Timepoint
  theme_bw() +
  xlab("Drugs") +
  ylab("Log Fold Change") +  # Updated y-axis label
  ggtitle("Log Fold Change for Late High Dose DOX-Specific Response Genes") +
  theme(
    plot.title = element_text(size = rel(1.5), hjust = 0.5),
    axis.title = element_text(size = 15, color = "black"),
    axis.line = element_line(linewidth = 1.5),
    strip.background = element_rect(fill = "gray"),  # Set strip background to gray
    strip.text = element_text(size = 12, color = "black", face = "bold"),  # Unified styling for all facet labels
    axis.text.x = element_text(size = 8, color = "black", angle = 15)
  )
```

## **📌 Corr-motif set mean logFC across timepoints**
```{r box19, echo=TRUE, message=FALSE, warning=FALSE, fig.height=12, fig.width=10}
# Combine all subset_toptables into a single data frame with additional metadata columns
all_toptables <- bind_rows(
  Toptable_CX_0.1_3_df %>% mutate(Drug = "CX.5461", Concentration = 0.1, Timepoint = "3"),
  Toptable_CX_0.1_24_df %>% mutate(Drug = "CX.5461", Concentration = 0.1, Timepoint = "24"),
  Toptable_CX_0.1_48_df %>% mutate(Drug = "CX.5461", Concentration = 0.1, Timepoint = "48"),
  Toptable_CX_0.5_3_df %>% mutate(Drug = "CX.5461", Concentration = 0.5, Timepoint = "3"),
  Toptable_CX_0.5_24_df %>% mutate(Drug = "CX.5461", Concentration = 0.5, Timepoint = "24"),
  Toptable_CX_0.5_48_df %>% mutate(Drug = "CX.5461", Concentration = 0.5, Timepoint = "48"),
  Toptable_DOX_0.1_3_df %>% mutate(Drug = "DOX", Concentration = 0.1, Timepoint = "3"),
  Toptable_DOX_0.1_24_df %>% mutate(Drug = "DOX", Concentration = 0.1, Timepoint = "24"),
  Toptable_DOX_0.1_48_df %>% mutate(Drug = "DOX", Concentration = 0.1, Timepoint = "48"),
  Toptable_DOX_0.5_3_df %>% mutate(Drug = "DOX", Concentration = 0.5, Timepoint = "3"),
  Toptable_DOX_0.5_24_df %>% mutate(Drug = "DOX", Concentration = 0.5, Timepoint = "24"),
  Toptable_DOX_0.5_48_df %>% mutate(Drug = "DOX", Concentration = 0.5, Timepoint = "48")
)

# Convert Entrez_ID in `all_toptables` to character for proper matching
all_toptables <- all_toptables %>% mutate(Entrez_ID = as.character(Entrez_ID))

# Define datasets for filtering (lists of Entrez IDs)
data_sets <- list(
  "Non Response" = prob_all_1,
  "CX_DOX Shared\nLate Response" = prob_all_2,
  "DOX-Specific Response" = prob_all_3,
  "Late High Dose\nDOX-Specific Response" = prob_all_4
)

# Filter datasets and add labels
data_combined <- bind_rows(lapply(names(data_sets), function(label) {
  all_toptables %>%
    filter(Entrez_ID %in% data_sets[[label]]) %>%  # Use list directly for filtering
    mutate(Dataset = label)
}))

# Ensure `logFC` column exists before computing mean
if (!"logFC" %in% colnames(data_combined)) {
  stop("Column `logFC` is missing in `data_combined`. Ensure that `all_toptables` has `logFC`.")
}

# Calculate the average logFC grouped by Dataset, Drug, Concentration, and Timepoint
data_summary <- data_combined %>%
  group_by(Dataset, Drug, Concentration, Timepoint) %>%
  dplyr::summarize(avg_logFC = mean(logFC, na.rm = TRUE), .groups = "drop") %>%
  as.data.frame()

# Convert factors for plotting
data_summary <- data_summary %>%
  mutate(
    Timepoint = factor(Timepoint, levels = c("3", "24", "48"), labels = c("3 hours", "24 hours", "48 hours")),
    Concentration = factor(Concentration, levels = c(0.1, 0.5)),
    Dataset = factor(Dataset, levels = c("Non Response", "CX_DOX Shared\nLate Response", "DOX-Specific Response", "Late High Dose\nDOX-Specific Response"))
  )

# Define custom drug palette
drug_palette <- c("CX.5461" = "blue", "DOX" = "red")

# Plot the data
ggplot(data_summary, aes(x = Timepoint, y = avg_logFC, group = Drug, color = Drug)) +
  geom_point(size = 3) +
  geom_line(size = 1.2) +
  scale_color_manual(values = drug_palette) +
  ylim(-2, 2) +  # Set the Y-axis scale
  facet_grid(Dataset ~ Concentration) +
  theme_bw() +
  labs(
    x = "Timepoints",
    y = "Mean Log Fold Change",
    title = "Average Log Fold Change Across Motifs and Concentrations",
    color = "Drug"
  ) +
  theme(
    plot.title = element_text(size = rel(1.5), hjust = 0.5),
    axis.title = element_text(size = 15, color = "black"),
    axis.text = element_text(size = 12, color = "black"),
    strip.text = element_text(size = 12, color = "black", face = "bold"),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12)
  )
```

## **📌 Save Corr-motif datasets for Gene Ontology analysis**
```{r box20, echo=TRUE, results='hide',message=FALSE, warning=FALSE}
write.csv(data.frame(Entrez_ID = prob_all_1), "data/prob_all_1.csv", row.names = FALSE)
write.csv(data.frame(Entrez_ID = prob_all_2), "data/prob_all_2.csv", row.names = FALSE)
write.csv(data.frame(Entrez_ID = prob_all_3), "data/prob_all_3.csv", row.names = FALSE)
write.csv(data.frame(Entrez_ID = prob_all_4), "data/prob_all_4.csv", row.names = FALSE)
```

