---
title: "Function and Pathway Enrichment Analysis of Corrmotifs (Concentration)"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---
## **📌 Load Required Libraries**
```{r load_libraries, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse) 
library(ggfortify)
library(ggplot2)
library(cluster)
library(edgeR)
library(limma)
library(Homo.sapiens)
library(BiocParallel)
library(qvalue)
library(pheatmap)
library(clusterProfiler)
library(AnnotationDbi)
library(org.Hs.eg.db)
library(RColorBrewer)
library(readr)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(DOSE)

# Load UCSC transcript database
txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene
```

## **📌 0.1 Micromolar**
## **📌 Read and Process Data**
```{r load_DEGs, echo=TRUE, results='hide', message=FALSE, warning=FALSE}
# Load the saved datasets
prob_1_0.1 <- read.csv("data/prob_1_0.1.csv")$Entrez_ID
prob_2_0.1 <- read.csv("data/prob_2_0.1.csv")$Entrez_ID
prob_3_0.1 <- read.csv("data/prob_3_0.1.csv")$Entrez_ID
prob_1_0.5 <- read.csv("data/prob_1_0.5.csv")$Entrez_ID
prob_2_0.5 <- read.csv("data/prob_2_0.5.csv")$Entrez_ID
prob_3_0.5 <- read.csv("data/prob_3_0.5.csv")$Entrez_ID
prob_4_0.5 <- read.csv("data/prob_4_0.5.csv")$Entrez_ID
prob_5_0.5 <- read.csv("data/prob_5_0.5.csv")$Entrez_ID


CX_0.1_3 <- read.csv("data/DEGs/Toptable_CX_0.1_3.csv")
background<-as.character(CX_0.1_3$Entrez_ID)
```

## **📌 Non response (0.1 µM)**
## **📌 GO Enrichment Clusterprofiler**
```{r GO1, echo=TRUE, message=FALSE, fig.width=12, fig.height=12}
# Perform GO enrichment analysis for BP, MF, and CC
go_enrichment_BP <- enrichGO(gene = prob_1_0.1,
                             OrgDb = org.Hs.eg.db,
                             keyType = "ENTREZID",
                             universe = background,
                             ont = "BP",
                             pvalueCutoff = 0.05)

go_enrichment_MF <- enrichGO(gene = prob_1_0.1,
                             OrgDb = org.Hs.eg.db,
                             keyType = "ENTREZID",
                             universe = background,
                             ont = "MF",
                             pvalueCutoff = 0.05)

go_enrichment_CC <- enrichGO(gene = prob_1_0.1,
                             OrgDb = org.Hs.eg.db,
                             keyType = "ENTREZID",
                             universe = background,
                             ont = "CC",
                             pvalueCutoff = 0.05)

# Convert each enrichment result to a tibble, add a category column, and select top 20 terms
process_enrichment_tibble <- function(enrichment, category) {
  if (is.null(enrichment) || nrow(as.data.frame(enrichment)) == 0) {
    return(tibble(Description = "No enriched terms", neglog = 0, Category = category))
  } else {
    enrichment %>%
      as_tibble() %>%
      mutate(Category = category,
             neglog = -log(p.adjust)) %>% # Add -log(p.adjust) column
      arrange(desc(neglog)) %>%          # Sort by -log(p.adjust)
      slice(1:20)                        # Select top 20 terms
  }
}

BP_Tibble <- process_enrichment_tibble(go_enrichment_BP, "Biological Process")
MF_Tibble <- process_enrichment_tibble(go_enrichment_MF, "Molecular Function")
CC_Tibble <- process_enrichment_tibble(go_enrichment_CC, "Cellular Component")

# Combine all tibbles
combined_GO_Tibble <- bind_rows(BP_Tibble, MF_Tibble, CC_Tibble)
# Function to generate enrichment plots
process_enrichment_plot <- function(tibble, title, color) {
  ggplot(data = tibble, aes(x = neglog, y = reorder(Description, neglog))) +
    geom_bar(stat = "identity", fill = color) +
    labs(x = "-log(p-adjust)",
         y = title,
         title = paste("Top 20", title, "GO Terms")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black"),
      panel.border = element_rect(colour = "black", fill = NA, size = 0.3)
    ) +
    xlim(c(0, max(tibble$neglog) + 1))
}

# Generate separate plots
plot_BP <- process_enrichment_plot(BP_Tibble, "Biological Process", "#2E86C1")
plot_MF <- process_enrichment_plot(MF_Tibble, "Molecular Function", "#28B463")
plot_CC <- process_enrichment_plot(CC_Tibble, "Cellular Component", "#D35400")

# Combine the plots using patchwork
combined_plot <- plot_BP / plot_MF / plot_CC

# Display the combined plot
combined_plot
```


## **📌 GO Enrichment g:Profiler**
```{r gprofiler_GO, echo=TRUE, message=FALSE, fig.width=12, fig.height=12}

# Load the gprofiler2 package
library(gprofiler2)
library(ggplot2)
library(dplyr)
library(patchwork)

# Perform GO enrichment analysis with gprofiler2
gost_results <- gost(
  query = prob_1_0.1,           # List of input genes (prob_all_1)
  organism = "hsapiens",   # Human organism
  user_threshold = 0.05,   # Adjusted p-value cutoff
  correction_method = "fdr", # Multiple testing correction
  domain_scope = "custom", # Use custom background
  custom_bg = background,    # Background set of genes
  sources = c("GO:BP", "GO:MF", "GO:CC") # Analyze GO categories
)

# Check if enrichment results exist
if (is.null(gost_results$result) || nrow(gost_results$result) == 0) {
  # If no enriched terms, create a placeholder dataframe
  combined_results <- tibble(
    term_name = "No enriched terms",
    p.adjust = NA,
    source = "N/A",
    Category = "N/A"
  )
} else {
  # Convert results to a data frame
  gost_results_df <- gost_results$result

  # Add a column for adjusted p-value
  gost_results_df <- gost_results_df %>%
    rename(p.adjust = p_value)

  # Separate results for BP, MF, and CC
  BP_results <- gost_results_df %>%
    filter(source == "GO:BP") %>%
    mutate(Category = "Biological Process")

  MF_results <- gost_results_df %>%
    filter(source == "GO:MF") %>%
    mutate(Category = "Molecular Function")

  CC_results <- gost_results_df %>%
    filter(source == "GO:CC") %>%
    mutate(Category = "Cellular Component")

  # Select the top 20 terms by adjusted p-value for each category
  top_BP <- BP_results %>%
    arrange(p.adjust) %>%
    slice_head(n = 20)

  top_MF <- MF_results %>%
    arrange(p.adjust) %>%
    slice_head(n = 20)

  top_CC <- CC_results %>%
    arrange(p.adjust) %>%
    slice_head(n = 20)

  # Combine all categories
  combined_results <- bind_rows(top_BP, top_MF, top_CC)
}

# Ensure all columns are atomic types for CSV export
combined_results_clean <- combined_results %>%
  mutate(across(everything(), ~ if (is.list(.)) sapply(., toString) else .))

# Function for plotting top terms
plot_gprofiler_results <- function(data, title, color) {
  ggplot(data, aes(x = -log10(p.adjust), y = reorder(term_name, -log10(p.adjust)))) +
    geom_bar(stat = "identity", fill = color) +
    labs(
      x = "-log10(Adjusted p-value)",
      y = title,
      title = paste("Top 20", title, "GO Terms")
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black"),
      panel.border = element_rect(colour = "black", fill = NA, size = 0.3)
    )
}

# Check if there are enrichment terms to plot
if (nrow(combined_results) == 1 && combined_results$term_name == "No enriched terms") {
  message("No enriched GO terms found for the input gene set.")
} else {
  # Plot the top 20 terms for each category
  plot_BP <- plot_gprofiler_results(top_BP, "Biological Process", "#2E86C1")
  plot_MF <- plot_gprofiler_results(top_MF, "Molecular Function", "#28B463")
  plot_CC <- plot_gprofiler_results(top_CC, "Cellular Component", "#D35400")

  # Combine the plots using patchwork
  combined_plot <- plot_BP / plot_MF / plot_CC

  # Display the combined plot
  combined_plot
}
```

## **📌 Pathway Enrichment**
```{r pathway_enrichment, echo=TRUE, message=FALSE, fig.width=15, fig.height=12}

# Load required libraries
library(clusterProfiler)
library(org.Hs.eg.db) # Required for enrichPathway
library(gprofiler2)
library(ggplot2)
library(dplyr)
library(patchwork)
library(ReactomePA)

# Function for ClusterProfiler Reactome & KEGG Analysis
process_clusterProfiler <- function(gene_set, background, category, color, y_title) {
  # Perform enrichment based on the selected category
  if (category == "Reactome") {
    enrichment <- enrichPathway(
      gene = gene_set,
      organism = "human",
      pvalueCutoff = 0.05,
      pAdjustMethod = "BH",
      universe = background
    )
  } else if (category == "KEGG") {
    enrichment <- enrichKEGG(
      gene = gene_set,
      organism = "hsa",
      pvalueCutoff = 0.05,
      pAdjustMethod = "BH",
      universe = background
    )
  }
  
  # Check if enrichment results exist
  if (is.null(enrichment) || nrow(as.data.frame(enrichment)) == 0) {
    message(paste("No significant enrichment found for", category, "in ClusterProfiler"))
    return(NULL)
  }
  
  # Convert results to tibble and process top 20 terms
  enrichment_tibble <- as_tibble(as.data.frame(enrichment)) %>%
    mutate(Category = category,
           neglog = -log10(p.adjust)) %>%  # Compute -log10(p.adjust)
    arrange(desc(neglog)) %>%
    slice_head(n = min(20, nrow(.)))  # Ensure safe slicing
  
  # Generate plot
  plot <- ggplot(enrichment_tibble, aes(x = neglog, y = reorder(Description, neglog))) +
    geom_bar(stat = "identity", fill = color) +
    labs(x = "-log10(Adjusted p-value)",
         y = y_title,
         title = paste("Enriched", category, "Pathways")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black")
    )
  
  return(plot)
}

# Function for gProfiler Reactome & KEGG Analysis
process_gprofiler <- function(gene_set, background, category, color, y_title) {
  # Perform enrichment using gprofiler2
  enrichment <- gost(
    query = gene_set,
    organism = "hsapiens",
    user_threshold = 0.05,
    correction_method = "fdr",
    domain_scope = "custom",
    custom_bg = background,
    sources = category # Either "REAC" or "KEGG"
  )
  
  # Check if enrichment results exist
  if (is.null(enrichment$result) || nrow(enrichment$result) == 0) {
    message(paste("No significant enrichment found for", category, "in gProfiler"))
    return(NULL)
  }
  
  # Convert results to tibble and process top 20 terms
  enrichment_tibble <- enrichment$result %>%
    as_tibble() %>%
    mutate(Category = category,
           neglog = -log10(p_value)) %>%  # Compute -log10(p-value)
    arrange(desc(neglog)) %>%
    slice_head(n = min(20, nrow(.)))  # Ensure safe slicing
  
  # Generate plot
  plot <- ggplot(enrichment_tibble, aes(x = neglog, y = reorder(term_name, neglog))) +
    geom_bar(stat = "identity", fill = color) +
    labs(x = "-log10(p-value)",
         y = y_title,
         title = paste("Enriched", category, "Pathways")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black")
    )
  
  return(plot)
}

# Perform analysis for Reactome and KEGG using ClusterProfiler
cluster_reactome <- process_clusterProfiler(
  gene_set = prob_1_0.1,
  background = background,
  category = "Reactome",
  color = "#2E86C1",
  y_title = "Reactome Pathways"
)

cluster_kegg <- process_clusterProfiler(
  gene_set = prob_1_0.1,
  background = background,
  category = "KEGG",
  color = "#28B463",
  y_title = "KEGG Pathways"
)

# Combine Reactome and KEGG for ClusterProfiler
if (!is.null(cluster_reactome) && !is.null(cluster_kegg)) {
  cluster_combined <- cluster_reactome / cluster_kegg
} else if (!is.null(cluster_reactome)) {
  cluster_combined <- cluster_reactome
} else if (!is.null(cluster_kegg)) {
  cluster_combined <- cluster_kegg
} else {
  cluster_combined <- NULL
}

# Perform analysis for Reactome and KEGG using GProfiler
gprofiler_reactome <- process_gprofiler(
  gene_set = prob_1_0.1,
  background = background,
  category = "REAC",  # Corrected category for Reactome in gProfiler
  color = "#D35400",
  y_title = "Reactome Pathways"
)

gprofiler_kegg <- process_gprofiler(
  gene_set = prob_1_0.1,
  background = background,
  category = "KEGG",
  color = "#F39C12",
  y_title = "KEGG Pathways"
)

# Combine Reactome and KEGG for GProfiler
if (!is.null(gprofiler_reactome) && !is.null(gprofiler_kegg)) {
  gprofiler_combined <- gprofiler_reactome / gprofiler_kegg
} else if (!is.null(gprofiler_reactome)) {
  gprofiler_combined <- gprofiler_reactome
} else if (!is.null(gprofiler_kegg)) {
  gprofiler_combined <- gprofiler_kegg
} else {
  gprofiler_combined <- NULL
}

# Display plots (if they are not NULL)
if (!is.null(cluster_combined)) print(cluster_combined)
if (!is.null(gprofiler_combined)) print(gprofiler_combined)
```

## **📌 CX-DOX mid-late response (0.1 µM)**
## **📌 GO Enrichment Clusterprofiler**
```{r GO2, echo=TRUE, message=FALSE, fig.width=12, fig.height=12}
# Perform GO enrichment analysis for BP, MF, and CC
go_enrichment_BP <- enrichGO(gene = prob_2_0.1,
                             OrgDb = org.Hs.eg.db,
                             keyType = "ENTREZID",
                             universe = background,
                             ont = "BP",
                             pvalueCutoff = 0.05)

go_enrichment_MF <- enrichGO(gene = prob_2_0.1,
                             OrgDb = org.Hs.eg.db,
                             keyType = "ENTREZID",
                             universe = background,
                             ont = "MF",
                             pvalueCutoff = 0.05)

go_enrichment_CC <- enrichGO(gene = prob_2_0.1,
                             OrgDb = org.Hs.eg.db,
                             keyType = "ENTREZID",
                             universe = background,
                             ont = "CC",
                             pvalueCutoff = 0.05)

# Convert each enrichment result to a tibble, add a category column, and select top 20 terms
process_enrichment_tibble <- function(enrichment, category) {
  if (is.null(enrichment) || nrow(as.data.frame(enrichment)) == 0) {
    return(tibble(Description = "No enriched terms", neglog = 0, Category = category))
  } else {
    enrichment %>%
      as_tibble() %>%
      mutate(Category = category,
             neglog = -log(p.adjust)) %>% # Add -log(p.adjust) column
      arrange(desc(neglog)) %>%          # Sort by -log(p.adjust)
      slice(1:20)                        # Select top 20 terms
  }
}

BP_Tibble <- process_enrichment_tibble(go_enrichment_BP, "Biological Process")
MF_Tibble <- process_enrichment_tibble(go_enrichment_MF, "Molecular Function")
CC_Tibble <- process_enrichment_tibble(go_enrichment_CC, "Cellular Component")

# Combine all tibbles
combined_GO_Tibble <- bind_rows(BP_Tibble, MF_Tibble, CC_Tibble)
# Function to generate enrichment plots
process_enrichment_plot <- function(tibble, title, color) {
  ggplot(data = tibble, aes(x = neglog, y = reorder(Description, neglog))) +
    geom_bar(stat = "identity", fill = color) +
    labs(x = "-log(p-adjust)",
         y = title,
         title = paste("Top 20", title, "GO Terms")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black"),
      panel.border = element_rect(colour = "black", fill = NA, size = 0.3)
    ) +
    xlim(c(0, max(tibble$neglog) + 1))
}

# Generate separate plots
plot_BP <- process_enrichment_plot(BP_Tibble, "Biological Process", "#2E86C1")
plot_MF <- process_enrichment_plot(MF_Tibble, "Molecular Function", "#28B463")
plot_CC <- process_enrichment_plot(CC_Tibble, "Cellular Component", "#D35400")

# Combine the plots using patchwork
combined_plot <- plot_BP / plot_MF / plot_CC

# Display the combined plot
combined_plot
```


## **📌 GO Enrichment g:Profiler**
```{r gprofiler_GO2, echo=TRUE, message=FALSE, fig.width=12, fig.height=12}

# Load the gprofiler2 package
library(gprofiler2)
library(ggplot2)
library(dplyr)
library(patchwork)

# Perform GO enrichment analysis with gprofiler2
gost_results <- gost(
  query = prob_2_0.1,       
  organism = "hsapiens",   # Human organism
  user_threshold = 0.05,   # Adjusted p-value cutoff
  correction_method = "fdr", # Multiple testing correction
  domain_scope = "custom", # Use custom background
  custom_bg = background,    # Background set of genes
  sources = c("GO:BP", "GO:MF", "GO:CC") # Analyze GO categories
)

# Check if enrichment results exist
if (is.null(gost_results$result) || nrow(gost_results$result) == 0) {
  # If no enriched terms, create a placeholder dataframe
  combined_results <- tibble(
    term_name = "No enriched terms",
    p.adjust = NA,
    source = "N/A",
    Category = "N/A"
  )
} else {
  # Convert results to a data frame
  gost_results_df <- gost_results$result

  # Add a column for adjusted p-value
  gost_results_df <- gost_results_df %>%
    rename(p.adjust = p_value)

  # Separate results for BP, MF, and CC
  BP_results <- gost_results_df %>%
    filter(source == "GO:BP") %>%
    mutate(Category = "Biological Process")

  MF_results <- gost_results_df %>%
    filter(source == "GO:MF") %>%
    mutate(Category = "Molecular Function")

  CC_results <- gost_results_df %>%
    filter(source == "GO:CC") %>%
    mutate(Category = "Cellular Component")

  # Select the top 20 terms by adjusted p-value for each category
  top_BP <- BP_results %>%
    arrange(p.adjust) %>%
    slice_head(n = 20)

  top_MF <- MF_results %>%
    arrange(p.adjust) %>%
    slice_head(n = 20)

  top_CC <- CC_results %>%
    arrange(p.adjust) %>%
    slice_head(n = 20)

  # Combine all categories
  combined_results <- bind_rows(top_BP, top_MF, top_CC)
}

# Ensure all columns are atomic types for CSV export
combined_results_clean <- combined_results %>%
  mutate(across(everything(), ~ if (is.list(.)) sapply(., toString) else .))

# Function for plotting top terms
plot_gprofiler_results <- function(data, title, color) {
  ggplot(data, aes(x = -log10(p.adjust), y = reorder(term_name, -log10(p.adjust)))) +
    geom_bar(stat = "identity", fill = color) +
    labs(
      x = "-log10(Adjusted p-value)",
      y = title,
      title = paste("Top 20", title, "GO Terms")
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black"),
      panel.border = element_rect(colour = "black", fill = NA, size = 0.3)
    )
}

# Check if there are enrichment terms to plot
if (nrow(combined_results) == 1 && combined_results$term_name == "No enriched terms") {
  message("No enriched GO terms found for the input gene set.")
} else {
  # Plot the top 20 terms for each category
  plot_BP <- plot_gprofiler_results(top_BP, "Biological Process", "#2E86C1")
  plot_MF <- plot_gprofiler_results(top_MF, "Molecular Function", "#28B463")
  plot_CC <- plot_gprofiler_results(top_CC, "Cellular Component", "#D35400")

  # Combine the plots using patchwork
  combined_plot <- plot_BP / plot_MF / plot_CC

  # Display the combined plot
  combined_plot
}
```

## **📌 Pathway Enrichment**
```{r pathway_enrichment2, echo=TRUE, message=FALSE, fig.width=15, fig.height=12}

# Load required libraries
library(clusterProfiler)
library(org.Hs.eg.db) # Required for enrichPathway
library(gprofiler2)
library(ggplot2)
library(dplyr)
library(patchwork)
library(ReactomePA)

# Function for ClusterProfiler Reactome & KEGG Analysis
process_clusterProfiler <- function(gene_set, background, category, color, y_title) {
  # Perform enrichment based on the selected category
  if (category == "Reactome") {
    enrichment <- enrichPathway(
      gene = gene_set,
      organism = "human",
      pvalueCutoff = 0.05,
      pAdjustMethod = "BH",
      universe = background
    )
  } else if (category == "KEGG") {
    enrichment <- enrichKEGG(
      gene = gene_set,
      organism = "hsa",
      pvalueCutoff = 0.05,
      pAdjustMethod = "BH",
      universe = background
    )
  }
  
  # Check if enrichment results exist
  if (is.null(enrichment) || nrow(as.data.frame(enrichment)) == 0) {
    message(paste("No significant enrichment found for", category, "in ClusterProfiler"))
    return(NULL)
  }
  
  # Convert results to tibble and process top 20 terms
  enrichment_tibble <- as_tibble(as.data.frame(enrichment)) %>%
    mutate(Category = category,
           neglog = -log10(p.adjust)) %>%  # Compute -log10(p.adjust)
    arrange(desc(neglog)) %>%
    slice_head(n = min(20, nrow(.)))  # Ensure safe slicing
  
  # Generate plot
  plot <- ggplot(enrichment_tibble, aes(x = neglog, y = reorder(Description, neglog))) +
    geom_bar(stat = "identity", fill = color) +
    labs(x = "-log10(Adjusted p-value)",
         y = y_title,
         title = paste("Enriched", category, "Pathways")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black")
    )
  
  return(plot)
}

# Function for gProfiler Reactome & KEGG Analysis
process_gprofiler <- function(gene_set, background, category, color, y_title) {
  # Perform enrichment using gprofiler2
  enrichment <- gost(
    query = gene_set,
    organism = "hsapiens",
    user_threshold = 0.05,
    correction_method = "fdr",
    domain_scope = "custom",
    custom_bg = background,
    sources = category # Either "REAC" or "KEGG"
  )
  
  # Check if enrichment results exist
  if (is.null(enrichment$result) || nrow(enrichment$result) == 0) {
    message(paste("No significant enrichment found for", category, "in gProfiler"))
    return(NULL)
  }
  
  # Convert results to tibble and process top 20 terms
  enrichment_tibble <- enrichment$result %>%
    as_tibble() %>%
    mutate(Category = category,
           neglog = -log10(p_value)) %>%  # Compute -log10(p-value)
    arrange(desc(neglog)) %>%
    slice_head(n = min(20, nrow(.)))  # Ensure safe slicing
  
  # Generate plot
  plot <- ggplot(enrichment_tibble, aes(x = neglog, y = reorder(term_name, neglog))) +
    geom_bar(stat = "identity", fill = color) +
    labs(x = "-log10(p-value)",
         y = y_title,
         title = paste("Enriched", category, "Pathways")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black")
    )
  
  return(plot)
}

# Perform analysis for Reactome and KEGG using ClusterProfiler
cluster_reactome <- process_clusterProfiler(
  gene_set = prob_2_0.1,
  background = background,
  category = "Reactome",
  color = "#2E86C1",
  y_title = "Reactome Pathways"
)

cluster_kegg <- process_clusterProfiler(
  gene_set = prob_2_0.1,
  background = background,
  category = "KEGG",
  color = "#28B463",
  y_title = "KEGG Pathways"
)

# Combine Reactome and KEGG for ClusterProfiler
if (!is.null(cluster_reactome) && !is.null(cluster_kegg)) {
  cluster_combined <- cluster_reactome / cluster_kegg
} else if (!is.null(cluster_reactome)) {
  cluster_combined <- cluster_reactome
} else if (!is.null(cluster_kegg)) {
  cluster_combined <- cluster_kegg
} else {
  cluster_combined <- NULL
}

# Perform analysis for Reactome and KEGG using GProfiler
gprofiler_reactome <- process_gprofiler(
  gene_set = prob_2_0.1,
  background = background,
  category = "REAC",  # Corrected category for Reactome in gProfiler
  color = "#D35400",
  y_title = "Reactome Pathways"
)

gprofiler_kegg <- process_gprofiler(
  gene_set = prob_2_0.1,
  background = background,
  category = "KEGG",
  color = "#F39C12",
  y_title = "KEGG Pathways"
)

# Combine Reactome and KEGG for GProfiler
if (!is.null(gprofiler_reactome) && !is.null(gprofiler_kegg)) {
  gprofiler_combined <- gprofiler_reactome / gprofiler_kegg
} else if (!is.null(gprofiler_reactome)) {
  gprofiler_combined <- gprofiler_reactome
} else if (!is.null(gprofiler_kegg)) {
  gprofiler_combined <- gprofiler_kegg
} else {
  gprofiler_combined <- NULL
}

# Display plots (if they are not NULL)
if (!is.null(cluster_combined)) print(cluster_combined)
if (!is.null(gprofiler_combined)) print(gprofiler_combined)
```

## **📌 DOX only mid-late (0.1 µM)**
## **📌 GO Enrichment Clusterprofiler**
```{r GO3, echo=TRUE, message=FALSE, fig.width=12, fig.height=12}
# Perform GO enrichment analysis for BP, MF, and CC
go_enrichment_BP <- enrichGO(gene = prob_3_0.1,
                             OrgDb = org.Hs.eg.db,
                             keyType = "ENTREZID",
                             universe = background,
                             ont = "BP",
                             pvalueCutoff = 0.05)

go_enrichment_MF <- enrichGO(gene = prob_3_0.1,
                             OrgDb = org.Hs.eg.db,
                             keyType = "ENTREZID",
                             universe = background,
                             ont = "MF",
                             pvalueCutoff = 0.05)

go_enrichment_CC <- enrichGO(gene = prob_3_0.1,
                             OrgDb = org.Hs.eg.db,
                             keyType = "ENTREZID",
                             universe = background,
                             ont = "CC",
                             pvalueCutoff = 0.05)

# Convert each enrichment result to a tibble, add a category column, and select top 20 terms
process_enrichment_tibble <- function(enrichment, category) {
  if (is.null(enrichment) || nrow(as.data.frame(enrichment)) == 0) {
    return(tibble(Description = "No enriched terms", neglog = 0, Category = category))
  } else {
    enrichment %>%
      as_tibble() %>%
      mutate(Category = category,
             neglog = -log(p.adjust)) %>% # Add -log(p.adjust) column
      arrange(desc(neglog)) %>%          # Sort by -log(p.adjust)
      slice(1:20)                        # Select top 20 terms
  }
}

BP_Tibble <- process_enrichment_tibble(go_enrichment_BP, "Biological Process")
MF_Tibble <- process_enrichment_tibble(go_enrichment_MF, "Molecular Function")
CC_Tibble <- process_enrichment_tibble(go_enrichment_CC, "Cellular Component")

# Combine all tibbles
combined_GO_Tibble <- bind_rows(BP_Tibble, MF_Tibble, CC_Tibble)
# Function to generate enrichment plots
process_enrichment_plot <- function(tibble, title, color) {
  ggplot(data = tibble, aes(x = neglog, y = reorder(Description, neglog))) +
    geom_bar(stat = "identity", fill = color) +
    labs(x = "-log(p-adjust)",
         y = title,
         title = paste("Top 20", title, "GO Terms")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black"),
      panel.border = element_rect(colour = "black", fill = NA, size = 0.3)
    ) +
    xlim(c(0, max(tibble$neglog) + 1))
}

# Generate separate plots
plot_BP <- process_enrichment_plot(BP_Tibble, "Biological Process", "#2E86C1")
plot_MF <- process_enrichment_plot(MF_Tibble, "Molecular Function", "#28B463")
plot_CC <- process_enrichment_plot(CC_Tibble, "Cellular Component", "#D35400")

# Combine the plots using patchwork
combined_plot <- plot_BP / plot_MF / plot_CC

# Display the combined plot
combined_plot
```


## **📌 GO Enrichment g:Profiler**
```{r gprofiler_GO3, echo=TRUE, message=FALSE, fig.width=12, fig.height=12}

# Load the gprofiler2 package
library(gprofiler2)
library(ggplot2)
library(dplyr)
library(patchwork)

# Perform GO enrichment analysis with gprofiler2
gost_results <- gost(
  query = prob_3_0.1,           # List of input genes (prob_all_1)
  organism = "hsapiens",   # Human organism
  user_threshold = 0.05,   # Adjusted p-value cutoff
  correction_method = "fdr", # Multiple testing correction
  domain_scope = "custom", # Use custom background
  custom_bg = background,    # Background set of genes
  sources = c("GO:BP", "GO:MF", "GO:CC") # Analyze GO categories
)

# Check if enrichment results exist
if (is.null(gost_results$result) || nrow(gost_results$result) == 0) {
  # If no enriched terms, create a placeholder dataframe
  combined_results <- tibble(
    term_name = "No enriched terms",
    p.adjust = NA,
    source = "N/A",
    Category = "N/A"
  )
} else {
  # Convert results to a data frame
  gost_results_df <- gost_results$result

  # Add a column for adjusted p-value
  gost_results_df <- gost_results_df %>%
    rename(p.adjust = p_value)

  # Separate results for BP, MF, and CC
  BP_results <- gost_results_df %>%
    filter(source == "GO:BP") %>%
    mutate(Category = "Biological Process")

  MF_results <- gost_results_df %>%
    filter(source == "GO:MF") %>%
    mutate(Category = "Molecular Function")

  CC_results <- gost_results_df %>%
    filter(source == "GO:CC") %>%
    mutate(Category = "Cellular Component")

  # Select the top 20 terms by adjusted p-value for each category
  top_BP <- BP_results %>%
    arrange(p.adjust) %>%
    slice_head(n = 20)

  top_MF <- MF_results %>%
    arrange(p.adjust) %>%
    slice_head(n = 20)

  top_CC <- CC_results %>%
    arrange(p.adjust) %>%
    slice_head(n = 20)

  # Combine all categories
  combined_results <- bind_rows(top_BP, top_MF, top_CC)
}

# Ensure all columns are atomic types for CSV export
combined_results_clean <- combined_results %>%
  mutate(across(everything(), ~ if (is.list(.)) sapply(., toString) else .))

# Function for plotting top terms
plot_gprofiler_results <- function(data, title, color) {
  ggplot(data, aes(x = -log10(p.adjust), y = reorder(term_name, -log10(p.adjust)))) +
    geom_bar(stat = "identity", fill = color) +
    labs(
      x = "-log10(Adjusted p-value)",
      y = title,
      title = paste("Top 20", title, "GO Terms")
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black"),
      panel.border = element_rect(colour = "black", fill = NA, size = 0.3)
    )
}

# Check if there are enrichment terms to plot
if (nrow(combined_results) == 1 && combined_results$term_name == "No enriched terms") {
  message("No enriched GO terms found for the input gene set.")
} else {
  # Plot the top 20 terms for each category
  plot_BP <- plot_gprofiler_results(top_BP, "Biological Process", "#2E86C1")
  plot_MF <- plot_gprofiler_results(top_MF, "Molecular Function", "#28B463")
  plot_CC <- plot_gprofiler_results(top_CC, "Cellular Component", "#D35400")

  # Combine the plots using patchwork
  combined_plot <- plot_BP / plot_MF / plot_CC

  # Display the combined plot
  combined_plot
}
```

## **📌 Pathway Enrichment**
```{r pathway_enrichment3, echo=TRUE, message=FALSE, fig.width=15, fig.height=12}

# Load required libraries
library(clusterProfiler)
library(org.Hs.eg.db) # Required for enrichPathway
library(gprofiler2)
library(ggplot2)
library(dplyr)
library(patchwork)
library(ReactomePA)

# Function for ClusterProfiler Reactome & KEGG Analysis
process_clusterProfiler <- function(gene_set, background, category, color, y_title) {
  # Perform enrichment based on the selected category
  if (category == "Reactome") {
    enrichment <- enrichPathway(
      gene = gene_set,
      organism = "human",
      pvalueCutoff = 0.05,
      pAdjustMethod = "BH",
      universe = background
    )
  } else if (category == "KEGG") {
    enrichment <- enrichKEGG(
      gene = gene_set,
      organism = "hsa",
      pvalueCutoff = 0.05,
      pAdjustMethod = "BH",
      universe = background
    )
  }
  
  # Check if enrichment results exist
  if (is.null(enrichment) || nrow(as.data.frame(enrichment)) == 0) {
    message(paste("No significant enrichment found for", category, "in ClusterProfiler"))
    return(NULL)
  }
  
  # Convert results to tibble and process top 20 terms
  enrichment_tibble <- as_tibble(as.data.frame(enrichment)) %>%
    mutate(Category = category,
           neglog = -log10(p.adjust)) %>%  # Compute -log10(p.adjust)
    arrange(desc(neglog)) %>%
    slice_head(n = min(20, nrow(.)))  # Ensure safe slicing
  
  # Generate plot
  plot <- ggplot(enrichment_tibble, aes(x = neglog, y = reorder(Description, neglog))) +
    geom_bar(stat = "identity", fill = color) +
    labs(x = "-log10(Adjusted p-value)",
         y = y_title,
         title = paste("Enriched", category, "Pathways")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black")
    )
  
  return(plot)
}

# Function for gProfiler Reactome & KEGG Analysis
process_gprofiler <- function(gene_set, background, category, color, y_title) {
  # Perform enrichment using gprofiler2
  enrichment <- gost(
    query = gene_set,
    organism = "hsapiens",
    user_threshold = 0.05,
    correction_method = "fdr",
    domain_scope = "custom",
    custom_bg = background,
    sources = category # Either "REAC" or "KEGG"
  )
  
  # Check if enrichment results exist
  if (is.null(enrichment$result) || nrow(enrichment$result) == 0) {
    message(paste("No significant enrichment found for", category, "in gProfiler"))
    return(NULL)
  }
  
  # Convert results to tibble and process top 20 terms
  enrichment_tibble <- enrichment$result %>%
    as_tibble() %>%
    mutate(Category = category,
           neglog = -log10(p_value)) %>%  # Compute -log10(p-value)
    arrange(desc(neglog)) %>%
    slice_head(n = min(20, nrow(.)))  # Ensure safe slicing
  
  # Generate plot
  plot <- ggplot(enrichment_tibble, aes(x = neglog, y = reorder(term_name, neglog))) +
    geom_bar(stat = "identity", fill = color) +
    labs(x = "-log10(p-value)",
         y = y_title,
         title = paste("Enriched", category, "Pathways")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black")
    )
  
  return(plot)
}

# Perform analysis for Reactome and KEGG using ClusterProfiler
cluster_reactome <- process_clusterProfiler(
  gene_set = prob_3_0.1,
  background = background,
  category = "Reactome",
  color = "#2E86C1",
  y_title = "Reactome Pathways"
)

cluster_kegg <- process_clusterProfiler(
  gene_set = prob_3_0.1,
  background = background,
  category = "KEGG",
  color = "#28B463",
  y_title = "KEGG Pathways"
)

# Combine Reactome and KEGG for ClusterProfiler
if (!is.null(cluster_reactome) && !is.null(cluster_kegg)) {
  cluster_combined <- cluster_reactome / cluster_kegg
} else if (!is.null(cluster_reactome)) {
  cluster_combined <- cluster_reactome
} else if (!is.null(cluster_kegg)) {
  cluster_combined <- cluster_kegg
} else {
  cluster_combined <- NULL
}

# Perform analysis for Reactome and KEGG using GProfiler
gprofiler_reactome <- process_gprofiler(
  gene_set = prob_3_0.1,
  background = background,
  category = "REAC",  # Corrected category for Reactome in gProfiler
  color = "#D35400",
  y_title = "Reactome Pathways"
)

gprofiler_kegg <- process_gprofiler(
  gene_set = prob_3_0.1,
  background = background,
  category = "KEGG",
  color = "#F39C12",
  y_title = "KEGG Pathways"
)

# Combine Reactome and KEGG for GProfiler
if (!is.null(gprofiler_reactome) && !is.null(gprofiler_kegg)) {
  gprofiler_combined <- gprofiler_reactome / gprofiler_kegg
} else if (!is.null(gprofiler_reactome)) {
  gprofiler_combined <- gprofiler_reactome
} else if (!is.null(gprofiler_kegg)) {
  gprofiler_combined <- gprofiler_kegg
} else {
  gprofiler_combined <- NULL
}

# Display plots (if they are not NULL)
if (!is.null(cluster_combined)) print(cluster_combined)
if (!is.null(gprofiler_combined)) print(gprofiler_combined)
```


## **📌 0.5 Micromolar**
## **📌 Read and Process Data**
```{r load_DEGs2, echo=TRUE, results='hide', message=FALSE, warning=FALSE}
# Load the saved datasets
prob_1_0.5 <- read.csv("data/prob_1_0.5.csv")$Entrez_ID
prob_2_0.5 <- read.csv("data/prob_2_0.5.csv")$Entrez_ID
prob_3_0.5 <- read.csv("data/prob_3_0.5.csv")$Entrez_ID
prob_4_0.5 <- read.csv("data/prob_4_0.5.csv")$Entrez_ID
prob_5_0.5 <- read.csv("data/prob_5_0.5.csv")$Entrez_ID


CX_0.1_3 <- read.csv("data/DEGs/Toptable_CX_0.1_3.csv")
background<-as.character(CX_0.1_3$Entrez_ID)
```

## **📌 Non response (0.5 µM)**
## **📌 GO Enrichment Clusterprofiler**
```{r GO4, echo=TRUE, message=FALSE, fig.width=12, fig.height=12}
# Perform GO enrichment analysis for BP, MF, and CC
go_enrichment_BP <- enrichGO(gene = prob_1_0.5,
                             OrgDb = org.Hs.eg.db,
                             keyType = "ENTREZID",
                             universe = background,
                             ont = "BP",
                             pvalueCutoff = 0.05)

go_enrichment_MF <- enrichGO(gene = prob_1_0.5,
                             OrgDb = org.Hs.eg.db,
                             keyType = "ENTREZID",
                             universe = background,
                             ont = "MF",
                             pvalueCutoff = 0.05)

go_enrichment_CC <- enrichGO(gene = prob_1_0.5,
                             OrgDb = org.Hs.eg.db,
                             keyType = "ENTREZID",
                             universe = background,
                             ont = "CC",
                             pvalueCutoff = 0.05)

# Convert each enrichment result to a tibble, add a category column, and select top 20 terms
process_enrichment_tibble <- function(enrichment, category) {
  if (is.null(enrichment) || nrow(as.data.frame(enrichment)) == 0) {
    return(tibble(Description = "No enriched terms", neglog = 0, Category = category))
  } else {
    enrichment %>%
      as_tibble() %>%
      mutate(Category = category,
             neglog = -log(p.adjust)) %>% # Add -log(p.adjust) column
      arrange(desc(neglog)) %>%          # Sort by -log(p.adjust)
      slice(1:20)                        # Select top 20 terms
  }
}

BP_Tibble <- process_enrichment_tibble(go_enrichment_BP, "Biological Process")
MF_Tibble <- process_enrichment_tibble(go_enrichment_MF, "Molecular Function")
CC_Tibble <- process_enrichment_tibble(go_enrichment_CC, "Cellular Component")

# Combine all tibbles
combined_GO_Tibble <- bind_rows(BP_Tibble, MF_Tibble, CC_Tibble)
# Function to generate enrichment plots
process_enrichment_plot <- function(tibble, title, color) {
  ggplot(data = tibble, aes(x = neglog, y = reorder(Description, neglog))) +
    geom_bar(stat = "identity", fill = color) +
    labs(x = "-log(p-adjust)",
         y = title,
         title = paste("Top 20", title, "GO Terms")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black"),
      panel.border = element_rect(colour = "black", fill = NA, size = 0.3)
    ) +
    xlim(c(0, max(tibble$neglog) + 1))
}

# Generate separate plots
plot_BP <- process_enrichment_plot(BP_Tibble, "Biological Process", "#2E86C1")
plot_MF <- process_enrichment_plot(MF_Tibble, "Molecular Function", "#28B463")
plot_CC <- process_enrichment_plot(CC_Tibble, "Cellular Component", "#D35400")

# Combine the plots using patchwork
combined_plot <- plot_BP / plot_MF / plot_CC

# Display the combined plot
combined_plot
```


## **📌 GO Enrichment g:Profiler**
```{r gprofiler_GO4, echo=TRUE, message=FALSE, fig.width=12, fig.height=12}

# Load the gprofiler2 package
library(gprofiler2)
library(ggplot2)
library(dplyr)
library(patchwork)

# Perform GO enrichment analysis with gprofiler2
gost_results <- gost(
  query = prob_1_0.5,           # List of input genes (prob_all_1)
  organism = "hsapiens",   # Human organism
  user_threshold = 0.05,   # Adjusted p-value cutoff
  correction_method = "fdr", # Multiple testing correction
  domain_scope = "custom", # Use custom background
  custom_bg = background,    # Background set of genes
  sources = c("GO:BP", "GO:MF", "GO:CC") # Analyze GO categories
)

# Check if enrichment results exist
if (is.null(gost_results$result) || nrow(gost_results$result) == 0) {
  # If no enriched terms, create a placeholder dataframe
  combined_results <- tibble(
    term_name = "No enriched terms",
    p.adjust = NA,
    source = "N/A",
    Category = "N/A"
  )
} else {
  # Convert results to a data frame
  gost_results_df <- gost_results$result

  # Add a column for adjusted p-value
  gost_results_df <- gost_results_df %>%
    rename(p.adjust = p_value)

  # Separate results for BP, MF, and CC
  BP_results <- gost_results_df %>%
    filter(source == "GO:BP") %>%
    mutate(Category = "Biological Process")

  MF_results <- gost_results_df %>%
    filter(source == "GO:MF") %>%
    mutate(Category = "Molecular Function")

  CC_results <- gost_results_df %>%
    filter(source == "GO:CC") %>%
    mutate(Category = "Cellular Component")

  # Select the top 20 terms by adjusted p-value for each category
  top_BP <- BP_results %>%
    arrange(p.adjust) %>%
    slice_head(n = 20)

  top_MF <- MF_results %>%
    arrange(p.adjust) %>%
    slice_head(n = 20)

  top_CC <- CC_results %>%
    arrange(p.adjust) %>%
    slice_head(n = 20)

  # Combine all categories
  combined_results <- bind_rows(top_BP, top_MF, top_CC)
}

# Ensure all columns are atomic types for CSV export
combined_results_clean <- combined_results %>%
  mutate(across(everything(), ~ if (is.list(.)) sapply(., toString) else .))

# Function for plotting top terms
plot_gprofiler_results <- function(data, title, color) {
  ggplot(data, aes(x = -log10(p.adjust), y = reorder(term_name, -log10(p.adjust)))) +
    geom_bar(stat = "identity", fill = color) +
    labs(
      x = "-log10(Adjusted p-value)",
      y = title,
      title = paste("Top 20", title, "GO Terms")
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black"),
      panel.border = element_rect(colour = "black", fill = NA, size = 0.3)
    )
}

# Check if there are enrichment terms to plot
if (nrow(combined_results) == 1 && combined_results$term_name == "No enriched terms") {
  message("No enriched GO terms found for the input gene set.")
} else {
  # Plot the top 20 terms for each category
  plot_BP <- plot_gprofiler_results(top_BP, "Biological Process", "#2E86C1")
  plot_MF <- plot_gprofiler_results(top_MF, "Molecular Function", "#28B463")
  plot_CC <- plot_gprofiler_results(top_CC, "Cellular Component", "#D35400")

  # Combine the plots using patchwork
  combined_plot <- plot_BP / plot_MF / plot_CC

  # Display the combined plot
  combined_plot
}
```

## **📌 Pathway Enrichment**
```{r pathway_enrichment4, echo=TRUE, message=FALSE, fig.width=15, fig.height=12}

# Load required libraries
library(clusterProfiler)
library(org.Hs.eg.db) # Required for enrichPathway
library(gprofiler2)
library(ggplot2)
library(dplyr)
library(patchwork)
library(ReactomePA)

# Function for ClusterProfiler Reactome & KEGG Analysis
process_clusterProfiler <- function(gene_set, background, category, color, y_title) {
  # Perform enrichment based on the selected category
  if (category == "Reactome") {
    enrichment <- enrichPathway(
      gene = gene_set,
      organism = "human",
      pvalueCutoff = 0.05,
      pAdjustMethod = "BH",
      universe = background
    )
  } else if (category == "KEGG") {
    enrichment <- enrichKEGG(
      gene = gene_set,
      organism = "hsa",
      pvalueCutoff = 0.05,
      pAdjustMethod = "BH",
      universe = background
    )
  }
  
  # Check if enrichment results exist
  if (is.null(enrichment) || nrow(as.data.frame(enrichment)) == 0) {
    message(paste("No significant enrichment found for", category, "in ClusterProfiler"))
    return(NULL)
  }
  
  # Convert results to tibble and process top 20 terms
  enrichment_tibble <- as_tibble(as.data.frame(enrichment)) %>%
    mutate(Category = category,
           neglog = -log10(p.adjust)) %>%  # Compute -log10(p.adjust)
    arrange(desc(neglog)) %>%
    slice_head(n = min(20, nrow(.)))  # Ensure safe slicing
  
  # Generate plot
  plot <- ggplot(enrichment_tibble, aes(x = neglog, y = reorder(Description, neglog))) +
    geom_bar(stat = "identity", fill = color) +
    labs(x = "-log10(Adjusted p-value)",
         y = y_title,
         title = paste("Enriched", category, "Pathways")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black")
    )
  
  return(plot)
}

# Function for gProfiler Reactome & KEGG Analysis
process_gprofiler <- function(gene_set, background, category, color, y_title) {
  # Perform enrichment using gprofiler2
  enrichment <- gost(
    query = gene_set,
    organism = "hsapiens",
    user_threshold = 0.05,
    correction_method = "fdr",
    domain_scope = "custom",
    custom_bg = background,
    sources = category # Either "REAC" or "KEGG"
  )
  
  # Check if enrichment results exist
  if (is.null(enrichment$result) || nrow(enrichment$result) == 0) {
    message(paste("No significant enrichment found for", category, "in gProfiler"))
    return(NULL)
  }
  
  # Convert results to tibble and process top 20 terms
  enrichment_tibble <- enrichment$result %>%
    as_tibble() %>%
    mutate(Category = category,
           neglog = -log10(p_value)) %>%  # Compute -log10(p-value)
    arrange(desc(neglog)) %>%
    slice_head(n = min(20, nrow(.)))  # Ensure safe slicing
  
  # Generate plot
  plot <- ggplot(enrichment_tibble, aes(x = neglog, y = reorder(term_name, neglog))) +
    geom_bar(stat = "identity", fill = color) +
    labs(x = "-log10(p-value)",
         y = y_title,
         title = paste("Enriched", category, "Pathways")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black")
    )
  
  return(plot)
}

# Perform analysis for Reactome and KEGG using ClusterProfiler
cluster_reactome <- process_clusterProfiler(
  gene_set = prob_1_0.5,
  background = background,
  category = "Reactome",
  color = "#2E86C1",
  y_title = "Reactome Pathways"
)

cluster_kegg <- process_clusterProfiler(
  gene_set = prob_1_0.5,
  background = background,
  category = "KEGG",
  color = "#28B463",
  y_title = "KEGG Pathways"
)

# Combine Reactome and KEGG for ClusterProfiler
if (!is.null(cluster_reactome) && !is.null(cluster_kegg)) {
  cluster_combined <- cluster_reactome / cluster_kegg
} else if (!is.null(cluster_reactome)) {
  cluster_combined <- cluster_reactome
} else if (!is.null(cluster_kegg)) {
  cluster_combined <- cluster_kegg
} else {
  cluster_combined <- NULL
}

# Perform analysis for Reactome and KEGG using GProfiler
gprofiler_reactome <- process_gprofiler(
  gene_set = prob_1_0.5,
  background = background,
  category = "REAC",  # Corrected category for Reactome in gProfiler
  color = "#D35400",
  y_title = "Reactome Pathways"
)

gprofiler_kegg <- process_gprofiler(
  gene_set = prob_1_0.5,
  background = background,
  category = "KEGG",
  color = "#F39C12",
  y_title = "KEGG Pathways"
)

# Combine Reactome and KEGG for GProfiler
if (!is.null(gprofiler_reactome) && !is.null(gprofiler_kegg)) {
  gprofiler_combined <- gprofiler_reactome / gprofiler_kegg
} else if (!is.null(gprofiler_reactome)) {
  gprofiler_combined <- gprofiler_reactome
} else if (!is.null(gprofiler_kegg)) {
  gprofiler_combined <- gprofiler_kegg
} else {
  gprofiler_combined <- NULL
}

# Display plots (if they are not NULL)
if (!is.null(cluster_combined)) print(cluster_combined)
if (!is.null(gprofiler_combined)) print(gprofiler_combined)
```

## **📌 DOX-specific response (0.5 µM)**
## **📌 GO Enrichment Clusterprofiler**
```{r GO5, echo=TRUE, message=FALSE, fig.width=12, fig.height=12}
# Perform GO enrichment analysis for BP, MF, and CC
go_enrichment_BP <- enrichGO(gene = prob_2_0.5,
                             OrgDb = org.Hs.eg.db,
                             keyType = "ENTREZID",
                             universe = background,
                             ont = "BP",
                             pvalueCutoff = 0.05)

go_enrichment_MF <- enrichGO(gene = prob_2_0.5,
                             OrgDb = org.Hs.eg.db,
                             keyType = "ENTREZID",
                             universe = background,
                             ont = "MF",
                             pvalueCutoff = 0.05)

go_enrichment_CC <- enrichGO(gene = prob_2_0.5,
                             OrgDb = org.Hs.eg.db,
                             keyType = "ENTREZID",
                             universe = background,
                             ont = "CC",
                             pvalueCutoff = 0.05)

# Convert each enrichment result to a tibble, add a category column, and select top 20 terms
process_enrichment_tibble <- function(enrichment, category) {
  if (is.null(enrichment) || nrow(as.data.frame(enrichment)) == 0) {
    return(tibble(Description = "No enriched terms", neglog = 0, Category = category))
  } else {
    enrichment %>%
      as_tibble() %>%
      mutate(Category = category,
             neglog = -log(p.adjust)) %>% # Add -log(p.adjust) column
      arrange(desc(neglog)) %>%          # Sort by -log(p.adjust)
      slice(1:20)                        # Select top 20 terms
  }
}

BP_Tibble <- process_enrichment_tibble(go_enrichment_BP, "Biological Process")
MF_Tibble <- process_enrichment_tibble(go_enrichment_MF, "Molecular Function")
CC_Tibble <- process_enrichment_tibble(go_enrichment_CC, "Cellular Component")

# Combine all tibbles
combined_GO_Tibble <- bind_rows(BP_Tibble, MF_Tibble, CC_Tibble)
# Function to generate enrichment plots
process_enrichment_plot <- function(tibble, title, color) {
  ggplot(data = tibble, aes(x = neglog, y = reorder(Description, neglog))) +
    geom_bar(stat = "identity", fill = color) +
    labs(x = "-log(p-adjust)",
         y = title,
         title = paste("Top 20", title, "GO Terms")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black"),
      panel.border = element_rect(colour = "black", fill = NA, size = 0.3)
    ) +
    xlim(c(0, max(tibble$neglog) + 1))
}

# Generate separate plots
plot_BP <- process_enrichment_plot(BP_Tibble, "Biological Process", "#2E86C1")
plot_MF <- process_enrichment_plot(MF_Tibble, "Molecular Function", "#28B463")
plot_CC <- process_enrichment_plot(CC_Tibble, "Cellular Component", "#D35400")

# Combine the plots using patchwork
combined_plot <- plot_BP / plot_MF / plot_CC

# Display the combined plot
combined_plot
```


## **📌 GO Enrichment g:Profiler**
```{r gprofiler_GO5, echo=TRUE, message=FALSE, fig.width=12, fig.height=12}

# Load the gprofiler2 package
library(gprofiler2)
library(ggplot2)
library(dplyr)
library(patchwork)

# Perform GO enrichment analysis with gprofiler2
gost_results <- gost(
  query = prob_2_0.5,           # List of input genes (prob_all_1)
  organism = "hsapiens",   # Human organism
  user_threshold = 0.05,   # Adjusted p-value cutoff
  correction_method = "fdr", # Multiple testing correction
  domain_scope = "custom", # Use custom background
  custom_bg = background,    # Background set of genes
  sources = c("GO:BP", "GO:MF", "GO:CC") # Analyze GO categories
)

# Check if enrichment results exist
if (is.null(gost_results$result) || nrow(gost_results$result) == 0) {
  # If no enriched terms, create a placeholder dataframe
  combined_results <- tibble(
    term_name = "No enriched terms",
    p.adjust = NA,
    source = "N/A",
    Category = "N/A"
  )
} else {
  # Convert results to a data frame
  gost_results_df <- gost_results$result

  # Add a column for adjusted p-value
  gost_results_df <- gost_results_df %>%
    rename(p.adjust = p_value)

  # Separate results for BP, MF, and CC
  BP_results <- gost_results_df %>%
    filter(source == "GO:BP") %>%
    mutate(Category = "Biological Process")

  MF_results <- gost_results_df %>%
    filter(source == "GO:MF") %>%
    mutate(Category = "Molecular Function")

  CC_results <- gost_results_df %>%
    filter(source == "GO:CC") %>%
    mutate(Category = "Cellular Component")

  # Select the top 20 terms by adjusted p-value for each category
  top_BP <- BP_results %>%
    arrange(p.adjust) %>%
    slice_head(n = 20)

  top_MF <- MF_results %>%
    arrange(p.adjust) %>%
    slice_head(n = 20)

  top_CC <- CC_results %>%
    arrange(p.adjust) %>%
    slice_head(n = 20)

  # Combine all categories
  combined_results <- bind_rows(top_BP, top_MF, top_CC)
}

# Ensure all columns are atomic types for CSV export
combined_results_clean <- combined_results %>%
  mutate(across(everything(), ~ if (is.list(.)) sapply(., toString) else .))

# Function for plotting top terms
plot_gprofiler_results <- function(data, title, color) {
  ggplot(data, aes(x = -log10(p.adjust), y = reorder(term_name, -log10(p.adjust)))) +
    geom_bar(stat = "identity", fill = color) +
    labs(
      x = "-log10(Adjusted p-value)",
      y = title,
      title = paste("Top 20", title, "GO Terms")
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black"),
      panel.border = element_rect(colour = "black", fill = NA, size = 0.3)
    )
}

# Check if there are enrichment terms to plot
if (nrow(combined_results) == 1 && combined_results$term_name == "No enriched terms") {
  message("No enriched GO terms found for the input gene set.")
} else {
  # Plot the top 20 terms for each category
  plot_BP <- plot_gprofiler_results(top_BP, "Biological Process", "#2E86C1")
  plot_MF <- plot_gprofiler_results(top_MF, "Molecular Function", "#28B463")
  plot_CC <- plot_gprofiler_results(top_CC, "Cellular Component", "#D35400")

  # Combine the plots using patchwork
  combined_plot <- plot_BP / plot_MF / plot_CC

  # Display the combined plot
  combined_plot
}
```

## **📌 Pathway Enrichment**
```{r pathway_enrichment5, echo=TRUE, message=FALSE, fig.width=15, fig.height=12}

# Load required libraries
library(clusterProfiler)
library(org.Hs.eg.db) # Required for enrichPathway
library(gprofiler2)
library(ggplot2)
library(dplyr)
library(patchwork)
library(ReactomePA)

# Function for ClusterProfiler Reactome & KEGG Analysis
process_clusterProfiler <- function(gene_set, background, category, color, y_title) {
  # Perform enrichment based on the selected category
  if (category == "Reactome") {
    enrichment <- enrichPathway(
      gene = gene_set,
      organism = "human",
      pvalueCutoff = 0.05,
      pAdjustMethod = "BH",
      universe = background
    )
  } else if (category == "KEGG") {
    enrichment <- enrichKEGG(
      gene = gene_set,
      organism = "hsa",
      pvalueCutoff = 0.05,
      pAdjustMethod = "BH",
      universe = background
    )
  }
  
  # Check if enrichment results exist
  if (is.null(enrichment) || nrow(as.data.frame(enrichment)) == 0) {
    message(paste("No significant enrichment found for", category, "in ClusterProfiler"))
    return(NULL)
  }
  
  # Convert results to tibble and process top 20 terms
  enrichment_tibble <- as_tibble(as.data.frame(enrichment)) %>%
    mutate(Category = category,
           neglog = -log10(p.adjust)) %>%  # Compute -log10(p.adjust)
    arrange(desc(neglog)) %>%
    slice_head(n = min(20, nrow(.)))  # Ensure safe slicing
  
  # Generate plot
  plot <- ggplot(enrichment_tibble, aes(x = neglog, y = reorder(Description, neglog))) +
    geom_bar(stat = "identity", fill = color) +
    labs(x = "-log10(Adjusted p-value)",
         y = y_title,
         title = paste("Enriched", category, "Pathways")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black")
    )
  
  return(plot)
}

# Function for gProfiler Reactome & KEGG Analysis
process_gprofiler <- function(gene_set, background, category, color, y_title) {
  # Perform enrichment using gprofiler2
  enrichment <- gost(
    query = gene_set,
    organism = "hsapiens",
    user_threshold = 0.05,
    correction_method = "fdr",
    domain_scope = "custom",
    custom_bg = background,
    sources = category # Either "REAC" or "KEGG"
  )
  
  # Check if enrichment results exist
  if (is.null(enrichment$result) || nrow(enrichment$result) == 0) {
    message(paste("No significant enrichment found for", category, "in gProfiler"))
    return(NULL)
  }
  
  # Convert results to tibble and process top 20 terms
  enrichment_tibble <- enrichment$result %>%
    as_tibble() %>%
    mutate(Category = category,
           neglog = -log10(p_value)) %>%  # Compute -log10(p-value)
    arrange(desc(neglog)) %>%
    slice_head(n = min(20, nrow(.)))  # Ensure safe slicing
  
  # Generate plot
  plot <- ggplot(enrichment_tibble, aes(x = neglog, y = reorder(term_name, neglog))) +
    geom_bar(stat = "identity", fill = color) +
    labs(x = "-log10(p-value)",
         y = y_title,
         title = paste("Enriched", category, "Pathways")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black")
    )
  
  return(plot)
}

# Perform analysis for Reactome and KEGG using ClusterProfiler
cluster_reactome <- process_clusterProfiler(
  gene_set = prob_2_0.5,
  background = background,
  category = "Reactome",
  color = "#2E86C1",
  y_title = "Reactome Pathways"
)

cluster_kegg <- process_clusterProfiler(
  gene_set = prob_2_0.5,
  background = background,
  category = "KEGG",
  color = "#28B463",
  y_title = "KEGG Pathways"
)

# Combine Reactome and KEGG for ClusterProfiler
if (!is.null(cluster_reactome) && !is.null(cluster_kegg)) {
  cluster_combined <- cluster_reactome / cluster_kegg
} else if (!is.null(cluster_reactome)) {
  cluster_combined <- cluster_reactome
} else if (!is.null(cluster_kegg)) {
  cluster_combined <- cluster_kegg
} else {
  cluster_combined <- NULL
}

# Perform analysis for Reactome and KEGG using GProfiler
gprofiler_reactome <- process_gprofiler(
  gene_set = prob_2_0.5,
  background = background,
  category = "REAC",  # Corrected category for Reactome in gProfiler
  color = "#D35400",
  y_title = "Reactome Pathways"
)

gprofiler_kegg <- process_gprofiler(
  gene_set = prob_2_0.5,
  background = background,
  category = "KEGG",
  color = "#F39C12",
  y_title = "KEGG Pathways"
)

# Combine Reactome and KEGG for GProfiler
if (!is.null(gprofiler_reactome) && !is.null(gprofiler_kegg)) {
  gprofiler_combined <- gprofiler_reactome / gprofiler_kegg
} else if (!is.null(gprofiler_reactome)) {
  gprofiler_combined <- gprofiler_reactome
} else if (!is.null(gprofiler_kegg)) {
  gprofiler_combined <- gprofiler_kegg
} else {
  gprofiler_combined <- NULL
}

# Display plots (if they are not NULL)
if (!is.null(cluster_combined)) print(cluster_combined)
if (!is.null(gprofiler_combined)) print(gprofiler_combined)
```

## **📌 DOX only mid-late response (0.5 µM)**
## **📌 GO Enrichment Clusterprofiler**
```{r GO6, echo=TRUE, message=FALSE, fig.width=12, fig.height=12}
# Perform GO enrichment analysis for BP, MF, and CC
go_enrichment_BP <- enrichGO(gene = prob_3_0.5,
                             OrgDb = org.Hs.eg.db,
                             keyType = "ENTREZID",
                             universe = background,
                             ont = "BP",
                             pvalueCutoff = 0.05)

go_enrichment_MF <- enrichGO(gene = prob_3_0.5,
                             OrgDb = org.Hs.eg.db,
                             keyType = "ENTREZID",
                             universe = background,
                             ont = "MF",
                             pvalueCutoff = 0.05)

go_enrichment_CC <- enrichGO(gene = prob_3_0.5,
                             OrgDb = org.Hs.eg.db,
                             keyType = "ENTREZID",
                             universe = background,
                             ont = "CC",
                             pvalueCutoff = 0.05)

# Convert each enrichment result to a tibble, add a category column, and select top 20 terms
process_enrichment_tibble <- function(enrichment, category) {
  if (is.null(enrichment) || nrow(as.data.frame(enrichment)) == 0) {
    return(tibble(Description = "No enriched terms", neglog = 0, Category = category))
  } else {
    enrichment %>%
      as_tibble() %>%
      mutate(Category = category,
             neglog = -log(p.adjust)) %>% # Add -log(p.adjust) column
      arrange(desc(neglog)) %>%          # Sort by -log(p.adjust)
      slice(1:20)                        # Select top 20 terms
  }
}

BP_Tibble <- process_enrichment_tibble(go_enrichment_BP, "Biological Process")
MF_Tibble <- process_enrichment_tibble(go_enrichment_MF, "Molecular Function")
CC_Tibble <- process_enrichment_tibble(go_enrichment_CC, "Cellular Component")

# Combine all tibbles
combined_GO_Tibble <- bind_rows(BP_Tibble, MF_Tibble, CC_Tibble)
# Function to generate enrichment plots
process_enrichment_plot <- function(tibble, title, color) {
  ggplot(data = tibble, aes(x = neglog, y = reorder(Description, neglog))) +
    geom_bar(stat = "identity", fill = color) +
    labs(x = "-log(p-adjust)",
         y = title,
         title = paste("Top 20", title, "GO Terms")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black"),
      panel.border = element_rect(colour = "black", fill = NA, size = 0.3)
    ) +
    xlim(c(0, max(tibble$neglog) + 1))
}

# Generate separate plots
plot_BP <- process_enrichment_plot(BP_Tibble, "Biological Process", "#2E86C1")
plot_MF <- process_enrichment_plot(MF_Tibble, "Molecular Function", "#28B463")
plot_CC <- process_enrichment_plot(CC_Tibble, "Cellular Component", "#D35400")

# Combine the plots using patchwork
combined_plot <- plot_BP / plot_MF / plot_CC

# Display the combined plot
combined_plot
```


## **📌 GO Enrichment g:Profiler**
```{r gprofiler_GO6, echo=TRUE, message=FALSE, fig.width=12, fig.height=12}

# Load the gprofiler2 package
library(gprofiler2)
library(ggplot2)
library(dplyr)
library(patchwork)

# Perform GO enrichment analysis with gprofiler2
gost_results <- gost(
  query = prob_3_0.5,           # List of input genes (prob_all_1)
  organism = "hsapiens",   # Human organism
  user_threshold = 0.05,   # Adjusted p-value cutoff
  correction_method = "fdr", # Multiple testing correction
  domain_scope = "custom", # Use custom background
  custom_bg = background,    # Background set of genes
  sources = c("GO:BP", "GO:MF", "GO:CC") # Analyze GO categories
)

# Check if enrichment results exist
if (is.null(gost_results$result) || nrow(gost_results$result) == 0) {
  # If no enriched terms, create a placeholder dataframe
  combined_results <- tibble(
    term_name = "No enriched terms",
    p.adjust = NA,
    source = "N/A",
    Category = "N/A"
  )
} else {
  # Convert results to a data frame
  gost_results_df <- gost_results$result

  # Add a column for adjusted p-value
  gost_results_df <- gost_results_df %>%
    rename(p.adjust = p_value)

  # Separate results for BP, MF, and CC
  BP_results <- gost_results_df %>%
    filter(source == "GO:BP") %>%
    mutate(Category = "Biological Process")

  MF_results <- gost_results_df %>%
    filter(source == "GO:MF") %>%
    mutate(Category = "Molecular Function")

  CC_results <- gost_results_df %>%
    filter(source == "GO:CC") %>%
    mutate(Category = "Cellular Component")

  # Select the top 20 terms by adjusted p-value for each category
  top_BP <- BP_results %>%
    arrange(p.adjust) %>%
    slice_head(n = 20)

  top_MF <- MF_results %>%
    arrange(p.adjust) %>%
    slice_head(n = 20)

  top_CC <- CC_results %>%
    arrange(p.adjust) %>%
    slice_head(n = 20)

  # Combine all categories
  combined_results <- bind_rows(top_BP, top_MF, top_CC)
}

# Ensure all columns are atomic types for CSV export
combined_results_clean <- combined_results %>%
  mutate(across(everything(), ~ if (is.list(.)) sapply(., toString) else .))

# Function for plotting top terms
plot_gprofiler_results <- function(data, title, color) {
  ggplot(data, aes(x = -log10(p.adjust), y = reorder(term_name, -log10(p.adjust)))) +
    geom_bar(stat = "identity", fill = color) +
    labs(
      x = "-log10(Adjusted p-value)",
      y = title,
      title = paste("Top 20", title, "GO Terms")
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black"),
      panel.border = element_rect(colour = "black", fill = NA, size = 0.3)
    )
}

# Check if there are enrichment terms to plot
if (nrow(combined_results) == 1 && combined_results$term_name == "No enriched terms") {
  message("No enriched GO terms found for the input gene set.")
} else {
  # Plot the top 20 terms for each category
  plot_BP <- plot_gprofiler_results(top_BP, "Biological Process", "#2E86C1")
  plot_MF <- plot_gprofiler_results(top_MF, "Molecular Function", "#28B463")
  plot_CC <- plot_gprofiler_results(top_CC, "Cellular Component", "#D35400")

  # Combine the plots using patchwork
  combined_plot <- plot_BP / plot_MF / plot_CC

  # Display the combined plot
  combined_plot
}
```

## **📌 Pathway Enrichment**
```{r pathway_enrichment6, echo=TRUE, message=FALSE, fig.width=15, fig.height=12}

# Load required libraries
library(clusterProfiler)
library(org.Hs.eg.db) # Required for enrichPathway
library(gprofiler2)
library(ggplot2)
library(dplyr)
library(patchwork)
library(ReactomePA)

# Function for ClusterProfiler Reactome & KEGG Analysis
process_clusterProfiler <- function(gene_set, background, category, color, y_title) {
  # Perform enrichment based on the selected category
  if (category == "Reactome") {
    enrichment <- enrichPathway(
      gene = gene_set,
      organism = "human",
      pvalueCutoff = 0.05,
      pAdjustMethod = "BH",
      universe = background
    )
  } else if (category == "KEGG") {
    enrichment <- enrichKEGG(
      gene = gene_set,
      organism = "hsa",
      pvalueCutoff = 0.05,
      pAdjustMethod = "BH",
      universe = background
    )
  }
  
  # Check if enrichment results exist
  if (is.null(enrichment) || nrow(as.data.frame(enrichment)) == 0) {
    message(paste("No significant enrichment found for", category, "in ClusterProfiler"))
    return(NULL)
  }
  
  # Convert results to tibble and process top 20 terms
  enrichment_tibble <- as_tibble(as.data.frame(enrichment)) %>%
    mutate(Category = category,
           neglog = -log10(p.adjust)) %>%  # Compute -log10(p.adjust)
    arrange(desc(neglog)) %>%
    slice_head(n = min(20, nrow(.)))  # Ensure safe slicing
  
  # Generate plot
  plot <- ggplot(enrichment_tibble, aes(x = neglog, y = reorder(Description, neglog))) +
    geom_bar(stat = "identity", fill = color) +
    labs(x = "-log10(Adjusted p-value)",
         y = y_title,
         title = paste("Enriched", category, "Pathways")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black")
    )
  
  return(plot)
}

# Function for gProfiler Reactome & KEGG Analysis
process_gprofiler <- function(gene_set, background, category, color, y_title) {
  # Perform enrichment using gprofiler2
  enrichment <- gost(
    query = gene_set,
    organism = "hsapiens",
    user_threshold = 0.05,
    correction_method = "fdr",
    domain_scope = "custom",
    custom_bg = background,
    sources = category # Either "REAC" or "KEGG"
  )
  
  # Check if enrichment results exist
  if (is.null(enrichment$result) || nrow(enrichment$result) == 0) {
    message(paste("No significant enrichment found for", category, "in gProfiler"))
    return(NULL)
  }
  
  # Convert results to tibble and process top 20 terms
  enrichment_tibble <- enrichment$result %>%
    as_tibble() %>%
    mutate(Category = category,
           neglog = -log10(p_value)) %>%  # Compute -log10(p-value)
    arrange(desc(neglog)) %>%
    slice_head(n = min(20, nrow(.)))  # Ensure safe slicing
  
  # Generate plot
  plot <- ggplot(enrichment_tibble, aes(x = neglog, y = reorder(term_name, neglog))) +
    geom_bar(stat = "identity", fill = color) +
    labs(x = "-log10(p-value)",
         y = y_title,
         title = paste("Enriched", category, "Pathways")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black")
    )
  
  return(plot)
}

# Perform analysis for Reactome and KEGG using ClusterProfiler
cluster_reactome <- process_clusterProfiler(
  gene_set = prob_3_0.5,
  background = background,
  category = "Reactome",
  color = "#2E86C1",
  y_title = "Reactome Pathways"
)

cluster_kegg <- process_clusterProfiler(
  gene_set = prob_3_0.5,
  background = background,
  category = "KEGG",
  color = "#28B463",
  y_title = "KEGG Pathways"
)

# Combine Reactome and KEGG for ClusterProfiler
if (!is.null(cluster_reactome) && !is.null(cluster_kegg)) {
  cluster_combined <- cluster_reactome / cluster_kegg
} else if (!is.null(cluster_reactome)) {
  cluster_combined <- cluster_reactome
} else if (!is.null(cluster_kegg)) {
  cluster_combined <- cluster_kegg
} else {
  cluster_combined <- NULL
}

# Perform analysis for Reactome and KEGG using GProfiler
gprofiler_reactome <- process_gprofiler(
  gene_set = prob_3_0.5,
  background = background,
  category = "REAC",  # Corrected category for Reactome in gProfiler
  color = "#D35400",
  y_title = "Reactome Pathways"
)

gprofiler_kegg <- process_gprofiler(
  gene_set = prob_3_0.5,
  background = background,
  category = "KEGG",
  color = "#F39C12",
  y_title = "KEGG Pathways"
)

# Combine Reactome and KEGG for GProfiler
if (!is.null(gprofiler_reactome) && !is.null(gprofiler_kegg)) {
  gprofiler_combined <- gprofiler_reactome / gprofiler_kegg
} else if (!is.null(gprofiler_reactome)) {
  gprofiler_combined <- gprofiler_reactome
} else if (!is.null(gprofiler_kegg)) {
  gprofiler_combined <- gprofiler_kegg
} else {
  gprofiler_combined <- NULL
}

# Display plots (if they are not NULL)
if (!is.null(cluster_combined)) print(cluster_combined)
if (!is.null(gprofiler_combined)) print(gprofiler_combined)
```

## **📌 CX + DOX (early) response (0.5 µM)**
## **📌 GO Enrichment Clusterprofiler**
```{r GO7, echo=TRUE, message=FALSE, fig.width=12, fig.height=12}
# Perform GO enrichment analysis for BP, MF, and CC
go_enrichment_BP <- enrichGO(gene = prob_4_0.5,
                             OrgDb = org.Hs.eg.db,
                             keyType = "ENTREZID",
                             universe = background,
                             ont = "BP",
                             pvalueCutoff = 0.05)

go_enrichment_MF <- enrichGO(gene = prob_4_0.5,
                             OrgDb = org.Hs.eg.db,
                             keyType = "ENTREZID",
                             universe = background,
                             ont = "MF",
                             pvalueCutoff = 0.05)

go_enrichment_CC <- enrichGO(gene = prob_4_0.5,
                             OrgDb = org.Hs.eg.db,
                             keyType = "ENTREZID",
                             universe = background,
                             ont = "CC",
                             pvalueCutoff = 0.05)

# Convert each enrichment result to a tibble, add a category column, and select top 20 terms
process_enrichment_tibble <- function(enrichment, category) {
  if (is.null(enrichment) || nrow(as.data.frame(enrichment)) == 0) {
    return(tibble(Description = "No enriched terms", neglog = 0, Category = category))
  } else {
    enrichment %>%
      as_tibble() %>%
      mutate(Category = category,
             neglog = -log(p.adjust)) %>% # Add -log(p.adjust) column
      arrange(desc(neglog)) %>%          # Sort by -log(p.adjust)
      slice(1:20)                        # Select top 20 terms
  }
}

BP_Tibble <- process_enrichment_tibble(go_enrichment_BP, "Biological Process")
MF_Tibble <- process_enrichment_tibble(go_enrichment_MF, "Molecular Function")
CC_Tibble <- process_enrichment_tibble(go_enrichment_CC, "Cellular Component")

# Combine all tibbles
combined_GO_Tibble <- bind_rows(BP_Tibble, MF_Tibble, CC_Tibble)
# Function to generate enrichment plots
process_enrichment_plot <- function(tibble, title, color) {
  ggplot(data = tibble, aes(x = neglog, y = reorder(Description, neglog))) +
    geom_bar(stat = "identity", fill = color) +
    labs(x = "-log(p-adjust)",
         y = title,
         title = paste("Top 20", title, "GO Terms")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black"),
      panel.border = element_rect(colour = "black", fill = NA, size = 0.3)
    ) +
    xlim(c(0, max(tibble$neglog) + 1))
}

# Generate separate plots
plot_BP <- process_enrichment_plot(BP_Tibble, "Biological Process", "#2E86C1")
plot_MF <- process_enrichment_plot(MF_Tibble, "Molecular Function", "#28B463")
plot_CC <- process_enrichment_plot(CC_Tibble, "Cellular Component", "#D35400")

# Combine the plots using patchwork
combined_plot <- plot_BP / plot_MF / plot_CC

# Display the combined plot
combined_plot
```


## **📌 GO Enrichment g:Profiler**
```{r gprofiler_GO7, echo=TRUE, message=FALSE, fig.width=12, fig.height=12}

# Load the gprofiler2 package
library(gprofiler2)
library(ggplot2)
library(dplyr)
library(patchwork)

# Perform GO enrichment analysis with gprofiler2
gost_results <- gost(
  query = prob_4_0.5,           # List of input genes (prob_all_1)
  organism = "hsapiens",   # Human organism
  user_threshold = 0.05,   # Adjusted p-value cutoff
  correction_method = "fdr", # Multiple testing correction
  domain_scope = "custom", # Use custom background
  custom_bg = background,    # Background set of genes
  sources = c("GO:BP", "GO:MF", "GO:CC") # Analyze GO categories
)

# Check if enrichment results exist
if (is.null(gost_results$result) || nrow(gost_results$result) == 0) {
  # If no enriched terms, create a placeholder dataframe
  combined_results <- tibble(
    term_name = "No enriched terms",
    p.adjust = NA,
    source = "N/A",
    Category = "N/A"
  )
} else {
  # Convert results to a data frame
  gost_results_df <- gost_results$result

  # Add a column for adjusted p-value
  gost_results_df <- gost_results_df %>%
    rename(p.adjust = p_value)

  # Separate results for BP, MF, and CC
  BP_results <- gost_results_df %>%
    filter(source == "GO:BP") %>%
    mutate(Category = "Biological Process")

  MF_results <- gost_results_df %>%
    filter(source == "GO:MF") %>%
    mutate(Category = "Molecular Function")

  CC_results <- gost_results_df %>%
    filter(source == "GO:CC") %>%
    mutate(Category = "Cellular Component")

  # Select the top 20 terms by adjusted p-value for each category
  top_BP <- BP_results %>%
    arrange(p.adjust) %>%
    slice_head(n = 20)

  top_MF <- MF_results %>%
    arrange(p.adjust) %>%
    slice_head(n = 20)

  top_CC <- CC_results %>%
    arrange(p.adjust) %>%
    slice_head(n = 20)

  # Combine all categories
  combined_results <- bind_rows(top_BP, top_MF, top_CC)
}

# Ensure all columns are atomic types for CSV export
combined_results_clean <- combined_results %>%
  mutate(across(everything(), ~ if (is.list(.)) sapply(., toString) else .))

# Function for plotting top terms
plot_gprofiler_results <- function(data, title, color) {
  ggplot(data, aes(x = -log10(p.adjust), y = reorder(term_name, -log10(p.adjust)))) +
    geom_bar(stat = "identity", fill = color) +
    labs(
      x = "-log10(Adjusted p-value)",
      y = title,
      title = paste("Top 20", title, "GO Terms")
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black"),
      panel.border = element_rect(colour = "black", fill = NA, size = 0.3)
    )
}

# Check if there are enrichment terms to plot
if (nrow(combined_results) == 1 && combined_results$term_name == "No enriched terms") {
  message("No enriched GO terms found for the input gene set.")
} else {
  # Plot the top 20 terms for each category
  plot_BP <- plot_gprofiler_results(top_BP, "Biological Process", "#2E86C1")
  plot_MF <- plot_gprofiler_results(top_MF, "Molecular Function", "#28B463")
  plot_CC <- plot_gprofiler_results(top_CC, "Cellular Component", "#D35400")

  # Combine the plots using patchwork
  combined_plot <- plot_BP / plot_MF / plot_CC

  # Display the combined plot
  combined_plot
}
```

## **📌 Pathway Enrichment**
```{r pathway_enrichment7, echo=TRUE, message=FALSE, fig.width=15, fig.height=12}

# Load required libraries
library(clusterProfiler)
library(org.Hs.eg.db) # Required for enrichPathway
library(gprofiler2)
library(ggplot2)
library(dplyr)
library(patchwork)
library(ReactomePA)

# Function for ClusterProfiler Reactome & KEGG Analysis
process_clusterProfiler <- function(gene_set, background, category, color, y_title) {
  # Perform enrichment based on the selected category
  if (category == "Reactome") {
    enrichment <- enrichPathway(
      gene = gene_set,
      organism = "human",
      pvalueCutoff = 0.05,
      pAdjustMethod = "BH",
      universe = background
    )
  } else if (category == "KEGG") {
    enrichment <- enrichKEGG(
      gene = gene_set,
      organism = "hsa",
      pvalueCutoff = 0.05,
      pAdjustMethod = "BH",
      universe = background
    )
  }
  
  # Check if enrichment results exist
  if (is.null(enrichment) || nrow(as.data.frame(enrichment)) == 0) {
    message(paste("No significant enrichment found for", category, "in ClusterProfiler"))
    return(NULL)
  }
  
  # Convert results to tibble and process top 20 terms
  enrichment_tibble <- as_tibble(as.data.frame(enrichment)) %>%
    mutate(Category = category,
           neglog = -log10(p.adjust)) %>%  # Compute -log10(p.adjust)
    arrange(desc(neglog)) %>%
    slice_head(n = min(20, nrow(.)))  # Ensure safe slicing
  
  # Generate plot
  plot <- ggplot(enrichment_tibble, aes(x = neglog, y = reorder(Description, neglog))) +
    geom_bar(stat = "identity", fill = color) +
    labs(x = "-log10(Adjusted p-value)",
         y = y_title,
         title = paste("Enriched", category, "Pathways")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black")
    )
  
  return(plot)
}

# Function for gProfiler Reactome & KEGG Analysis
process_gprofiler <- function(gene_set, background, category, color, y_title) {
  # Perform enrichment using gprofiler2
  enrichment <- gost(
    query = gene_set,
    organism = "hsapiens",
    user_threshold = 0.05,
    correction_method = "fdr",
    domain_scope = "custom",
    custom_bg = background,
    sources = category # Either "REAC" or "KEGG"
  )
  
  # Check if enrichment results exist
  if (is.null(enrichment$result) || nrow(enrichment$result) == 0) {
    message(paste("No significant enrichment found for", category, "in gProfiler"))
    return(NULL)
  }
  
  # Convert results to tibble and process top 20 terms
  enrichment_tibble <- enrichment$result %>%
    as_tibble() %>%
    mutate(Category = category,
           neglog = -log10(p_value)) %>%  # Compute -log10(p-value)
    arrange(desc(neglog)) %>%
    slice_head(n = min(20, nrow(.)))  # Ensure safe slicing
  
  # Generate plot
  plot <- ggplot(enrichment_tibble, aes(x = neglog, y = reorder(term_name, neglog))) +
    geom_bar(stat = "identity", fill = color) +
    labs(x = "-log10(p-value)",
         y = y_title,
         title = paste("Enriched", category, "Pathways")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black")
    )
  
  return(plot)
}

# Perform analysis for Reactome and KEGG using ClusterProfiler
cluster_reactome <- process_clusterProfiler(
  gene_set = prob_4_0.5,
  background = background,
  category = "Reactome",
  color = "#2E86C1",
  y_title = "Reactome Pathways"
)

cluster_kegg <- process_clusterProfiler(
  gene_set = prob_4_0.5,
  background = background,
  category = "KEGG",
  color = "#28B463",
  y_title = "KEGG Pathways"
)

# Combine Reactome and KEGG for ClusterProfiler
if (!is.null(cluster_reactome) && !is.null(cluster_kegg)) {
  cluster_combined <- cluster_reactome / cluster_kegg
} else if (!is.null(cluster_reactome)) {
  cluster_combined <- cluster_reactome
} else if (!is.null(cluster_kegg)) {
  cluster_combined <- cluster_kegg
} else {
  cluster_combined <- NULL
}

# Perform analysis for Reactome and KEGG using GProfiler
gprofiler_reactome <- process_gprofiler(
  gene_set = prob_4_0.5,
  background = background,
  category = "REAC",  # Corrected category for Reactome in gProfiler
  color = "#D35400",
  y_title = "Reactome Pathways"
)

gprofiler_kegg <- process_gprofiler(
  gene_set = prob_4_0.5,
  background = background,
  category = "KEGG",
  color = "#F39C12",
  y_title = "KEGG Pathways"
)

# Combine Reactome and KEGG for GProfiler
if (!is.null(gprofiler_reactome) && !is.null(gprofiler_kegg)) {
  gprofiler_combined <- gprofiler_reactome / gprofiler_kegg
} else if (!is.null(gprofiler_reactome)) {
  gprofiler_combined <- gprofiler_reactome
} else if (!is.null(gprofiler_kegg)) {
  gprofiler_combined <- gprofiler_kegg
} else {
  gprofiler_combined <- NULL
}

# Display plots (if they are not NULL)
if (!is.null(cluster_combined)) print(cluster_combined)
if (!is.null(gprofiler_combined)) print(gprofiler_combined)
```

## **📌 DOX + CX (mid-late) response (0.5 µM)**
## **📌 GO Enrichment Clusterprofiler**
```{r GO8, echo=TRUE, message=FALSE, fig.width=12, fig.height=12}
# Perform GO enrichment analysis for BP, MF, and CC
go_enrichment_BP <- enrichGO(gene = prob_5_0.5,
                             OrgDb = org.Hs.eg.db,
                             keyType = "ENTREZID",
                             universe = background,
                             ont = "BP",
                             pvalueCutoff = 0.05)

go_enrichment_MF <- enrichGO(gene = prob_5_0.5,
                             OrgDb = org.Hs.eg.db,
                             keyType = "ENTREZID",
                             universe = background,
                             ont = "MF",
                             pvalueCutoff = 0.05)

go_enrichment_CC <- enrichGO(gene = prob_5_0.5,
                             OrgDb = org.Hs.eg.db,
                             keyType = "ENTREZID",
                             universe = background,
                             ont = "CC",
                             pvalueCutoff = 0.05)

# Convert each enrichment result to a tibble, add a category column, and select top 20 terms
process_enrichment_tibble <- function(enrichment, category) {
  if (is.null(enrichment) || nrow(as.data.frame(enrichment)) == 0) {
    return(tibble(Description = "No enriched terms", neglog = 0, Category = category))
  } else {
    enrichment %>%
      as_tibble() %>%
      mutate(Category = category,
             neglog = -log(p.adjust)) %>% # Add -log(p.adjust) column
      arrange(desc(neglog)) %>%          # Sort by -log(p.adjust)
      slice(1:20)                        # Select top 20 terms
  }
}

BP_Tibble <- process_enrichment_tibble(go_enrichment_BP, "Biological Process")
MF_Tibble <- process_enrichment_tibble(go_enrichment_MF, "Molecular Function")
CC_Tibble <- process_enrichment_tibble(go_enrichment_CC, "Cellular Component")

# Combine all tibbles
combined_GO_Tibble <- bind_rows(BP_Tibble, MF_Tibble, CC_Tibble)
# Function to generate enrichment plots
process_enrichment_plot <- function(tibble, title, color) {
  ggplot(data = tibble, aes(x = neglog, y = reorder(Description, neglog))) +
    geom_bar(stat = "identity", fill = color) +
    labs(x = "-log(p-adjust)",
         y = title,
         title = paste("Top 20", title, "GO Terms")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black"),
      panel.border = element_rect(colour = "black", fill = NA, size = 0.3)
    ) +
    xlim(c(0, max(tibble$neglog) + 1))
}

# Generate separate plots
plot_BP <- process_enrichment_plot(BP_Tibble, "Biological Process", "#2E86C1")
plot_MF <- process_enrichment_plot(MF_Tibble, "Molecular Function", "#28B463")
plot_CC <- process_enrichment_plot(CC_Tibble, "Cellular Component", "#D35400")

# Combine the plots using patchwork
combined_plot <- plot_BP / plot_MF / plot_CC

# Display the combined plot
combined_plot
```


## **📌 GO Enrichment g:Profiler**
```{r gprofiler_GO8, echo=TRUE, message=FALSE, fig.width=12, fig.height=12}

# Load the gprofiler2 package
library(gprofiler2)
library(ggplot2)
library(dplyr)
library(patchwork)

# Perform GO enrichment analysis with gprofiler2
gost_results <- gost(
  query = prob_5_0.5,           # List of input genes (prob_all_1)
  organism = "hsapiens",   # Human organism
  user_threshold = 0.05,   # Adjusted p-value cutoff
  correction_method = "fdr", # Multiple testing correction
  domain_scope = "custom", # Use custom background
  custom_bg = background,    # Background set of genes
  sources = c("GO:BP", "GO:MF", "GO:CC") # Analyze GO categories
)

# Check if enrichment results exist
if (is.null(gost_results$result) || nrow(gost_results$result) == 0) {
  # If no enriched terms, create a placeholder dataframe
  combined_results <- tibble(
    term_name = "No enriched terms",
    p.adjust = NA,
    source = "N/A",
    Category = "N/A"
  )
} else {
  # Convert results to a data frame
  gost_results_df <- gost_results$result

  # Add a column for adjusted p-value
  gost_results_df <- gost_results_df %>%
    rename(p.adjust = p_value)

  # Separate results for BP, MF, and CC
  BP_results <- gost_results_df %>%
    filter(source == "GO:BP") %>%
    mutate(Category = "Biological Process")

  MF_results <- gost_results_df %>%
    filter(source == "GO:MF") %>%
    mutate(Category = "Molecular Function")

  CC_results <- gost_results_df %>%
    filter(source == "GO:CC") %>%
    mutate(Category = "Cellular Component")

  # Select the top 20 terms by adjusted p-value for each category
  top_BP <- BP_results %>%
    arrange(p.adjust) %>%
    slice_head(n = 20)

  top_MF <- MF_results %>%
    arrange(p.adjust) %>%
    slice_head(n = 20)

  top_CC <- CC_results %>%
    arrange(p.adjust) %>%
    slice_head(n = 20)

  # Combine all categories
  combined_results <- bind_rows(top_BP, top_MF, top_CC)
}

# Ensure all columns are atomic types for CSV export
combined_results_clean <- combined_results %>%
  mutate(across(everything(), ~ if (is.list(.)) sapply(., toString) else .))

# Function for plotting top terms
plot_gprofiler_results <- function(data, title, color) {
  ggplot(data, aes(x = -log10(p.adjust), y = reorder(term_name, -log10(p.adjust)))) +
    geom_bar(stat = "identity", fill = color) +
    labs(
      x = "-log10(Adjusted p-value)",
      y = title,
      title = paste("Top 20", title, "GO Terms")
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black"),
      panel.border = element_rect(colour = "black", fill = NA, size = 0.3)
    )
}

# Check if there are enrichment terms to plot
if (nrow(combined_results) == 1 && combined_results$term_name == "No enriched terms") {
  message("No enriched GO terms found for the input gene set.")
} else {
  # Plot the top 20 terms for each category
  plot_BP <- plot_gprofiler_results(top_BP, "Biological Process", "#2E86C1")
  plot_MF <- plot_gprofiler_results(top_MF, "Molecular Function", "#28B463")
  plot_CC <- plot_gprofiler_results(top_CC, "Cellular Component", "#D35400")

  # Combine the plots using patchwork
  combined_plot <- plot_BP / plot_MF / plot_CC

  # Display the combined plot
  combined_plot
}
```

## **📌 Pathway Enrichment**
```{r pathway_enrichment8, echo=TRUE, message=FALSE, fig.width=15, fig.height=12}

# Load required libraries
library(clusterProfiler)
library(org.Hs.eg.db) # Required for enrichPathway
library(gprofiler2)
library(ggplot2)
library(dplyr)
library(patchwork)
library(ReactomePA)

# Function for ClusterProfiler Reactome & KEGG Analysis
process_clusterProfiler <- function(gene_set, background, category, color, y_title) {
  # Perform enrichment based on the selected category
  if (category == "Reactome") {
    enrichment <- enrichPathway(
      gene = gene_set,
      organism = "human",
      pvalueCutoff = 0.05,
      pAdjustMethod = "BH",
      universe = background
    )
  } else if (category == "KEGG") {
    enrichment <- enrichKEGG(
      gene = gene_set,
      organism = "hsa",
      pvalueCutoff = 0.05,
      pAdjustMethod = "BH",
      universe = background
    )
  }
  
  # Check if enrichment results exist
  if (is.null(enrichment) || nrow(as.data.frame(enrichment)) == 0) {
    message(paste("No significant enrichment found for", category, "in ClusterProfiler"))
    return(NULL)
  }
  
  # Convert results to tibble and process top 20 terms
  enrichment_tibble <- as_tibble(as.data.frame(enrichment)) %>%
    mutate(Category = category,
           neglog = -log10(p.adjust)) %>%  # Compute -log10(p.adjust)
    arrange(desc(neglog)) %>%
    slice_head(n = min(20, nrow(.)))  # Ensure safe slicing
  
  # Generate plot
  plot <- ggplot(enrichment_tibble, aes(x = neglog, y = reorder(Description, neglog))) +
    geom_bar(stat = "identity", fill = color) +
    labs(x = "-log10(Adjusted p-value)",
         y = y_title,
         title = paste("Enriched", category, "Pathways")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black")
    )
  
  return(plot)
}

# Function for gProfiler Reactome & KEGG Analysis
process_gprofiler <- function(gene_set, background, category, color, y_title) {
  # Perform enrichment using gprofiler2
  enrichment <- gost(
    query = gene_set,
    organism = "hsapiens",
    user_threshold = 0.05,
    correction_method = "fdr",
    domain_scope = "custom",
    custom_bg = background,
    sources = category # Either "REAC" or "KEGG"
  )
  
  # Check if enrichment results exist
  if (is.null(enrichment$result) || nrow(enrichment$result) == 0) {
    message(paste("No significant enrichment found for", category, "in gProfiler"))
    return(NULL)
  }
  
  # Convert results to tibble and process top 20 terms
  enrichment_tibble <- enrichment$result %>%
    as_tibble() %>%
    mutate(Category = category,
           neglog = -log10(p_value)) %>%  # Compute -log10(p-value)
    arrange(desc(neglog)) %>%
    slice_head(n = min(20, nrow(.)))  # Ensure safe slicing
  
  # Generate plot
  plot <- ggplot(enrichment_tibble, aes(x = neglog, y = reorder(term_name, neglog))) +
    geom_bar(stat = "identity", fill = color) +
    labs(x = "-log10(p-value)",
         y = y_title,
         title = paste("Enriched", category, "Pathways")) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(size = 12, face = "bold", colour = "black", angle = 45, hjust = 1),
      axis.text.y = element_text(size = 12, face = "bold", colour = "black"),
      axis.title.x = element_text(size = 14, face = "bold", colour = "black"),
      axis.title.y = element_text(size = 14, face = "bold", colour = "black"),
      plot.title = element_text(size = 14, face = "bold", colour = "black")
    )
  
  return(plot)
}

# Perform analysis for Reactome and KEGG using ClusterProfiler
cluster_reactome <- process_clusterProfiler(
  gene_set = prob_5_0.5,
  background = background,
  category = "Reactome",
  color = "#2E86C1",
  y_title = "Reactome Pathways"
)

cluster_kegg <- process_clusterProfiler(
  gene_set = prob_5_0.5,
  background = background,
  category = "KEGG",
  color = "#28B463",
  y_title = "KEGG Pathways"
)

# Combine Reactome and KEGG for ClusterProfiler
if (!is.null(cluster_reactome) && !is.null(cluster_kegg)) {
  cluster_combined <- cluster_reactome / cluster_kegg
} else if (!is.null(cluster_reactome)) {
  cluster_combined <- cluster_reactome
} else if (!is.null(cluster_kegg)) {
  cluster_combined <- cluster_kegg
} else {
  cluster_combined <- NULL
}

# Perform analysis for Reactome and KEGG using GProfiler
gprofiler_reactome <- process_gprofiler(
  gene_set = prob_5_0.5,
  background = background,
  category = "REAC",  # Corrected category for Reactome in gProfiler
  color = "#D35400",
  y_title = "Reactome Pathways"
)

gprofiler_kegg <- process_gprofiler(
  gene_set = prob_5_0.5,
  background = background,
  category = "KEGG",
  color = "#F39C12",
  y_title = "KEGG Pathways"
)

# Combine Reactome and KEGG for GProfiler
if (!is.null(gprofiler_reactome) && !is.null(gprofiler_kegg)) {
  gprofiler_combined <- gprofiler_reactome / gprofiler_kegg
} else if (!is.null(gprofiler_reactome)) {
  gprofiler_combined <- gprofiler_reactome
} else if (!is.null(gprofiler_kegg)) {
  gprofiler_combined <- gprofiler_kegg
} else {
  gprofiler_combined <- NULL
}

# Display plots (if they are not NULL)
if (!is.null(cluster_combined)) print(cluster_combined)
if (!is.null(gprofiler_combined)) print(gprofiler_combined)
```


## **📌 Top BP (Cluster Profiler)**
```{r GO comp, echo=TRUE, message=FALSE, fig.width=14, fig.height=6}

# Load necessary libraries
library(clusterProfiler)
library(ggplot2)
library(dplyr)
library(tibble)

# Define response groups and their corresponding gene lists
response_groups_0.1 <- list(
  "Non response (0.1)" = prob_1_0.1,
  "CX-DOX mid-late response (0.1)" = prob_2_0.1,
  "DOX only mid-late (0.1)" = prob_3_0.1
)

response_groups_0.5 <- list(
  "Non response (0.5)" = prob_1_0.5,
  "DOX specific response (0.5)" = prob_2_0.5,
  "DOX only mid-late response (0.5)" = prob_3_0.5,
  "CX + DOX (early) response (0.5)" = prob_4_0.5,
  "DOX + CX (mid-late) response (0.5)" = prob_5_0.5
)

# Function to perform GO enrichment for BP
perform_BP_enrichment <- function(gene_set, sample_name, concentration) {
  enrichment <- enrichGO(
    gene = gene_set,
    OrgDb = org.Hs.eg.db,
    keyType = "ENTREZID",
    universe = background,
    ont = "BP",
    pvalueCutoff = 0.05
  )
  
  if (!is.null(enrichment) && nrow(as.data.frame(enrichment)) > 0) {
    enrichment_df <- as_tibble(as.data.frame(enrichment)) %>%
      mutate(
        Sample = sample_name,
        Concentration = concentration,  # Assign correct concentration here
        GeneRatio = Count / as.numeric(sub("/.*", "", BgRatio)),  # Compute Gene Ratio
        neglog_padj = -log10(p.adjust)  # Compute -log10(adjusted p-value)
      ) %>%
      arrange(p.adjust) %>%
      slice_head(n = 5)  # Select top 5 terms
    
    return(enrichment_df)
  } else {
    return(NULL)
  }
}

# Perform enrichment for 0.1-specific response groups
BP_results_0.1 <- lapply(names(response_groups_0.1), function(sample) {
  perform_BP_enrichment(response_groups_0.1[[sample]], sample, "0.1")
})

# Perform enrichment for 0.5-specific response groups
BP_results_0.5 <- lapply(names(response_groups_0.5), function(sample) {
  perform_BP_enrichment(response_groups_0.5[[sample]], sample, "0.5")
})

# Combine results into a single dataframe
BP_combined_df <- bind_rows(BP_results_0.1, BP_results_0.5)

# Ensure all response groups appear in the x-axis, even if they have no enriched terms
all_groups_0.1 <- names(response_groups_0.1)
all_groups_0.5 <- names(response_groups_0.5)

missing_groups_0.1 <- setdiff(all_groups_0.1, unique(BP_combined_df$Sample))
missing_groups_0.5 <- setdiff(all_groups_0.5, unique(BP_combined_df$Sample))

# Create placeholder rows for missing samples, using "No enriched terms" instead of NA
if (length(missing_groups_0.1) > 0) {
  missing_df_0.1 <- tibble(
    Sample = missing_groups_0.1,
    Description = "No enriched terms",  # Show this in the y-axis
    GeneRatio = NA,
    neglog_padj = NA,
    Concentration = "0.1"  # Assign to 0.1 group
  )
  BP_combined_df <- bind_rows(BP_combined_df, missing_df_0.1)
}

if (length(missing_groups_0.5) > 0) {
  missing_df_0.5 <- tibble(
    Sample = missing_groups_0.5,
    Description = "No enriched terms",  # Show this in the y-axis
    GeneRatio = NA,
    neglog_padj = NA,
    Concentration = "0.5"  # Assign to 0.5 group
  )
  BP_combined_df <- bind_rows(BP_combined_df, missing_df_0.5)
}

# Ensure all response groups appear in the correct order
BP_combined_df$Sample <- factor(BP_combined_df$Sample, levels = c(all_groups_0.1, all_groups_0.5))

# Generate dot plot
dotplot_BP <- ggplot(BP_combined_df, aes(x = Sample, y = Description, size = GeneRatio, color = neglog_padj)) +
  geom_point(na.rm = TRUE) +  # Ignore NA values
  scale_color_gradient(low = "blue", high = "red", na.value = "grey") +  # NA values in grey
  labs(
    x = "Response Groups",
    y = "Biological Process",
    color = "-log10(Adjusted P-Value)",
    size = "Gene Ratio",
    title = "Top 5 BP Enrichments Across Response Groups (0.1 & 0.5)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10, face = "bold"),
    axis.text.y = element_text(size = 10, face = "bold"),
    legend.text = element_text(size = 10)
  ) +
  facet_wrap(~ Concentration, scales = "free_x")  # Ensures 0.1 and 0.5 groups don't overlap

# Display the dot plot
print(dotplot_BP)
```

## **📌 Top BP (GProfiler)**
```{r GO comp2, echo=TRUE, message=FALSE, fig.width=14, fig.height=6}
# Load necessary libraries
library(gprofiler2)
library(ggplot2)
library(dplyr)
library(tibble)

# Define response groups and their corresponding gene lists
response_groups_0.1 <- list(
  "Non response (0.1)" = prob_1_0.1,
  "CX-DOX mid-late response (0.1)" = prob_2_0.1,
  "DOX only mid-late (0.1)" = prob_3_0.1
)

response_groups_0.5 <- list(
  "Non response (0.5)" = prob_1_0.5,
  "DOX specific response (0.5)" = prob_2_0.5,
  "DOX only mid-late response (0.5)" = prob_3_0.5,
  "CX + DOX (early) response (0.5)" = prob_4_0.5,
  "DOX + CX (mid-late) response (0.5)" = prob_5_0.5
)

# Function to perform GO enrichment for BP using g:Profiler
perform_BP_enrichment <- function(gene_set, sample_name, concentration) {
  enrichment <- gost(
    query = gene_set,
    organism = "hsapiens",
    ordered_query = FALSE,
    significant = TRUE,
    correction_method = "fdr",
    sources = "GO:BP",
    custom_bg = background
  )
  
  if (!is.null(enrichment) && nrow(enrichment$result) > 0) {
    enrichment_df <- as_tibble(enrichment$result) %>%
      mutate(
        Sample = sample_name,
        Concentration = concentration,  # Assign correct concentration here
        GeneRatio = intersection_size / effective_domain_size,  # Compute Gene Ratio
        neglog_padj = -log10(p_value)  # Compute -log10(adjusted p-value)
      ) %>%
      arrange(p_value) %>%
      slice_head(n = 5)  # Select top 5 terms
    
    return(enrichment_df)
  } else {
    return(NULL)
  }
}

# Perform enrichment for 0.1-specific response groups
BP_results_0.1 <- lapply(names(response_groups_0.1), function(sample) {
  perform_BP_enrichment(response_groups_0.1[[sample]], sample, "0.1")
})

# Perform enrichment for 0.5-specific response groups
BP_results_0.5 <- lapply(names(response_groups_0.5), function(sample) {
  perform_BP_enrichment(response_groups_0.5[[sample]], sample, "0.5")
})

# Combine results into a single dataframe
BP_combined_df <- bind_rows(BP_results_0.1, BP_results_0.5)

# Ensure all response groups appear in the x-axis, even if they have no enriched terms
all_groups_0.1 <- names(response_groups_0.1)
all_groups_0.5 <- names(response_groups_0.5)

missing_groups_0.1 <- setdiff(all_groups_0.1, unique(BP_combined_df$Sample))
missing_groups_0.5 <- setdiff(all_groups_0.5, unique(BP_combined_df$Sample))

# Create placeholder rows for missing samples, using "No enriched terms" instead of NA
if (length(missing_groups_0.1) > 0) {
  missing_df_0.1 <- tibble(
    Sample = missing_groups_0.1,
    term_name = "No enriched terms",  # Show this in the y-axis
    GeneRatio = NA,
    neglog_padj = NA,
    Concentration = "0.1"  # Assign to 0.1 group
  )
  BP_combined_df <- bind_rows(BP_combined_df, missing_df_0.1)
}

if (length(missing_groups_0.5) > 0) {
  missing_df_0.5 <- tibble(
    Sample = missing_groups_0.5,
    term_name = "No enriched terms",  # Show this in the y-axis
    GeneRatio = NA,
    neglog_padj = NA,
    Concentration = "0.5"  # Assign to 0.5 group
  )
  BP_combined_df <- bind_rows(BP_combined_df, missing_df_0.5)
}

# Ensure all response groups appear in the correct order
BP_combined_df$Sample <- factor(BP_combined_df$Sample, levels = c(all_groups_0.1, all_groups_0.5))

# Generate dot plot
dotplot_BP <- ggplot(BP_combined_df, aes(x = Sample, y = term_name, size = GeneRatio, color = neglog_padj)) +
  geom_point(na.rm = TRUE) +  # Ignore NA values
  scale_color_gradient(low = "blue", high = "red", na.value = "grey") +  # NA values in grey
  labs(
    x = "Response Groups",
    y = "Biological Process",
    color = "-log10(Adjusted P-Value)",
    size = "Gene Ratio",
    title = "Top 5 BP Enrichments Across Response Groups (0.1 & 0.5)"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10, face = "bold"),
    axis.text.y = element_text(size = 10, face = "bold"),
    legend.text = element_text(size = 10)
  ) +
  facet_wrap(~ Concentration, scales = "free_x")  # Ensures 0.1 and 0.5 groups don't overlap

# Display the dot plot
print(dotplot_BP)
```

