---
title: "Count Matrix Analysis"
author: "Sayan Paul"
date: "`r Sys.Date()`"
output: workflowr::wflow_html
---

### **Load Required Libraries**
```{r setup, include=FALSE}
library(edgeR)
library(ggplot2)
library(reshape2)
library(dplyr)
library(Biobase)
library(limma)
library(RColorBrewer)
library(tidyverse)
library(scales)
library(biomaRt)
library(Homo.sapiens)
library(cowplot)
library(ggrepel)
library(corrplot)
library(Hmisc)
library(org.Hs.eg.db)
library(AnnotationDbi)
```

📍 Load Count Matrix
```{r Read mapping counts, echo=TRUE, message=FALSE}
# Load the count matrix from the TSV file
counts_matrix <- read.csv("data/counts_matrix.csv", header=TRUE, check.names=FALSE)
# Remove "X" prefix from column names (if present)
colnames(counts_matrix) <- gsub("^X", "", colnames(counts_matrix))
# Display first few rows
head(counts_matrix)
```

📍 Boxplot of Raw Counts (Unnormalized)
```{r boxplot unnormalized, echo=TRUE, message=FALSE}
# Boxplot of unnormalized counts
boxplot(counts_matrix, 
        main = "Boxplots of samples (Unnormalized)", 
        names = colnames(counts_matrix), 
        las = 2, 
        cex.axis = 0.7)
```

📍 Log-Transformed Counts (CPM)
```{r cpm, echo=TRUE, message=FALSE}
# Compute CPM values
cpm <- cpm(counts_matrix)
lcpm <- cpm(counts_matrix, log=TRUE)

# Dimensions
dim(lcpm)
```

📍 Histogram of Unfiltered Log Counts
```{r hist_cpm, echo=TRUE, message=FALSE,fig.width=28,fig.height=8}
hist(lcpm,  
     main = "Histogram of total counts (unfiltered)", 
     xlab = expression("Log"[2]*" counts-per-million"), 
     col = 4)
```

📍 Filtering Count Matrix (rowMeans > 0)
```{r row0, echo=TRUE, message=FALSE}
filcpm_matrix <- subset(lcpm, (rowMeans(lcpm) > 0))
dim(filcpm_matrix)
```

📍 Histogram of filtered Log Counts (rowMeans > 0)
```{r hist_row0, echo=TRUE, message=FALSE}
hist(filcpm_matrix,  
     main = "Histogram of filtered counts using rowMeans > 0 method", 
     xlab = expression("Log"[2]*" counts-per-million"), 
     col = 2)
```

📍 Filtering Count Matrix (rowMeans > 0.5)
```{r row0.5, echo=TRUE, message=FALSE}
filcpm_matrix1 <- subset(lcpm, (rowMeans(lcpm) > 0.5))
dim(filcpm_matrix1)
```

📍 Histogram of filtered Log Counts (rowMeans > 0.5)
```{r hist_row0.5, echo=TRUE, message=FALSE}
hist(filcpm_matrix1,  
     main = "Histogram of filtered counts using rowMeans > 0.5 method", 
     xlab = expression("Log"[2]*" counts-per-million"), 
     col = 5)
```

📍 Filtering Count Matrix (rowMeans > 1)
```{r row1, echo=TRUE, message=FALSE}
filcpm_matrix2 <- subset(lcpm, (rowMeans(lcpm) > 1))
dim(filcpm_matrix2)
```

📍 Histogram of filtered Log Counts (rowMeans > 0.5)
```{r hist_row1, echo=TRUE, message=FALSE}
hist(filcpm_matrix2,  
     main = "Histogram of filtered counts using rowMeans > 1 method", 
     xlab = expression("Log"[2]*" counts-per-million"), 
     col = 6)
```


