---
title: "Count Matrix Analysis"
author: "Sayan Paul"
date: "`r Sys.Date()`"
output: workflowr::wflow_html
---

### **Load Required Libraries**
```{r setup, include=FALSE}
library(edgeR)
library(ggplot2)
library(reshape2)
library(dplyr)
library(Biobase)
library(limma)
library(RColorBrewer)
library(tidyverse)
library(scales)
library(biomaRt)
library(Homo.sapiens)
library(cowplot)
library(ggrepel)
library(corrplot)
library(Hmisc)
library(org.Hs.eg.db)
library(AnnotationDbi)
```

📍 Load Count Matrix
```{r Read mapping counts, echo=TRUE, message=FALSE}
# Load the count matrix from the TSV file
counts_matrix <- read.csv("data/counts_matrix.csv", header=TRUE, check.names=FALSE)
# Remove "X" prefix from column names (if present)
colnames(counts_matrix) <- gsub("^X", "", colnames(counts_matrix))
# Display first few rows
head(counts_matrix)
```

### **📌 Boxplot of Raw Counts (Unnormalized)**
```{r boxplot_unnormalized, echo=TRUE, message=FALSE, fig.width=20, fig.height=8}
# Adjust graphical parameters for better spacing
par(mar = c(12, 5, 5, 2))  # Bottom margin increased to 12

# Boxplot of unnormalized counts
boxplot(counts_matrix, 
        main = "Boxplots of samples (Unnormalized)", 
        names = colnames(counts_matrix), 
        las = 2,      # Make labels vertical
        cex.axis = 0.6)  # Reduce text size for better fit
```

📍 Log-Transformed Counts (CPM)
```{r cpm, echo=TRUE, message=FALSE}
# Compute CPM values
cpm <- cpm(counts_matrix)
lcpm <- cpm(counts_matrix, log=TRUE)

# Dimensions
dim(lcpm)
```

### **📌 Histogram of Unfiltered Log Counts**
```{r hist_cpm, echo=TRUE, message=FALSE}
hist(lcpm,  
     main = "Histogram of total counts (unfiltered)", 
     xlab = expression("Log"[2]*" counts-per-million"), 
     col = 4)
```

📍 Filtering Count Matrix (rowMeans > 0)
```{r row0, echo=TRUE, message=FALSE}
filcpm_matrix <- subset(lcpm, (rowMeans(lcpm) > 0))
dim(filcpm_matrix)
```

### **📌 Histogram of filtered Log Counts (rowMeans > 0)**
```{r hist_row0, echo=TRUE, message=FALSE}
hist(filcpm_matrix,  
     main = "Histogram of filtered counts using rowMeans > 0 method", 
     xlab = expression("Log"[2]*" counts-per-million"), 
     col = 2)
```

📍 Filtering Count Matrix (rowMeans > 0.5)
```{r row0.5, echo=TRUE, message=FALSE}
filcpm_matrix1 <- subset(lcpm, (rowMeans(lcpm) > 0.5))
dim(filcpm_matrix1)
```

### **📌 Histogram of filtered Log Counts (rowMeans > 0.5)**
```{r hist_row0.5, echo=TRUE, message=FALSE}
hist(filcpm_matrix1,  
     main = "Histogram of filtered counts using rowMeans > 0.5 method", 
     xlab = expression("Log"[2]*" counts-per-million"), 
     col = 5)
```

📍 Filtering Count Matrix (rowMeans > 1)
```{r row1, echo=TRUE, message=FALSE}
filcpm_matrix2 <- subset(lcpm, (rowMeans(lcpm) > 1))
dim(filcpm_matrix2)
```

### **📌 Histogram of filtered Log Counts (rowMeans > 0.5)**
```{r hist_row1, echo=TRUE, message=FALSE}
hist(filcpm_matrix2,  
     main = "Histogram of filtered counts using rowMeans > 1 method", 
     xlab = expression("Log"[2]*" counts-per-million"), 
     col = 6)
```


### **📌 Density Plots for Log2 CPM Values**
```{r den_lCPM, echo=TRUE, message=FALSE}
## Convert log2 CPM matrix into long format for visualization
log2_cpm_long <- as.data.frame(as.table(lcpm))
colnames(log2_cpm_long) <- c("Feature", "Sample", "Log2_CPM")

## Density Plot of Log2 CPM Values (Unfiltered)
ggplot(log2_cpm_long, aes(x = Log2_CPM)) +
  geom_density(fill = "blue", alpha = 0.5) +
  labs(title = "Density Plot of Log2 CPM Values (Unfiltered)", x = "Log2 CPM", y = "Density") +
  theme_minimal()

## Histogram + Density Plot for Log2 CPM Values (Unfiltered)
ggplot(log2_cpm_long, aes(x = Log2_CPM)) +
  geom_histogram(aes(y = ..density..),
                 bins = 40,
                 fill = "lightblue",
                 color = "black") +
  geom_density(alpha = 0.5, fill = "lightgreen") +
  labs(title = "Log2 CPM Values (Unfiltered) (Density+Histogram)",
       x = "Log2 CPM",
       y = "Density") +
  theme_minimal()
```

### **📌 Density Plots for Filtered Log2 CPM Values (rowMeans > 0)**
```{r den_lCPM0, echo=TRUE, message=FALSE}
log2_cpm_long1 <- as.data.frame(as.table(filcpm_matrix))
colnames(log2_cpm_long1) <- c("Feature", "Sample", "Log2_CPM")

## Density Plot for Log2 CPM (rowMeans > 0)
ggplot(log2_cpm_long1, aes(x = Log2_CPM)) +
  geom_density(fill = "blue", alpha = 0.5) +
  labs(title = "Density Plot of Log2 CPM Values (rowMeans > 0)", x = "Log2 CPM", y = "Density") +
  theme_minimal()

## Histogram + Density Plot for Log2 CPM (rowMeans > 0)
ggplot(log2_cpm_long1, aes(x = Log2_CPM)) +
  geom_histogram(aes(y = ..density..),
                 bins = 40,
                 fill = "lightblue",
                 color = "black") +
  geom_density(alpha = 0.5, fill = "lightgreen") +
  labs(title = "Log2 CPM Values (rowMeans > 0) (Density+Histogram)",
       x = "Log2 CPM",
       y = "Density") +
  theme_minimal()
```

### **📌 Density Plots for Filtered Log2 CPM Values (rowMeans > 0.5)**
```{r den_lCPM0.5, echo=TRUE, message=FALSE}
log2_cpm_long2 <- as.data.frame(as.table(filcpm_matrix1))
colnames(log2_cpm_long2) <- c("Feature", "Sample", "Log2_CPM")

## Density Plot for Log2 CPM (rowMeans > 0.5)
ggplot(log2_cpm_long2, aes(x = Log2_CPM)) +
  geom_density(fill = "blue", alpha = 0.5) +
  labs(title = "Density Plot of Log2 CPM Values (rowMeans > 0.5)", x = "Log2 CPM", y = "Density") +
  theme_minimal()

## Histogram + Density Plot for Log2 CPM (rowMeans > 0.5)
ggplot(log2_cpm_long2, aes(x = Log2_CPM)) +
  geom_histogram(aes(y = ..density..),
                 bins = 40,
                 fill = "lightblue",
                 color = "black") +
  geom_density(alpha = 0.5, fill = "lightgreen") +
  labs(title = "Log2 CPM Values (rowMeans > 0.5) (Density+Histogram)",
       x = "Log2 CPM",
       y = "Density") +
  theme_minimal()
```

### **📌 Density Plots for Filtered Log2 CPM Values (rowMeans > 1)**
```{r den_lCPM1, echo=TRUE, message=FALSE}
log2_cpm_long3 <- as.data.frame(as.table(filcpm_matrix2))

colnames(log2_cpm_long3) <- c("Feature", "Sample", "Log2_CPM")
## Density Plot for Log2 CPM (rowMeans > 1)
ggplot(log2_cpm_long3, aes(x = Log2_CPM)) +
  geom_density(fill = "blue", alpha = 0.5) +
  labs(title = "Density Plot of Log2 CPM Values (rowMeans > 1)", x = "Log2 CPM", y = "Density") +
  theme_minimal()
## Histogram + Density Plot for Log2 CPM (rowMeans > 1)
ggplot(log2_cpm_long3, aes(x = Log2_CPM)) +
  geom_histogram(aes(y = ..density..),
                 bins = 40,
                 fill = "lightblue",
                 color = "black") +
  geom_density(alpha = 0.5, fill = "lightgreen") +
  labs(title = "Log2 CPM Values (rowMeans > 1) (Density+Histogram)",
       x = "Log2 CPM",
       y = "Density") +
  theme_minimal()
```
