---
title: "Log2CPM Boxplots of Most Significant DE Genes"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## **📌 Log2CPM Boxplots of Most Significant DE Genes**

This analysis generates boxplots for the log2CPM values of the most significant DE genes across CX-5461 and DOX treatments.

---

## **📌 Load Required Libraries**
```{r load_libraries, echo=TRUE, message=FALSE}
library(dplyr)
library(ggplot2)
library(ggsignif)
library(tidyr)
```

## **📌 CX_0.1_3 vs VEH_0.1_3**

## **📌 Define File Paths**
```{r load_path_CX_3, echo=TRUE,results='hide',message=FALSE}
# Define file paths
input_file <- "data/Feature_count_Matrix_Log2CPM_filtered.csv"
toptable_file <- "data/DEGs/Toptable_CX_0.1_3.csv"
output_boxplot <- "data/Boxplot_CX_0.1_3.csv"
output_log2CPM <- "data/log2CPM_filtered.csv"
```


## **📌 Read and Clean Dataset**
```{r load_data_CX3, echo=TRUE,results='hide',message=FALSE}
# Read Log2CPM Matrix
boxplot1 <- read.csv(input_file) %>%
  as.data.frame()
colnames(boxplot1) <- trimws(gsub("^X", "", colnames(boxplot1)))  # Clean column names

# Select relevant columns
matching_columns <- grep("ENTREZID|SYMBOL|CX.5461_0.1_3|VEH_0.1_3", colnames(boxplot1), value = TRUE)

if (!"SYMBOL" %in% colnames(boxplot1)) stop("ERROR: 'SYMBOL' column is missing in boxplot1! Check dataset.")

new_boxplot1 <- boxplot1[, matching_columns]

# Save cleaned datasets
write.csv(boxplot1, output_log2CPM, row.names = FALSE)
write.csv(new_boxplot1, output_boxplot, row.names = FALSE)

# Display first few rows
head(new_boxplot1)
```

## **📌 Define Color Palettes**
```{r load_color_CX3, echo=TRUE,message=FALSE}
drug_palc <- c("CX.5461" = "#8B006D", "VEH" = "#F1B72B")
```


## **📌 Read and Process Toptable Data**
```{r Read_CX3, echo=TRUE,message=FALSE}
toptable <- read.csv(toptable_file)
colnames(toptable)[1] <- "Entrezid"

filtered_toptable <- toptable %>%
  filter(adj.P.Val < 0.05) %>%
  arrange(adj.P.Val) %>%
  slice_head(n = 10)

# Display first few rows of the filtered Toptable
head(filtered_toptable)
```

## **📌 Filter Boxplot Data for Selected Genes**
```{r Filter_CX3, echo=TRUE,message=FALSE}
filtered_boxplot <- new_boxplot1 %>%
  filter(ENTREZID %in% filtered_toptable$Entrezid)

if (!"SYMBOL" %in% colnames(filtered_boxplot)) stop("ERROR: 'SYMBOL' column is missing in filtered_boxplot!")

# Pivot data to long format
long_boxplot <- filtered_boxplot %>%
  pivot_longer(cols = -c(ENTREZID, SYMBOL), names_to = "Sample", values_to = "log2CPM") %>%
  as.data.frame()

if (!"SYMBOL" %in% colnames(long_boxplot)) stop("ERROR: 'SYMBOL' column is missing after pivoting!")

long_boxplot$SYMBOL <- as.character(long_boxplot$SYMBOL)
```


## **📌 Define Individual Mapping**
```{r Ind_CX3, echo=TRUE, message=FALSE}
indv_mapping <- c("75.1" = 1, "78.1" = 2, "87.1" = 3, "17.3" = 4, "84.1" = 5, "90.1" = 6)

# Debug: Print column names before mutate
print("Columns in long_boxplot before mutate():")
print(colnames(long_boxplot))
```

## **📌Process Data with Metadata**
```{r meta_CX3, echo=TRUE, message=FALSE}
formatted_data <- long_boxplot %>%
  mutate(
    Indv_id = sub("_.*", "", Sample),  # Extract individual ID
    Indv = ifelse(Indv_id %in% names(indv_mapping), as.character(indv_mapping[Indv_id]), NA),
    Drug = ifelse(grepl("CX.5461", Sample), "CX.5461", "VEH"),
    Conc = "0.1",
    Timepoint = "3"
  ) %>%
  dplyr::select(any_of(c("ENTREZID", "SYMBOL", "Sample", "Indv", "Drug", "Conc", "Timepoint", "log2CPM"))) %>%
  as.data.frame()

# Rename SYMBOL → Gene
colnames(formatted_data) <- trimws(colnames(formatted_data))
if ("SYMBOL" %in% colnames(formatted_data)) {
  colnames(formatted_data)[colnames(formatted_data) == "SYMBOL"] <- "Gene"
} else {
  stop("ERROR: 'SYMBOL' column is missing before renaming.")
}

if (!"Gene" %in% colnames(formatted_data)) stop("ERROR: 'Gene' column is missing after renaming.")

formatted_data$Indv <- as.character(formatted_data$Indv)

# Display first few rows
head(formatted_data)
```

## **📌Generate Boxplot**
```{r box_CX3, echo=TRUE, message=FALSE}
ggplot(formatted_data, aes(x = Drug, y = log2CPM, fill = Drug)) +
  geom_boxplot(outlier.shape = NA) +
  scale_fill_manual(values = drug_palc) +
  facet_wrap(~ Gene, ncol = 5) +
  geom_point(aes(color = Indv), size = 2, alpha = 0.5, position = position_jitter(width = -1, height = 0)) +
  ggtitle("log2CPM CX.5461_0.1 vs Vehicle_0.1 (3hrs)") +
  labs(x = "Drugs", y = "log2CPM") +
  theme_bw() +
  theme(
    plot.title = element_text(size = rel(2), hjust = 0.5),
    axis.title = element_text(size = 15, color = "black"),
    axis.ticks = element_line(linewidth = 1.5),
    axis.line = element_line(linewidth = 1.5),
    axis.text.y = element_text(size = 10, color = "black"),
    axis.text.x = element_text(size = 10, color = "black"),
    strip.text.y = element_text(color = "white")
  )
```


## **📌 CX_0.1_24 vs VEH_0.1_24**

## **📌 Define File Paths**
```{r load_path_CX24, echo=TRUE,results='hide',message=FALSE}
# Define file paths
input_file <- "data/Feature_count_Matrix_Log2CPM_filtered.csv"
toptable_file <- "data/DEGs/Toptable_CX_0.1_24.csv"
output_boxplot <- "data/Boxplot_CX_0.1_24.csv"
output_log2CPM <- "data/log2CPM_filtered.csv"
```


## **📌 Read and Clean Dataset**
```{r load_data_CX24, echo=TRUE,results='hide',message=FALSE}
# Read Log2CPM Matrix
boxplot1 <- read.csv(input_file) %>%
  as.data.frame()
colnames(boxplot1) <- trimws(gsub("^X", "", colnames(boxplot1)))  # Clean column names

# Select relevant columns
matching_columns <- grep("ENTREZID|SYMBOL|CX.5461_0.1_24|VEH_0.1_24", colnames(boxplot1), value = TRUE)

if (!"SYMBOL" %in% colnames(boxplot1)) stop("ERROR: 'SYMBOL' column is missing in boxplot1! Check dataset.")

new_boxplot1 <- boxplot1[, matching_columns]

# Save cleaned datasets
write.csv(boxplot1, output_log2CPM, row.names = FALSE)
write.csv(new_boxplot1, output_boxplot, row.names = FALSE)

# Display first few rows
head(new_boxplot1)
```

## **📌 Define Color Palettes**
```{r load_color_CX24, echo=TRUE,message=FALSE}
drug_palc <- c("CX.5461" = "#8B006D", "VEH" = "#F1B72B")
```


## **📌 Read and Process Toptable Data**
```{r Read_CX24, echo=TRUE,message=FALSE}
toptable <- read.csv(toptable_file)
colnames(toptable)[1] <- "Entrezid"

filtered_toptable <- toptable %>%
  filter(adj.P.Val < 0.05) %>%
  arrange(adj.P.Val) %>%
  slice_head(n = 10)

# Display first few rows of the filtered Toptable
head(filtered_toptable)
```

## **📌 Filter Boxplot Data for Selected Genes**
```{r Filter_CX24, echo=TRUE,message=FALSE}
filtered_boxplot <- new_boxplot1 %>%
  filter(ENTREZID %in% filtered_toptable$Entrezid)

if (!"SYMBOL" %in% colnames(filtered_boxplot)) stop("ERROR: 'SYMBOL' column is missing in filtered_boxplot!")

# Pivot data to long format
long_boxplot <- filtered_boxplot %>%
  pivot_longer(cols = -c(ENTREZID, SYMBOL), names_to = "Sample", values_to = "log2CPM") %>%
  as.data.frame()

if (!"SYMBOL" %in% colnames(long_boxplot)) stop("ERROR: 'SYMBOL' column is missing after pivoting!")

long_boxplot$SYMBOL <- as.character(long_boxplot$SYMBOL)
```


## **📌 Define Individual Mapping**
```{r Ind_CX24, echo=TRUE, message=FALSE}
indv_mapping <- c("75.1" = 1, "78.1" = 2, "87.1" = 3, "17.3" = 4, "84.1" = 5, "90.1" = 6)

# Debug: Print column names before mutate
print("Columns in long_boxplot before mutate():")
print(colnames(long_boxplot))
```

## **📌Process Data with Metadata**
```{r meta_CX24, echo=TRUE, message=FALSE}
formatted_data <- long_boxplot %>%
  mutate(
    Indv_id = sub("_.*", "", Sample),  # Extract individual ID
    Indv = ifelse(Indv_id %in% names(indv_mapping), as.character(indv_mapping[Indv_id]), NA),
    Drug = ifelse(grepl("CX.5461", Sample), "CX.5461", "VEH"),
    Conc = "0.1",
    Timepoint = "24"
  ) %>%
  dplyr::select(any_of(c("ENTREZID", "SYMBOL", "Sample", "Indv", "Drug", "Conc", "Timepoint", "log2CPM"))) %>%
  as.data.frame()

# Rename SYMBOL → Gene
colnames(formatted_data) <- trimws(colnames(formatted_data))
if ("SYMBOL" %in% colnames(formatted_data)) {
  colnames(formatted_data)[colnames(formatted_data) == "SYMBOL"] <- "Gene"
} else {
  stop("ERROR: 'SYMBOL' column is missing before renaming.")
}

if (!"Gene" %in% colnames(formatted_data)) stop("ERROR: 'Gene' column is missing after renaming.")

formatted_data$Indv <- as.character(formatted_data$Indv)

# Display first few rows
head(formatted_data)
```

## **📌Generate Boxplot**
```{r box_CX24, echo=TRUE, message=FALSE, fig.width=11, fig.height=8}
ggplot(formatted_data, aes(x = Drug, y = log2CPM, fill = Drug)) +
  geom_boxplot(outlier.shape = NA) +
  scale_fill_manual(values = drug_palc) +
  facet_wrap(~ Gene, ncol = 5) +
  geom_point(aes(color = Indv), size = 2, alpha = 0.5, position = position_jitter(width = -1, height = 0)) +
  ggtitle("log2CPM CX.5461_0.1 vs Vehicle_0.1 (24hrs)") +
  labs(x = "Drugs", y = "log2CPM") +
  theme_bw() +
  theme(
    plot.title = element_text(size = rel(2), hjust = 0.5),
    axis.title = element_text(size = 15, color = "black"),
    axis.ticks = element_line(linewidth = 1.5),
    axis.line = element_line(linewidth = 1.5),
    axis.text.y = element_text(size = 10, color = "black"),
    axis.text.x = element_text(size = 10, color = "black"),
    strip.text.y = element_text(color = "white")
  )
```



## **📌 CX_0.1_48 vs VEH_0.1_48**

## **📌 Define File Paths**
```{r load_path_CX48, echo=TRUE,results='hide',message=FALSE}
# Define file paths
input_file <- "data/Feature_count_Matrix_Log2CPM_filtered.csv"
toptable_file <- "data/DEGs/Toptable_CX_0.1_48.csv"
output_boxplot <- "data/Boxplot_CX_0.1_48.csv"
output_log2CPM <- "data/log2CPM_filtered.csv"
```


## **📌 Read and Clean Dataset**
```{r load_data_CX48, echo=TRUE,results='hide',message=FALSE}
# Read Log2CPM Matrix
boxplot1 <- read.csv(input_file) %>%
  as.data.frame()
colnames(boxplot1) <- trimws(gsub("^X", "", colnames(boxplot1)))  # Clean column names

# Select relevant columns
matching_columns <- grep("ENTREZID|SYMBOL|CX.5461_0.1_48|VEH_0.1_48", colnames(boxplot1), value = TRUE)

if (!"SYMBOL" %in% colnames(boxplot1)) stop("ERROR: 'SYMBOL' column is missing in boxplot1! Check dataset.")

new_boxplot1 <- boxplot1[, matching_columns]

# Save cleaned datasets
write.csv(boxplot1, output_log2CPM, row.names = FALSE)
write.csv(new_boxplot1, output_boxplot, row.names = FALSE)

# Display first few rows
head(new_boxplot1)
```

## **📌 Define Color Palettes**
```{r load_color_CX48, echo=TRUE,message=FALSE}
drug_palc <- c("CX.5461" = "#8B006D", "VEH" = "#F1B72B")
```


## **📌 Read and Process Toptable Data**
```{r Read_CX48, echo=TRUE,message=FALSE}
toptable <- read.csv(toptable_file)
colnames(toptable)[1] <- "Entrezid"

filtered_toptable <- toptable %>%
  filter(adj.P.Val < 0.05) %>%
  arrange(adj.P.Val) %>%
  slice_head(n = 10)

# Display first few rows of the filtered Toptable
head(filtered_toptable)
```

## **📌 Filter Boxplot Data for Selected Genes**
```{r Filter_CX48, echo=TRUE,message=FALSE}
filtered_boxplot <- new_boxplot1 %>%
  filter(ENTREZID %in% filtered_toptable$Entrezid)

if (!"SYMBOL" %in% colnames(filtered_boxplot)) stop("ERROR: 'SYMBOL' column is missing in filtered_boxplot!")

# Pivot data to long format
long_boxplot <- filtered_boxplot %>%
  pivot_longer(cols = -c(ENTREZID, SYMBOL), names_to = "Sample", values_to = "log2CPM") %>%
  as.data.frame()

if (!"SYMBOL" %in% colnames(long_boxplot)) stop("ERROR: 'SYMBOL' column is missing after pivoting!")

long_boxplot$SYMBOL <- as.character(long_boxplot$SYMBOL)
```


## **📌 Define Individual Mapping**
```{r Ind_CX48, echo=TRUE, message=FALSE}
indv_mapping <- c("75.1" = 1, "78.1" = 2, "87.1" = 3, "17.3" = 4, "84.1" = 5, "90.1" = 6)

# Debug: Print column names before mutate
print("Columns in long_boxplot before mutate():")
print(colnames(long_boxplot))
```

## **📌Process Data with Metadata**
```{r meta_CX48, echo=TRUE, message=FALSE}
formatted_data <- long_boxplot %>%
  mutate(
    Indv_id = sub("_.*", "", Sample),  # Extract individual ID
    Indv = ifelse(Indv_id %in% names(indv_mapping), as.character(indv_mapping[Indv_id]), NA),
    Drug = ifelse(grepl("CX.5461", Sample), "CX.5461", "VEH"),
    Conc = "0.1",
    Timepoint = "48"
  ) %>%
  dplyr::select(any_of(c("ENTREZID", "SYMBOL", "Sample", "Indv", "Drug", "Conc", "Timepoint", "log2CPM"))) %>%
  as.data.frame()

# Rename SYMBOL → Gene
colnames(formatted_data) <- trimws(colnames(formatted_data))
if ("SYMBOL" %in% colnames(formatted_data)) {
  colnames(formatted_data)[colnames(formatted_data) == "SYMBOL"] <- "Gene"
} else {
  stop("ERROR: 'SYMBOL' column is missing before renaming.")
}

if (!"Gene" %in% colnames(formatted_data)) stop("ERROR: 'Gene' column is missing after renaming.")

formatted_data$Indv <- as.character(formatted_data$Indv)

# Display first few rows
head(formatted_data)
```

## **📌Generate Boxplot**
```{r box_CX48, echo=TRUE, message=FALSE, fig.width=11, fig.height=8}
ggplot(formatted_data, aes(x = Drug, y = log2CPM, fill = Drug)) +
  geom_boxplot(outlier.shape = NA) +
  scale_fill_manual(values = drug_palc) +
  facet_wrap(~ Gene, ncol = 5) +
  geom_point(aes(color = Indv), size = 2, alpha = 0.5, position = position_jitter(width = -1, height = 0)) +
  ggtitle("log2CPM CX.5461_0.1 vs Vehicle_0.1 (48hrs)") +
  labs(x = "Drugs", y = "log2CPM") +
  theme_bw() +
  theme(
    plot.title = element_text(size = rel(2), hjust = 0.5),
    axis.title = element_text(size = 15, color = "black"),
    axis.ticks = element_line(linewidth = 1.5),
    axis.line = element_line(linewidth = 1.5),
    axis.text.y = element_text(size = 10, color = "black"),
    axis.text.x = element_text(size = 10, color = "black"),
    strip.text.y = element_text(color = "white")
  )
```


## **📌 CX_0.5_3 vs VEH_0.5_3**

## **📌 Define File Paths**
```{r load_path_CX0.53, echo=TRUE,results='hide',message=FALSE}
# Define file paths
input_file <- "data/Feature_count_Matrix_Log2CPM_filtered.csv"
toptable_file <- "data/DEGs/Toptable_CX_0.5_3.csv"
output_boxplot <- "data/Boxplot_CX_0.5_3.csv"
output_log2CPM <- "data/log2CPM_filtered.csv"
```


## **📌 Read and Clean Dataset**
```{r load_data_CX0.53, echo=TRUE,results='hide',message=FALSE}
# Read Log2CPM Matrix
boxplot1 <- read.csv(input_file) %>%
  as.data.frame()
colnames(boxplot1) <- trimws(gsub("^X", "", colnames(boxplot1)))  # Clean column names

# Select relevant columns
matching_columns <- grep("ENTREZID|SYMBOL|CX.5461_0.5_3|VEH_0.5_3", colnames(boxplot1), value = TRUE)

if (!"SYMBOL" %in% colnames(boxplot1)) stop("ERROR: 'SYMBOL' column is missing in boxplot1! Check dataset.")

new_boxplot1 <- boxplot1[, matching_columns]

# Save cleaned datasets
write.csv(boxplot1, output_log2CPM, row.names = FALSE)
write.csv(new_boxplot1, output_boxplot, row.names = FALSE)

# Display first few rows
head(new_boxplot1)
```

## **📌 Define Color Palettes**
```{r load_color_CX0.53, echo=TRUE,message=FALSE}
drug_palc <- c("CX.5461" = "#8B006D", "VEH" = "#F1B72B")
```


## **📌 Read and Process Toptable Data**
```{r Read_CX0.53, echo=TRUE,message=FALSE}
toptable <- read.csv(toptable_file)
colnames(toptable)[1] <- "Entrezid"

filtered_toptable <- toptable %>%
  filter(adj.P.Val < 0.05) %>%
  arrange(adj.P.Val) %>%
  slice_head(n = 10)

# Display first few rows of the filtered Toptable
head(filtered_toptable)
```

## **📌 Filter Boxplot Data for Selected Genes**
```{r Filter_0.53, echo=TRUE,message=FALSE}
filtered_boxplot <- new_boxplot1 %>%
  filter(ENTREZID %in% filtered_toptable$Entrezid)

if (!"SYMBOL" %in% colnames(filtered_boxplot)) stop("ERROR: 'SYMBOL' column is missing in filtered_boxplot!")

# Pivot data to long format
long_boxplot <- filtered_boxplot %>%
  pivot_longer(cols = -c(ENTREZID, SYMBOL), names_to = "Sample", values_to = "log2CPM") %>%
  as.data.frame()

if (!"SYMBOL" %in% colnames(long_boxplot)) stop("ERROR: 'SYMBOL' column is missing after pivoting!")

long_boxplot$SYMBOL <- as.character(long_boxplot$SYMBOL)
```


## **📌 Define Individual Mapping**
```{r Ind_CX0.53, echo=TRUE, message=FALSE}
indv_mapping <- c("75.1" = 1, "78.1" = 2, "87.1" = 3, "17.3" = 4, "84.1" = 5, "90.1" = 6)

# Debug: Print column names before mutate
print("Columns in long_boxplot before mutate():")
print(colnames(long_boxplot))
```

## **📌Process Data with Metadata**
```{r meta_CX0.53, echo=TRUE, message=FALSE}
formatted_data <- long_boxplot %>%
  mutate(
    Indv_id = sub("_.*", "", Sample),  # Extract individual ID
    Indv = ifelse(Indv_id %in% names(indv_mapping), as.character(indv_mapping[Indv_id]), NA),
    Drug = ifelse(grepl("CX.5461", Sample), "CX.5461", "VEH"),
    Conc = "0.5",
    Timepoint = "3"
  ) %>%
  dplyr::select(any_of(c("ENTREZID", "SYMBOL", "Sample", "Indv", "Drug", "Conc", "Timepoint", "log2CPM"))) %>%
  as.data.frame()

# Rename SYMBOL → Gene
colnames(formatted_data) <- trimws(colnames(formatted_data))
if ("SYMBOL" %in% colnames(formatted_data)) {
  colnames(formatted_data)[colnames(formatted_data) == "SYMBOL"] <- "Gene"
} else {
  stop("ERROR: 'SYMBOL' column is missing before renaming.")
}

if (!"Gene" %in% colnames(formatted_data)) stop("ERROR: 'Gene' column is missing after renaming.")

formatted_data$Indv <- as.character(formatted_data$Indv)

# Display first few rows
head(formatted_data)
```

## **📌Generate Boxplot**
```{r box_CX0.53, echo=TRUE, message=FALSE, fig.width=11, fig.height=8}
ggplot(formatted_data, aes(x = Drug, y = log2CPM, fill = Drug)) +
  geom_boxplot(outlier.shape = NA) +
  scale_fill_manual(values = drug_palc) +
  facet_wrap(~ Gene, ncol = 5) +
  geom_point(aes(color = Indv), size = 2, alpha = 0.5, position = position_jitter(width = -1, height = 0)) +
  ggtitle("log2CPM CX.5461_0.5 vs Vehicle_0.5 (3hrs)") +
  labs(x = "Drugs", y = "log2CPM") +
  theme_bw() +
  theme(
    plot.title = element_text(size = rel(2), hjust = 0.5),
    axis.title = element_text(size = 15, color = "black"),
    axis.ticks = element_line(linewidth = 1.5),
    axis.line = element_line(linewidth = 1.5),
    axis.text.y = element_text(size = 10, color = "black"),
    axis.text.x = element_text(size = 10, color = "black"),
    strip.text.y = element_text(color = "white")
  )
```


## **📌 CX_0.5_24 vs VEH_0.5_24**

## **📌 Define File Paths**
```{r load_path_CX0.524, echo=TRUE,results='hide',message=FALSE}
# Define file paths
input_file <- "data/Feature_count_Matrix_Log2CPM_filtered.csv"
toptable_file <- "data/DEGs/Toptable_CX_0.5_24.csv"
output_boxplot <- "data/Boxplot_CX_0.5_24.csv"
output_log2CPM <- "data/log2CPM_filtered.csv"
```


## **📌 Read and Clean Dataset**
```{r load_data_CX0.524, echo=TRUE,results='hide',message=FALSE}
# Read Log2CPM Matrix
boxplot1 <- read.csv(input_file) %>%
  as.data.frame()
colnames(boxplot1) <- trimws(gsub("^X", "", colnames(boxplot1)))  # Clean column names

# Select relevant columns
matching_columns <- grep("ENTREZID|SYMBOL|CX.5461_0.5_24|VEH_0.5_24", colnames(boxplot1), value = TRUE)

if (!"SYMBOL" %in% colnames(boxplot1)) stop("ERROR: 'SYMBOL' column is missing in boxplot1! Check dataset.")

new_boxplot1 <- boxplot1[, matching_columns]

# Save cleaned datasets
write.csv(boxplot1, output_log2CPM, row.names = FALSE)
write.csv(new_boxplot1, output_boxplot, row.names = FALSE)

# Display first few rows
head(new_boxplot1)
```

## **📌 Define Color Palettes**
```{r load_color_CX0.524, echo=TRUE,message=FALSE}
drug_palc <- c("CX.5461" = "#8B006D", "VEH" = "#F1B72B")
```


## **📌 Read and Process Toptable Data**
```{r Read_CX0.524, echo=TRUE,message=FALSE}
toptable <- read.csv(toptable_file)
colnames(toptable)[1] <- "Entrezid"

filtered_toptable <- toptable %>%
  filter(adj.P.Val < 0.05) %>%
  arrange(adj.P.Val) %>%
  slice_head(n = 10)

# Display first few rows of the filtered Toptable
head(filtered_toptable)
```

## **📌 Filter Boxplot Data for Selected Genes**
```{r Filter_0.524, echo=TRUE,message=FALSE}
filtered_boxplot <- new_boxplot1 %>%
  filter(ENTREZID %in% filtered_toptable$Entrezid)

if (!"SYMBOL" %in% colnames(filtered_boxplot)) stop("ERROR: 'SYMBOL' column is missing in filtered_boxplot!")

# Pivot data to long format
long_boxplot <- filtered_boxplot %>%
  pivot_longer(cols = -c(ENTREZID, SYMBOL), names_to = "Sample", values_to = "log2CPM") %>%
  as.data.frame()

if (!"SYMBOL" %in% colnames(long_boxplot)) stop("ERROR: 'SYMBOL' column is missing after pivoting!")

long_boxplot$SYMBOL <- as.character(long_boxplot$SYMBOL)
```


## **📌 Define Individual Mapping**
```{r Ind_CX0.524, echo=TRUE, message=FALSE}
indv_mapping <- c("75.1" = 1, "78.1" = 2, "87.1" = 3, "17.3" = 4, "84.1" = 5, "90.1" = 6)

# Debug: Print column names before mutate
print("Columns in long_boxplot before mutate():")
print(colnames(long_boxplot))
```

## **📌Process Data with Metadata**
```{r meta_CX0.524, echo=TRUE, message=FALSE}
formatted_data <- long_boxplot %>%
  mutate(
    Indv_id = sub("_.*", "", Sample),  # Extract individual ID
    Indv = ifelse(Indv_id %in% names(indv_mapping), as.character(indv_mapping[Indv_id]), NA),
    Drug = ifelse(grepl("CX.5461", Sample), "CX.5461", "VEH"),
    Conc = "0.5",
    Timepoint = "24"
  ) %>%
  dplyr::select(any_of(c("ENTREZID", "SYMBOL", "Sample", "Indv", "Drug", "Conc", "Timepoint", "log2CPM"))) %>%
  as.data.frame()

# Rename SYMBOL → Gene
colnames(formatted_data) <- trimws(colnames(formatted_data))
if ("SYMBOL" %in% colnames(formatted_data)) {
  colnames(formatted_data)[colnames(formatted_data) == "SYMBOL"] <- "Gene"
} else {
  stop("ERROR: 'SYMBOL' column is missing before renaming.")
}

if (!"Gene" %in% colnames(formatted_data)) stop("ERROR: 'Gene' column is missing after renaming.")

formatted_data$Indv <- as.character(formatted_data$Indv)

# Display first few rows
head(formatted_data)
```

## **📌Generate Boxplot**
```{r box_CX0.524, echo=TRUE, message=FALSE, fig.width=11, fig.height=8}
ggplot(formatted_data, aes(x = Drug, y = log2CPM, fill = Drug)) +
  geom_boxplot(outlier.shape = NA) +
  scale_fill_manual(values = drug_palc) +
  facet_wrap(~ Gene, ncol = 5) +
  geom_point(aes(color = Indv), size = 2, alpha = 0.5, position = position_jitter(width = -1, height = 0)) +
  ggtitle("log2CPM CX.5461_0.5 vs Vehicle_0.5 (24hrs)") +
  labs(x = "Drugs", y = "log2CPM") +
  theme_bw() +
  theme(
    plot.title = element_text(size = rel(2), hjust = 0.5),
    axis.title = element_text(size = 15, color = "black"),
    axis.ticks = element_line(linewidth = 1.5),
    axis.line = element_line(linewidth = 1.5),
    axis.text.y = element_text(size = 10, color = "black"),
    axis.text.x = element_text(size = 10, color = "black"),
    strip.text.y = element_text(color = "white")
  )
```


## **📌 CX_0.5_48 vs VEH_0.5_48**

## **📌 Define File Paths**
```{r load_path_CX0.548, echo=TRUE,results='hide',message=FALSE}
# Define file paths
input_file <- "data/Feature_count_Matrix_Log2CPM_filtered.csv"
toptable_file <- "data/DEGs/Toptable_CX_0.5_48.csv"
output_boxplot <- "data/Boxplot_CX_0.5_48.csv"
output_log2CPM <- "data/log2CPM_filtered.csv"
```


## **📌 Read and Clean Dataset**
```{r load_data_CX0.548, echo=TRUE,results='hide',message=FALSE}
# Read Log2CPM Matrix
boxplot1 <- read.csv(input_file) %>%
  as.data.frame()
colnames(boxplot1) <- trimws(gsub("^X", "", colnames(boxplot1)))  # Clean column names

# Select relevant columns
matching_columns <- grep("ENTREZID|SYMBOL|CX.5461_0.5_48|VEH_0.5_48", colnames(boxplot1), value = TRUE)

if (!"SYMBOL" %in% colnames(boxplot1)) stop("ERROR: 'SYMBOL' column is missing in boxplot1! Check dataset.")

new_boxplot1 <- boxplot1[, matching_columns]

# Save cleaned datasets
write.csv(boxplot1, output_log2CPM, row.names = FALSE)
write.csv(new_boxplot1, output_boxplot, row.names = FALSE)

# Display first few rows
head(new_boxplot1)
```

## **📌 Define Color Palettes**
```{r load_color_CX0.548, echo=TRUE,message=FALSE}
drug_palc <- c("CX.5461" = "#8B006D", "VEH" = "#F1B72B")
```


## **📌 Read and Process Toptable Data**
```{r Read_CX0.548, echo=TRUE,message=FALSE}
toptable <- read.csv(toptable_file)
colnames(toptable)[1] <- "Entrezid"

filtered_toptable <- toptable %>%
  filter(adj.P.Val < 0.05) %>%
  arrange(adj.P.Val) %>%
  slice_head(n = 10)

# Display first few rows of the filtered Toptable
head(filtered_toptable)
```

## **📌 Filter Boxplot Data for Selected Genes**
```{r Filter_0.548, echo=TRUE,message=FALSE}
filtered_boxplot <- new_boxplot1 %>%
  filter(ENTREZID %in% filtered_toptable$Entrezid)

if (!"SYMBOL" %in% colnames(filtered_boxplot)) stop("ERROR: 'SYMBOL' column is missing in filtered_boxplot!")

# Pivot data to long format
long_boxplot <- filtered_boxplot %>%
  pivot_longer(cols = -c(ENTREZID, SYMBOL), names_to = "Sample", values_to = "log2CPM") %>%
  as.data.frame()

if (!"SYMBOL" %in% colnames(long_boxplot)) stop("ERROR: 'SYMBOL' column is missing after pivoting!")

long_boxplot$SYMBOL <- as.character(long_boxplot$SYMBOL)
```


## **📌 Define Individual Mapping**
```{r Ind_CX0.548, echo=TRUE, message=FALSE}
indv_mapping <- c("75.1" = 1, "78.1" = 2, "87.1" = 3, "17.3" = 4, "84.1" = 5, "90.1" = 6)

# Debug: Print column names before mutate
print("Columns in long_boxplot before mutate():")
print(colnames(long_boxplot))
```

## **📌Process Data with Metadata**
```{r meta_CX0.548, echo=TRUE, message=FALSE}
formatted_data <- long_boxplot %>%
  mutate(
    Indv_id = sub("_.*", "", Sample),  # Extract individual ID
    Indv = ifelse(Indv_id %in% names(indv_mapping), as.character(indv_mapping[Indv_id]), NA),
    Drug = ifelse(grepl("CX.5461", Sample), "CX.5461", "VEH"),
    Conc = "0.5",
    Timepoint = "48"
  ) %>%
  dplyr::select(any_of(c("ENTREZID", "SYMBOL", "Sample", "Indv", "Drug", "Conc", "Timepoint", "log2CPM"))) %>%
  as.data.frame()

# Rename SYMBOL → Gene
colnames(formatted_data) <- trimws(colnames(formatted_data))
if ("SYMBOL" %in% colnames(formatted_data)) {
  colnames(formatted_data)[colnames(formatted_data) == "SYMBOL"] <- "Gene"
} else {
  stop("ERROR: 'SYMBOL' column is missing before renaming.")
}

if (!"Gene" %in% colnames(formatted_data)) stop("ERROR: 'Gene' column is missing after renaming.")

formatted_data$Indv <- as.character(formatted_data$Indv)

# Display first few rows
head(formatted_data)
```

## **📌Generate Boxplot**
```{r box_CX0.548, echo=TRUE, message=FALSE, fig.width=11, fig.height=8}
ggplot(formatted_data, aes(x = Drug, y = log2CPM, fill = Drug)) +
  geom_boxplot(outlier.shape = NA) +
  scale_fill_manual(values = drug_palc) +
  facet_wrap(~ Gene, ncol = 5) +
  geom_point(aes(color = Indv), size = 2, alpha = 0.5, position = position_jitter(width = -1, height = 0)) +
  ggtitle("log2CPM CX.5461_0.5 vs Vehicle_0.5 (48hrs)") +
  labs(x = "Drugs", y = "log2CPM") +
  theme_bw() +
  theme(
    plot.title = element_text(size = rel(2), hjust = 0.5),
    axis.title = element_text(size = 15, color = "black"),
    axis.ticks = element_line(linewidth = 1.5),
    axis.line = element_line(linewidth = 1.5),
    axis.text.y = element_text(size = 10, color = "black"),
    axis.text.x = element_text(size = 10, color = "black"),
    strip.text.y = element_text(color = "white")
  )
```



## **📌 DOX_0.1_3 vs VEH_0.1_3**

## **📌 Define File Paths**
```{r load_path_DOX0.13, echo=TRUE,results='hide',message=FALSE}
# Define file paths
input_file <- "data/Feature_count_Matrix_Log2CPM_filtered.csv"
toptable_file <- "data/DEGs/Toptable_DOX_0.1_3.csv"
output_boxplot <- "data/Boxplot_DOX_0.1_3.csv"
output_log2CPM <- "data/log2CPM_filtered.csv"
```


## **📌 Read and Clean Dataset**
```{r load_data_DOX0.13, echo=TRUE,results='hide',message=FALSE}
# Read Log2CPM Matrix
boxplot1 <- read.csv(input_file) %>%
  as.data.frame()
colnames(boxplot1) <- trimws(gsub("^X", "", colnames(boxplot1)))  # Clean column names

# Select relevant columns
matching_columns <- grep("ENTREZID|SYMBOL|DOX_0.1_3|VEH_0.1_3", colnames(boxplot1), value = TRUE)

if (!"SYMBOL" %in% colnames(boxplot1)) stop("ERROR: 'SYMBOL' column is missing in boxplot1! Check dataset.")

new_boxplot1 <- boxplot1[, matching_columns]

# Save cleaned datasets
write.csv(boxplot1, output_log2CPM, row.names = FALSE)
write.csv(new_boxplot1, output_boxplot, row.names = FALSE)

# Display first few rows
head(new_boxplot1)
```

## **📌 Define Color Palettes**
```{r load_color_DOX0.13, echo=TRUE,message=FALSE}
drug_palc <- c("DOX" = "#8B006D", "VEH" = "#F1B72B")
```


## **📌 Read and Process Toptable Data**
```{r Read_DOX0.13, echo=TRUE,message=FALSE}
toptable <- read.csv(toptable_file)
colnames(toptable)[1] <- "Entrezid"

filtered_toptable <- toptable %>%
  filter(adj.P.Val < 0.05) %>%
  arrange(adj.P.Val) %>%
  slice_head(n = 10)

# Display first few rows of the filtered Toptable
head(filtered_toptable)
```

## **📌 Filter Boxplot Data for Selected Genes**
```{r Filter_DOX0.13, echo=TRUE,message=FALSE}
filtered_boxplot <- new_boxplot1 %>%
  filter(ENTREZID %in% filtered_toptable$Entrezid)

if (!"SYMBOL" %in% colnames(filtered_boxplot)) stop("ERROR: 'SYMBOL' column is missing in filtered_boxplot!")

# Pivot data to long format
long_boxplot <- filtered_boxplot %>%
  pivot_longer(cols = -c(ENTREZID, SYMBOL), names_to = "Sample", values_to = "log2CPM") %>%
  as.data.frame()

if (!"SYMBOL" %in% colnames(long_boxplot)) stop("ERROR: 'SYMBOL' column is missing after pivoting!")

long_boxplot$SYMBOL <- as.character(long_boxplot$SYMBOL)
```


## **📌 Define Individual Mapping**
```{r Ind_DOX0.13, echo=TRUE, message=FALSE}
indv_mapping <- c("75.1" = 1, "78.1" = 2, "87.1" = 3, "17.3" = 4, "84.1" = 5, "90.1" = 6)

# Debug: Print column names before mutate
print("Columns in long_boxplot before mutate():")
print(colnames(long_boxplot))
```

## **📌Process Data with Metadata**
```{r meta_DOX0.13, echo=TRUE, message=FALSE}
formatted_data <- long_boxplot %>%
  mutate(
    Indv_id = sub("_.*", "", Sample),  # Extract individual ID
    Indv = ifelse(Indv_id %in% names(indv_mapping), as.character(indv_mapping[Indv_id]), NA),
    Drug = ifelse(grepl("DOX", Sample), "DOX", "VEH"),
    Conc = "0.1",
    Timepoint = "3"
  ) %>%
  dplyr::select(any_of(c("ENTREZID", "SYMBOL", "Sample", "Indv", "Drug", "Conc", "Timepoint", "log2CPM"))) %>%
  as.data.frame()

# Rename SYMBOL → Gene
colnames(formatted_data) <- trimws(colnames(formatted_data))
if ("SYMBOL" %in% colnames(formatted_data)) {
  colnames(formatted_data)[colnames(formatted_data) == "SYMBOL"] <- "Gene"
} else {
  stop("ERROR: 'SYMBOL' column is missing before renaming.")
}

if (!"Gene" %in% colnames(formatted_data)) stop("ERROR: 'Gene' column is missing after renaming.")

formatted_data$Indv <- as.character(formatted_data$Indv)

# Display first few rows
head(formatted_data)
```

## **📌Generate Boxplot**
```{r box_DOX0.13, echo=TRUE, message=FALSE, fig.width=11, fig.height=8}
ggplot(formatted_data, aes(x = Drug, y = log2CPM, fill = Drug)) +
  geom_boxplot(outlier.shape = NA) +
  scale_fill_manual(values = drug_palc) +
  facet_wrap(~ Gene, ncol = 5) +
  geom_point(aes(color = Indv), size = 2, alpha = 0.5, position = position_jitter(width = -1, height = 0)) +
  ggtitle("log2CPM DOX_0.1 vs Vehicle_0.1 (3hrs)") +
  labs(x = "Drugs", y = "log2CPM") +
  theme_bw() +
  theme(
    plot.title = element_text(size = rel(2), hjust = 0.5),
    axis.title = element_text(size = 15, color = "black"),
    axis.ticks = element_line(linewidth = 1.5),
    axis.line = element_line(linewidth = 1.5),
    axis.text.y = element_text(size = 10, color = "black"),
    axis.text.x = element_text(size = 10, color = "black"),
    strip.text.y = element_text(color = "white")
  )
```



## **📌 DOX_0.1_24 vs VEH_0.1_24**

## **📌 Define File Paths**
```{r load_path_DOX0.124, echo=TRUE,results='hide',message=FALSE}
# Define file paths
input_file <- "data/Feature_count_Matrix_Log2CPM_filtered.csv"
toptable_file <- "data/DEGs/Toptable_DOX_0.1_24.csv"
output_boxplot <- "data/Boxplot_DOX_0.1_24.csv"
output_log2CPM <- "data/log2CPM_filtered.csv"
```


## **📌 Read and Clean Dataset**
```{r load_data_DOX0.124, echo=TRUE,results='hide',message=FALSE}
# Read Log2CPM Matrix
boxplot1 <- read.csv(input_file) %>%
  as.data.frame()
colnames(boxplot1) <- trimws(gsub("^X", "", colnames(boxplot1)))  # Clean column names

# Select relevant columns
matching_columns <- grep("ENTREZID|SYMBOL|DOX_0.1_24|VEH_0.1_24", colnames(boxplot1), value = TRUE)

if (!"SYMBOL" %in% colnames(boxplot1)) stop("ERROR: 'SYMBOL' column is missing in boxplot1! Check dataset.")

new_boxplot1 <- boxplot1[, matching_columns]

# Save cleaned datasets
write.csv(boxplot1, output_log2CPM, row.names = FALSE)
write.csv(new_boxplot1, output_boxplot, row.names = FALSE)

# Display first few rows
head(new_boxplot1)
```

## **📌 Define Color Palettes**
```{r load_color_DOX0.124, echo=TRUE,message=FALSE}
drug_palc <- c("DOX" = "#8B006D", "VEH" = "#F1B72B")
```


## **📌 Read and Process Toptable Data**
```{r Read_DOX0.124, echo=TRUE,message=FALSE}
toptable <- read.csv(toptable_file)
colnames(toptable)[1] <- "Entrezid"

filtered_toptable <- toptable %>%
  filter(adj.P.Val < 0.05) %>%
  arrange(adj.P.Val) %>%
  slice_head(n = 10)

# Display first few rows of the filtered Toptable
head(filtered_toptable)
```

## **📌 Filter Boxplot Data for Selected Genes**
```{r Filter_DOX0.124, echo=TRUE,message=FALSE}
filtered_boxplot <- new_boxplot1 %>%
  filter(ENTREZID %in% filtered_toptable$Entrezid)

if (!"SYMBOL" %in% colnames(filtered_boxplot)) stop("ERROR: 'SYMBOL' column is missing in filtered_boxplot!")

# Pivot data to long format
long_boxplot <- filtered_boxplot %>%
  pivot_longer(cols = -c(ENTREZID, SYMBOL), names_to = "Sample", values_to = "log2CPM") %>%
  as.data.frame()

if (!"SYMBOL" %in% colnames(long_boxplot)) stop("ERROR: 'SYMBOL' column is missing after pivoting!")

long_boxplot$SYMBOL <- as.character(long_boxplot$SYMBOL)
```


## **📌 Define Individual Mapping**
```{r Ind_DOX0.124, echo=TRUE, message=FALSE}
indv_mapping <- c("75.1" = 1, "78.1" = 2, "87.1" = 3, "17.3" = 4, "84.1" = 5, "90.1" = 6)

# Debug: Print column names before mutate
print("Columns in long_boxplot before mutate():")
print(colnames(long_boxplot))
```

## **📌Process Data with Metadata**
```{r meta_DOX0.124, echo=TRUE, message=FALSE}
formatted_data <- long_boxplot %>%
  mutate(
    Indv_id = sub("_.*", "", Sample),  # Extract individual ID
    Indv = ifelse(Indv_id %in% names(indv_mapping), as.character(indv_mapping[Indv_id]), NA),
    Drug = ifelse(grepl("DOX", Sample), "DOX", "VEH"),
    Conc = "0.1",
    Timepoint = "24"
  ) %>%
  dplyr::select(any_of(c("ENTREZID", "SYMBOL", "Sample", "Indv", "Drug", "Conc", "Timepoint", "log2CPM"))) %>%
  as.data.frame()

# Rename SYMBOL → Gene
colnames(formatted_data) <- trimws(colnames(formatted_data))
if ("SYMBOL" %in% colnames(formatted_data)) {
  colnames(formatted_data)[colnames(formatted_data) == "SYMBOL"] <- "Gene"
} else {
  stop("ERROR: 'SYMBOL' column is missing before renaming.")
}

if (!"Gene" %in% colnames(formatted_data)) stop("ERROR: 'Gene' column is missing after renaming.")

formatted_data$Indv <- as.character(formatted_data$Indv)

# Display first few rows
head(formatted_data)
```

## **📌Generate Boxplot**
```{r box_DOX0.124, echo=TRUE, message=FALSE, fig.width=11, fig.height=8}
ggplot(formatted_data, aes(x = Drug, y = log2CPM, fill = Drug)) +
  geom_boxplot(outlier.shape = NA) +
  scale_fill_manual(values = drug_palc) +
  facet_wrap(~ Gene, ncol = 5) +
  geom_point(aes(color = Indv), size = 2, alpha = 0.5, position = position_jitter(width = -1, height = 0)) +
  ggtitle("log2CPM DOX_0.1 vs Vehicle_0.1 (24hrs)") +
  labs(x = "Drugs", y = "log2CPM") +
  theme_bw() +
  theme(
    plot.title = element_text(size = rel(2), hjust = 0.5),
    axis.title = element_text(size = 15, color = "black"),
    axis.ticks = element_line(linewidth = 1.5),
    axis.line = element_line(linewidth = 1.5),
    axis.text.y = element_text(size = 10, color = "black"),
    axis.text.x = element_text(size = 10, color = "black"),
    strip.text.y = element_text(color = "white")
  )
```


## **📌 DOX_0.1_48 vs VEH_0.1_48**

## **📌 Define File Paths**
```{r load_path_DOX0.148, echo=TRUE,results='hide',message=FALSE}
# Define file paths
input_file <- "data/Feature_count_Matrix_Log2CPM_filtered.csv"
toptable_file <- "data/DEGs/Toptable_DOX_0.1_48.csv"
output_boxplot <- "data/Boxplot_DOX_0.1_48.csv"
output_log2CPM <- "data/log2CPM_filtered.csv"
```


## **📌 Read and Clean Dataset**
```{r load_data_DOX0.148, echo=TRUE,results='hide',message=FALSE}
# Read Log2CPM Matrix
boxplot1 <- read.csv(input_file) %>%
  as.data.frame()
colnames(boxplot1) <- trimws(gsub("^X", "", colnames(boxplot1)))  # Clean column names

# Select relevant columns
matching_columns <- grep("ENTREZID|SYMBOL|DOX_0.1_48|VEH_0.1_48", colnames(boxplot1), value = TRUE)

if (!"SYMBOL" %in% colnames(boxplot1)) stop("ERROR: 'SYMBOL' column is missing in boxplot1! Check dataset.")

new_boxplot1 <- boxplot1[, matching_columns]

# Save cleaned datasets
write.csv(boxplot1, output_log2CPM, row.names = FALSE)
write.csv(new_boxplot1, output_boxplot, row.names = FALSE)

# Display first few rows
head(new_boxplot1)
```

## **📌 Define Color Palettes**
```{r load_color_DOX0.148, echo=TRUE,message=FALSE}
drug_palc <- c("DOX" = "#8B006D", "VEH" = "#F1B72B")
```


## **📌 Read and Process Toptable Data**
```{r Read_DOX0.148, echo=TRUE,message=FALSE}
toptable <- read.csv(toptable_file)
colnames(toptable)[1] <- "Entrezid"

filtered_toptable <- toptable %>%
  filter(adj.P.Val < 0.05) %>%
  arrange(adj.P.Val) %>%
  slice_head(n = 10)

# Display first few rows of the filtered Toptable
head(filtered_toptable)
```

## **📌 Filter Boxplot Data for Selected Genes**
```{r Filter_DOX0.148, echo=TRUE,message=FALSE}
filtered_boxplot <- new_boxplot1 %>%
  filter(ENTREZID %in% filtered_toptable$Entrezid)

if (!"SYMBOL" %in% colnames(filtered_boxplot)) stop("ERROR: 'SYMBOL' column is missing in filtered_boxplot!")

# Pivot data to long format
long_boxplot <- filtered_boxplot %>%
  pivot_longer(cols = -c(ENTREZID, SYMBOL), names_to = "Sample", values_to = "log2CPM") %>%
  as.data.frame()

if (!"SYMBOL" %in% colnames(long_boxplot)) stop("ERROR: 'SYMBOL' column is missing after pivoting!")

long_boxplot$SYMBOL <- as.character(long_boxplot$SYMBOL)
```


## **📌 Define Individual Mapping**
```{r Ind_DOX0.148, echo=TRUE, message=FALSE}
indv_mapping <- c("75.1" = 1, "78.1" = 2, "87.1" = 3, "17.3" = 4, "84.1" = 5, "90.1" = 6)

# Debug: Print column names before mutate
print("Columns in long_boxplot before mutate():")
print(colnames(long_boxplot))
```

## **📌Process Data with Metadata**
```{r meta_DOX0.148, echo=TRUE, message=FALSE}
formatted_data <- long_boxplot %>%
  mutate(
    Indv_id = sub("_.*", "", Sample),  # Extract individual ID
    Indv = ifelse(Indv_id %in% names(indv_mapping), as.character(indv_mapping[Indv_id]), NA),
    Drug = ifelse(grepl("DOX", Sample), "DOX", "VEH"),
    Conc = "0.1",
    Timepoint = "48"
  ) %>%
  dplyr::select(any_of(c("ENTREZID", "SYMBOL", "Sample", "Indv", "Drug", "Conc", "Timepoint", "log2CPM"))) %>%
  as.data.frame()

# Rename SYMBOL → Gene
colnames(formatted_data) <- trimws(colnames(formatted_data))
if ("SYMBOL" %in% colnames(formatted_data)) {
  colnames(formatted_data)[colnames(formatted_data) == "SYMBOL"] <- "Gene"
} else {
  stop("ERROR: 'SYMBOL' column is missing before renaming.")
}

if (!"Gene" %in% colnames(formatted_data)) stop("ERROR: 'Gene' column is missing after renaming.")

formatted_data$Indv <- as.character(formatted_data$Indv)

# Display first few rows
head(formatted_data)
```

## **📌Generate Boxplot**
```{r box_DOX0.148, echo=TRUE, message=FALSE, fig.width=11, fig.height=8}
ggplot(formatted_data, aes(x = Drug, y = log2CPM, fill = Drug)) +
  geom_boxplot(outlier.shape = NA) +
  scale_fill_manual(values = drug_palc) +
  facet_wrap(~ Gene, ncol = 5) +
  geom_point(aes(color = Indv), size = 2, alpha = 0.5, position = position_jitter(width = -1, height = 0)) +
  ggtitle("log2CPM DOX_0.1 vs Vehicle_0.1 (48hrs)") +
  labs(x = "Drugs", y = "log2CPM") +
  theme_bw() +
  theme(
    plot.title = element_text(size = rel(2), hjust = 0.5),
    axis.title = element_text(size = 15, color = "black"),
    axis.ticks = element_line(linewidth = 1.5),
    axis.line = element_line(linewidth = 1.5),
    axis.text.y = element_text(size = 10, color = "black"),
    axis.text.x = element_text(size = 10, color = "black"),
    strip.text.y = element_text(color = "white")
  )
```


## **📌 DOX_0.5_3 vs VEH_0.5_3**

## **📌 Define File Paths**
```{r load_path_DOX0.53, echo=TRUE,results='hide',message=FALSE}
# Define file paths
input_file <- "data/Feature_count_Matrix_Log2CPM_filtered.csv"
toptable_file <- "data/DEGs/Toptable_DOX_0.5_3.csv"
output_boxplot <- "data/Boxplot_DOX_0.5_3.csv"
output_log2CPM <- "data/log2CPM_filtered.csv"
```


## **📌 Read and Clean Dataset**
```{r load_data_DOX0.53, echo=TRUE,results='hide',message=FALSE}
# Read Log2CPM Matrix
boxplot1 <- read.csv(input_file) %>%
  as.data.frame()
colnames(boxplot1) <- trimws(gsub("^X", "", colnames(boxplot1)))  # Clean column names

# Select relevant columns
matching_columns <- grep("ENTREZID|SYMBOL|DOX_0.5_3|VEH_0.5_3", colnames(boxplot1), value = TRUE)

if (!"SYMBOL" %in% colnames(boxplot1)) stop("ERROR: 'SYMBOL' column is missing in boxplot1! Check dataset.")

new_boxplot1 <- boxplot1[, matching_columns]

# Save cleaned datasets
write.csv(boxplot1, output_log2CPM, row.names = FALSE)
write.csv(new_boxplot1, output_boxplot, row.names = FALSE)

# Display first few rows
head(new_boxplot1)
```

## **📌 Define Color Palettes**
```{r load_color_DOX0.53, echo=TRUE,message=FALSE}
drug_palc <- c("DOX" = "#8B006D", "VEH" = "#F1B72B")
```


## **📌 Read and Process Toptable Data**
```{r Read_DOX0.53, echo=TRUE,message=FALSE}
toptable <- read.csv(toptable_file)
colnames(toptable)[1] <- "Entrezid"

filtered_toptable <- toptable %>%
  filter(adj.P.Val < 0.05) %>%
  arrange(adj.P.Val) %>%
  slice_head(n = 10)

# Display first few rows of the filtered Toptable
head(filtered_toptable)
```

## **📌 Filter Boxplot Data for Selected Genes**
```{r Filter_DOX0.53, echo=TRUE,message=FALSE}
filtered_boxplot <- new_boxplot1 %>%
  filter(ENTREZID %in% filtered_toptable$Entrezid)

if (!"SYMBOL" %in% colnames(filtered_boxplot)) stop("ERROR: 'SYMBOL' column is missing in filtered_boxplot!")

# Pivot data to long format
long_boxplot <- filtered_boxplot %>%
  pivot_longer(cols = -c(ENTREZID, SYMBOL), names_to = "Sample", values_to = "log2CPM") %>%
  as.data.frame()

if (!"SYMBOL" %in% colnames(long_boxplot)) stop("ERROR: 'SYMBOL' column is missing after pivoting!")

long_boxplot$SYMBOL <- as.character(long_boxplot$SYMBOL)
```


## **📌 Define Individual Mapping**
```{r Ind_DOX0.53, echo=TRUE, message=FALSE}
indv_mapping <- c("75.1" = 1, "78.1" = 2, "87.1" = 3, "17.3" = 4, "84.1" = 5, "90.1" = 6)

# Debug: Print column names before mutate
print("Columns in long_boxplot before mutate():")
print(colnames(long_boxplot))
```

## **📌Process Data with Metadata**
```{r meta_DOX0.53, echo=TRUE, message=FALSE}
formatted_data <- long_boxplot %>%
  mutate(
    Indv_id = sub("_.*", "", Sample),  # Extract individual ID
    Indv = ifelse(Indv_id %in% names(indv_mapping), as.character(indv_mapping[Indv_id]), NA),
    Drug = ifelse(grepl("DOX", Sample), "DOX", "VEH"),
    Conc = "0.5",
    Timepoint = "3"
  ) %>%
  dplyr::select(any_of(c("ENTREZID", "SYMBOL", "Sample", "Indv", "Drug", "Conc", "Timepoint", "log2CPM"))) %>%
  as.data.frame()

# Rename SYMBOL → Gene
colnames(formatted_data) <- trimws(colnames(formatted_data))
if ("SYMBOL" %in% colnames(formatted_data)) {
  colnames(formatted_data)[colnames(formatted_data) == "SYMBOL"] <- "Gene"
} else {
  stop("ERROR: 'SYMBOL' column is missing before renaming.")
}

if (!"Gene" %in% colnames(formatted_data)) stop("ERROR: 'Gene' column is missing after renaming.")

formatted_data$Indv <- as.character(formatted_data$Indv)

# Display first few rows
head(formatted_data)
```

## **📌Generate Boxplot**
```{r box_DOX0.53, echo=TRUE, message=FALSE, fig.width=11, fig.height=8}
ggplot(formatted_data, aes(x = Drug, y = log2CPM, fill = Drug)) +
  geom_boxplot(outlier.shape = NA) +
  scale_fill_manual(values = drug_palc) +
  facet_wrap(~ Gene, ncol = 5) +
  geom_point(aes(color = Indv), size = 2, alpha = 0.5, position = position_jitter(width = -1, height = 0)) +
  ggtitle("log2CPM DOX_0.5 vs Vehicle_0.5 (3hrs)") +
  labs(x = "Drugs", y = "log2CPM") +
  theme_bw() +
  theme(
    plot.title = element_text(size = rel(2), hjust = 0.5),
    axis.title = element_text(size = 15, color = "black"),
    axis.ticks = element_line(linewidth = 1.5),
    axis.line = element_line(linewidth = 1.5),
    axis.text.y = element_text(size = 10, color = "black"),
    axis.text.x = element_text(size = 10, color = "black"),
    strip.text.y = element_text(color = "white")
  )
```

## **📌 DOX_0.5_24 vs VEH_0.5_24**

## **📌 Define File Paths**
```{r load_path_DOX0.524, echo=TRUE,results='hide',message=FALSE}
# Define file paths
input_file <- "data/Feature_count_Matrix_Log2CPM_filtered.csv"
toptable_file <- "data/DEGs/Toptable_DOX_0.5_24.csv"
output_boxplot <- "data/Boxplot_DOX_0.5_24.csv"
output_log2CPM <- "data/log2CPM_filtered.csv"
```


## **📌 Read and Clean Dataset**
```{r load_data_DOX0.524, echo=TRUE,results='hide',message=FALSE}
# Read Log2CPM Matrix
boxplot1 <- read.csv(input_file) %>%
  as.data.frame()
colnames(boxplot1) <- trimws(gsub("^X", "", colnames(boxplot1)))  # Clean column names

# Select relevant columns
matching_columns <- grep("ENTREZID|SYMBOL|DOX_0.5_24|VEH_0.5_24", colnames(boxplot1), value = TRUE)

if (!"SYMBOL" %in% colnames(boxplot1)) stop("ERROR: 'SYMBOL' column is missing in boxplot1! Check dataset.")

new_boxplot1 <- boxplot1[, matching_columns]

# Save cleaned datasets
write.csv(boxplot1, output_log2CPM, row.names = FALSE)
write.csv(new_boxplot1, output_boxplot, row.names = FALSE)

# Display first few rows
head(new_boxplot1)
```

## **📌 Define Color Palettes**
```{r load_color_DOX0.524, echo=TRUE,message=FALSE}
drug_palc <- c("DOX" = "#8B006D", "VEH" = "#F1B72B")
```


## **📌 Read and Process Toptable Data**
```{r Read_DOX0.524, echo=TRUE,message=FALSE}
toptable <- read.csv(toptable_file)
colnames(toptable)[1] <- "Entrezid"

filtered_toptable <- toptable %>%
  filter(adj.P.Val < 0.05) %>%
  arrange(adj.P.Val) %>%
  slice_head(n = 10)

# Display first few rows of the filtered Toptable
head(filtered_toptable)
```

## **📌 Filter Boxplot Data for Selected Genes**
```{r Filter_DOX0.524, echo=TRUE,message=FALSE}
filtered_boxplot <- new_boxplot1 %>%
  filter(ENTREZID %in% filtered_toptable$Entrezid)

if (!"SYMBOL" %in% colnames(filtered_boxplot)) stop("ERROR: 'SYMBOL' column is missing in filtered_boxplot!")

# Pivot data to long format
long_boxplot <- filtered_boxplot %>%
  pivot_longer(cols = -c(ENTREZID, SYMBOL), names_to = "Sample", values_to = "log2CPM") %>%
  as.data.frame()

if (!"SYMBOL" %in% colnames(long_boxplot)) stop("ERROR: 'SYMBOL' column is missing after pivoting!")

long_boxplot$SYMBOL <- as.character(long_boxplot$SYMBOL)
```


## **📌 Define Individual Mapping**
```{r Ind_DOX0.524, echo=TRUE, message=FALSE}
indv_mapping <- c("75.1" = 1, "78.1" = 2, "87.1" = 3, "17.3" = 4, "84.1" = 5, "90.1" = 6)

# Debug: Print column names before mutate
print("Columns in long_boxplot before mutate():")
print(colnames(long_boxplot))
```

## **📌Process Data with Metadata**
```{r meta_DOX0.524, echo=TRUE, message=FALSE}
formatted_data <- long_boxplot %>%
  mutate(
    Indv_id = sub("_.*", "", Sample),  # Extract individual ID
    Indv = ifelse(Indv_id %in% names(indv_mapping), as.character(indv_mapping[Indv_id]), NA),
    Drug = ifelse(grepl("DOX", Sample), "DOX", "VEH"),
    Conc = "0.5",
    Timepoint = "24"
  ) %>%
  dplyr::select(any_of(c("ENTREZID", "SYMBOL", "Sample", "Indv", "Drug", "Conc", "Timepoint", "log2CPM"))) %>%
  as.data.frame()

# Rename SYMBOL → Gene
colnames(formatted_data) <- trimws(colnames(formatted_data))
if ("SYMBOL" %in% colnames(formatted_data)) {
  colnames(formatted_data)[colnames(formatted_data) == "SYMBOL"] <- "Gene"
} else {
  stop("ERROR: 'SYMBOL' column is missing before renaming.")
}

if (!"Gene" %in% colnames(formatted_data)) stop("ERROR: 'Gene' column is missing after renaming.")

formatted_data$Indv <- as.character(formatted_data$Indv)

# Display first few rows
head(formatted_data)
```

## **📌Generate Boxplot**
```{r box_DOX0.524, echo=TRUE, message=FALSE, fig.width=11, fig.height=8}
ggplot(formatted_data, aes(x = Drug, y = log2CPM, fill = Drug)) +
  geom_boxplot(outlier.shape = NA) +
  scale_fill_manual(values = drug_palc) +
  facet_wrap(~ Gene, ncol = 5) +
  geom_point(aes(color = Indv), size = 2, alpha = 0.5, position = position_jitter(width = -1, height = 0)) +
  ggtitle("log2CPM DOX_0.5 vs Vehicle_0.5 (24hrs)") +
  labs(x = "Drugs", y = "log2CPM") +
  theme_bw() +
  theme(
    plot.title = element_text(size = rel(2), hjust = 0.5),
    axis.title = element_text(size = 15, color = "black"),
    axis.ticks = element_line(linewidth = 1.5),
    axis.line = element_line(linewidth = 1.5),
    axis.text.y = element_text(size = 10, color = "black"),
    axis.text.x = element_text(size = 10, color = "black"),
    strip.text.y = element_text(color = "white")
  )
```


## **📌 DOX_0.5_48 vs VEH_0.5_48**

## **📌 Define File Paths**
```{r load_path_DOX0.548, echo=TRUE,results='hide',message=FALSE}
# Define file paths
input_file <- "data/Feature_count_Matrix_Log2CPM_filtered.csv"
toptable_file <- "data/DEGs/Toptable_DOX_0.5_48.csv"
output_boxplot <- "data/Boxplot_DOX_0.5_48.csv"
output_log2CPM <- "data/log2CPM_filtered.csv"
```


## **📌 Read and Clean Dataset**
```{r load_data_DOX0.548, echo=TRUE,results='hide',message=FALSE}
# Read Log2CPM Matrix
boxplot1 <- read.csv(input_file) %>%
  as.data.frame()
colnames(boxplot1) <- trimws(gsub("^X", "", colnames(boxplot1)))  # Clean column names

# Select relevant columns
matching_columns <- grep("ENTREZID|SYMBOL|DOX_0.5_48|VEH_0.5_48", colnames(boxplot1), value = TRUE)

if (!"SYMBOL" %in% colnames(boxplot1)) stop("ERROR: 'SYMBOL' column is missing in boxplot1! Check dataset.")

new_boxplot1 <- boxplot1[, matching_columns]

# Save cleaned datasets
write.csv(boxplot1, output_log2CPM, row.names = FALSE)
write.csv(new_boxplot1, output_boxplot, row.names = FALSE)

# Display first few rows
head(new_boxplot1)
```

## **📌 Define Color Palettes**
```{r load_color_DOX0.548, echo=TRUE,message=FALSE}
drug_palc <- c("DOX" = "#8B006D", "VEH" = "#F1B72B")
```


## **📌 Read and Process Toptable Data**
```{r Read_DOX0.548, echo=TRUE,message=FALSE}
toptable <- read.csv(toptable_file)
colnames(toptable)[1] <- "Entrezid"

filtered_toptable <- toptable %>%
  filter(adj.P.Val < 0.05) %>%
  arrange(adj.P.Val) %>%
  slice_head(n = 10)

# Display first few rows of the filtered Toptable
head(filtered_toptable)
```

## **📌 Filter Boxplot Data for Selected Genes**
```{r Filter_DOX0.548, echo=TRUE,message=FALSE}
filtered_boxplot <- new_boxplot1 %>%
  filter(ENTREZID %in% filtered_toptable$Entrezid)

if (!"SYMBOL" %in% colnames(filtered_boxplot)) stop("ERROR: 'SYMBOL' column is missing in filtered_boxplot!")

# Pivot data to long format
long_boxplot <- filtered_boxplot %>%
  pivot_longer(cols = -c(ENTREZID, SYMBOL), names_to = "Sample", values_to = "log2CPM") %>%
  as.data.frame()

if (!"SYMBOL" %in% colnames(long_boxplot)) stop("ERROR: 'SYMBOL' column is missing after pivoting!")

long_boxplot$SYMBOL <- as.character(long_boxplot$SYMBOL)
```


## **📌 Define Individual Mapping**
```{r Ind_DOX0.548, echo=TRUE, message=FALSE}
indv_mapping <- c("75.1" = 1, "78.1" = 2, "87.1" = 3, "17.3" = 4, "84.1" = 5, "90.1" = 6)

# Debug: Print column names before mutate
print("Columns in long_boxplot before mutate():")
print(colnames(long_boxplot))
```

## **📌Process Data with Metadata**
```{r meta_DOX0.548, echo=TRUE, message=FALSE}
formatted_data <- long_boxplot %>%
  mutate(
    Indv_id = sub("_.*", "", Sample),  # Extract individual ID
    Indv = ifelse(Indv_id %in% names(indv_mapping), as.character(indv_mapping[Indv_id]), NA),
    Drug = ifelse(grepl("DOX", Sample), "DOX", "VEH"),
    Conc = "0.5",
    Timepoint = "48"
  ) %>%
  dplyr::select(any_of(c("ENTREZID", "SYMBOL", "Sample", "Indv", "Drug", "Conc", "Timepoint", "log2CPM"))) %>%
  as.data.frame()

# Rename SYMBOL → Gene
colnames(formatted_data) <- trimws(colnames(formatted_data))
if ("SYMBOL" %in% colnames(formatted_data)) {
  colnames(formatted_data)[colnames(formatted_data) == "SYMBOL"] <- "Gene"
} else {
  stop("ERROR: 'SYMBOL' column is missing before renaming.")
}

if (!"Gene" %in% colnames(formatted_data)) stop("ERROR: 'Gene' column is missing after renaming.")

formatted_data$Indv <- as.character(formatted_data$Indv)

# Display first few rows
head(formatted_data)
```

## **📌Generate Boxplot**
```{r box_DOX0.548, echo=TRUE, message=FALSE, fig.width=11, fig.height=8}
ggplot(formatted_data, aes(x = Drug, y = log2CPM, fill = Drug)) +
  geom_boxplot(outlier.shape = NA) +
  scale_fill_manual(values = drug_palc) +
  facet_wrap(~ Gene, ncol = 5) +
  geom_point(aes(color = Indv), size = 2, alpha = 0.5, position = position_jitter(width = -1, height = 0)) +
  ggtitle("log2CPM DOX_0.5 vs Vehicle_0.5 (48hrs)") +
  labs(x = "Drugs", y = "log2CPM") +
  theme_bw() +
  theme(
    plot.title = element_text(size = rel(2), hjust = 0.5),
    axis.title = element_text(size = 15, color = "black"),
    axis.ticks = element_line(linewidth = 1.5),
    axis.line = element_line(linewidth = 1.5),
    axis.text.y = element_text(size = 10, color = "black"),
    axis.text.x = element_text(size = 10, color = "black"),
    strip.text.y = element_text(color = "white")
  )
```


