---
title: "LogFC Correlation"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## **📌LogFC Correlation Analysis**

This analysis generates correlation heatmaps of log fold change (logFC) values across different comparisons.

### **📌Load Required Libraries**
```{r load_libraries, echo=TRUE, message=FALSE}
library(ComplexHeatmap)
library(tidyverse)
library(edgeR)
library(limma)
library(data.table)
```

### **📌Prepare LogFC Data for Correlation Analysis**

```{r load_data, echo=TRUE, results='hide', message=FALSE}
# Load Toptables directly from the DGE_Analysis Rmd environment
source("analysis/DGE_Analysis.Rmd")

# Ensure that all required Toptables are available
toptable_list <- list(
  CX_0.1_3 = Toptable_CX_0.1_3,
  CX_0.1_24 = Toptable_CX_0.1_24,
  CX_0.1_48 = Toptable_CX_0.1_48,
  CX_0.5_3 = Toptable_CX_0.5_3,
  CX_0.5_24 = Toptable_CX_0.5_24,
  CX_0.5_48 = Toptable_CX_0.5_48,
  DOX_0.1_3 = Toptable_DOX_0.1_3,
  DOX_0.1_24 = Toptable_DOX_0.1_24,
  DOX_0.1_48 = Toptable_DOX_0.1_48,
  DOX_0.5_3 = Toptable_DOX_0.5_3,
  DOX_0.5_24 = Toptable_DOX_0.5_24,
  DOX_0.5_48 = Toptable_DOX_0.5_48
)

# Extract Entrez_IDs from row names and logFC values
logFC_list <- lapply(names(toptable_list), function(name) {
  df <- toptable_list[[name]]
  df <- data.frame(Entrez_IDs = rownames(df), logFC = df$logFC)
  colnames(df)[2] <- name  # Rename the logFC column to the respective toptable name
  return(df)
})

# Merge all logFC data frames by Entrez_IDs
final_df <- Reduce(function(x, y) merge(x, y, by = "Entrez_IDs", all = TRUE), logFC_list)

# Print or save the final dataframe
print(head(final_df))  # Display first few rows
```

### **📌Load Metadata**
```{r load_Metadata, echo=TRUE, message=FALSE}
meta <- read.csv("data/Metadata.csv")

colnames(final_df)[-1] <- meta$Sample_ID
Drug <- meta$Drug
time <- meta$Time
conc <- as.character(meta$Conc.)
```

### **📌Define Color Annotations for Heatmap**
```{r Color Annotations, echo=TRUE, message=FALSE}
time_colors <- c("3" = "purple", "24" = "pink", "48" = "tomato3")
drug_colors <- c("CX-5461" = "yellow", "DOX" = "magenta4")
conc_colors <- c("0.1" = "lightblue", "0.5" = "lightcoral")

# Create annotations
top_annotation1 <- HeatmapAnnotation(
  timepoints = time,
  drugs = Drug,
  concentrations = conc,
  col = list(
    timepoints = time_colors,
    drugs = drug_colors,
    concentrations = conc_colors
  )
)
```

### **📌Compute Pearson and Spearman Correlation Matrices**
```{r Correlation, echo=TRUE, message=FALSE}
log2corr <- final_df[, -1]

cor_matrix1 <- cor(log2corr, method = "pearson")
cor_matrix2 <- cor(log2corr, method = "spearman")
```


### **📌Generate Heatmap (Pearson Correlation)**
```{r Pearson, echo=TRUE, message=FALSE}
heatmap1 <- Heatmap(
  cor_matrix1,
  name = "Correlation",
  top_annotation = top_annotation1,
  rect_gp = gpar(col = "black", lwd = 1),
  show_row_names = TRUE,
  show_column_names = TRUE,
  cell_fun = function(j, i, x, y, width, height, fill) {
    grid.text(sprintf("%.3f", cor_matrix1[i, j]), x, y, gp = gpar(fontsize = 10, col = "black"))
  }
)

# Draw the heatmap
draw(heatmap1)
```


### **📌Generate Heatmap (Spearman Correlation)**
```{r Spearman, echo=TRUE, message=FALSE}
heatmap2 <- Heatmap(
  cor_matrix2,
  name = "Correlation",
  top_annotation = top_annotation1,
  rect_gp = gpar(col = "black", lwd = 1),
  show_row_names = TRUE,
  show_column_names = TRUE,
  cell_fun = function(j, i, x, y, width, height, fill) {
    grid.text(sprintf("%.3f", cor_matrix2[i, j]), x, y, gp = gpar(fontsize = 10, col = "black"))
  }
)

# Draw the heatmap
draw(heatmap2)
```





