---
title: "LogFC Correlation"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## **📌 LogFC Correlation Analysis**

This analysis generates correlation heatmaps of log fold change (logFC) values across different comparisons.

### **📌 Load Required Libraries**
```{r load_libraries, echo=TRUE, message=FALSE}
library(ComplexHeatmap)
library(tidyverse)
library(data.table)
```

### **📌 Load LogFC Data**
```{r load_data, echo=TRUE, message=FALSE}
# Load logFC data from CSV
logFC_corr <- read.csv("data/LOG2FC.csv")

# Convert to dataframe
logFC_corr_df <- data.frame(logFC_corr)

# Remove 'X' prefix from the first column
names(logFC_corr_df)[1] <- sub("^X", "", names(logFC_corr_df)[1])

# Convert to matrix format for correlation analysis
log2corr <- as.matrix(logFC_corr_df[, -1])

# Display first few rows
print(head(log2corr))
```

### **📌 Load Metadata**
```{r load_metadata, echo=TRUE, message=FALSE}
# Load metadata
meta <- read.csv("data/Meta.csv")

# Assign column names based on sample metadata
colnames(log2corr) <- meta$Sample
Drug <- meta$Drug
time <- meta$Time
conc <- as.character(meta$Conc.)
```

### **📌 Define Color Annotations for Heatmap**
```{r color_annotations, echo=TRUE, message=FALSE}
time_colors <- c("3" = "purple", "24" = "pink", "48" = "tomato3")
drug_colors <- c("CX-5461" = "yellow", "DOX" = "magenta4")
conc_colors <- c("0.1" = "lightblue", "0.5" = "lightcoral")

# Create annotations
top_annotation1 <- HeatmapAnnotation(
  timepoints = time,
  drugs = Drug,
  concentrations = conc,
  col = list(
    timepoints = time_colors,
    drugs = drug_colors,
    concentrations = conc_colors
  )
)
```

### **📌 Compute Pearson and Spearman Correlation Matrices**
```{r correlation_matrices, echo=TRUE, message=FALSE}
cor_matrix1 <- cor(log2corr, method = "pearson")
cor_matrix2 <- cor(log2corr, method = "spearman")
```

### **📌 Generate Heatmap (Pearson Correlation)**
```{r pearson_heatmap, echo=TRUE, message=FALSE, fig.width=11, fig.height=8}
heatmap1 <- Heatmap(
  cor_matrix1,
  name = "Correlation",
  top_annotation = top_annotation1,
  rect_gp = gpar(col = "black", lwd = 1),
  show_row_names = TRUE,
  show_column_names = TRUE,
  cell_fun = function(j, i, x, y, width, height, fill) {
    grid.text(sprintf("%.3f", cor_matrix1[i, j]), x, y, gp = gpar(fontsize = 10, col = "black"))
  }
)

# Draw the heatmap
draw(heatmap1)
```

### **📌 Generate Heatmap (Spearman Correlation)**
```{r spearman_heatmap, echo=TRUE, message=FALSE,fig.width=11, fig.height=8}
heatmap2 <- Heatmap(
  cor_matrix2,
  name = "Correlation",
  top_annotation = top_annotation1,
  rect_gp = gpar(col = "black", lwd = 1),
  show_row_names = TRUE,
  show_column_names = TRUE,
  cell_fun = function(j, i, x, y, width, height, fill) {
    grid.text(sprintf("%.3f", cor_matrix2[i, j]), x, y, gp = gpar(fontsize = 10, col = "black"))
  }
)

# Draw the heatmap
draw(heatmap2)
```

