---
title: "Macrophage dataset comparison"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## **📌 Proportion of Macrophage DE genes in Corrmotif clusters**

## **📌 Load Required Libraries**
```{r load_libraries, echo=TRUE, message=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(org.Hs.eg.db)
library(clusterProfiler)
library(biomaRt)
library(gprofiler2)
library(AnnotationDbi)
```

## **📌 Load Data**
```{r load_File Paths, echo=TRUE,results='hide',message=FALSE}
# Read the file
file_path <- "data/Macrophage/IFN_gamma_primed_macrophages_DEGs.csv"  

IFN_gamma_primed_macrophages_DEGs <- read.csv(file_path, header = TRUE)

# Extract rat gene symbols (replace "Symbol" with the actual column name for gene symbols)
rat_genes <- IFN_gamma_primed_macrophages_DEGs$Symbol

# Map rat gene symbols to human homologs
homologs <- gorth(query = rat_genes,
                  source_organism = "rnorvegicus",  # Rat organism code
                  target_organism = "hsapiens")  # Human organism code


macrophage_new<- data.frame(homologs$ortholog_name)

# Map gene symbols to Entrez IDs using org.Hs.eg.db
macrophage_new <- macrophage_new %>%
  mutate(Entrez_ID = mapIds(org.Hs.eg.db,
                            keys = homologs.ortholog_name,
                            column = "ENTREZID",
                            keytype = "SYMBOL",
                            multiVals = "first"))
```

## **📌 Proportion of DE genes across response groups**
```{r load_data_tissue, echo=TRUE,results='hide', message=FALSE, fig.width=14,fig.height=8}

# Load the saved datasets
prob_all_1 <- read.csv("data/prob_all_1.csv")$Entrez_ID
prob_all_2 <- read.csv("data/prob_all_2.csv")$Entrez_ID
prob_all_3 <- read.csv("data/prob_all_3.csv")$Entrez_ID
prob_all_4 <- read.csv("data/prob_all_4.csv")$Entrez_ID

# Example Response Groups Data (Replace with actual data)
response_groups <- list(
  "Non Response" = prob_all_1, # Replace 'prob_all_1', 'prob_all_2', etc. with your actual response group dataframes
  "CX_DOX Shared Late Response" = prob_all_2,
  "DOX-Specific Response" = prob_all_3,
  "Late High Dose DOX-Specific Response" = prob_all_4
)

# Combine response groups into a single dataframe
response_groups_df <- bind_rows(
  lapply(response_groups, function(ids) {
    data.frame(Entrez_ID = ids)
  }),
  .id = "Set"
)

# Step 2: Match Overlap Genes with Response Groups
# Classify genes as DEG (match) or Non-DEG (no match)
response_groups_df <- response_groups_df %>%
  mutate(
    DEG_Status = ifelse(Entrez_ID %in% macrophage_new$Entrez_ID, "DEG", "Non-DEG")
  )

# Step 3: Calculate Proportions
proportion_data <- response_groups_df %>%
  group_by(Set, DEG_Status) %>%
  summarize(Count = n(), .groups = "drop") %>%
  group_by(Set) %>%
  mutate(Percentage = (Count / sum(Count)) * 100)

# Step 4: Perform Chi-Square Tests (Refactored Version)
# Get counts for the Non Response group
non_response_counts <- proportion_data %>%
  filter(Set == "Non Response") %>%
  dplyr::select(DEG_Status, Count) %>%
  {setNames(.$Count, .$DEG_Status)}  # Create named vector for Non Response counts

# Perform chi-square test for selected response groups
chi_results <- proportion_data %>%
  filter(Set != "Non Response") %>% # Exclude "Non Response"
  group_by(Set) %>%
  summarize(
    p_value = {
      # Extract counts for the current response group
      group_counts <- Count[DEG_Status %in% c("DEG", "Non-DEG")]
      # Ensure there are no missing categories, fill with 0 if missing
      if (!"DEG" %in% DEG_Status) group_counts <- c(group_counts, 0)
      if (!"Non-DEG" %in% DEG_Status) group_counts <- c(0, group_counts)
      # Create contingency table
      contingency_table <- matrix(c(
        group_counts[1], group_counts[2],
        non_response_counts["DEG"], non_response_counts["Non-DEG"]
      ), nrow = 2, byrow = TRUE)
      # Debugging: Print the contingency table
      print(paste("Set:", unique(Set)))
      print("Contingency Table:")
      print(contingency_table)
      # Perform chi-square test
      if (all(contingency_table >= 0 & is.finite(contingency_table))) {
        chisq.test(contingency_table)$p.value
      } else {
        NA
      }
    },
    .groups = "drop"
  ) %>%
  mutate(Significance = ifelse(!is.na(p_value) & p_value < 0.05, "*", ""))

# Step 5: Merge Results and Plot Proportions
# Merge chi-square results back into proportion data
proportion_data <- proportion_data %>%
  left_join(chi_results %>% dplyr::select(Set, Significance), by = "Set")

# Define the correct order for response groups
response_order <- c(
  "Non Response",
  "CX_DOX Shared Late Response",
  "DOX-Specific Response",
  "Late High Dose DOX-Specific Response"
)
proportion_data$Set <- factor(proportion_data$Set, levels = response_order)

# Plot proportions with significance stars
ggplot(proportion_data, aes(x = Set, y = Percentage, fill = DEG_Status)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(
    data = proportion_data %>% distinct(Set, Significance),
    aes(x = Set, y = 105, label = Significance), # Position stars above bars
    inherit.aes = FALSE,
    size = 6,
    color = "black",
    hjust = 0.5
  ) +
  scale_fill_manual(values = c("DEG" = "#e41a1c", "Non-DEG" = "#377eb8")) +
  labs(
    title = "Proportion of DEGs and Non-DEGs Across Response Groups\n(IFN_gamma_primed_macrophages_DEGs)",
    x = "Response Groups",
    y = "Percentage",
    fill = "Category"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = rel(1.5), hjust = 0.5),
    axis.title = element_text(size = 15, color = "black"),
    axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
    legend.title = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1.2)
  )
```

## **📌 Proportion of Macrophage genes in CX and DOX DEGs**

## **📌 Read and Process Data**
```{r load_DEGs, echo=TRUE, results='hide', message=FALSE, warning=FALSE}
# Load DEGs Data
CX_0.1_3 <- read.csv("data/DEGs/Toptable_CX_0.1_3.csv")
CX_0.1_24 <- read.csv("data/DEGs/Toptable_CX_0.1_24.csv")
CX_0.1_48 <- read.csv("data/DEGs/Toptable_CX_0.1_48.csv")
CX_0.5_3 <- read.csv("data/DEGs/Toptable_CX_0.5_3.csv")
CX_0.5_24 <- read.csv("data/DEGs/Toptable_CX_0.5_24.csv")
CX_0.5_48 <- read.csv("data/DEGs/Toptable_CX_0.5_48.csv")

DOX_0.1_3 <- read.csv("data/DEGs/Toptable_DOX_0.1_3.csv")
DOX_0.1_24 <- read.csv("data/DEGs/Toptable_DOX_0.1_24.csv")
DOX_0.1_48 <- read.csv("data/DEGs/Toptable_DOX_0.1_48.csv")
DOX_0.5_3 <- read.csv("data/DEGs/Toptable_DOX_0.5_3.csv")
DOX_0.5_24 <- read.csv("data/DEGs/Toptable_DOX_0.5_24.csv")
DOX_0.5_48 <- read.csv("data/DEGs/Toptable_DOX_0.5_48.csv")

# Extract Significant DEGs
DEG1 <- as.character(CX_0.1_3$Entrez_ID[CX_0.1_3$adj.P.Val < 0.05])
DEG2 <- as.character(CX_0.1_24$Entrez_ID[CX_0.1_24$adj.P.Val < 0.05])
DEG3 <- as.character(CX_0.1_48$Entrez_ID[CX_0.1_48$adj.P.Val < 0.05])
DEG4 <- as.character(CX_0.5_3$Entrez_ID[CX_0.5_3$adj.P.Val < 0.05])
DEG5 <- as.character(CX_0.5_24$Entrez_ID[CX_0.5_24$adj.P.Val < 0.05])
DEG6 <- as.character(CX_0.5_48$Entrez_ID[CX_0.5_48$adj.P.Val < 0.05])
DEG7 <- as.character(DOX_0.1_3$Entrez_ID[DOX_0.1_3$adj.P.Val < 0.05])
DEG8 <- as.character(DOX_0.1_24$Entrez_ID[DOX_0.1_24$adj.P.Val < 0.05])
DEG9 <- as.character(DOX_0.1_48$Entrez_ID[DOX_0.1_48$adj.P.Val < 0.05])
DEG10 <- as.character(DOX_0.5_3$Entrez_ID[DOX_0.5_3$adj.P.Val < 0.05])
DEG11 <- as.character(DOX_0.5_24$Entrez_ID[DOX_0.5_24$adj.P.Val < 0.05])
DEG12 <- as.character(DOX_0.5_48$Entrez_ID[DOX_0.5_48$adj.P.Val < 0.05])
```

## **📌 Proportion of Macrophage genes in CX and DOX DEGs datasets**
```{r proportion1, echo=TRUE,results='hide', message=FALSE, fig.width=14,fig.height=8}

# Define CX-5461 DEG lists
CX_DEGs <- list(
  "CX_0.1_3" = DEG1, "CX_0.1_24" = DEG2, "CX_0.1_48" = DEG3,
  "CX_0.5_3" = DEG4, "CX_0.5_24" = DEG5, "CX_0.5_48" = DEG6
)

# Define DOX DEG lists
DOX_DEGs <- list(
  "DOX_0.1_3" = DEG7, "DOX_0.1_24" = DEG8, "DOX_0.1_48" = DEG9,
  "DOX_0.5_3" = DEG10, "DOX_0.5_24" = DEG11, "DOX_0.5_48" = DEG12
)

# Load Macrophage_new dataset (Use `Entrez_ID` for matching)
Macrophage_genes <- na.omit(macrophage_new$Entrez_ID)  # Keep only Entrez_IDs

# **Process CX-5461 Samples**
CX_DEGs_df <- bind_rows(
  lapply(CX_DEGs, function(ids) data.frame(Entrez_ID = ids, Sample_Type = "CX-5461")),
  .id = "Sample"
) %>%
  mutate(Category = ifelse(Entrez_ID %in% Macrophage_genes, "Yes", "No")) %>%
  group_by(Sample, Sample_Type, Category) %>%
  summarise(Count = n(), .groups = "drop") %>%
  group_by(Sample, Sample_Type) %>%
  mutate(Percentage = (Count / sum(Count)) * 100)

# **Process DOX Samples**
DOX_DEGs_df <- bind_rows(
  lapply(DOX_DEGs, function(ids) data.frame(Entrez_ID = ids, Sample_Type = "DOX")),
  .id = "Sample"
) %>%
  mutate(Category = ifelse(Entrez_ID %in% Macrophage_genes, "Yes", "No")) %>%
  group_by(Sample, Sample_Type, Category) %>%
  summarise(Count = n(), .groups = "drop") %>%
  group_by(Sample, Sample_Type) %>%
  mutate(Percentage = (Count / sum(Count)) * 100)

# **Merge CX and DOX Dataframes**
proportion_data <- bind_rows(CX_DEGs_df, DOX_DEGs_df)



# **Fix: Normalize Percentages to Sum Exactly 100%**
proportion_data <- proportion_data %>%
  group_by(Sample, Sample_Type) %>%
  mutate(Percentage = round(Percentage, 2)) %>%  # Round percentages
  mutate(Adjustment = 100 - sum(Percentage, na.rm = TRUE)) %>%  # Compute rounding difference
  mutate(Percentage = ifelse(Category == "No", Percentage + Adjustment, Percentage)) %>%  # Apply adjustment to "No"
  mutate(Percentage = ifelse(Percentage < 0, 0, Percentage)) %>%  # Prevent negatives
  mutate(Percentage = ifelse(Percentage > 100, 100, Percentage)) %>%  # Prevent values > 100
  ungroup() %>%
  replace_na(list(Percentage = 0))  # Ensure no NA values

# **Ensure "Yes" is at the Bottom and "No" is at the Top**
proportion_data$Category <- factor(proportion_data$Category, levels = c("Yes", "No"))


# **Ensure "Yes" is at the Bottom and "No" is at the Top**
proportion_data$Category <- factor(proportion_data$Category, levels = c("Yes", "No"))

# Define correct order for samples
sample_order <- c(
  "CX_0.1_3", "CX_0.1_24", "CX_0.1_48", "CX_0.5_3", "CX_0.5_24", "CX_0.5_48",
  "DOX_0.1_3", "DOX_0.1_24", "DOX_0.1_48", "DOX_0.5_3", "DOX_0.5_24", "DOX_0.5_48"
)
proportion_data$Sample <- factor(proportion_data$Sample, levels = sample_order)


# **Generate Proportion Plot for CX-5461 and DOX Separately**
ggplot(proportion_data, aes(x = Sample, y = Percentage, fill = Category)) +
  geom_bar(stat = "identity", position = "stack") +  # Stacked bars
  facet_wrap(~Sample_Type, scales = "free_x") +  # Separate CX-5461 and DOX
  scale_y_continuous(labels = scales::percent_format(scale = 1), limits = c(0, 100)) +  # Y-axis as percentage
  scale_fill_manual(values = c("Yes" = "#e41a1c", "No" = "#377eb8")) +  # Yes (Red) at Bottom, No (Blue) on Top
  labs(
    title = "Proportion of IFN gamma primed macrophage Genes in CX-5461 and DOX DEGs",
    x = "Samples (CX-5461 and DOX)",
    y = "Percentage",
    fill = "Category"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = rel(1.5), hjust = 0.5),
    axis.title = element_text(size = 15, color = "black"),
    axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
    legend.title = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1.2),  # Updated for ggplot2 3.4.0+
    strip.background = element_blank(),
    strip.text = element_text(size = 12, face = "bold")
  )
```

## **📌 Correlation of Macrophage genes with CX and DOX expressed genes**
## **📌 Correlation of Macrophage genes with CX expressed genes**
```{r Data_prep, echo=TRUE, message=FALSE, fig.height=8,fig.width=12}
macrophage <- read.csv("data/Macrophage/IFN_gamma_primed_macrophages_DEGs.csv")

# **Step 2: Extract Rat Gene Symbols**
rat_genes <- unique(macrophage$Symbol)

# **Step 3: Map Rat Gene Symbols to Human Orthologs**
homologs <- gorth(query = rat_genes,
                  source_organism = "rnorvegicus",  # Rat organism code
                  target_organism = "hsapiens",  # Human organism code
                  numeric_ns = "")  # Ensure proper namespace handling

# **Step 4: Check Column Names**
print(colnames(homologs))  # Check available columns

# **Step 5: Add Human Orthologs Directly to Macrophage Dataset**
# Create a lookup table for faster mapping
homolog_map <- setNames(homologs$ortholog_name, homologs$input)

# Assign the mapped human genes directly into a new column in macrophage dataset
macrophage$Human_Symbol <- homolog_map[macrophage$Symbol]

# **Step 6: Remove Unmapped Entries (Optional)**
macrophage <- macrophage %>% filter(!is.na(Human_Symbol))

# **Step 7: Save Final Dataset**
write.csv(macrophage, "C:/Work/Postdoc_UTMB/CX-5461 Project/Transcriptome literatures/Macrophage/macrophage_Mapped.csv", row.names = FALSE)

# **Step 8: Display Output**
print(head(macrophage))  # Show first few rows of updated dataset


# **Step 6: Filter Out NA Values (Only Mapped Genes)**
macrophage <- macrophage %>% filter(!is.na(Human_Symbol))

# Ensure Human_Symbol is the first column and remove Symbol
macrophage <- macrophage %>%
  dplyr::select(Human_Symbol, everything(), -Symbol)

# **Step 1: Map Gene Symbols to Entrez IDs using org.Hs.eg.db**
macrophage <- macrophage %>%
  mutate(Entrez_ID = mapIds(org.Hs.eg.db,
                            keys = Human_Symbol,
                            column = "ENTREZID",
                            keytype = "SYMBOL",
                            multiVals = "first"))

# **Step 2: Convert Entrez_ID to character to avoid merge issues**
macrophage$Entrez_ID <- as.character(macrophage$Entrez_ID)

CX_0.1_3$Entrez_ID <- as.character(CX_0.1_3$Entrez_ID)
CX_0.5_3$Entrez_ID <- as.character(CX_0.5_3$Entrez_ID)
CX_0.1_24$Entrez_ID <- as.character(CX_0.1_24$Entrez_ID)
CX_0.5_24$Entrez_ID <- as.character(CX_0.5_24$Entrez_ID)
CX_0.1_48$Entrez_ID <- as.character(CX_0.1_48$Entrez_ID)
CX_0.5_48$Entrez_ID <- as.character(CX_0.5_48$Entrez_ID)

# **Step 3: Merge Macrophage dataset with CX at different concentrations & timepoints**
merged_CX_0.1_3 <- merge(macrophage, CX_0.1_3, by = "Entrez_ID")
merged_CX_0.5_3 <- merge(macrophage, CX_0.5_3, by = "Entrez_ID")
merged_CX_0.1_24 <- merge(macrophage, CX_0.1_24, by = "Entrez_ID")
merged_CX_0.5_24 <- merge(macrophage, CX_0.5_24, by = "Entrez_ID")
merged_CX_0.1_48 <- merge(macrophage, CX_0.1_48, by = "Entrez_ID")
merged_CX_0.5_48 <- merge(macrophage, CX_0.5_48, by = "Entrez_ID")

# **Step 4: Remove NA values**
merged_CX_0.1_3 <- na.omit(merged_CX_0.1_3)
merged_CX_0.5_3 <- na.omit(merged_CX_0.5_3)
merged_CX_0.1_24 <- na.omit(merged_CX_0.1_24)
merged_CX_0.5_24 <- na.omit(merged_CX_0.5_24)
merged_CX_0.1_48 <- na.omit(merged_CX_0.1_48)
merged_CX_0.5_48 <- na.omit(merged_CX_0.5_48)

# **Step 5: Rename columns to avoid conflicts**
colnames(merged_CX_0.1_3) <- colnames(merged_CX_0.5_3) <-
  colnames(merged_CX_0.1_24) <- colnames(merged_CX_0.5_24) <-
  colnames(merged_CX_0.1_48) <- colnames(merged_CX_0.5_48) <-
  c("Entrez_ID", "Symbol_Macrophage", "logFC_Macrophage", "logFC_CX", "AveExpr_CX", "t_CX", "P.Value_CX", "adj.P.Val_CX", "B_CX")

# **Step 6: Add timepoint and concentration labels for faceting**
merged_CX_0.1_3$Timepoint <- "3hr"
merged_CX_0.5_3$Timepoint <- "3hr"
merged_CX_0.1_24$Timepoint <- "24hr"
merged_CX_0.5_24$Timepoint <- "24hr"
merged_CX_0.1_48$Timepoint <- "48hr"
merged_CX_0.5_48$Timepoint <- "48hr"

merged_CX_0.1_3$Concentration <- "0.1 µM"
merged_CX_0.5_3$Concentration <- "0.5 µM"
merged_CX_0.1_24$Concentration <- "0.1 µM"
merged_CX_0.5_24$Concentration <- "0.5 µM"
merged_CX_0.1_48$Concentration <- "0.1 µM"
merged_CX_0.5_48$Concentration <- "0.5 µM"

# **Step 7: Combine all datasets into a single data frame**
merged_data_macrophage <- rbind(
  merged_CX_0.1_3[, c("Entrez_ID", "logFC_CX", "logFC_Macrophage", "Timepoint", "Concentration")],
  merged_CX_0.5_3[, c("Entrez_ID", "logFC_CX", "logFC_Macrophage", "Timepoint", "Concentration")],
  merged_CX_0.1_24[, c("Entrez_ID", "logFC_CX", "logFC_Macrophage", "Timepoint", "Concentration")],
  merged_CX_0.5_24[, c("Entrez_ID", "logFC_CX", "logFC_Macrophage", "Timepoint", "Concentration")],
  merged_CX_0.1_48[, c("Entrez_ID", "logFC_CX", "logFC_Macrophage", "Timepoint", "Concentration")],
  merged_CX_0.5_48[, c("Entrez_ID", "logFC_CX", "logFC_Macrophage", "Timepoint", "Concentration")]
)

# **Ensure timepoints are in correct order**
merged_data_macrophage$Timepoint <- factor(merged_data_macrophage$Timepoint, levels = c("3hr", "24hr", "48hr"))

# **Step 8: Compute correlations for each facet**
correlations <- merged_data_macrophage %>%
  group_by(Concentration, Timepoint) %>%
  summarise(
    r_value = cor(logFC_CX, logFC_Macrophage, method = "pearson"),
    p_value = cor.test(logFC_CX, logFC_Macrophage, method = "pearson")$p.value,
    .groups = "drop"
  )

# **Step 9: Create correlation annotation data**
correlation_data <- correlations %>%
  mutate(
    x = 1.5,  # Adjusted to fit within fixed axis range (-5 to 2)
    y = max(merged_data_macrophage$logFC_Macrophage, na.rm = TRUE) * 0.85,
    label = paste0("r = ", round(r_value, 3), "\np = ", signif(p_value, 3))
  )

# **Step 10: Create styled scatter plot with fixed X-axis range and ordered timepoints**
scatter_plot_macrophage <- ggplot(merged_data_macrophage, aes(x = logFC_CX, y = logFC_Macrophage)) +
  geom_point(alpha = 0.6, color = "black") +
  geom_smooth(method = "lm", color = "black", se = FALSE) +
  scale_x_continuous(limits = c(-6, 2)) +  # Fixed X-axis range
  labs(
    title = "Correlation between CX and IFN_gamma_primed_macrophages logFC",
    x = "logFC (CX)",
    y = "logFC (Macrophage)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 2),  # Outer border
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1.5),
    strip.text = element_text(size = 12, face = "bold", color = "black")
  ) +
  facet_grid(Timepoint ~ Concentration, scales = "fixed") +  # Ensures correct timepoint order
  geom_text(data = correlation_data,
            aes(x = x, y = y, label = label),
            inherit.aes = FALSE, size = 3, fontface = "bold")

# **Step 11: Display the plot**
print(scatter_plot_macrophage)
```

## **📌 Correlation of Myeloma genes with DOX expressed genes**
```{r Data_DOX, echo=TRUE, message=FALSE, fig.height=8,fig.width=12}
# **Step 2: Convert Entrez_ID to character to avoid merge issues**
macrophage$Entrez_ID <- as.character(macrophage$Entrez_ID)

DOX_0.1_3$Entrez_ID <- as.character(DOX_0.1_3$Entrez_ID)
DOX_0.5_3$Entrez_ID <- as.character(DOX_0.5_3$Entrez_ID)
DOX_0.1_24$Entrez_ID <- as.character(DOX_0.1_24$Entrez_ID)
DOX_0.5_24$Entrez_ID <- as.character(DOX_0.5_24$Entrez_ID)
DOX_0.1_48$Entrez_ID <- as.character(DOX_0.1_48$Entrez_ID)
DOX_0.5_48$Entrez_ID <- as.character(DOX_0.5_48$Entrez_ID)

# **Step 3: Merge Macrophage dataset with DOX at different concentrations & timepoints**
merged_DOX_0.1_3 <- merge(macrophage, DOX_0.1_3, by = "Entrez_ID")
merged_DOX_0.5_3 <- merge(macrophage, DOX_0.5_3, by = "Entrez_ID")
merged_DOX_0.1_24 <- merge(macrophage, DOX_0.1_24, by = "Entrez_ID")
merged_DOX_0.5_24 <- merge(macrophage, DOX_0.5_24, by = "Entrez_ID")
merged_DOX_0.1_48 <- merge(macrophage, DOX_0.1_48, by = "Entrez_ID")
merged_DOX_0.5_48 <- merge(macrophage, DOX_0.5_48, by = "Entrez_ID")

# **Step 4: Remove NA values**
merged_DOX_0.1_3 <- na.omit(merged_DOX_0.1_3)
merged_DOX_0.5_3 <- na.omit(merged_DOX_0.5_3)
merged_DOX_0.1_24 <- na.omit(merged_DOX_0.1_24)
merged_DOX_0.5_24 <- na.omit(merged_DOX_0.5_24)
merged_DOX_0.1_48 <- na.omit(merged_DOX_0.1_48)
merged_DOX_0.5_48 <- na.omit(merged_DOX_0.5_48)

# **Step 5: Rename columns to avoid conflicts**
colnames(merged_DOX_0.1_3) <- colnames(merged_DOX_0.5_3) <-
  colnames(merged_DOX_0.1_24) <- colnames(merged_DOX_0.5_24) <-
  colnames(merged_DOX_0.1_48) <- colnames(merged_DOX_0.5_48) <-
  c("Entrez_ID", "Symbol_Macrophage", "logFC_Macrophage", "logFC_DOX", "AveExpr_DOX", "t_DOX", "P.Value_DOX", "adj.P.Val_DOX", "B_DOX")

# **Step 6: Add timepoint and concentration labels for faceting**
merged_DOX_0.1_3$Timepoint <- "3hr"
merged_DOX_0.5_3$Timepoint <- "3hr"
merged_DOX_0.1_24$Timepoint <- "24hr"
merged_DOX_0.5_24$Timepoint <- "24hr"
merged_DOX_0.1_48$Timepoint <- "48hr"
merged_DOX_0.5_48$Timepoint <- "48hr"

merged_DOX_0.1_3$Concentration <- "0.1 µM"
merged_DOX_0.5_3$Concentration <- "0.5 µM"
merged_DOX_0.1_24$Concentration <- "0.1 µM"
merged_DOX_0.5_24$Concentration <- "0.5 µM"
merged_DOX_0.1_48$Concentration <- "0.1 µM"
merged_DOX_0.5_48$Concentration <- "0.5 µM"

# **Step 7: Combine all datasets into a single data frame**
merged_data_macrophage_dox <- rbind(
  merged_DOX_0.1_3[, c("Entrez_ID", "logFC_DOX", "logFC_Macrophage", "Timepoint", "Concentration")],
  merged_DOX_0.5_3[, c("Entrez_ID", "logFC_DOX", "logFC_Macrophage", "Timepoint", "Concentration")],
  merged_DOX_0.1_24[, c("Entrez_ID", "logFC_DOX", "logFC_Macrophage", "Timepoint", "Concentration")],
  merged_DOX_0.5_24[, c("Entrez_ID", "logFC_DOX", "logFC_Macrophage", "Timepoint", "Concentration")],
  merged_DOX_0.1_48[, c("Entrez_ID", "logFC_DOX", "logFC_Macrophage", "Timepoint", "Concentration")],
  merged_DOX_0.5_48[, c("Entrez_ID", "logFC_DOX", "logFC_Macrophage", "Timepoint", "Concentration")]
)

# **Ensure timepoints are in correct order**
merged_data_macrophage_dox$Timepoint <- factor(merged_data_macrophage_dox$Timepoint, levels = c("3hr", "24hr", "48hr"))

# **Step 8: Compute correlations for each facet**
correlations <- merged_data_macrophage_dox %>%
  group_by(Concentration, Timepoint) %>%
  summarise(
    r_value = cor(logFC_DOX, logFC_Macrophage, method = "pearson"),
    p_value = cor.test(logFC_DOX, logFC_Macrophage, method = "pearson")$p.value,
    .groups = "drop"
  )

# **Step 9: Create correlation annotation data**
correlation_data <- correlations %>%
  mutate(
    x = 1.5,  # Adjusted to fit within fixed axis range (-5 to 2)
    y = max(merged_data_macrophage_dox$logFC_Macrophage, na.rm = TRUE) * 0.85,
    label = paste0("r = ", round(r_value, 3), "\np = ", signif(p_value, 3))
  )

# **Step 10: Create styled scatter plot with fixed X-axis range and ordered timepoints**
scatter_plot_macrophage_dox <- ggplot(merged_data_macrophage_dox, aes(x = logFC_DOX, y = logFC_Macrophage)) +
  geom_point(alpha = 0.6, color = "black") +
  geom_smooth(method = "lm", color = "black", se = FALSE) +
  #scale_x_continuous(limits = c(-5, 2)) +  # Fixed X-axis range
  labs(
    title = "Correlation between DOX and IFN_gamma_primed_macrophages logFC",
    x = "logFC (DOX)",
    y = "logFC (Macrophage)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 2),  # Outer border
    strip.background = element_rect(fill = "white", color = "black", linewidth = 1.5),
    strip.text = element_text(size = 12, face = "bold", color = "black")
  ) +
  facet_grid(Timepoint ~ Concentration, scales = "fixed") +  # Ensures correct timepoint order
  geom_text(data = correlation_data,
            aes(x = x, y = y, label = label),
            inherit.aes = FALSE, size = 3, fontface = "bold")

# **Step 11: Display the plot**
print(scatter_plot_macrophage_dox)
```


## **📌 Proportion of Macrophage Genes in corrmotif 0.1 micromolar**
```{r prop DNA, echo=TRUE, message=FALSE}
# Load necessary libraries
library(dplyr)
library(ggplot2)
library(tidyr)
library(org.Hs.eg.db)
library(gprofiler2)

# Read Macrophage Dataset
file_path <- "data/Macrophage/IFN_gamma_primed_macrophages_DEGs.csv"  
IFN_gamma_primed_macrophages_DEGs <- read.csv(file_path, header = TRUE)

# Extract rat gene symbols
rat_genes <- IFN_gamma_primed_macrophages_DEGs$Symbol

# Map rat gene symbols to human homologs
homologs <- gorth(query = rat_genes,
                  source_organism = "rnorvegicus",  # Rat organism code
                  target_organism = "hsapiens")  # Human organism code

# Convert to dataframe and map gene symbols to Entrez IDs
macrophage_new <- data.frame(homologs$ortholog_name) %>%
  mutate(Entrez_ID = mapIds(org.Hs.eg.db,
                            keys = homologs.ortholog_name,
                            column = "ENTREZID",
                            keytype = "SYMBOL",
                            multiVals = "first")) %>%
  na.omit()

# Load Corrmotif Groups for 0.1 Concentration
prob_groups_0.1 <- list(
  "Non Response (0.1)" = read.csv("data/prob_1_0.1.csv")$Entrez_ID,
  "DOX only mid-late (0.1)" = read.csv("data/prob_2_0.1.csv")$Entrez_ID,
  "CX_DOX mid-late (0.1)" = read.csv("data/prob_3_0.1.csv")$Entrez_ID
)

# Function to calculate proportions in macrophage dataset
calculate_proportion <- function(deg_list, group_name) {
  total_group_genes <- length(deg_list)
  matched_genes <- sum(deg_list %in% macrophage_new$Entrez_ID)
  
  data.frame(
    Group = group_name,
    Macrophage_DEGs = matched_genes,
    Non_Macrophage_DEGs = total_group_genes - matched_genes
  ) %>%
    mutate(
      Yes_Proportion = (Macrophage_DEGs / total_group_genes) * 100,
      No_Proportion = (Non_Macrophage_DEGs / total_group_genes) * 100
    )
}

# Calculate Proportions for Each Group
proportion_data <- bind_rows(
  calculate_proportion(prob_groups_0.1[["Non Response (0.1)"]], "Non Response (0.1)"),
  calculate_proportion(prob_groups_0.1[["DOX only mid-late (0.1)"]], "DOX only mid-late (0.1)"),
  calculate_proportion(prob_groups_0.1[["CX_DOX mid-late (0.1)"]], "CX_DOX mid-late (0.1)")
)

# Convert Data to Long Format
proportion_long <- proportion_data %>%
  pivot_longer(cols = c(Yes_Proportion, No_Proportion), names_to = "Category", values_to = "Percentage") %>%
  mutate(Category = ifelse(Category == "Yes_Proportion", "Yes", "No"))

# Perform Chi-Square Test Comparing Groups to Non-Response
test_results <- data.frame(Group = character(), P_Value = numeric())

target_groups <- c("DOX only mid-late (0.1)", "CX_DOX mid-late (0.1)")
for (grp in target_groups) {
  contingency_table <- matrix(
    c(
      sum(prob_groups_0.1[[grp]] %in% macrophage_new$Entrez_ID),
      length(prob_groups_0.1[[grp]]) - sum(prob_groups_0.1[[grp]] %in% macrophage_new$Entrez_ID),
      sum(prob_groups_0.1[["Non Response (0.1)"]] %in% macrophage_new$Entrez_ID),
      length(prob_groups_0.1[["Non Response (0.1)"]]) - sum(prob_groups_0.1[["Non Response (0.1)"]] %in% macrophage_new$Entrez_ID)
    ),
    nrow = 2, byrow = TRUE
  )
  test_result <- chisq.test(contingency_table)
  test_results <- rbind(test_results, data.frame(Group = grp, P_Value = test_result$p.value))
}

# Add Significance Stars
test_results$Significant <- ifelse(test_results$P_Value < 0.05, "*", "")
proportion_long <- left_join(proportion_long, test_results, by = "Group")

# Ensure Correct Ordering
proportion_long$Group <- factor(proportion_long$Group, levels = c("Non Response (0.1)", "DOX only mid-late (0.1)", "CX_DOX mid-late (0.1)"))
proportion_long$Category <- factor(proportion_long$Category, levels = c("Yes", "No"))

# Generate Stacked Bar Plot with Significance Annotations
ggplot(proportion_long, aes(x = Group, y = Percentage, fill = Category)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(data = subset(proportion_long, Significant == "*"),
            aes(x = Group, y = 102, label = "*"),
            size = 6, color = "black", fontface = "bold") +
  scale_y_continuous(labels = scales::percent_format(scale = 1), limits = c(0, 105)) +
  scale_fill_manual(values = c("Yes" = "#e41a1c", "No" = "#377eb8")) +
  labs(
    title = "Proportion of Macrophage Genes in\n0.1 Corrmotif Response Groups",
    x = "Group",
    y = "Percentage",
    fill = "Category"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = rel(1.5), hjust = 0.5),
    axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
    legend.title = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1.2),
    strip.background = element_blank(),
    strip.text = element_text(size = 12, face = "bold")
  )
```


## **📌 Proportion of Macrophage Genes in corrmotif 0.5 micromolar**
```{r prop 0.5, echo=TRUE, message=FALSE}
# Load necessary libraries
library(dplyr)
library(ggplot2)
library(tidyr)
library(org.Hs.eg.db)
library(gprofiler2)

# Read Macrophage Dataset
file_path <- "data/Macrophage/IFN_gamma_primed_macrophages_DEGs.csv"  
IFN_gamma_primed_macrophages_DEGs <- read.csv(file_path, header = TRUE)

# Extract rat gene symbols
rat_genes <- IFN_gamma_primed_macrophages_DEGs$Symbol

# Map rat gene symbols to human homologs
homologs <- gorth(query = rat_genes,
                  source_organism = "rnorvegicus",  # Rat organism code
                  target_organism = "hsapiens")  # Human organism code

# Convert to dataframe and map gene symbols to Entrez IDs
macrophage_new <- data.frame(homologs$ortholog_name) %>%
  mutate(Entrez_ID = mapIds(org.Hs.eg.db,
                            keys = homologs.ortholog_name,
                            column = "ENTREZID",
                            keytype = "SYMBOL",
                            multiVals = "first")) %>%
  na.omit()

# Load Corrmotif Groups for 0.5 Concentration
prob_groups_0.5 <- list(
  "Non Response (0.5)" = read.csv("data/prob_1_0.5.csv")$Entrez_ID,
  "DOX-specific response (0.5)" = read.csv("data/prob_2_0.5.csv")$Entrez_ID,
  "DOX only mid-late response (0.5)" = read.csv("data/prob_3_0.5.csv")$Entrez_ID,
  "CX DOX (early) response (0.5)" = read.csv("data/prob_4_0.5.csv")$Entrez_ID,
  "DOX + CX (mid-late) response (0.5)" = read.csv("data/prob_5_0.5.csv")$Entrez_ID
)

# Function to calculate proportions in macrophage dataset
calculate_proportion <- function(deg_list, group_name) {
  total_group_genes <- length(deg_list)
  matched_genes <- sum(deg_list %in% macrophage_new$Entrez_ID)
  
  data.frame(
    Group = group_name,
    Macrophage_DEGs = matched_genes,
    Non_Macrophage_DEGs = total_group_genes - matched_genes
  ) %>%
    mutate(
      Yes_Proportion = (Macrophage_DEGs / total_group_genes) * 100,
      No_Proportion = (Non_Macrophage_DEGs / total_group_genes) * 100
    )
}

# Calculate Proportions for Each Group
proportion_data <- bind_rows(
  calculate_proportion(prob_groups_0.5[["Non Response (0.5)"]], "Non Response (0.5)"),
  calculate_proportion(prob_groups_0.5[["DOX-specific response (0.5)"]], "DOX-specific response (0.5)"),
  calculate_proportion(prob_groups_0.5[["DOX only mid-late response (0.5)"]], "DOX only mid-late response (0.5)"),
  calculate_proportion(prob_groups_0.5[["CX DOX (early) response (0.5)"]], "CX DOX (early) response (0.5)"),
  calculate_proportion(prob_groups_0.5[["DOX + CX (mid-late) response (0.5)"]], "DOX + CX (mid-late) response (0.5)")
)

# Convert Data to Long Format
proportion_long <- proportion_data %>%
  pivot_longer(cols = c(Yes_Proportion, No_Proportion), names_to = "Category", values_to = "Percentage") %>%
  mutate(Category = ifelse(Category == "Yes_Proportion", "Yes", "No"))

# Perform Chi-Square Test Comparing Groups to Non-Response
test_results <- data.frame(Group = character(), P_Value = numeric())

target_groups <- c("DOX-specific response (0.5)", "DOX only mid-late response (0.5)", "CX DOX (early) response (0.5)", "DOX + CX (mid-late) response (0.5)")
for (grp in target_groups) {
  contingency_table <- matrix(
    c(
      sum(prob_groups_0.5[[grp]] %in% macrophage_new$Entrez_ID),
      length(prob_groups_0.5[[grp]]) - sum(prob_groups_0.5[[grp]] %in% macrophage_new$Entrez_ID),
      sum(prob_groups_0.5[["Non Response (0.5)"]] %in% macrophage_new$Entrez_ID),
      length(prob_groups_0.5[["Non Response (0.5)"]]) - sum(prob_groups_0.5[["Non Response (0.5)"]] %in% macrophage_new$Entrez_ID)
    ),
    nrow = 2, byrow = TRUE
  )
  test_result <- chisq.test(contingency_table)
  test_results <- rbind(test_results, data.frame(Group = grp, P_Value = test_result$p.value))
}

# Add Significance Stars
test_results$Significant <- ifelse(test_results$P_Value < 0.05, "*", "")
proportion_long <- left_join(proportion_long, test_results, by = "Group")

# Ensure Correct Ordering
proportion_long$Group <- factor(proportion_long$Group, levels = c("Non Response (0.5)", "DOX-specific response (0.5)", "DOX only mid-late response (0.5)", "CX DOX (early) response (0.5)", "DOX + CX (mid-late) response (0.5)"))
proportion_long$Category <- factor(proportion_long$Category, levels = c("Yes", "No"))

# Generate Stacked Bar Plot with Significance Annotations
ggplot(proportion_long, aes(x = Group, y = Percentage, fill = Category)) +
  geom_bar(stat = "identity", position = "stack") +
  geom_text(data = subset(proportion_long, Significant == "*"),
            aes(x = Group, y = 102, label = "*"),
            size = 6, color = "black", fontface = "bold") +
  scale_y_continuous(labels = scales::percent_format(scale = 1), limits = c(0, 105)) +
  scale_fill_manual(values = c("Yes" = "#e41a1c", "No" = "#377eb8")) +
  labs(
    title = "Proportion of Macrophage Genes in\n0.5 Corrmotif Response Groups",
    x = "Group",
    y = "Percentage",
    fill = "Category"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = rel(1.5), hjust = 0.5),
    axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
    legend.title = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1.2),
    strip.background = element_blank(),
    strip.text = element_text(size = 12, face = "bold")
  )
```


