---
title: "Venn Diagram Overlap of DE Genes"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## **📌 overlap of DE genes**

## **📌 Read and Process DEG Data**
```{r load_DEGs, echo=TRUE, results='hide', message=FALSE, warning=FALSE}
# Load DEGs Data
CX_0.1_3 <- read.csv("data/DEGs/Toptable_CX_0.1_3.csv")
CX_0.1_24 <- read.csv("data/DEGs/Toptable_CX_0.1_24.csv")
CX_0.1_48 <- read.csv("data/DEGs/Toptable_CX_0.1_48.csv")
CX_0.5_3 <- read.csv("data/DEGs/Toptable_CX_0.5_3.csv")
CX_0.5_24 <- read.csv("data/DEGs/Toptable_CX_0.5_24.csv")
CX_0.5_48 <- read.csv("data/DEGs/Toptable_CX_0.5_48.csv")

DOX_0.1_3 <- read.csv("data/DEGs/Toptable_DOX_0.1_3.csv")
DOX_0.1_24 <- read.csv("data/DEGs/Toptable_DOX_0.1_24.csv")
DOX_0.1_48 <- read.csv("data/DEGs/Toptable_DOX_0.1_48.csv")
DOX_0.5_3 <- read.csv("data/DEGs/Toptable_DOX_0.5_3.csv")
DOX_0.5_24 <- read.csv("data/DEGs/Toptable_DOX_0.5_24.csv")
DOX_0.5_48 <- read.csv("data/DEGs/Toptable_DOX_0.5_48.csv")

# Extract Significant DEGs
DEG1 <- CX_0.1_3$Entrez_ID[CX_0.1_3$adj.P.Val < 0.05]
DEG2 <- CX_0.1_24$Entrez_ID[CX_0.1_24$adj.P.Val < 0.05]
DEG3 <- CX_0.1_48$Entrez_ID[CX_0.1_48$adj.P.Val < 0.05]
DEG4 <- CX_0.5_3$Entrez_ID[CX_0.5_3$adj.P.Val < 0.05]
DEG5 <- CX_0.5_24$Entrez_ID[CX_0.5_24$adj.P.Val < 0.05]
DEG6 <- CX_0.5_48$Entrez_ID[CX_0.5_48$adj.P.Val < 0.05]

DEG7 <- DOX_0.1_3$Entrez_ID[DOX_0.1_3$adj.P.Val < 0.05]
DEG8 <- DOX_0.1_24$Entrez_ID[DOX_0.1_24$adj.P.Val < 0.05]
DEG9 <- DOX_0.1_48$Entrez_ID[DOX_0.1_48$adj.P.Val < 0.05]
DEG10 <- DOX_0.5_3$Entrez_ID[DOX_0.5_3$adj.P.Val < 0.05]
DEG11 <- DOX_0.5_24$Entrez_ID[DOX_0.5_24$adj.P.Val < 0.05]
DEG12 <- DOX_0.5_48$Entrez_ID[DOX_0.5_48$adj.P.Val < 0.05]
```

## **📌 Load Libraries**
```{r load_lib, echo=TRUE, results='hide', message=FALSE, warning=FALSE}
library(ggplot2)
library(ggVennDiagram)
library(UpSetR)
```

## **📌 Across Drugs**

## **📌Venn Diagram: CX-5461 vs VEH (0.1 µM)**
```{r venn1, echo=TRUE, results='hide', message=FALSE, fig.width=10, fig.height=8}
venntest <- list(DEG1, DEG2, DEG3)
ggVennDiagram(
  venntest,
  category.names = c("CX_0.1_3", "CX_0.1_24", "CX_0.1_48"),
  fill = c("red", "blue", "green")
) + ggtitle("CX-5461 Vs VEH 0.1 micromolar")+
  theme(
    plot.title = element_text(size = 16, face = "bold"),  # Increase title size
    text = element_text(size = 16)  # Increase text size globally
  )

```

## **📌Venn Diagram: CX-5461 vs VEH (0.5 µM)**
```{r venn2, echo=TRUE, results='hide', message=FALSE, fig.width=10, fig.height=8}
venntest1 <- list(DEG4, DEG5, DEG6)
ggVennDiagram(
  venntest1,
  category.names = c("CX_0.5_3", "CX_0.5_24", "CX_0.5_48"),
  fill = c("red", "blue", "green")
) + ggtitle("CX-5461 Vs VEH 0.5 micromolar")+
  theme(
    plot.title = element_text(size = 16, face = "bold"),  # Increase title size
    text = element_text(size = 16)  # Increase text size globally
  )
```

## **📌Venn Diagram: DOX vs VEH (0.1 µM)**
```{r venn3, echo=TRUE, results='hide', message=FALSE, fig.width=10, fig.height=8}
venntest2 <- list(DEG7, DEG8, DEG9)
ggVennDiagram(
  venntest2,
  category.names = c("DOX_0.1_3", "DOX_0.1_24", "DOX_0.1_48"),
  fill = c("red", "blue", "green")
) + ggtitle("DOX Vs VEH 0.1 micromolar")+
  theme(
    plot.title = element_text(size = 16, face = "bold"),  # Increase title size
    text = element_text(size = 16)  # Increase text size globally
  )
```

## **📌Venn Diagram: DOX vs VEH (0.5 µM)**
```{r venn4, echo=TRUE, results='hide', message=FALSE, fig.width=10, fig.height=8}
venntest3 <- list(DEG10, DEG11, DEG12)
ggVennDiagram(
  venntest3,
  category.names = c("DOX_0.5_3", "DOX_0.5_24", "DOX_0.5_48"),
  fill = c("red", "blue", "green")
) + ggtitle("DOX Vs VEH 0.5 micromolar")+
  theme(
    plot.title = element_text(size = 16, face = "bold"),  # Increase title size
    text = element_text(size = 16)  # Increase text size globally
  )
```


## **📌 Across Concentrations**

## **📌Venn Diagram: 0.1 Micromolar**
```{r venn5, echo=TRUE, results='hide', message=FALSE, fig.width=10, fig.height=8}
venntest7 <- list(DEG1, DEG2, DEG3, DEG7, DEG8, DEG9)
ggVennDiagram(
  venntest7, label = "count",
  category.names = c("CX_0.1_3", "CX_0.1_24", "CX_0.1_48", "DOX_0.1_3", "DOX_0.1_24", "DOX_0.1_48")
) + ggtitle("0.1 micromolar")+
  theme(
    plot.title = element_text(size = 16, face = "bold"),  # Increase title size
    text = element_text(size = 16)  # Increase text size globally
  )
```


## **📌Venn Diagram: 0.5 Micromolar**
```{r venn6, echo=TRUE, results='hide', message=FALSE, fig.width=10, fig.height=8}
venntest8 <- list(DEG4, DEG5, DEG6, DEG10, DEG11, DEG12)
ggVennDiagram(
  venntest8, label = "count",
  category.names = c("CX_0.5_3", "CX_0.5_24", "CX_0.5_48", "DOX_0.5_3", "DOX_0.5_24", "DOX_0.5_48")
) + ggtitle("0.5 micromolar")+
  theme(
    plot.title = element_text(size = 16, face = "bold"),  # Increase title size
    text = element_text(size = 16)  # Increase text size globally
  )
```


## **📌 Across Timepoints**


## **📌Venn Diagram: 3-hour Timepoint**
```{r venn7, echo=TRUE, results='hide', message=FALSE, fig.width=10, fig.height=8}
venntest4 <- list(DEG1, DEG4, DEG7, DEG10)
ggVennDiagram(
  venntest4, label_percent_digit = 2,
  category.names = c("CX_0.1_3", "CX_0.5_3", "DOX_0.1_3", "DOX_0.5_3")
) + ggtitle("3hr")+
  theme(
    plot.title = element_text(size = 16, face = "bold"),  # Increase title size
    text = element_text(size = 16)  # Increase text size globally
  )
```


## **📌Venn Diagram: 24-hour Timepoint**
```{r venn8, echo=TRUE, results='hide', message=FALSE, fig.width=10, fig.height=8}
venntest5 <- list(DEG2, DEG5, DEG8, DEG11)
ggVennDiagram(
  venntest5, label_percent_digit = 2,
  category.names = c("CX_0.1_24", "CX_0.5_24", "DOX_0.1_24", "DOX_0.5_24")
) + ggtitle("24hr")+
  theme(
    plot.title = element_text(size = 16, face = "bold"),  # Increase title size
    text = element_text(size = 16)  # Increase text size globally
  )
```


## **📌Venn Diagram: 48-hour Timepoint**
```{r venn9, echo=TRUE, results='hide', message=FALSE, fig.width=10, fig.height=8}
venntest6 <- list(DEG3, DEG6, DEG9, DEG12)
ggVennDiagram(
  venntest6, label_percent_digit = 2,
  category.names = c("CX_0.1_48", "CX_0.5_48", "DOX_0.1_48", "DOX_0.5_48")
) + ggtitle("48hr")+
  theme(
    plot.title = element_text(size = 16, face = "bold"),  # Increase title size
    text = element_text(size = 16)  # Increase text size globally
  )
```


## **📌 Overlap of DEGs (Upset Plot)**

## **📌 Loading data**
```{r upset1, echo=TRUE, results='hide', message=FALSE}
# Extract Significant DEGs
# Create a list of DEGs for each sample
DEG_list <- list(
  CX_0.1_3 = CX_0.1_3$Entrez_ID[CX_0.1_3$adj.P.Val < 0.05],
  CX_0.1_24 = CX_0.1_24$Entrez_ID[CX_0.1_24$adj.P.Val < 0.05],
  CX_0.1_48 = CX_0.1_48$Entrez_ID[CX_0.1_48$adj.P.Val < 0.05],
  CX_0.5_3 = CX_0.5_3$Entrez_ID[CX_0.5_3$adj.P.Val < 0.05],
  CX_0.5_24 = CX_0.5_24$Entrez_ID[CX_0.5_24$adj.P.Val < 0.05],
  CX_0.5_48 = CX_0.5_48$Entrez_ID[CX_0.5_48$adj.P.Val < 0.05],
  DOX_0.1_3 = DOX_0.1_3$Entrez_ID[DOX_0.1_3$adj.P.Val < 0.05],
  DOX_0.1_24 = DOX_0.1_24$Entrez_ID[DOX_0.1_24$adj.P.Val < 0.05],
  DOX_0.1_48 = DOX_0.1_48$Entrez_ID[DOX_0.1_48$adj.P.Val < 0.05],
  DOX_0.5_3 = DOX_0.5_3$Entrez_ID[DOX_0.5_3$adj.P.Val < 0.05],
  DOX_0.5_24 = DOX_0.5_24$Entrez_ID[DOX_0.5_24$adj.P.Val < 0.05],
  DOX_0.5_48 = DOX_0.5_48$Entrez_ID[DOX_0.5_48$adj.P.Val < 0.05]
)

# Convert list to binary matrix
DEG_matrix <- fromList(DEG_list)

# Define order of sets
set_order <- names(DEG_list)
```

## **📌 UpSet Plot of DEGs Across Samples\n(Show all intersections till lowest size 5)**
```{r upset2, echo=TRUE, message=FALSE, fig.width=12, fig.height=8}
upset(
  DEG_matrix,
  sets = set_order,  # Specify the exact order of sets
  order.by = "freq",  # Order intersections by frequency
  main.bar.color = "blue",  # Color for the intersection bars
  matrix.color = "black",  # Color for matrix dots
  sets.bar.color = rainbow(length(DEG_list)),  # Assign different colors to set size bars
  keep.order = TRUE,  # Keep the specified order of sets
  number.angles = 0,  # Angle of numbers in intersection size bars
  point.size = 2.5,  # Size of points in the matrix
  text.scale = 1,  # Scale for text elements
  show.numbers = "yes"  # Show intersection size numbers directly
)
```

## **📌 UpSet Plot of DEGs Across Samples\n(Show all intersections till last lowest size)**
```{r upset3, echo=TRUE, message=FALSE,fig.width=16, fig.height=9}
upset(
  DEG_matrix,
  sets = set_order,  # Specify the exact order of sets
  order.by = "freq",  # Order intersections by frequency
  main.bar.color = "blue",  # Color for the intersection bars
  matrix.color = "black",  # Color for matrix dots
  sets.bar.color = rainbow(length(DEG_list)),  # Assign different colors to set size bars
  keep.order = TRUE,  # Keep the specified order of sets
  number.angles = 0,  # Angle of numbers in intersection size bars
  point.size = 2.5,  # Size of points in the matrix
  text.scale = 1,  # Scale for text elements
  show.numbers = "yes",  # Show intersection size numbers directly
  nintersects = NA  # Show all intersections including those with lowest size
)
```

## **📌 Drug Specific DE genes**

## **📌 UpSet Plot: CX-5461 vs DOX**
```{r upset4, echo=TRUE, message=FALSE}
# Combine DEGs under CX-5461 and DOX
CX_DEGs <- unique(unlist(DEG_list[1:6]))
DOX_DEGs <- unique(unlist(DEG_list[7:12]))

# Create binary matrix for drug-specific DEGs
DEG_matrix_drug <- fromList(list(CX_5461 = CX_DEGs, DOX = DOX_DEGs))

# Generate the UpSet plot for drugs
upset(
  DEG_matrix_drug,
  sets = c("CX_5461", "DOX"),
  order.by = "freq",
  main.bar.color = "darkgreen",
  point.size = 3,
  text.scale = 1.5,
  matrix.color = "purple",
  sets.bar.color = c("blue", "red")
)
```

## **📌 Conc. Specific DE genes**
## **📌 UpSet Plot: Concentration-Specific DEGs**
```{r upset5, echo=TRUE, message=FALSE}
# Combine DEGs under concentrations 0.1 and 0.5
DEG_0.1 <- unique(unlist(DEG_list[c(1, 2, 3, 7, 8, 9)]))
DEG_0.5 <- unique(unlist(DEG_list[c(4, 5, 6, 10, 11, 12)]))

# Create binary matrix for concentration-specific DEGs
DEG_matrix_concentration <- fromList(list(Concentration_0.1 = DEG_0.1, Concentration_0.5 = DEG_0.5))

# Generate the UpSet plot for concentration
upset(
  DEG_matrix_concentration,
  sets = c("Concentration_0.1", "Concentration_0.5"),
  order.by = "freq",
  main.bar.color = "darkorange",
  matrix.color = "darkblue",
  sets.bar.color = c("cyan", "magenta"),
  text.scale = 1.5,
  keep.order = TRUE
)
```


## **📌 Timepoints Specific DE genes**
## **📌 UpSet Plot: Timepoint-Specific DEGs**
```{r upset6, echo=TRUE, message=FALSE}
# Combine DEGs under timepoints 3hr, 24hr, and 48hr
DEG_3hr <- unique(unlist(DEG_list[c(1, 4, 7, 10)]))
DEG_24hr <- unique(unlist(DEG_list[c(2, 5, 8, 11)]))
DEG_48hr <- unique(unlist(DEG_list[c(3, 6, 9, 12)]))

# Create binary matrix for timepoint-specific DEGs
DEG_matrix_timepoint <- fromList(list(Timepoint_3hr = DEG_3hr, Timepoint_24hr = DEG_24hr, Timepoint_48hr = DEG_48hr))

# Generate the UpSet plot for timepoints
upset(
  DEG_matrix_timepoint,
  sets = c("Timepoint_3hr", "Timepoint_24hr", "Timepoint_48hr"),
  order.by = "freq",
  main.bar.color = "darkgreen",
  matrix.color = "darkred",
  sets.bar.color = c("blue", "orange", "purple"),
  text.scale = 1.5,
  keep.order = TRUE
)
```


