---
title: "Overlap of GO Terms in DEGs"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## **ðŸ“Œ Load Required Libraries**
```{r load_libraries, echo=TRUE, results='hide',message=FALSE, warning=FALSE}
library(ggVennDiagram)
library(ggplot2)
library(dplyr)
```

## **ðŸ“Œ Read and Process GO Terms**
```{r load_GO, echo=TRUE, results='hide',message=FALSE, warning=FALSE}
# Load GO term data for different conditions
# Load GO term data for different conditions
CX_0.1_24_GO <- read.csv("data/CX_0.1_24 (Combined).csv")
CX_0.1_48_GO <- read.csv("data/CX_0.1_48 (Combined).csv")
CX_0.5_3_GO  <- read.csv("data/CX_0.5_3 (Combined).csv")
CX_0.5_24_GO <- read.csv("data/CX_0.5_24 (Combined).csv")
CX_0.5_48_GO <- read.csv("data/CX_0.5_48 (Combined).csv")

DOX_0.1_3_GO  <- read.csv("data/DOX_0.1_3 (Combined).csv")
DOX_0.1_24_GO <- read.csv("data/DOX_0.1_24 (Combined).csv")
DOX_0.1_48_GO <- read.csv("data/DOX_0.1_48 (Combined).csv")
DOX_0.5_3_GO  <- read.csv("data/DOX_0.5_3 (Combined).csv")
DOX_0.5_24_GO <- read.csv("data/DOX_0.5_24 (Combined).csv")
DOX_0.5_48_GO <- read.csv("data/DOX_0.5_48 (Combined).csv")

# Extract GO term IDs for each condition
DEG2_GO <- CX_0.1_24_GO$ID
DEG3_GO <- CX_0.1_48_GO$ID
DEG4_GO <- CX_0.5_3_GO$ID
DEG5_GO <- CX_0.5_24_GO$ID
DEG6_GO <- CX_0.5_48_GO$ID
DEG7_GO <- DOX_0.1_3_GO$ID
DEG8_GO <- DOX_0.1_24_GO$ID
DEG9_GO <- DOX_0.1_48_GO$ID
DEG10_GO <- DOX_0.5_3_GO$ID
DEG11_GO <- DOX_0.5_24_GO$ID
DEG12_GO <- DOX_0.5_48_GO$term_id
```

## **ðŸ“Œ Overlap of GO Terms across the drugs**

## **ðŸ“Œ Overlap of GO Terms in CX-5461 vs VEH**
```{r Venn1, echo=TRUE, message=FALSE, warning=FALSE, fig.width=10, fig.height=8}
CX_Datasets <- list(
  "CX_0.1_24" = DEG2_GO,
  "CX_0.1_48" = DEG3_GO,
  "CX_0.5_3" = DEG4_GO,
  "CX_0.5_24" = DEG5_GO,
  "CX_0.5_48" = DEG6_GO
)

ggVennDiagram(CX_Datasets, label = "count") +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 10)
  ) +
  labs(title = "CX-5461 GO Term Overlap")
```

## **ðŸ“Œ Overlap of GO Terms in DOX vs VEH**
```{r Venn2, echo=TRUE, message=FALSE, warning=FALSE, fig.width=10, fig.height=8}
DOX_Datasets <- list(
  "DOX_0.1_3" = DEG7_GO,
  "DOX_0.1_24" = DEG8_GO,
  "DOX_0.1_48" = DEG9_GO,
  "DOX_0.5_3" = DEG10_GO,
  "DOX_0.5_24" = DEG11_GO,
  "DOX_0.5_48" = DEG12_GO
)

ggVennDiagram(DOX_Datasets, label = "count") +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 10)
  ) +
  labs(title = "DOX GO Term Overlap")
```

## **ðŸ“Œ Overlap of GO Terms across concentrations**
## **ðŸ“Œ 0.1 ÂµM Concentration**
```{r Venn3, echo=TRUE, message=FALSE, warning=FALSE, fig.width=10, fig.height=8}
Concentration_0_1 <- list(
  "CX_0.1_24" = DEG2_GO,
  "CX_0.1_48" = DEG3_GO,
  "DOX_0.1_3" = DEG7_GO,
  "DOX_0.1_24" = DEG8_GO,
  "DOX_0.1_48" = DEG9_GO
)

ggVennDiagram(Concentration_0_1, label = "count") +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 10)
  ) +
  labs(title = "0.1 ÂµM GO Term Overlap")
```

## **ðŸ“Œ 0.5 ÂµM Concentration**
```{r Venn4, echo=TRUE, message=FALSE, warning=FALSE, fig.width=10, fig.height=8}
Concentration_0_5 <- list(
  "CX_0.5_3" = DEG4_GO,
  "CX_0.5_24" = DEG5_GO,
  "CX_0.5_48" = DEG6_GO,
  "DOX_0.5_3" = DEG10_GO,
  "DOX_0.5_24" = DEG11_GO,
  "DOX_0.5_48" = DEG12_GO
)

ggVennDiagram(Concentration_0_5, label = "count") +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 10)
  ) +
  labs(title = "0.5 ÂµM GO Term Overlap")
```

## **ðŸ“Œ Overlap of GO Terms Across Timepoints**
## **ðŸ“Œ3 Hours**
```{r Venn5, echo=TRUE, message=FALSE, warning=FALSE, fig.width=10, fig.height=8}
Timepoint_3hrs <- list(
  "CX_0.5_3" = DEG4_GO,
  "DOX_0.1_3" = DEG7_GO,
  "DOX_0.5_3" = DEG10_GO
)

ggVennDiagram(Timepoint_3hrs, label = "count") +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 10)
  ) +
  labs(title = "3 Hours GO Term Overlap")
```

## **ðŸ“Œ24 Hours**
```{r Venn6, echo=TRUE, message=FALSE, warning=FALSE, fig.width=10, fig.height=8}
Timepoint_24hrs <- list(
  "CX_0.1_24" = DEG2_GO,
  "CX_0.5_24" = DEG5_GO,
  "DOX_0.1_24" = DEG8_GO,
  "DOX_0.5_24" = DEG11_GO
)

ggVennDiagram(Timepoint_24hrs, label = "count") +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 10)
  ) +
  labs(title = "24 Hours GO Term Overlap")
```

## **ðŸ“Œ48 Hours**
```{r Venn7, echo=TRUE, message=FALSE, warning=FALSE, fig.width=10, fig.height=8}
Timepoint_48hrs <- list(
  "CX_0.1_48" = DEG3_GO,
  "CX_0.5_48" = DEG6_GO,
  "DOX_0.1_48" = DEG9_GO,
  "DOX_0.5_48" = DEG12_GO
)

ggVennDiagram(Timepoint_48hrs, label = "count") +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    legend.title = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 10)
  ) +
  labs(title = "48 Hours GO Term Overlap")
```

## **ðŸ“ŒOverlapping of GO terms across all samples Upset plot**
```{r Upset1, echo=TRUE, message=FALSE, warning=FALSE, fig.width=8, fig.height=6}

# ðŸ“¦ Load Required Libraries
library(UpSetR)
library(dplyr)
library(tidyr)

# âœ… Load all GO term datasets
CX_0.1_24_GO <- read.csv("data/BP/CX_0.1_24 (Combined).csv")
CX_0.1_48_GO <- read.csv("data/BP/CX_0.1_48 (Combined).csv")
CX_0.5_3_GO  <- read.csv("data/BP/CX_0.5_3 (Combined).csv")
CX_0.5_24_GO <- read.csv("data/BP/CX_0.5_24 (Combined).csv")
CX_0.5_48_GO <- read.csv("data/BP/CX_0.5_48 (Combined).csv")
DOX_0.1_3_GO <- read.csv("data/BP/DOX_0.1_3 (Combined).csv")
DOX_0.1_24_GO <- read.csv("data/BP/DOX_0.1_24 (Combined).csv")
DOX_0.1_48_GO <- read.csv("data/BP/DOX_0.1_48 (Combined).csv")
DOX_0.5_3_GO <- read.csv("data/BP/DOX_0.5_3 (Combined).csv")

# âœ… Prepare list of GO IDs
gene_sets <- list(
  `CX_0.1_24` = as.character(CX_0.1_24_GO$ID),
  `CX_0.1_48` = as.character(CX_0.1_48_GO$ID),
  `CX_0.5_3`  = as.character(CX_0.5_3_GO$ID),
  `CX_0.5_24` = as.character(CX_0.5_24_GO$ID),
  `CX_0.5_48` = as.character(CX_0.5_48_GO$ID),
  `DOX_0.1_3` = as.character(DOX_0.1_3_GO$ID),
  `DOX_0.1_24` = as.character(DOX_0.1_24_GO$ID),
  `DOX_0.1_48` = as.character(DOX_0.1_48_GO$ID),
  `DOX_0.5_3` = as.character(DOX_0.5_3_GO$ID)
)

# ðŸ§® Build binary matrix
all_ids <- unique(unlist(gene_sets))
binary_matrix <- data.frame(GO_ID = all_ids)

for (name in names(gene_sets)) {
  binary_matrix[[name]] <- as.integer(binary_matrix$GO_ID %in% gene_sets[[name]])
}

# âœ… Remove GO_ID for upset input
upset_input <- binary_matrix[, -1]
colnames(upset_input) <- names(gene_sets)  # ensure consistent names

# âœ… Plot
par(mar = c(10, 4, 2, 2))  # bottom, left, top, right
upset(upset_input,
      sets = colnames(upset_input),
      order.by = "freq",
      sets.bar.color = "#56B4E9",
      mainbar.y.label = "Number of Shared BP",
      sets.x.label = "BP per Condition",
      text.scale = 1.2,
      nintersects = 30)
```

## **ðŸ“ŒIdentifying Unique GO terms**
```{r Unique, echo=TRUE, message=FALSE, warning=FALSE}
# ðŸ“¦ Load Required Libraries
library(dplyr)
library(readr)

# âœ… Step 1: Load GO Enrichment Files
CX_0.1_24_GO <- read_csv("data/BP/CX_0.1_24 (Combined).csv")
CX_0.1_48_GO <- read_csv("data/BP/CX_0.1_48 (Combined).csv")
CX_0.5_3_GO  <- read_csv("data/BP/CX_0.5_3 (Combined).csv")
CX_0.5_24_GO <- read_csv("data/BP/CX_0.5_24 (Combined).csv")
CX_0.5_48_GO <- read_csv("data/BP/CX_0.5_48 (Combined).csv")
DOX_0.1_3_GO <- read_csv("data/BP/DOX_0.1_3 (Combined).csv")
DOX_0.1_24_GO <- read_csv("data/BP/DOX_0.1_24 (Combined).csv")
DOX_0.1_48_GO <- read_csv("data/BP/DOX_0.1_48 (Combined).csv")
DOX_0.5_3_GO <- read_csv("data/BP/DOX_0.5_3 (Combined).csv")

# âœ… Step 2: Create Named List of Data Frames
go_files <- list(
  `CX_0.1_24` = CX_0.1_24_GO,
  `CX_0.1_48` = CX_0.1_48_GO,
  `CX_0.5_3`  = CX_0.5_3_GO,
  `CX_0.5_24` = CX_0.5_24_GO,
  `CX_0.5_48` = CX_0.5_48_GO,
  `DOX_0.1_3` = DOX_0.1_3_GO,
  `DOX_0.1_24` = DOX_0.1_24_GO,
  `DOX_0.1_48` = DOX_0.1_48_GO,
  `DOX_0.5_3` = DOX_0.5_3_GO
)

# âœ… Step 3: Find Unique GO IDs for Each Condition
unique_go_terms <- list()

for (set_name in names(go_files)) {
  current_ids <- go_files[[set_name]]$ID
  other_ids <- unlist(go_files[names(go_files) != set_name], use.names = FALSE)
  unique_ids <- setdiff(current_ids, other_ids)
  
  if (length(unique_ids) > 0) {
    unique_go_terms[[set_name]] <- unique_ids
  }
}

# âœ… Step 4: Map Unique GO IDs to Descriptions and p.adjust
mapped_unique_go_terms <- list()

for (set_name in names(unique_go_terms)) {
  source_df <- go_files[[set_name]]
  unique_ids <- unique_go_terms[[set_name]]
  
  mapped_df <- source_df %>%
    filter(ID %in% unique_ids) %>%
    dplyr::select(GO_ID = ID, Function = Description, p.adjust)
  
  mapped_unique_go_terms[[set_name]] <- mapped_df
}

# ðŸŽ¯ Final Output: List of data.frames with GO_ID, Description, and p.adjust
mapped_unique_go_terms
```

## **ðŸ“ŒTOP Unique GO terms**
```{r Unique1, echo=TRUE, message=FALSE, warning=FALSE, fig.width=12, fig.height=10}

# ðŸ“¦ Load Required Libraries
library(dplyr)
library(ggplot2)
library(ggpubr)
library(stringr)

# ðŸ§¾ mapped_unique_go_terms must already be defined from your previous step:
# List of data frames with GO_ID, Function, p.adjust

# ðŸ” Step 1: Prepare top 10 unique terms per sample
plot_data <- list()

for (set_name in names(mapped_unique_go_terms)) {
  unique_df <- mapped_unique_go_terms[[set_name]]
  
  if (nrow(unique_df) > 0) {
    unique_df$Sample <- set_name
    unique_df$NegLog10Padj <- -log10(unique_df$p.adjust)
    
    top10 <- unique_df %>%
      dplyr::slice_min(order_by = p.adjust, n = 10, with_ties = FALSE) %>%
      dplyr::select(Sample, Function, NegLog10Padj)
    
    plot_data[[set_name]] <- top10
  }
}

# ðŸ”— Step 2: Combine all and format
plot_df <- bind_rows(plot_data)
plot_df$Function <- str_trunc(plot_df$Function, 60)

# ðŸŽ¨ Step 3: Plot
ggplot(plot_df, aes(x = NegLog10Padj, y = reorder(Function, NegLog10Padj))) +
  geom_bar(stat = "identity", fill = "#3182bd") +
  facet_wrap(~ Sample, scales = "free_y", ncol = 2, strip.position = "top") +
  labs(x = "-log10(p.adjust)", y = "GO Term", title = "Top Unique GO BP Terms per Sample") +
  theme_pubr(base_size = 14) +
  theme(strip.background = element_rect(colour = "black", fill = "grey90", size = 1),
        strip.text = element_text(face = "bold"),
        axis.text.y = element_text(size = 10),
        panel.border = element_rect(color = "black", fill = NA, size = 1))
```

