---
title: "PCA Analysis"
author: "Your Name"
date: "`r Sys.Date()`"
output:
  workflowr::wflow_html:
    toc: true
    toc_float: true
    number_sections: false
---

### **Load Required Libraries**
```{r setup, echo=TRUE,message=FALSE}
library(edgeR)
library(ggplot2)
library(reshape2)
library(dplyr)
library(Biobase)
library(limma)
library(tidyverse)
library(scales)
library(biomaRt)
library(ggrepel)
library(corrplot)
library(Hmisc)
library(org.Hs.eg.db)
library(AnnotationDbi)
library(tidyr)
library(ggfortify)
```

📍 Load the Count Matrix CSV file

```{r Read mapping counts, echo=FALSE, include=FALSE}
# Load the count matrix from the TSV file
counts_matrix <- read.csv("data/counts_matrix.csv", header=TRUE, check.names=FALSE)

# Compute log2 Counts Per Million (CPM)
cpm <- cpm(counts_matrix)
lcpm <- cpm(counts_matrix, log=TRUE)  # Log2 CPM matrix

# Apply filtering thresholds
filcpm_matrix <- subset(lcpm, rowMeans(lcpm) > 0)
filcpm_matrix1 <- subset(lcpm, rowMeans(lcpm) > 0.5)
filcpm_matrix2 <- subset(lcpm, rowMeans(lcpm) > 1)
```


## **📌Color palettes**

```{r color_palettes, echo=TRUE, message=FALSE}
drug_palc <- c("#8B006D","#DF707E","#F1B72B", "#3386DD","#707031","#41B333")
drug_palc1 <- c("#8B006D","#F1B72B", "#3386DD","#707031")
drug_palc2 <- c("#8B006D","#F1B72B", "#3386DD")
```

## **📌Load Metadata**

```{r load_metadata, echo=FALSE, include=FALSE}
Metadata <- read.csv("data/Metadata.csv")
dim(Metadata)
head(Metadata)
Metadata$Time <- as.character(Metadata$Time)
Metadata$Ind <- as.character(Metadata$Ind)
Indiv <- as.character(Metadata$Ind)
# Convert to matrix
matrix <- as.matrix(lcpm)
```


## **📌PCA of Unfiltered log2(CPM)**

```{r pca_unfiltered, echo=TRUE, message=FALSE}
prcomp_res <- prcomp(t(lcpm %>% as.matrix()), center = TRUE)

ggplot2::autoplot(prcomp_res, data = Metadata, colour = "Condition", shape = "Time", size =4, x=1, y=2) +
  ggrepel::geom_text_repel(label=Indiv) +
  scale_color_manual(values=drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Unfiltered")) +
  theme_bw()
```


```{r pca_unfiltered2, echo=TRUE, message=FALSE}
prcomp_res <- prcomp(t(lcpm %>% as.matrix()), center = TRUE)

ggplot2::autoplot(prcomp_res, data = Metadata, colour = "Condition", shape = "Time", size =4, x=2, y=3) +
  ggrepel::geom_text_repel(label=Indiv) +
  scale_color_manual(values=drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Unfiltered")) +
  theme_bw()
```


```{r pca_unfiltered3, echo=TRUE, message=FALSE}
prcomp_res <- prcomp(t(lcpm %>% as.matrix()), center = TRUE)

ggplot2::autoplot(prcomp_res, data = Metadata, colour = "Condition", shape = "Time", size =4, x=3, y=4) +
  ggrepel::geom_text_repel(label=Indiv) +
  scale_color_manual(values=drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Unfiltered")) +
  theme_bw()
```


## **📌PCA of Filtered log2(CPM) (RowMeans > 0)**

```{r pca_filtered_0, echo=TRUE, message=FALSE}
prcomp_res1 <- prcomp(t(filcpm_matrix %>% as.matrix()), center = TRUE)

ggplot2::autoplot(prcomp_res1, data = Metadata, colour = "Condition", shape = "Time", size =4, x=1, y=2) +
  ggrepel::geom_text_repel(label=Indiv) +
  scale_color_manual(values=drug_palc) +
  ggtitle(expression("PCA of gene expression (log2 cpm)")) +
  theme_bw()
```

```{r pca_filtered_0_1, echo=TRUE, message=FALSE}
prcomp_res1 <- prcomp(t(filcpm_matrix %>% as.matrix()), center = TRUE)

ggplot2::autoplot(prcomp_res1, data = Metadata, colour = "Condition", shape = "Time", size =4, x=2, y=3) +
  ggrepel::geom_text_repel(label=Indiv) +
  scale_color_manual(values=drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) filtered (Rowmeans >0)")) +
  theme_bw()
```

```{r pca_filtered_0_2, echo=TRUE, message=FALSE}
prcomp_res1 <- prcomp(t(filcpm_matrix %>% as.matrix()), center = TRUE)

ggplot2::autoplot(prcomp_res1, data = Metadata, colour = "Condition", shape = "Time", size =4, x=3, y=4) +
  ggrepel::geom_text_repel(label=Indiv) +
  scale_color_manual(values=drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) filtered (Rowmeans >0)")) +
  theme_bw()
```

## **📌PCA of Filtered log2(CPM) (RowMeans > 0.5)**

```{r pca_filtered_0.5, echo=TRUE, message=FALSE}
prcomp_res2 <- prcomp(t(filcpm_matrix1 %>% as.matrix()), center = TRUE)

ggplot2::autoplot(prcomp_res2, data = Metadata, colour = "Condition", shape = "Time", size =4, x=1, y=2) +
  ggrepel::geom_text_repel(label=Indiv) +
  scale_color_manual(values=drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) filtered (Rowmeans >0.5)")) +
  theme_bw()
```


```{r pca_filtered_0.5_1, echo=TRUE, message=FALSE}
prcomp_res2 <- prcomp(t(filcpm_matrix1 %>% as.matrix()), center = TRUE)

ggplot2::autoplot(prcomp_res2, data = Metadata, colour = "Condition", shape = "Time", size =4, x=2, y=3) +
  ggrepel::geom_text_repel(label=Indiv) +
  scale_color_manual(values=drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) filtered (Rowmeans >0.5)")) +
  theme_bw()
```


```{r pca_filtered_0.5_2, echo=TRUE, message=FALSE}
prcomp_res2 <- prcomp(t(filcpm_matrix1 %>% as.matrix()), center = TRUE)

ggplot2::autoplot(prcomp_res2, data = Metadata, colour = "Condition", shape = "Time", size =4, x=3, y=4) +
  ggrepel::geom_text_repel(label=Indiv) +
  scale_color_manual(values=drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) filtered (Rowmeans >0.5)")) +
  theme_bw()
```


## **📌PCA of Filtered log2(CPM) (RowMeans > 1)**

```{r pca_filtered_1, echo=TRUE, message=FALSE}
prcomp_res3 <- prcomp(t(filcpm_matrix2 %>% as.matrix()), center = TRUE)

ggplot2::autoplot(prcomp_res3, data = Metadata, colour = "Condition", shape = "Time", size =4, x=1, y=2) +
  ggrepel::geom_text_repel(label=Indiv) +
  scale_color_manual(values=drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) filtered (Rowmeans >1)")) +
  theme_bw()
```


```{r pca_filtered_1_1, echo=TRUE, message=FALSE}
prcomp_res3 <- prcomp(t(filcpm_matrix2 %>% as.matrix()), center = TRUE)

ggplot2::autoplot(prcomp_res3, data = Metadata, colour = "Condition", shape = "Time", size =4, x=2, y=3) +
  ggrepel::geom_text_repel(label=Indiv) +
  scale_color_manual(values=drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) filtered (Rowmeans >1)")) +
  theme_bw()
```


```{r pca_filtered_1_2, echo=TRUE, message=FALSE}
prcomp_res3 <- prcomp(t(filcpm_matrix2 %>% as.matrix()), center = TRUE)

ggplot2::autoplot(prcomp_res3, data = Metadata, colour = "Condition", shape = "Time", size =4, x=3, y=4) +
  ggrepel::geom_text_repel(label=Indiv) +
  scale_color_manual(values=drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) filtered (Rowmeans >1)")) +
  theme_bw()
```


## **📌 PCA Analysis by Drugs**

## **📌 PCA Analysis: CX-5461 & VEH **
```{r pca_drug, echo=TRUE, message=FALSE}
selected_columns <- grepl("VEH|CX.5461", colnames(matrix))
subset_matrix_CX <- matrix[, selected_columns]

subset_meta <- subset(Metadata, Metadata$Drug %in% c("VEH", "CX-5461"))

prcomp_res4 <- prcomp(t(subset_matrix_CX), center = TRUE)

ggplot2::autoplot(prcomp_res4, data = as.data.frame(subset_meta), colour = "Condition", shape = "Time", size = 4) +
  ggrepel::geom_text_repel(label = subset_meta$Ind) +  # ✅ Corrected label
  scale_color_manual(values = drug_palc1) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Unfiltered (CX-5461 vs VEH)")) +
  theme_bw()
```


```{r pca_drug_1, echo=TRUE, message=FALSE}
prcomp_res5 <- prcomp(t(subset_matrix_CX[rowMeans(subset_matrix_CX) > 0, ]), center = TRUE)

ggplot2::autoplot(prcomp_res5, data = as.data.frame(subset_meta), colour = "Condition", shape = "Time", size = 4) +
  ggrepel::geom_text_repel(label = subset_meta$Ind) +
  scale_color_manual(values = drug_palc1) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0 (CX-5461 vs VEH)")) +
  theme_bw()
```

```{r pca_drug_2, echo=TRUE, message=FALSE}
prcomp_res6 <- prcomp(t(subset_matrix_CX[rowMeans(subset_matrix_CX) > 0.5, ]), center = TRUE)

ggplot2::autoplot(prcomp_res6, data = as.data.frame(subset_meta), colour = "Condition", shape = "Time", size = 4) +
  ggrepel::geom_text_repel(label = subset_meta$Ind) +
  scale_color_manual(values = drug_palc1) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0.5 (CX-5461 vs VEH)")) +
  theme_bw()
```

```{r pca_drug_3, echo=TRUE, message=FALSE}
prcomp_res7 <- prcomp(t(subset_matrix_CX[rowMeans(subset_matrix_CX) > 1, ]), center = TRUE)

ggplot2::autoplot(prcomp_res7, data = as.data.frame(subset_meta), colour = "Condition", shape = "Time", size = 4) +
  ggrepel::geom_text_repel(label = subset_meta$Ind) +
  scale_color_manual(values = drug_palc1) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 1 (CX-5461 vs VEH)")) +
  theme_bw()
```


## **📌 PCA Analysis: DOX & VEH **

```{r pca_drug_dox_veh, echo=TRUE, message=FALSE}
selected_columns <- grepl("VEH|DOX", colnames(matrix))
subset_matrix_DOX <- matrix[, selected_columns]

subset_meta_dox <- subset(Metadata, Metadata$Drug %in% c("VEH", "DOX"))

prcomp_res8 <- prcomp(t(subset_matrix_DOX), center = TRUE)

ggplot2::autoplot(prcomp_res8, data = as.data.frame(subset_meta_dox), colour = "Condition", shape = "Time", size = 4) +
  ggrepel::geom_text_repel(label = subset_meta_dox$Ind) +
  scale_color_manual(values = drug_palc1) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Unfiltered (DOX vs VEH)")) +
  theme_bw()
```


```{r pca_drug_dox_veh_1, echo=TRUE, message=FALSE}
prcomp_res9 <- prcomp(t(subset_matrix_DOX[rowMeans(subset_matrix_DOX) > 0, ]), center = TRUE)

ggplot2::autoplot(prcomp_res9, data = as.data.frame(subset_meta_dox), colour = "Condition", shape = "Time", size = 4) +
  ggrepel::geom_text_repel(label = subset_meta_dox$Ind) +
  scale_color_manual(values = drug_palc1) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0 (DOX vs VEH)")) +
  theme_bw()
```


```{r pca_drug_dox_veh_2, echo=TRUE, message=FALSE}
prcomp_res10 <- prcomp(t(subset_matrix_DOX[rowMeans(subset_matrix_DOX) > 0.5, ]), center = TRUE)

ggplot2::autoplot(prcomp_res10, data = as.data.frame(subset_meta_dox), colour = "Condition", shape = "Time", size = 4) +
  ggrepel::geom_text_repel(label = subset_meta_dox$Ind) +
  scale_color_manual(values = drug_palc1) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0.5 (DOX vs VEH)")) +
  theme_bw()
```



```{r pca_drug_dox_veh_3, echo=TRUE, message=FALSE}
prcomp_res11 <- prcomp(t(subset_matrix_DOX[rowMeans(subset_matrix_DOX) > 1, ]), center = TRUE)

ggplot2::autoplot(prcomp_res11, data = as.data.frame(subset_meta_dox), colour = "Condition", shape = "Time", size = 4) +
  ggrepel::geom_text_repel(label = subset_meta_dox$Ind) +
  scale_color_manual(values = drug_palc1) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 1 (DOX vs VEH)")) +
  theme_bw()
```


## **📌 PCA Analysis: CX-5461 & DOX**

```{r pca_drug_cx_dox, echo=TRUE, message=FALSE}
selected_columns <- grepl("CX.5461|DOX", colnames(matrix))
subset_matrix_CX_DOX <- matrix[, selected_columns]

subset_meta_cx_dox <- subset(Metadata, Metadata$Drug %in% c("CX-5461", "DOX"))

prcomp_res12 <- prcomp(t(subset_matrix_CX_DOX), center = TRUE)

ggplot2::autoplot(prcomp_res12, data = as.data.frame(subset_meta_cx_dox), colour = "Condition", shape = "Time", size = 4) +
  ggrepel::geom_text_repel(label = subset_meta_cx_dox$Ind) +
  scale_color_manual(values = drug_palc1) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Unfiltered (CX-5461 vs DOX)")) +
  theme_bw()
```


```{r pca_drug_cx_dox_1, echo=TRUE, message=FALSE}
prcomp_res13 <- prcomp(t(subset_matrix_CX_DOX[rowMeans(subset_matrix_CX_DOX) > 0, ]), center = TRUE)

ggplot2::autoplot(prcomp_res13, data = as.data.frame(subset_meta_cx_dox), colour = "Condition", shape = "Time", size = 4) +
  ggrepel::geom_text_repel(label = subset_meta_cx_dox$Ind) +
  scale_color_manual(values = drug_palc1) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0 (CX-5461 vs DOX)")) +
  theme_bw()
```



```{r pca_drug_cx_dox_2, echo=TRUE, message=FALSE}
prcomp_res14 <- prcomp(t(subset_matrix_CX_DOX[rowMeans(subset_matrix_CX_DOX) > 0.5, ]), center = TRUE)

ggplot2::autoplot(prcomp_res14, data = as.data.frame(subset_meta_cx_dox), colour = "Condition", shape = "Time", size = 4) +
  ggrepel::geom_text_repel(label = subset_meta_cx_dox$Ind) +
  scale_color_manual(values = drug_palc1) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0.5 (CX-5461 vs DOX)")) +
  theme_bw()
```


```{r pca_drug_cx_dox_3, echo=TRUE, message=FALSE}
prcomp_res15 <- prcomp(t(subset_matrix_CX_DOX[rowMeans(subset_matrix_CX_DOX) > 1, ]), center = TRUE)

ggplot2::autoplot(prcomp_res15, data = as.data.frame(subset_meta_cx_dox), colour = "Condition", shape = "Time", size = 4) +
  ggrepel::geom_text_repel(label = subset_meta_cx_dox$Ind) +
  scale_color_manual(values = drug_palc1) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 1 (CX-5461 vs DOX)")) +
  theme_bw()
```

## **📌 PCA Analysis by Timepoints**

## **📌 3-Hour Timepoint**

```{r pca_3hr, echo=TRUE, message=FALSE}
selected_columns <- grepl("_3", colnames(matrix))
subset_matrix_3hr <- matrix[, selected_columns]

subset_meta_3hr <- subset(Metadata, Metadata$Time == 3)

prcomp_res_3hr <- prcomp(t(subset_matrix_3hr[rowMeans(subset_matrix_3hr) > 0, ]), center = TRUE)

ggplot2::autoplot(prcomp_res_3hr, data = as.data.frame(subset_meta_3hr), colour = "Condition", shape = "Drug", size = 4) +
  ggrepel::geom_text_repel(label = subset_meta_3hr$Ind) +
  scale_color_manual(values = drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0 (3 Hours)")) +
  theme_bw()
```

```{r pca_3hr_2, echo=TRUE, message=FALSE}
prcomp_res_3hr <- prcomp(t(subset_matrix_3hr[rowMeans(subset_matrix_3hr) > 0, ]), center = TRUE)

ggplot2::autoplot(prcomp_res_3hr, data = as.data.frame(subset_meta_3hr), colour = "Condition", shape = "Drug", size = 4, x=2, y=3) +
  ggrepel::geom_text_repel(label = subset_meta_3hr$Ind) +
  scale_color_manual(values = drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0 (3 Hours)")) +
  theme_bw()
```


```{r pca_3hr_3, echo=TRUE, message=FALSE}
prcomp_res_3hr <- prcomp(t(subset_matrix_3hr[rowMeans(subset_matrix_3hr) > 0, ]), center = TRUE)

ggplot2::autoplot(prcomp_res_3hr, data = as.data.frame(subset_meta_3hr), colour = "Condition", shape = "Drug", size = 4, x=3, y=4) +
  ggrepel::geom_text_repel(label = subset_meta_3hr$Ind) +
  scale_color_manual(values = drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0 (3 Hours)")) +
  theme_bw()
```


## **📌 24-Hour Timepoint**

```{r pca_24hr, echo=TRUE, message=FALSE}
selected_columns <- grepl("_24", colnames(matrix))
subset_matrix_24hr <- matrix[, selected_columns]

subset_meta_24hr <- subset(Metadata, Metadata$Time == 24)

prcomp_res_24hr <- prcomp(t(subset_matrix_24hr[rowMeans(subset_matrix_24hr) > 0, ]), center = TRUE)

ggplot2::autoplot(prcomp_res_24hr, data = as.data.frame(subset_meta_24hr), colour = "Condition", shape = "Drug", size = 4) +
  ggrepel::geom_text_repel(label = subset_meta_24hr$Ind) +
  scale_color_manual(values = drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0 (24 Hours)")) +
  theme_bw()
```


```{r pca_24hr_1, echo=TRUE, message=FALSE}
prcomp_res_24hr <- prcomp(t(subset_matrix_24hr[rowMeans(subset_matrix_24hr) > 0, ]), center = TRUE)

ggplot2::autoplot(prcomp_res_24hr, data = as.data.frame(subset_meta_24hr), colour = "Condition", shape = "Drug", size = 4, x=2, y=3) +
  ggrepel::geom_text_repel(label = subset_meta_24hr$Ind) +
  scale_color_manual(values = drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0 (24 Hours)")) +
  theme_bw()
```


```{r pca_24hr_2, echo=TRUE, message=FALSE}
prcomp_res_24hr <- prcomp(t(subset_matrix_24hr[rowMeans(subset_matrix_24hr) > 0, ]), center = TRUE)

ggplot2::autoplot(prcomp_res_24hr, data = as.data.frame(subset_meta_24hr), colour = "Condition", shape = "Drug", size = 4, x=3, y=4) +
  ggrepel::geom_text_repel(label = subset_meta_24hr$Ind) +
  scale_color_manual(values = drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0 (24 Hours)")) +
  theme_bw()
```



## **📌 48-Hour Timepoint**
```{r pca_48hr, echo=TRUE, message=FALSE}
selected_columns <- grepl("_48", colnames(matrix))
subset_matrix_48hr <- matrix[, selected_columns]
subset_meta_48hr <- subset(Metadata, Metadata$Time == 48)

prcomp_res_48hr_1 <- prcomp(t(subset_matrix_48hr[rowMeans(subset_matrix_48hr) > 0, ]), center = TRUE)

ggplot2::autoplot(prcomp_res_48hr_1, data = as.data.frame(subset_meta_48hr), colour = "Condition", shape = "Drug", size = 4) +
  ggrepel::geom_text_repel(label = subset_meta_48hr$Ind) +
  scale_color_manual(values = drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0 (48 Hours)")) +
  theme_bw()
```


```{r pca_48hr_1, echo=TRUE, message=FALSE}

prcomp_res_48hr_1 <- prcomp(t(subset_matrix_48hr[rowMeans(subset_matrix_48hr) > 0, ]), center = TRUE)

ggplot2::autoplot(prcomp_res_48hr_1, data = as.data.frame(subset_meta_48hr), colour = "Condition", shape = "Drug", size = 4, x=2, y=3) +
  ggrepel::geom_text_repel(label = subset_meta_48hr$Ind) +
  scale_color_manual(values = drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0 (48 Hours)")) +
  theme_bw()
```


```{r pca_48hr_2, echo=TRUE, message=FALSE}

prcomp_res_48hr_1 <- prcomp(t(subset_matrix_48hr[rowMeans(subset_matrix_48hr) > 0, ]), center = TRUE)

ggplot2::autoplot(prcomp_res_48hr_1, data = as.data.frame(subset_meta_48hr), colour = "Condition", shape = "Drug", size = 4, x=3, y=4) +
  ggrepel::geom_text_repel(label = subset_meta_48hr$Ind) +
  scale_color_manual(values = drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0 (48 Hours)")) +
  theme_bw()
```


## **📌 PCA Analysis by Concentrations**

## **📌0.1 µM Concentration**

```{r pca_0.1, echo=TRUE, message=FALSE}
selected_columns <- grepl("_0.1_", colnames(matrix))
subset_matrix_0.1 <- matrix[, selected_columns]

subset_meta_0.1 <- subset(Metadata, Metadata$Conc. == 0.1)

prcomp_res_0.1 <- prcomp(t(subset_matrix_0.1[rowMeans(subset_matrix_0.1) > 0, ]), center = TRUE)

ggplot2::autoplot(prcomp_res_0.1, data = as.data.frame(subset_meta_0.1), colour = "Drug", shape = "Time", size = 4) +
  ggrepel::geom_text_repel(label = subset_meta_0.1$Ind) +
  scale_color_manual(values = drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0 (0.1 µM)")) +
  theme_bw()
```


```{r pca_0.1_1, echo=TRUE, message=FALSE}
prcomp_res_0.1 <- prcomp(t(subset_matrix_0.1[rowMeans(subset_matrix_0.1) > 0, ]), center = TRUE)

ggplot2::autoplot(prcomp_res_0.1, data = as.data.frame(subset_meta_0.1), colour = "Drug", shape = "Time", size = 4, x=2, y=3) +
  ggrepel::geom_text_repel(label = subset_meta_0.1$Ind) +
  scale_color_manual(values = drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0 (0.1 µM)")) +
  theme_bw()
```


```{r pca_0.1_2, echo=TRUE, message=FALSE}
prcomp_res_0.1 <- prcomp(t(subset_matrix_0.1[rowMeans(subset_matrix_0.1) > 0, ]), center = TRUE)

ggplot2::autoplot(prcomp_res_0.1, data = as.data.frame(subset_meta_0.1), colour = "Drug", shape = "Time", size = 4, x=3, y=4) +
  ggrepel::geom_text_repel(label = subset_meta_0.1$Ind) +
  scale_color_manual(values = drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0 (0.1 µM)")) +
  theme_bw()
```


## **📌0.5 µM Concentration**
```{r pca_0.5, echo=TRUE, message=FALSE}
selected_columns <- grepl("_0.5_", colnames(matrix))
subset_matrix_0.5 <- matrix[, selected_columns]

subset_meta_0.5 <- subset(Metadata, Metadata$Conc. == 0.5)

prcomp_res_0.5 <- prcomp(t(subset_matrix_0.5[rowMeans(subset_matrix_0.5) > 0, ]), center = TRUE)

ggplot2::autoplot(prcomp_res_0.5, data = as.data.frame(subset_meta_0.5), colour = "Drug", shape = "Time", size = 4) +
  ggrepel::geom_text_repel(label = subset_meta_0.5$Ind) +
  scale_color_manual(values = drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0 (0.5 µM)")) +
  theme_bw()
```


```{r pca_0.5_1, echo=TRUE, message=FALSE}
prcomp_res_0.5 <- prcomp(t(subset_matrix_0.5[rowMeans(subset_matrix_0.5) > 0, ]), center = TRUE)

ggplot2::autoplot(prcomp_res_0.5, data = as.data.frame(subset_meta_0.5), colour = "Drug", shape = "Time", size = 4, x=2, y=3) +
  ggrepel::geom_text_repel(label = subset_meta_0.5$Ind) +
  scale_color_manual(values = drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0 (0.5 µM)")) +
  theme_bw()
```



```{r pca_0.5_2, echo=TRUE, message=FALSE}
prcomp_res_0.5 <- prcomp(t(subset_matrix_0.5[rowMeans(subset_matrix_0.5) > 0, ]), center = TRUE)

ggplot2::autoplot(prcomp_res_0.5, data = as.data.frame(subset_meta_0.5), colour = "Drug", shape = "Time", size = 4, x=3, y=4) +
  ggrepel::geom_text_repel(label = subset_meta_0.5$Ind) +
  scale_color_manual(values = drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0 (0.5 µM)")) +
  theme_bw()
```


