---
title: "PCA Analysis"
author: "Your Name"
date: "`r Sys.Date()`"
output:
  workflowr::wflow_html:
    toc: true
    toc_float: true
    number_sections: false
---

### **Load Required Libraries**
```{r setup, echo=TRUE,message=FALSE}
library(edgeR)
library(ggplot2)
library(reshape2)
library(dplyr)
library(Biobase)
library(limma)
library(tidyverse)
library(scales)
library(biomaRt)
library(ggrepel)
library(corrplot)
library(Hmisc)
library(org.Hs.eg.db)
library(AnnotationDbi)
library(tidyr)
library(ggfortify)
```

📍 Load the Count Matrix CSV file

```{r Read mapping counts, echo=FALSE, include=FALSE}
### 📍 Load the Count Matrix CSV file
counts_matrix <- read.csv("data/counts_matrix.csv", header = TRUE, check.names = FALSE)

# Compute log2 Counts Per Million (CPM)
cpm <- cpm(counts_matrix)
lcpm <- cpm(counts_matrix, log = TRUE)

# Apply filtering thresholds
filcpm_matrix  <- subset(lcpm, rowMeans(lcpm) > 0)
filcpm_matrix1 <- subset(lcpm, rowMeans(lcpm) > 0.5)
filcpm_matrix2 <- subset(lcpm, rowMeans(lcpm) > 1)
```


## **📌Color palettes**

```{r color_palettes, echo=TRUE, message=FALSE}
### 📌 Color palettes (updated)
drug_conc_palette <- c(
  "CX-5461_0.1" = "gold",  # light green
  "CX-5461_0.5" = "green4",  # dark green
  "DOX_0.1"     = "salmon2",  # peach
  "DOX_0.5"     = "red3",  # burnt orange
  "VEH_0.1"     = "lightblue3",  # sky blue
  "VEH_0.5"     = "darkblue"   # navy blue
)
drug_palc <- c("#8B006D","#DF707E","#F1B72B", "#3386DD","#707031","#41B333")
drug_palc1 <- c("#8B006D","#F1B72B", "#3386DD","#707031")
drug_palc2 <- c("#8B006D","#F1B72B", "#3386DD")
```

## **📌Load Metadata**

```{r load_metadata, echo=FALSE, include=FALSE}
Metadata <- read.csv("data/Metadata.csv")
dim(Metadata)
head(Metadata)

# Time relabeling
Metadata$Time <- factor(Metadata$Time, levels = c(3, 24, 48),
                        labels = c("3hr", "24hr", "48hr"))
Metadata$Ind <- as.character(Metadata$Ind)
Metadata$Drug <- as.character(Metadata$Drug)
Metadata$Conc <- as.character(Metadata$Conc)
Metadata$Drug_Conc <- paste(Metadata$Drug, Metadata$Conc, sep = "_")

Metadata$Indiv <- factor(Metadata$Ind, levels = c("75-1", "78-1", "87-1", "17-3", "84-1", "90-1"),
                         labels = c("1 (Female)", "2 (Female)", "3 (Female)", 
                                    "4 (Male)", "5 (Male)", "6 (Male)"))

Indiv <- Metadata$Ind
matrix <- as.matrix(lcpm)
```


## **📌PCA of Unfiltered log2(CPM)**

```{r pca_unfiltered, echo=TRUE, message=FALSE}
prcomp_res <- prcomp(t(matrix), center = TRUE)

ggplot2::autoplot(prcomp_res, data = Metadata,
                  colour = "Drug_Conc", shape = "Time", size = 4, x = 1, y = 2) +
  ggrepel::geom_text_repel(label = Indiv) +
  scale_color_manual(values = drug_conc_palette) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Unfiltered")) +
  theme_bw()
```


```{r pca_unfiltered2, echo=TRUE, message=FALSE}
prcomp_res <- prcomp(t(lcpm %>% as.matrix()), center = TRUE)

ggplot2::autoplot(prcomp_res, data = Metadata, colour = "Condition", shape = "Time", size =4, x=2, y=3) +
  ggrepel::geom_text_repel(label=Indiv) +
  scale_color_manual(values=drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Unfiltered")) +
  theme_bw()
```


```{r pca_unfiltered3, echo=TRUE, message=FALSE}
prcomp_res <- prcomp(t(lcpm %>% as.matrix()), center = TRUE)

ggplot2::autoplot(prcomp_res, data = Metadata, colour = "Condition", shape = "Time", size =4, x=3, y=4) +
  ggrepel::geom_text_repel(label=Indiv) +
  scale_color_manual(values=drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Unfiltered")) +
  theme_bw()
```


## **📌PCA of Filtered log2(CPM) (RowMeans > 0)**

```{r pca_filtered_0, echo=TRUE, message=FALSE}
prcomp_res1 <- prcomp(t(filcpm_matrix %>% as.matrix()), center = TRUE)

ggplot2::autoplot(prcomp_res1, data = Metadata, colour = "Drug_Conc", shape = "Time", size =4, x=1, y=2) +
  ggrepel::geom_text_repel(label=Indiv) +
  scale_color_manual(values=drug_conc_palette) +
  ggtitle(expression("PCA of gene expression (log2 cpm)")) +
  theme_bw()
```

```{r pca_filtered_0_1, echo=TRUE, message=FALSE}
prcomp_res1 <- prcomp(t(filcpm_matrix %>% as.matrix()), center = TRUE)

ggplot2::autoplot(prcomp_res1, data = Metadata, colour = "Drug_Conc", shape = "Time", size =4, x=2, y=3) +
  ggrepel::geom_text_repel(label=Indiv) +
  scale_color_manual(values=drug_conc_palette) +
  ggtitle(expression("PCA of log"[2]*"(cpm) filtered (Rowmeans >0)")) +
  theme_bw()
```

```{r pca_filtered_0_2, echo=TRUE, message=FALSE}
prcomp_res1 <- prcomp(t(filcpm_matrix %>% as.matrix()), center = TRUE)

ggplot2::autoplot(prcomp_res1, data = Metadata, colour = "Condition", shape = "Time", size =4, x=3, y=4) +
  ggrepel::geom_text_repel(label=Indiv) +
  scale_color_manual(values=drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) filtered (Rowmeans >0)")) +
  theme_bw()
```

## **📌PC1–PC3 Gene Expression Variance Across Individual, Drug, Concentration, and Timepoint log2(CPM) (RowMeans > 0)**
```{r pca_box, echo=TRUE, message=FALSE}

# Load Required Libraries
library(edgeR)
library(ggplot2)
library(dplyr)
library(tidyr)
library(ggrepel)
library(patchwork)

# Load Count Matrix
counts_matrix <- read.csv("data/counts_matrix.csv", header = TRUE, check.names = FALSE)
cpm <- cpm(counts_matrix)
lcpm <- cpm(counts_matrix, log = TRUE)
filcpm_matrix <- subset(lcpm, rowMeans(lcpm) > 0)
matrix <- as.matrix(filcpm_matrix)

# Load and clean Metadata
Metadata <- read.csv("data/Metadata.csv")
Metadata$Time <- factor(Metadata$Time, levels = c(3, 24, 48), labels = c("3hr", "24hr", "48hr"))
Metadata$Ind <- factor(Metadata$Ind, levels = 1:6, labels = as.character(1:6))
Metadata$Drug <- as.character(Metadata$Drug)
Metadata$`Conc.` <- factor(Metadata$`Conc.`, levels = c(0.1, 0.5))  # access using backticks

# PCA
prcomp_res <- prcomp(t(matrix), center = TRUE)
pca_df <- as.data.frame(prcomp_res$x[, 1:3])  # PC1–PC3
pca_df$Ind <- Metadata$Ind
pca_df$Drug <- Metadata$Drug
pca_df$Conc <- Metadata$`Conc.`
pca_df$Time <- Metadata$Time

# Renee-style p-value from linear model
get_regr_pval <- function(mod) {
  stopifnot(class(mod) == "lm")
  fstat <- summary(mod)$fstatistic
  pval <- 1 - pf(fstat[1], fstat[2], fstat[3])
  return(pval)
}

# Boxplot function with lm() p-value
plot_pc_box <- function(df, group_var, pc) {
  group_data <- df[[group_var]]
  n_groups <- length(unique(group_data))
  
  if (n_groups > 1) {
    model <- lm(df[[pc]] ~ group_data)
    pval <- get_regr_pval(model)
    pval_label <- paste0("p-value: ", signif(pval, 3))
  } else {
    pval_label <- "p-value: NA"
  }
  
  ggplot(df, aes_string(x = group_var, y = pc, fill = group_var)) +
    geom_boxplot(color = "black") +
    theme_bw(base_size = 11) +
    ylab(pc) + xlab(group_var) +
    ggtitle(paste(pc, "by", group_var), subtitle = pval_label) +
    theme(legend.position = "none",
          plot.title = element_text(face = "bold", size = 12),
          plot.subtitle = element_text(size = 10))
}

# 🔄 Generate all 12 plots: PC1–3 × Ind, Drug, Conc, Time
pcs <- c("PC1", "PC2", "PC3")
group_vars <- c("Ind", "Drug", "Conc", "Time")
plots <- list()

for (pc in pcs) {
  for (group in group_vars) {
    key <- paste(pc, group, sep = "_")
    
    base_plot <- plot_pc_box(pca_df, group, pc)
    
    if (pc == "PC1") {
      upper_limit <- max(pca_df[[pc]], na.rm = TRUE) * 1.1
      plots[[key]] <- base_plot +
        scale_y_continuous(limits = c(-60, upper_limit),
                           breaks = c(-60, -30, 0, 30, 60, 90, 120))
    } else {
      plots[[key]] <- base_plot
    }
  }
}

# Remove titles (we’ll add column headers instead)
plots <- lapply(plots, function(p) {
  p + theme(plot.title = element_blank())  # removes main title only, keeps subtitle
})

# Create column headers as empty plots with titles
header_ind  <- ggplot() + theme_void() + ggtitle("Ind")  + theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"))
header_drug <- ggplot() + theme_void() + ggtitle("Drug") + theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"))
header_conc <- ggplot() + theme_void() + ggtitle("Conc") + theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"))
header_time <- ggplot() + theme_void() + ggtitle("Time") + theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"))

# Assemble the full 4-column layout with 3 PC rows
final_plot <- (
  (header_ind | header_drug | header_conc | header_time) /
    (plots[["PC1_Ind"]] | plots[["PC1_Drug"]] | plots[["PC1_Conc"]] | plots[["PC1_Time"]]) /
    (plots[["PC2_Ind"]] | plots[["PC2_Drug"]] | plots[["PC2_Conc"]] | plots[["PC2_Time"]]) /
    (plots[["PC3_Ind"]] | plots[["PC3_Drug"]] | plots[["PC3_Conc"]] | plots[["PC3_Time"]])
) + plot_layout(heights = c(0.07, 1, 1, 1))  # header height

# Show the plot
print(final_plot)

```

## **📌PCA of Filtered log2(CPM) (RowMeans > 0.5)**

```{r pca_filtered_0.5, echo=TRUE, message=FALSE}
prcomp_res2 <- prcomp(t(filcpm_matrix1 %>% as.matrix()), center = TRUE)

ggplot2::autoplot(prcomp_res2, data = Metadata, colour = "Condition", shape = "Time", size =4, x=1, y=2) +
  ggrepel::geom_text_repel(label=Indiv) +
  scale_color_manual(values=drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) filtered (Rowmeans >0.5)")) +
  theme_bw()
```


```{r pca_filtered_0.5_1, echo=TRUE, message=FALSE}
prcomp_res2 <- prcomp(t(filcpm_matrix1 %>% as.matrix()), center = TRUE)

ggplot2::autoplot(prcomp_res2, data = Metadata, colour = "Condition", shape = "Time", size =4, x=2, y=3) +
  ggrepel::geom_text_repel(label=Indiv) +
  scale_color_manual(values=drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) filtered (Rowmeans >0.5)")) +
  theme_bw()
```


```{r pca_filtered_0.5_2, echo=TRUE, message=FALSE}
prcomp_res2 <- prcomp(t(filcpm_matrix1 %>% as.matrix()), center = TRUE)

ggplot2::autoplot(prcomp_res2, data = Metadata, colour = "Condition", shape = "Time", size =4, x=3, y=4) +
  ggrepel::geom_text_repel(label=Indiv) +
  scale_color_manual(values=drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) filtered (Rowmeans >0.5)")) +
  theme_bw()
```


## **📌PCA of Filtered log2(CPM) (RowMeans > 1)**

```{r pca_filtered_1, echo=TRUE, message=FALSE}
prcomp_res3 <- prcomp(t(filcpm_matrix2 %>% as.matrix()), center = TRUE)

ggplot2::autoplot(prcomp_res3, data = Metadata, colour = "Condition", shape = "Time", size =4, x=1, y=2) +
  ggrepel::geom_text_repel(label=Indiv) +
  scale_color_manual(values=drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) filtered (Rowmeans >1)")) +
  theme_bw()
```


```{r pca_filtered_1_1, echo=TRUE, message=FALSE}
prcomp_res3 <- prcomp(t(filcpm_matrix2 %>% as.matrix()), center = TRUE)

ggplot2::autoplot(prcomp_res3, data = Metadata, colour = "Condition", shape = "Time", size =4, x=2, y=3) +
  ggrepel::geom_text_repel(label=Indiv) +
  scale_color_manual(values=drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) filtered (Rowmeans >1)")) +
  theme_bw()
```


```{r pca_filtered_1_2, echo=TRUE, message=FALSE}
prcomp_res3 <- prcomp(t(filcpm_matrix2 %>% as.matrix()), center = TRUE)

ggplot2::autoplot(prcomp_res3, data = Metadata, colour = "Condition", shape = "Time", size =4, x=3, y=4) +
  ggrepel::geom_text_repel(label=Indiv) +
  scale_color_manual(values=drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) filtered (Rowmeans >1)")) +
  theme_bw()
```


## **📌 PCA Analysis by Drugs**

## **📌 PCA Analysis: CX-5461 & VEH **
```{r pca_drug, echo=TRUE, message=FALSE}
selected_columns <- grepl("VEH|CX.5461", colnames(matrix))
subset_matrix_CX <- matrix[, selected_columns]

subset_meta <- subset(Metadata, Metadata$Drug %in% c("VEH", "CX-5461"))

prcomp_res4 <- prcomp(t(subset_matrix_CX), center = TRUE)

ggplot2::autoplot(prcomp_res4, data = as.data.frame(subset_meta), colour = "Condition", shape = "Time", size = 4) +
  ggrepel::geom_text_repel(label = subset_meta$Ind) +  # ✅ Corrected label
  scale_color_manual(values = drug_palc1) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Unfiltered (CX-5461 vs VEH)")) +
  theme_bw()
```


```{r pca_drug_1, echo=TRUE, message=FALSE}
prcomp_res5 <- prcomp(t(subset_matrix_CX[rowMeans(subset_matrix_CX) > 0, ]), center = TRUE)

ggplot2::autoplot(prcomp_res5, data = as.data.frame(subset_meta), colour = "Condition", shape = "Time", size = 4) +
  ggrepel::geom_text_repel(label = subset_meta$Ind) +
  scale_color_manual(values = drug_palc1) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0 (CX-5461 vs VEH)")) +
  theme_bw()
```

```{r pca_drug_2, echo=TRUE, message=FALSE}
prcomp_res6 <- prcomp(t(subset_matrix_CX[rowMeans(subset_matrix_CX) > 0.5, ]), center = TRUE)

ggplot2::autoplot(prcomp_res6, data = as.data.frame(subset_meta), colour = "Condition", shape = "Time", size = 4) +
  ggrepel::geom_text_repel(label = subset_meta$Ind) +
  scale_color_manual(values = drug_palc1) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0.5 (CX-5461 vs VEH)")) +
  theme_bw()
```

```{r pca_drug_3, echo=TRUE, message=FALSE}
prcomp_res7 <- prcomp(t(subset_matrix_CX[rowMeans(subset_matrix_CX) > 1, ]), center = TRUE)

ggplot2::autoplot(prcomp_res7, data = as.data.frame(subset_meta), colour = "Condition", shape = "Time", size = 4) +
  ggrepel::geom_text_repel(label = subset_meta$Ind) +
  scale_color_manual(values = drug_palc1) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 1 (CX-5461 vs VEH)")) +
  theme_bw()
```


## **📌 PCA Analysis: DOX & VEH **

```{r pca_drug_dox_veh, echo=TRUE, message=FALSE}
selected_columns <- grepl("VEH|DOX", colnames(matrix))
subset_matrix_DOX <- matrix[, selected_columns]

subset_meta_dox <- subset(Metadata, Metadata$Drug %in% c("VEH", "DOX"))

prcomp_res8 <- prcomp(t(subset_matrix_DOX), center = TRUE)

ggplot2::autoplot(prcomp_res8, data = as.data.frame(subset_meta_dox), colour = "Condition", shape = "Time", size = 4) +
  ggrepel::geom_text_repel(label = subset_meta_dox$Ind) +
  scale_color_manual(values = drug_palc1) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Unfiltered (DOX vs VEH)")) +
  theme_bw()
```


```{r pca_drug_dox_veh_1, echo=TRUE, message=FALSE}
prcomp_res9 <- prcomp(t(subset_matrix_DOX[rowMeans(subset_matrix_DOX) > 0, ]), center = TRUE)

ggplot2::autoplot(prcomp_res9, data = as.data.frame(subset_meta_dox), colour = "Condition", shape = "Time", size = 4) +
  ggrepel::geom_text_repel(label = subset_meta_dox$Ind) +
  scale_color_manual(values = drug_palc1) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0 (DOX vs VEH)")) +
  theme_bw()
```


```{r pca_drug_dox_veh_2, echo=TRUE, message=FALSE}
prcomp_res10 <- prcomp(t(subset_matrix_DOX[rowMeans(subset_matrix_DOX) > 0.5, ]), center = TRUE)

ggplot2::autoplot(prcomp_res10, data = as.data.frame(subset_meta_dox), colour = "Condition", shape = "Time", size = 4) +
  ggrepel::geom_text_repel(label = subset_meta_dox$Ind) +
  scale_color_manual(values = drug_palc1) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0.5 (DOX vs VEH)")) +
  theme_bw()
```



```{r pca_drug_dox_veh_3, echo=TRUE, message=FALSE}
prcomp_res11 <- prcomp(t(subset_matrix_DOX[rowMeans(subset_matrix_DOX) > 1, ]), center = TRUE)

ggplot2::autoplot(prcomp_res11, data = as.data.frame(subset_meta_dox), colour = "Condition", shape = "Time", size = 4) +
  ggrepel::geom_text_repel(label = subset_meta_dox$Ind) +
  scale_color_manual(values = drug_palc1) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 1 (DOX vs VEH)")) +
  theme_bw()
```


## **📌 PCA Analysis: CX-5461 & DOX**

```{r pca_drug_cx_dox, echo=TRUE, message=FALSE}
selected_columns <- grepl("CX.5461|DOX", colnames(matrix))
subset_matrix_CX_DOX <- matrix[, selected_columns]

subset_meta_cx_dox <- subset(Metadata, Metadata$Drug %in% c("CX-5461", "DOX"))

prcomp_res12 <- prcomp(t(subset_matrix_CX_DOX), center = TRUE)

ggplot2::autoplot(prcomp_res12, data = as.data.frame(subset_meta_cx_dox), colour = "Condition", shape = "Time", size = 4) +
  ggrepel::geom_text_repel(label = subset_meta_cx_dox$Ind) +
  scale_color_manual(values = drug_palc1) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Unfiltered (CX-5461 vs DOX)")) +
  theme_bw()
```


```{r pca_drug_cx_dox_1, echo=TRUE, message=FALSE}
prcomp_res13 <- prcomp(t(subset_matrix_CX_DOX[rowMeans(subset_matrix_CX_DOX) > 0, ]), center = TRUE)

ggplot2::autoplot(prcomp_res13, data = as.data.frame(subset_meta_cx_dox), colour = "Condition", shape = "Time", size = 4) +
  ggrepel::geom_text_repel(label = subset_meta_cx_dox$Ind) +
  scale_color_manual(values = drug_palc1) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0 (CX-5461 vs DOX)")) +
  theme_bw()
```



```{r pca_drug_cx_dox_2, echo=TRUE, message=FALSE}
prcomp_res14 <- prcomp(t(subset_matrix_CX_DOX[rowMeans(subset_matrix_CX_DOX) > 0.5, ]), center = TRUE)

ggplot2::autoplot(prcomp_res14, data = as.data.frame(subset_meta_cx_dox), colour = "Condition", shape = "Time", size = 4) +
  ggrepel::geom_text_repel(label = subset_meta_cx_dox$Ind) +
  scale_color_manual(values = drug_palc1) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0.5 (CX-5461 vs DOX)")) +
  theme_bw()
```


```{r pca_drug_cx_dox_3, echo=TRUE, message=FALSE}
prcomp_res15 <- prcomp(t(subset_matrix_CX_DOX[rowMeans(subset_matrix_CX_DOX) > 1, ]), center = TRUE)

ggplot2::autoplot(prcomp_res15, data = as.data.frame(subset_meta_cx_dox), colour = "Condition", shape = "Time", size = 4) +
  ggrepel::geom_text_repel(label = subset_meta_cx_dox$Ind) +
  scale_color_manual(values = drug_palc1) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 1 (CX-5461 vs DOX)")) +
  theme_bw()
```

## **📌 PCA Analysis by Timepoints**

## **📌 3-Hour Timepoint**

```{r pca_3hr, echo=TRUE, message=FALSE}
# 🎯 3-Hour PCA Subset
Metadata$Drug_Conc <- paste(Metadata$Drug, Metadata$Conc, sep = "_")

selected_columns <- grepl("_3", colnames(matrix))
subset_matrix_3hr <- matrix[, selected_columns]
subset_meta_3hr <- subset(Metadata, Metadata$Time == "3hr")

# Filter genes by rowMeans > 0 for 3hr samples
filtered_matrix_3hr <- subset_matrix_3hr[rowMeans(subset_matrix_3hr) > 0, ]

if (nrow(filtered_matrix_3hr) > 2) {
  # Run PCA
  prcomp_res_3hr <- prcomp(t(filtered_matrix_3hr), center = TRUE)

  # Combine PCA coordinates with metadata
  pca_3hr_df <- as.data.frame(prcomp_res_3hr$x[, 1:2])  # PC1 and PC2
  pca_3hr_df$Ind <- subset_meta_3hr$Ind
  pca_3hr_df$Drug <- subset_meta_3hr$Drug
  pca_3hr_df$Drug_Conc <- subset_meta_3hr$Drug_Conc

  # Plot
  ggplot(pca_3hr_df, aes(x = PC1, y = PC2, colour = Drug_Conc, shape = Drug)) +
    geom_point(size = 4) +
    ggrepel::geom_text_repel(aes(label = Ind)) +
    scale_color_manual(values = drug_conc_palette) +
    ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0 (3 Hours)")) +
    theme_bw()
} else {
  print("No genes passed the rowMeans > 0 filter for 3hr samples.")
}
```

```{r pca_3hr_2, echo=TRUE, message=FALSE}
prcomp_res_3hr <- prcomp(t(subset_matrix_3hr[rowMeans(subset_matrix_3hr) > 0, ]), center = TRUE)

ggplot2::autoplot(prcomp_res_3hr, data = as.data.frame(subset_meta_3hr), colour = "Condition", shape = "Drug", size = 4, x=2, y=3) +
  ggrepel::geom_text_repel(label = subset_meta_3hr$Ind) +
  scale_color_manual(values = drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0 (3 Hours)")) +
  theme_bw()
```


```{r pca_3hr_3, echo=TRUE, message=FALSE}
prcomp_res_3hr <- prcomp(t(subset_matrix_3hr[rowMeans(subset_matrix_3hr) > 0, ]), center = TRUE)

ggplot2::autoplot(prcomp_res_3hr, data = as.data.frame(subset_meta_3hr), colour = "Condition", shape = "Drug", size = 4, x=3, y=4) +
  ggrepel::geom_text_repel(label = subset_meta_3hr$Ind) +
  scale_color_manual(values = drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0 (3 Hours)")) +
  theme_bw()
```


## **📌 24-Hour Timepoint**

```{r pca_24hr, echo=TRUE, message=FALSE}
# 📌 Subset for 24hr samples
selected_columns <- grepl("_24", colnames(matrix))
subset_matrix_24hr <- matrix[, selected_columns]
subset_meta_24hr <- subset(Metadata, Metadata$Time == "24hr")  # match your relabeled timepoints

# 📌 Filter low-expression genes
filtered_matrix_24hr <- subset_matrix_24hr[rowMeans(subset_matrix_24hr) > 0, ]

# 📌 Run PCA if genes remain
if (nrow(filtered_matrix_24hr) > 2) {
  prcomp_res_24hr <- prcomp(t(filtered_matrix_24hr), center = TRUE)

  ggplot2::autoplot(prcomp_res_24hr, data = as.data.frame(subset_meta_24hr),
                    colour = "Drug_Conc", shape = "Drug", size = 4) +
    ggrepel::geom_text_repel(aes(label = Ind)) +
    scale_color_manual(values = drug_conc_palette) +
    ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0 (24 Hours)")) +
    theme_bw()
} else {
  message("⚠️ No genes passed the rowMeans > 0 filter for 24-hour samples.")
}
```


```{r pca_24hr_1, echo=TRUE, message=FALSE}
prcomp_res_24hr <- prcomp(t(subset_matrix_24hr[rowMeans(subset_matrix_24hr) > 0, ]), center = TRUE)

ggplot2::autoplot(prcomp_res_24hr, data = as.data.frame(subset_meta_24hr), colour = "Condition", shape = "Drug", size = 4, x=2, y=3) +
  ggrepel::geom_text_repel(label = subset_meta_24hr$Ind) +
  scale_color_manual(values = drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0 (24 Hours)")) +
  theme_bw()
```


```{r pca_24hr_2, echo=TRUE, message=FALSE}
prcomp_res_24hr <- prcomp(t(subset_matrix_24hr[rowMeans(subset_matrix_24hr) > 0, ]), center = TRUE)

ggplot2::autoplot(prcomp_res_24hr, data = as.data.frame(subset_meta_24hr), colour = "Condition", shape = "Drug", size = 4, x=3, y=4) +
  ggrepel::geom_text_repel(label = subset_meta_24hr$Ind) +
  scale_color_manual(values = drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0 (24 Hours)")) +
  theme_bw()
```



## **📌 48-Hour Timepoint**
```{r pca_48hr, echo=TRUE, message=FALSE}
# 📌 Subset for 48hr samples
selected_columns <- grepl("_48", colnames(matrix))
subset_matrix_48hr <- matrix[, selected_columns]
subset_meta_48hr <- subset(Metadata, Metadata$Time == "48hr")  # must match relabeled levels

# 📌 Filter low-expression genes
filtered_matrix_48hr <- subset_matrix_48hr[rowMeans(subset_matrix_48hr) > 0, ]

# 📌 Run PCA only if data is valid
if (nrow(filtered_matrix_48hr) > 2) {
  prcomp_res_48hr_1 <- prcomp(t(filtered_matrix_48hr), center = TRUE)

  ggplot2::autoplot(prcomp_res_48hr_1, data = as.data.frame(subset_meta_48hr),
                    colour = "Drug_Conc", shape = "Drug", size = 4) +
    ggrepel::geom_text_repel(aes(label = Ind)) +
    scale_color_manual(values = drug_conc_palette) +
    ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0 (48 Hours)")) +
    theme_bw()
} else {
  message("⚠️ No genes passed the rowMeans > 0 filter for 48-hour samples.")
}
```


```{r pca_48hr_1, echo=TRUE, message=FALSE}

prcomp_res_48hr_1 <- prcomp(t(subset_matrix_48hr[rowMeans(subset_matrix_48hr) > 0, ]), center = TRUE)

ggplot2::autoplot(prcomp_res_48hr_1, data = as.data.frame(subset_meta_48hr), colour = "Condition", shape = "Drug", size = 4, x=2, y=3) +
  ggrepel::geom_text_repel(label = subset_meta_48hr$Ind) +
  scale_color_manual(values = drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0 (48 Hours)")) +
  theme_bw()
```


```{r pca_48hr_2, echo=TRUE, message=FALSE}

prcomp_res_48hr_1 <- prcomp(t(subset_matrix_48hr[rowMeans(subset_matrix_48hr) > 0, ]), center = TRUE)

ggplot2::autoplot(prcomp_res_48hr_1, data = as.data.frame(subset_meta_48hr), colour = "Condition", shape = "Drug", size = 4, x=3, y=4) +
  ggrepel::geom_text_repel(label = subset_meta_48hr$Ind) +
  scale_color_manual(values = drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0 (48 Hours)")) +
  theme_bw()
```


## **📌 PCA Analysis by Concentrations**

## **📌0.1 µM Concentration**

```{r pca_0.1, echo=TRUE, message=FALSE}
selected_columns <- grepl("_0.1_", colnames(matrix))
subset_matrix_0.1 <- matrix[, selected_columns]

subset_meta_0.1 <- subset(Metadata, Metadata$Conc. == 0.1)

prcomp_res_0.1 <- prcomp(t(subset_matrix_0.1[rowMeans(subset_matrix_0.1) > 0, ]), center = TRUE)

ggplot2::autoplot(prcomp_res_0.1, data = as.data.frame(subset_meta_0.1), colour = "Drug", shape = "Time", size = 4) +
  ggrepel::geom_text_repel(label = subset_meta_0.1$Ind) +
  scale_color_manual(values = drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0 (0.1 µM)")) +
  theme_bw()
```


```{r pca_0.1_1, echo=TRUE, message=FALSE}
prcomp_res_0.1 <- prcomp(t(subset_matrix_0.1[rowMeans(subset_matrix_0.1) > 0, ]), center = TRUE)

ggplot2::autoplot(prcomp_res_0.1, data = as.data.frame(subset_meta_0.1), colour = "Drug", shape = "Time", size = 4, x=2, y=3) +
  ggrepel::geom_text_repel(label = subset_meta_0.1$Ind) +
  scale_color_manual(values = drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0 (0.1 µM)")) +
  theme_bw()
```


```{r pca_0.1_2, echo=TRUE, message=FALSE}
prcomp_res_0.1 <- prcomp(t(subset_matrix_0.1[rowMeans(subset_matrix_0.1) > 0, ]), center = TRUE)

ggplot2::autoplot(prcomp_res_0.1, data = as.data.frame(subset_meta_0.1), colour = "Drug", shape = "Time", size = 4, x=3, y=4) +
  ggrepel::geom_text_repel(label = subset_meta_0.1$Ind) +
  scale_color_manual(values = drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0 (0.1 µM)")) +
  theme_bw()
```


## **📌0.5 µM Concentration**
```{r pca_0.5, echo=TRUE, message=FALSE}
selected_columns <- grepl("_0.5_", colnames(matrix))
subset_matrix_0.5 <- matrix[, selected_columns]

subset_meta_0.5 <- subset(Metadata, Metadata$Conc. == 0.5)

prcomp_res_0.5 <- prcomp(t(subset_matrix_0.5[rowMeans(subset_matrix_0.5) > 0, ]), center = TRUE)

ggplot2::autoplot(prcomp_res_0.5, data = as.data.frame(subset_meta_0.5), colour = "Drug", shape = "Time", size = 4) +
  ggrepel::geom_text_repel(label = subset_meta_0.5$Ind) +
  scale_color_manual(values = drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0 (0.5 µM)")) +
  theme_bw()
```


```{r pca_0.5_1, echo=TRUE, message=FALSE}
prcomp_res_0.5 <- prcomp(t(subset_matrix_0.5[rowMeans(subset_matrix_0.5) > 0, ]), center = TRUE)

ggplot2::autoplot(prcomp_res_0.5, data = as.data.frame(subset_meta_0.5), colour = "Drug", shape = "Time", size = 4, x=2, y=3) +
  ggrepel::geom_text_repel(label = subset_meta_0.5$Ind) +
  scale_color_manual(values = drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0 (0.5 µM)")) +
  theme_bw()
```



```{r pca_0.5_2, echo=TRUE, message=FALSE}
prcomp_res_0.5 <- prcomp(t(subset_matrix_0.5[rowMeans(subset_matrix_0.5) > 0, ]), center = TRUE)

ggplot2::autoplot(prcomp_res_0.5, data = as.data.frame(subset_meta_0.5), colour = "Drug", shape = "Time", size = 4, x=3, y=4) +
  ggrepel::geom_text_repel(label = subset_meta_0.5$Ind) +
  scale_color_manual(values = drug_palc) +
  ggtitle(expression("PCA of log"[2]*"(cpm) Filtered rowMeans > 0 (0.5 µM)")) +
  theme_bw()
```


